### 2018-12-17 rrbiome.Rmd
rmarkdown

---
title: "Microbiome and hypertension"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-12-17 init.R

@@ -16,3 +16,6 @@ install.packages("dplyr")
 
 install.packages("ggplot2")
 install.packages("tidyr")
+
+install.packages("Rcpp")
+install.packages("rmarkdown")

### 2018-12-18 rrbiome.Rmd
Packages installed

@@
 
-## R Markdown
+```{r libraries}
+library(dplyr)
+library(phyloseq)
+library(microbiome)
+```
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+```{r data}
+pseq <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+```
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+Pseq has readcount `r readcount(pseq)`.
 
 ```{r cars}
-summary(cars)
+print(pseq)
+summarize_phyloseq(pseq)
+meta <- meta(pseq)
+taxonomy <- tax_table(pseq)
+otu.absolute <- abundances(pseq)
+otu.relative <- abundances(pseq, "compositional")
+
+df <- psmelt(pseq)
+head(sample_names(pseq))
+sample_sums(pseq)
+s <- rownames(subset(meta(pseq), nationality == "AFR" & sex == "Female"))
+prune_samples(s, pseq)
+
+sample_variables(pseq)
+
+# Calculate diversity for samples
+div <- alpha(pseq, index = "shannon")
+
+# Assign the estimated diversity to sample metadata
+sample_data(pseq)$diversity <- div
+
+n <- ntaxa(pseq)
+topx <- top_taxa(pseq, n = 10)
+ranks <- rank_names(pseq)  # Taxonomic levels
+taxa  <- taxa(pseq)        # Taxa names at the analysed level
+pseq.bac <- subset_taxa(pseq, Phylum == "Bacteroidetes")
+# List of Genera in the Bacteroideted Phylum
+taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacteroidetes
+
+# With given taxon names
+ex2 <- prune_taxa(taxa, pseq)
+
+# Taxa with positive sum across samples
+ex3 <- prune_taxa(taxa_sums(pseq) > 0 , pseq)
+
+head(get_taxa_unique(pseq, "Phylum"))
+
+samplename <- sample_names(pseq)[[1]]
+
+# Pick abundances for a particular taxon
+tax.abundances <- abundances(pseq)[, samplename]
 ```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-12-19 rrbiome.Rmd
Beta diversity

--- 2018-12-17 rrbiome.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, results='asis', cache=TRUE, warning=FALSE)
+knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
+                      file.path = 'cache/', fig.path = 'cache/')
 ```
 
-## R Markdown
+Importing libraries
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+```{r libraries}
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+library(dplyr)
+library(tibble)
+library(phyloseq)
+library(microbiome)
+library(finriskmetagcommon)
+library(knitr)
+library(tidyr)
+library(vegan)
 
-```{r cars}
-summary(cars)
 ```
 
-## Including Plots
+```{r functions, include = FALSE}
 
-You can also embed plots, for example:
+print.dset <- function(dset) {
+    dset %>%
+        gather(Covariate, Value) %>%
+        group_by(Covariate) %>%
+        summarise_all(., .funs = funs(mean(as.numeric(.), na.rm = TRUE),
+                                         median(as.numeric(.), na.rm = TRUE),
+                                         sd(as.numeric(.), na.rm = TRUE),
+                                         min(as.numeric(.), na.rm = TRUE),
+                                         max(as.numeric(.), na.rm = TRUE),
+                                         NAs = sum(is.na(.)),
+                                         n = n(),
+                                         n_distinct)) %>%
+            mutate_if(is.numeric, round, 2) %>%
+            kable
+}
 
-```{r pressure, echo=FALSE}
-plot(pressure)
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+Loading phyloseq object
+
+```{r data}
+
+pseq <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+
+```
+
+Loading descriptions for clinical data
+
+```{r variables}
+
+if (!exists('phfinrisk_metadatadesc')) {
+  # this is a hand-compiled data.frame that contains metadata information for variables utilized in our analyses
+  data("phfinrisk_metadatadesc", package="finriskmetagcommon")
+}
+names.dset <- filter(phfinrisk_metadatadesc, Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
+
+```
+
+Filtering clinical data
+
+```{r filtering}
+
+dset <- meta(pseq) %>%
+    tibble::rownames_to_column(var = "rowname") %>%
+    dplyr::select(rowname, DNA_OK, MEN, BL_AGE, SYSTM, DIASM, BP_TREAT, BMI, CURR_SMOKE, DIAB, BL_USE_RX_C09,
+        Q57X, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
+    dplyr::mutate(PULSEPRESSURE = SYSTM - DIASM,
+                  HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | BP_TREAT == 1, 1, 0)),
+                  SEX = factor(ifelse(MEN == "Female", 1, 0)),
+                  MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
+    dplyr::select(-MEN) %>%
+    stats::na.omit() %>%
+    tibble::remove_rownames() %>%
+    tibble::column_to_rownames(var = "rowname")
+
+print.dset(dset)
+
+```
+
+Continuous variables
+
+```{r distribution plots, fig.width=10, fig.height=4}
+ggplot((dset %>%
+        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP) %>%
+        gather(Covariate, Value)), aes_string(x = "Value")) +
+    facet_wrap(~Covariate, ncol=4, scales = "free") +
+    geom_density(aes(y=..scaled..)) + 
+    ggplot2::theme_minimal() 
+```
+
+Categorical variables
+
+```{r histogramsy, fig.width=10, fig.height=4}
+ggplot((dset %>%
+        select(-BL_AGE, -BMI, -SYSTM, -DIASM, -PULSEPRESSURE, -MAP) %>%
+        gather(Covariate, Value)), aes_string(x = "Value")) +
+    facet_wrap(~Covariate, ncol=8, scales = "free") +
+    geom_bar() + 
+    ggplot2::theme_minimal() 
+```
+
+```{r Updating meta data}
+
+pseq_filtered <- pseq
+sample_data(pseq_filtered) <- sample_data(dset)
+
+```
+
+# Alpha diversity
+
+Dset with alpha diversity
+
+```{r alpha diversity}
+
+alphadiversity <- microbiome::global(pseq_filtered, index="shannon")
+dset.alpha <- merge(dset, alphadiversity, by=0, all=TRUE)
+
+```
+
+```{r alpha diversity scatterplot, fig.width=8, fig.height=4}
+
+ggplot(dset.alpha, aes(x=diversities_shannon, y=MAP)) +
+    geom_point(size=1, shape=1) +
+    coord_fixed() +
+    ggplot2::theme_minimal() 
+```
+
+Linear models
+
+```{r lm model}
+lm(MAP ~ diversities_shannon, data=dset.alpha) %>%
+    summary(.) %>% coef %>% kable(., digits=4)
+lm(MAP ~ BL_AGE + SEX + diversities_shannon, data=dset.alpha) %>%
+    summary(.) %>% coef %>% kable(., digits=4)
+lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT + diversities_shannon, data=dset.alpha) %>%
+    summary(.) %>% coef %>% kable(., digits=4)
+
+```
+
+# Beta diversity
+
+Matriisin teko _bc_dissimilarity_matrix.R ja käyttö comp_adonis_covariates.R
+
+```{r vegdist}
+
+phf_profile.rel <- microbiome::transform(pseq_filtered, 'compositional')
+otu <- microbiome::abundances(phf_profile.rel)
+
+if (!file.exists("rds/bray.dist.m.rds")) {
+    bray.dist.m <- vegdist(t(otu), method="bray")
+    saveRDS(bray.dist.m, file = "rds/bray.dist.m.rds")
+} else {
+    bray.dist.m <- readRDS("rds/bray.dist.m.rds")
+}
+
+```
+
+```{r adonis}
+
+attr(bray.dist.m, "method")
+dist.matrix <- as.matrix(bray.dist.m)
+attr(dist.matrix, "method") <- "bray"
+
+if (!file.exists("rds/phf.adonis.rds")) {
+phf.adonis <- adonis(dist.matrix ~ BL_AGE + SEX + MAP,
+                     data = meta(pseq_filtered),
+                     permutations = 99)
+    saveRDS(phf.adonis, file = "rds/phf.adonis.rds")
+} else {
+    phf.adonis <- readRDS("rds/phf.adonis.rds")
+}
+
+phf.adonis$aov.tab %>% as.data.frame %>% kable
+
+```
+
+```{r beta heatmap, eval = FALSE}
+
+ggplot(data = dist.matrix, aes(x = metrics, y = teams)) +
+      geom_tile(aes(fill = performance))
+
+```

### 2018-12-20 rrbiome.Rmd
PCA-plots

@@
 library(vegan)
+library(reshape)
+library(ggpmisc)
 
@@
 
-if (!exists('phfinrisk_metadatadesc')) {
-  # this is a hand-compiled data.frame that contains metadata information for variables utilized in our analyses
 data("phfinrisk_metadatadesc", package="finriskmetagcommon")
-}
-names.dset <- filter(phfinrisk_metadatadesc, Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
+names.dset <- filter(phfinrisk_metadatadesc,
+                     Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
 
@@
 
-```{r alpha diversity scatterplot, fig.width=8, fig.height=4}
+```{r alpha diversity scatterplot, fig.width=10, fig.height=4}
 
-ggplot(dset.alpha, aes(x=diversities_shannon, y=MAP)) +
-    geom_point(size=1, shape=1) +
-    coord_fixed() +
+ggplot((dset.alpha %>%
+        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP, SEX, diversities_shannon) %>%
+        gather(Covariate, Value, -SEX, -diversities_shannon)), 
+       aes(x=diversities_shannon, y=Value)) +
+    facet_wrap(~Covariate, ncol=4, scales = "free") +
+    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
+    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
+    stat_poly_eq(formula = "y ~ x",
+                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
+                 label.y = 0,
+                 parse = TRUE,
+                 size = 3) +   
+    ylim(0, 200) +
     ggplot2::theme_minimal()
+
 ```
@@
 
-Matriisin teko _bc_dissimilarity_matrix.R ja käyttö comp_adonis_covariates.R
+Beetadiversiteetti-matriisin lasku. Tiedosista comp_bc_dissimilarity_matrix.R ja käyttö comp_adonis_covariates.R.
 
@@
 
+dist.matrix <- as.matrix(bray.dist.m)
+attr(dist.matrix, "method") <- "bray"
+
 ```
 
+```{r beta hclust, fig.width = 5, fig.height = 4, dpi = 300}
+gplots::heatmap.2(dist.matrix,
+                  distfun = dist,
+                  trace = "none",
+                  hclustfun = hclust,
+                  key.title = NA,
+                  dendrogram = "column",
+                  density.info = "none",
+                  labRow = NA,
+                  labCol = NA)
+```
+
+Mallin lasku Adoniksella
+
 ```{r adonis}
-
-attr(bray.dist.m, "method")
-dist.matrix <- as.matrix(bray.dist.m)
-attr(dist.matrix, "method") <- "bray"
 
@@
 
-```{r beta heatmap, eval = FALSE}
+# PCA
 
-ggplot(data = dist.matrix, aes(x = metrics, y = teams)) +
-      geom_tile(aes(fill = performance))
+```{r pca}
+pseq.clr <- microbiome::transform(pseq_filtered, "clr")
+X <- as.matrix(t(abundances(pseq.clr)))
+prcomp <- stats::prcomp(X)
+```
+
+```{r pca labels}
+percentage <- round(prcomp$sdev/sum(prcomp$sdev) * 100, 2)
+pca.labels <- paste0(colnames(prcomp$x), " (", paste0(as.character(percentage), "%", ")"))
+```
+
+```{r PCA plots categorical, fig.width = 10, fig.height = 10, dpi = 300}
+ggplot((cbind(meta(pseq_filtered), prcomp$x[,1:2]) %>%
+        select(PC1, PC2, BP_TREAT, CURR_SMOKE, DIAB, BL_USE_RX_C09, Q57X,
+               BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08, HYPERTENSION, SEX) %>%
+        gather(Covariate, Value, -PC1, -PC2)), 
+       aes(x=PC1, y=PC2)) +
+    facet_wrap(~Covariate, ncol=2, scales = "free") +
+    geom_point(aes(color = Value), size = 0.5) +
+    scale_color_brewer(palette="Dark2") +
+    xlab(pca.labels[1]) +
+    ylab(pca.labels[2]) +
+    ggplot2::theme_minimal()
 
 ```
+
+```{r PCA plots continuous, fig.width = 10, fig.height = 10, dpi = 300}
+
+ggplot((cbind(meta(pseq_filtered), prcomp$x[,1:2]) %>%
+        select(PC1, PC2, BL_AGE, SYSTM, DIASM, BMI, MAP, PULSEPRESSURE) %>%
+        gather(Covariate, Value, -PC1, -PC2)), 
+       aes(x=PC1, y=PC2)) +
+    facet_wrap(~Covariate, ncol=2, scales = "free") +
+    geom_point(aes(color = Value), size = 0.5) +
+    scale_color_gradientn(colours = rainbow(12)) +
+    xlab(pca.labels[1]) +
+    ylab(pca.labels[2]) +
+    ggplot2::theme_minimal()
+
+```

### 2018-12-20 rrbiome.Rmd
First maaslin calculation

@@
 library(ggpmisc)
+library(Maaslin)
 
@@
 
-```{r adonis}
-
-if (!file.exists("rds/phf.adonis.rds")) {
-    phf.adonis <- adonis(dist.matrix ~ BL_AGE + SEX + MAP,
+```{r adonis1}
+adonis(dist.matrix ~ MAP, 
        data = meta(pseq_filtered),
-                         permutations = 99)
-    saveRDS(phf.adonis, file = "rds/phf.adonis.rds")
-} else {
-    phf.adonis <- readRDS("rds/phf.adonis.rds")
-}
-
-phf.adonis$aov.tab %>% as.data.frame %>% kable
-
+       permutations = 99)$aov.tab %>%
+                            as.data.frame %>%
+                            kable
 ```
 
-# PCA
+```{r adonis2}
+adonis(dist.matrix ~ BL_AGE + SEX + MAP,
+       data = meta(pseq_filtered),
+       permutations = 99)$aov.tab %>%
+                            as.data.frame %>%
+                            kable
+```
 
-```{r pca}
-pseq.clr <- microbiome::transform(pseq_filtered, "clr")
-X <- as.matrix(t(abundances(pseq.clr)))
-prcomp <- stats::prcomp(X)
+```{r adonis3}
+adonis(dist.matrix ~ MAP + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT,
+       data = meta(pseq_filtered),
+       permutations = 99)$aov.tab %>%
+                            as.data.frame %>%
+                            kable
 ```
@@
 ```
+
+# Genus
+
+Aaron work-december-2018 prepare_maaslin_input.R
+
+```{r prepare maaslin}
+
+pheno <- t(meta(sample_data(pseq_filtered)))
+sample.ids <- colnames(pheno)
+phfg.pseq.rel <- microbiome::transform(pseq_filtered, "compositional")
+abud.rel <- abundances(phfg.pseq.rel)
+
+rownames(abud.rel) <- sapply(rownames(abud.rel),
+                             function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
+
+merged <- rbind(sample.ids, pheno, abud.rel)
+write.table(t(merged), file="merged_phfinrisk_genus.tsv", sep='\t', 
+            quote=FALSE, row.names=FALSE, col.names=TRUE)
+
+variables <- rownames(pheno)
+variables_wo_ctrls <- variables[!variables %in% c("BL_AGE","BMI","SEX")]
+data <- file("maaslin_config.config" , open='wt')
+pcl.contents <- c("Matrix: Metadata",
+                  paste(c("Read_PCL_Rows: BL_AGE,BMI,SEX",variables_wo_ctrls), collapse=","),
+                  "",
+                  "Matrix: Adundance",
+                  "Read_PCL_Rows: Archaea.Archaea-")
+writeLines(pcl.contents, con=data)
+close(data)
+```
+
+```{r run maaslin, eval = FALSE} 
+Maaslin("merged_phfinrisk_genus.tsv",
+        "maaslin",
+        strInputConfig = "maaslin_config.config",
+        strModelSelection="none",
+        dSignificanceLevel = 1.0, 
+        fAllvAll = TRUE,
+        strForcedPredictors = c("BL_AGE", "BMI", "SEX"))
+
+```
+
+Data framen luku *comp_maaslin_results_from_single_job.R* ja plottaus käyttäen *comp_maaslin_plots.R* ja *edgeR_plot_functions.R*.
+
+```{r Reading maaslins output}
+
+maas.results.df <- read.table("maaslin/merged_phfinrisk_genus.txt", header=TRUE, sep='\t')
+maas.results.df[,'Covariate'] <- maas.results.df[,'Variable']
+maas.results.df <- merge(maas.results.df,
+                         names.dset[,c('Covariate','Category', 'Name')],
+                         by=c('Covariate'))
+
+maas.results.df[,'Maaslin.Q.value'] <- maas.results.df[,'Q.value']
+maas.results.df[,'Qval'] <- p.adjust(maas.results.df[,'P.value'], method="BH")
+maas.results.df[,'beta'] <- maas.results.df[,'Coefficient']
+maas.results.df.signf <- dplyr::filter(maas.results.df, Qval < 0.05)
+
+```
+
+```{r scatter plots}
+sysm.signf.features <- maas.results.df.signf %>% dplyr::filter(Covariate == "SYSTM") %>% pull(Feature)
+```
+
+```{r maaslin plots, fig.width = 10, fig.height = 4, dpi = 300}
+source("edgeR_plot_functions.R")
+
+plot_maaslin_heatmap_final(maas.results.df.signf,
+                      xlab.text="Genera",
+                      beta.maxabs=max(abs(maas.results.df.signf[,'beta'])))
+
+```

### 2018-12-20 edgeR_plot_functions.R

@@ -1,0 +1,384 @@
+library(reshape2)
+library(scales)
+library(ggpubr)
+# plot heatmap output of edgeR heatmap
+plot_edgeR_fc_heatmap <- function (tables.df, xlab.text = "Phyla", logFC.maxabs=NULL) {
+
+  if (nrow(tables.df) == 0) {
+    cat("empty heatmap\n")
+    p <- ggplot(data.frame()) + geom_blank()
+    return(p)
+  }
+
+  p <- ggplot(tables.df, aes(x = OTU, y = Name, fill=logFC)) + geom_tile()
+
+  if (is.null(logFC.maxabs)) {
+    p <- p + scale_fill_distiller(palette = "Spectral", name = 'log fold-change')
+  } else {
+    p <- p + scale_fill_distiller(palette = "Spectral", name = 'log fold-change',
+                                  limits=c(-logFC.maxabs, logFC.maxabs))
+  }
+  p <- p + xlab(xlab.text) + ylab("")
+  p <- p + theme_classic()
+  p <- p + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
+  p <- p + theme(axis.text.y = element_text(size=5))
+
+  p
+}
+
+plot_edgeR_fc_heatmap_reps <- function (tables.df, xlab.text = "Phyla", logFC.maxabs=NULL) {
+
+  if (nrow(tables.df) == 0) {
+    cat("empty heatmap\n")
+    p <- ggplot(data.frame()) + geom_blank()
+    return(p)
+  }
+
+  p <- ggplot(tables.df, aes(x = OTU, y = Name, fill=logFC)) + geom_tile()
+
+  p <- p + geom_text(aes(label = paste0("P ",
+                                        sprintf("%0.4f", round(PValue, digits=4)),
+                                        " rep ",
+                                        sprintf("%0.4f", round(PValue.reps, digits=4)),
+                                        "\nQ ",
+                                        sprintf("%0.4f", round(Qval, digits=4)),
+                                        " rep ",
+                                        sprintf("%0.4f", round(Qval.reps, digits=4)),
+                                        "\nlog FC ",
+                                        sprintf("%0.4f", round(logFC, digits=4)),
+                                        " rep ",
[…]

### 2018-12-20 lactobacillus.Rmd
Data exploration v1 ready

--- 2018-12-19 rrbiome.Rmd
@@
 ---
-title: "Microbiome and hypertension"
+title: "Hypertension and microbiome"
 author: "Joonatan Palmu"
@@
 library(vegan)
+library(reshape)
+library(ggpmisc)
+library(Maaslin)
 
 ```
-
-```{r functions, include = FALSE}
-
-print.dset <- function(dset) {
-    dset %>%
-        gather(Covariate, Value) %>%
-        group_by(Covariate) %>%
-        summarise_all(., .funs = funs(mean(as.numeric(.), na.rm = TRUE),
-                                         median(as.numeric(.), na.rm = TRUE),
-                                         sd(as.numeric(.), na.rm = TRUE),
-                                         min(as.numeric(.), na.rm = TRUE),
-                                         max(as.numeric(.), na.rm = TRUE),
-                                         NAs = sum(is.na(.)),
-                                         n = n(),
-                                         n_distinct)) %>%
-            mutate_if(is.numeric, round, 2) %>%
-            kable
-}
-
-```
-
-Loading phyloseq object
 
@@
 
-Loading descriptions for clinical data
-
-```{r variables}
-
-if (!exists('phfinrisk_metadatadesc')) {
-  # this is a hand-compiled data.frame that contains metadata information for variables utilized in our analyses
-  data("phfinrisk_metadatadesc", package="finriskmetagcommon")
-}
-names.dset <- filter(phfinrisk_metadatadesc, Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
+```{r}
+summarize_phyloseq(pseq_filtered)
+ntaxa(pseq_filtered)
+pseq_filtered %>% taxa() %>% sort
+phfg.pseq.rel %>% rank_names()
+subset_taxa(pseq_filtered, Genus == "Lac")
+taxa_unique(phfg.pseq.rel, "Phylum")
+abundances(phfg.pseq.rel)
 
@@
 
-Filtering clinical data
+```{r}
 
-```{r filtering}
-
-dset <- meta(pseq) %>%
-    tibble::rownames_to_column(var = "rowname") %>%
-    dplyr::select(rowname, DNA_OK, MEN, BL_AGE, SYSTM, DIASM, BP_TREAT, BMI, CURR_SMOKE, DIAB, BL_USE_RX_C09,
-        Q57X, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
-    dplyr::mutate(PULSEPRESSURE = SYSTM - DIASM,
-                  HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | BP_TREAT == 1, 1, 0)),
-                  SEX = factor(ifelse(MEN == "Female", 1, 0)),
-                  MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
-    dplyr::select(-MEN) %>%
-    stats::na.omit() %>%
-    tibble::remove_rownames() %>%
-    tibble::column_to_rownames(var = "rowname")
-
-print.dset(dset)
+taxa <- names(subset_taxa(pseq.s, Genus=="Lactobacillus"))
 
 ```
-
-Continuous variables
-
-```{r distribution plots, fig.width=10, fig.height=4}
-ggplot((dset %>%
-        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP) %>%
-        gather(Covariate, Value)), aes_string(x = "Value")) +
-    facet_wrap(~Covariate, ncol=4, scales = "free") +
-    geom_density(aes(y=..scaled..)) + 
-    ggplot2::theme_minimal() 
-```
-
-Categorical variables
-
-```{r histogramsy, fig.width=10, fig.height=4}
-ggplot((dset %>%
-        select(-BL_AGE, -BMI, -SYSTM, -DIASM, -PULSEPRESSURE, -MAP) %>%
-        gather(Covariate, Value)), aes_string(x = "Value")) +
-    facet_wrap(~Covariate, ncol=8, scales = "free") +
-    geom_bar() + 
-    ggplot2::theme_minimal() 
-```
-
-```{r Updating meta data}
-
-pseq_filtered <- pseq
-sample_data(pseq_filtered) <- sample_data(dset)
-
-```
-
-# Alpha diversity
-
-Dset with alpha diversity
-
-```{r alpha diversity}
-
-alphadiversity <- microbiome::global(pseq_filtered, index="shannon")
-dset.alpha <- merge(dset, alphadiversity, by=0, all=TRUE)
-
-```
-
-```{r alpha diversity scatterplot, fig.width=8, fig.height=4}
-
-ggplot(dset.alpha, aes(x=diversities_shannon, y=MAP)) +
-    geom_point(size=1, shape=1) +
-    coord_fixed() +
-    ggplot2::theme_minimal() 
-```
-
-Linear models
-
-```{r lm model}
-lm(MAP ~ diversities_shannon, data=dset.alpha) %>%
-    summary(.) %>% coef %>% kable(., digits=4)
-lm(MAP ~ BL_AGE + SEX + diversities_shannon, data=dset.alpha) %>%
-    summary(.) %>% coef %>% kable(., digits=4)
-lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT + diversities_shannon, data=dset.alpha) %>%
-    summary(.) %>% coef %>% kable(., digits=4)
-
-```
-
-# Beta diversity
-
-Matriisin teko _bc_dissimilarity_matrix.R ja käyttö comp_adonis_covariates.R
-
-```{r vegdist}
-
-phf_profile.rel <- microbiome::transform(pseq_filtered, 'compositional')
-otu <- microbiome::abundances(phf_profile.rel)
-
-if (!file.exists("rds/bray.dist.m.rds")) {
-    bray.dist.m <- vegdist(t(otu), method="bray")
-    saveRDS(bray.dist.m, file = "rds/bray.dist.m.rds")
-} else {
-    bray.dist.m <- readRDS("rds/bray.dist.m.rds")
-}
-
-```
-
-```{r adonis}
-
-attr(bray.dist.m, "method")
-dist.matrix <- as.matrix(bray.dist.m)
-attr(dist.matrix, "method") <- "bray"
-
-if (!file.exists("rds/phf.adonis.rds")) {
-phf.adonis <- adonis(dist.matrix ~ BL_AGE + SEX + MAP,
-                     data = meta(pseq_filtered),
-                     permutations = 99)
-    saveRDS(phf.adonis, file = "rds/phf.adonis.rds")
-} else {
-    phf.adonis <- readRDS("rds/phf.adonis.rds")
-}
-
-phf.adonis$aov.tab %>% as.data.frame %>% kable
-
-```
-
-```{r beta heatmap, eval = FALSE}
-
-ggplot(data = dist.matrix, aes(x = metrics, y = teams)) +
-      geom_tile(aes(fill = performance))
-
-```

### 2018-12-20 rrbiome.Rmd
Data exploration v1 ready

@@
 
+# PCA
+
+```{r pca}
+pseq.clr <- microbiome::transform(pseq_filtered, "clr")
+X <- as.matrix(t(abundances(pseq.clr)))
+prcomp <- stats::prcomp(X)
+```
+
 ```{r pca labels}
@@
 ```{r scatter plots}
-sysm.signf.features <- maas.results.df.signf %>% dplyr::filter(Covariate == "SYSTM") %>% pull(Feature)
+
+sysm.signf.features <- maas.results.df.signf %>%
+    dplyr::filter(Covariate == "SYSTM") %>%
+    pull(Feature) %>%
+    as.character() %>%
+    gsub("_", ".", .)
+df.merged <- t(merged) %>% as.data.frame
+```
+
+```{r maaslin scatter plot, fig.width = 10, fig.height = 20, dpi = 300}
+
+#df.merged %>% pull("Rikenella.Bacteria") %>% as.character I%>% as.numeric %>% sqrt %>% asin
+
+ggplot(df.merged %>%
+       select(sysm.signf.features, SYSTM, SEX) %>%
+       gather(Covariate, Value, sysm.signf.features) %>%
+       mutate(SYSTM = as.numeric(as.character(SYSTM)),
+              Value = as.numeric(as.character(Value))),
+       aes(x = asin(sqrt(Value)), y = SYSTM)) +
+    facet_wrap(~Covariate, ncol=2, scales = "free") +
+    geom_point(aes(color = SEX), size = 0.5) +
+    scale_y_continuous(breaks=seq(0,200,40)) +
+    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
+    stat_poly_eq(formula = "y ~ x",
+                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
+                 label.y = 0,
+                 parse = TRUE,
+                 size = 3) +   
+    ggplot2::theme_minimal()
+
 ```

### 2018-12-26 rrbiome.Rmd
Emacs

@@
 ```{r libraries}
-
 library(dplyr)
@@
 library(Maaslin)
-
 ```
@@
 ```{r functions, include = FALSE}
-
 print.dset <- function(dset) {
@@
 ```{r data}
-
 pseq <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
-
 ```
@@
 ```{r variables}
-
 data("phfinrisk_metadatadesc", package="finriskmetagcommon")
@@
                      Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
-
 ```
@@
 ```{r filtering}
-
 dset <- meta(pseq) %>%
@@
 print.dset(dset)
-
 ```
@@
 ```{r Updating meta data}
-
 pseq_filtered <- pseq
 sample_data(pseq_filtered) <- sample_data(dset)
-
 ```
@@
 ```
+
+```{r scp html}
+system('scp rrbiome.html utu:articletwo')
+```

### 2018-12-31 lactobacillus.Rmd
Emacs
V2 mod start

@@
 library(Maaslin)
-
 ```
 
-```{r data}
+```{r clear session}
+remove(list = ls())
+```
 
-pseq <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
-
+```{R data}
+pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+pseq.species <- readRDS("data/phfinrisk_species_ok_drop50k_2018-11-16.RDs")
 ```
@@
 ```{r}
-summarize_phyloseq(pseq_filtered)
-ntaxa(pseq_filtered)
-pseq_filtered %>% taxa() %>% sort
+summarize_phyloseq(pseq)
+ntaxa(pseq)
+pseq %>% taxa() %>% sort
 phfg.pseq.rel %>% rank_names()
-subset_taxa(pseq_filtered, Genus == "Lac")
+subset_taxa(pseq.species, Genus == "Lactobacillus")
 taxa_unique(phfg.pseq.rel, "Phylum")
@@
 
-taxa <- names(subset_taxa(pseq.s, Genus=="Lactobacillus"))
+taxa <- subset_taxa(pseq.species, Genus=="Lactobacillus") %>% taxa()
+taxa
+taxa %>% kable
 

### 2018-12-31 rrbiome.Rmd
V2 mod start

@@
 }
-
 ```
@@
 ```{r data}
-pseq <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+pseq.species <- readRDS("data/phfinrisk_species_ok_drop50k_2018-11-16.RDs")
 ```

### 2018-12-31 rrbiome.Rmd
Group comparison

--- 2018-12-26 rrbiome.Rmd
@@
 
-```
-
-Loading phyloseq object
-
-```{r data}
-pseq <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+filter.phenotype.data <- function(pseq,
+                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
+                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
+                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
+    meta(pseq) %>%
+        tibble::rownames_to_column(var = "rowname") %>%
+        dplyr::select(rowname, included) %>%
+        dplyr::mutate(PULSEPRESSURE = SYSTM - DIASM,
+                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | BP_TREAT == 1, 1, 0)),
+                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
+                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
+ #       dplyr::filter(DNA_OK == 1) %>%
+        stats::na.omit() %>%
+        dplyr::select(-MEN, -DNA_OK) %>%
+        tibble::remove_rownames() %>%
+        tibble::column_to_rownames(var = "rowname")
 ```
@@
 
+Loading phyloseq object
+
+```{r data}
+pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
+```
+
 Filtering clinical data
@@
 ```{r filtering}
-dset <- meta(pseq) %>%
-    tibble::rownames_to_column(var = "rowname") %>%
-    dplyr::select(rowname, DNA_OK, MEN, BL_AGE, SYSTM, DIASM, BP_TREAT, BMI, CURR_SMOKE, DIAB, BL_USE_RX_C09,
-        Q57X, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
-    dplyr::mutate(PULSEPRESSURE = SYSTM - DIASM,
-                  HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | BP_TREAT == 1, 1, 0)),
-                  SEX = factor(ifelse(MEN == "Female", 1, 0)),
-                  MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
-    dplyr::select(-MEN) %>%
-    stats::na.omit() %>%
-    tibble::remove_rownames() %>%
-    tibble::column_to_rownames(var = "rowname")
-
-print.dset(dset)
+dset.genus <- filter.phenotype.data(pseq.genus)
+dset.species <- filter.phenotype.data(pseq.species)
 ```
 
-Continuous variables
+Phenotypical data between genus and species all.equal `r
+isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data
+frames are `r dim(dset.genus)` and `r dim(dset.species)`.
 
-```{r distribution plots, fig.width=10, fig.height=4}
-ggplot((dset %>%
-        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP) %>%
-        gather(Covariate, Value)), aes_string(x = "Value")) +
-    facet_wrap(~Covariate, ncol=4, scales = "free") +
-    geom_density(aes(y=..scaled..)) + 
-    ggplot2::theme_minimal() 
+Updating phyloseq objects
+
+```{r updating phyloseq objects}
+pseq.genus <- sample_data(dset.genus)
+pseq.species <- sample_data(dset.species)
 ```
 
-Categorical variables
+# How large subgroups are?
 
-```{r histogramsy, fig.width=10, fig.height=4}
-ggplot((dset %>%
-        select(-BL_AGE, -BMI, -SYSTM, -DIASM, -PULSEPRESSURE, -MAP) %>%
-        gather(Covariate, Value)), aes_string(x = "Value")) +
-    facet_wrap(~Covariate, ncol=8, scales = "free") +
-    geom_bar() + 
-    ggplot2::theme_minimal() 
+Number of participants in each combination of categorical variables
+
+```{r numbers by group}
+dset.species %>%
+    group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION, Q57X,
+             BL_USE_RX_C09, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
+    select(-SYSTM, -DIASM, -BL_AGE, -BMI, -PULSEPRESSURE, -MAP) %>%
+    summarise(n = n()) %>%
+    arrange(desc(n)) %>%
+    kable            
 ```
 
-```{r Updating meta data}
-pseq_filtered <- pseq
-sample_data(pseq_filtered) <- sample_data(dset)
+Number of particiapants in each combination of subset of categorical variables
+
+```{r numbers by group reduced}
+dset.species %>% group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION) %>%
+    select(SEX, BP_TREAT, CURR_SMOKE, HYPERTENSION) %>%
+    summarise(n = n()) %>%
+    arrange(desc(n)) %>%
+    kable
 ```

### 2018-12-31 rrbiome.Rmd
better subgrouping

@@
 
-Phenotypical data between genus and species all.equal `r
-isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data
-frames are `r dim(dset.genus)` and `r dim(dset.species)`.
+Phenotypical data between genus and species all.equal `r isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data frames are (`r dim(dset.genus)`) and (`r dim(dset.species)`).
 
@@
 
-# Alpha diversity
-
-Dset with alpha diversity
-
-```{r alpha diversity}
-
-alphadiversity <- microbiome::global(pseq_filtered, index="shannon")
-dset.alpha <- merge(dset, alphadiversity, by=0, all=TRUE)
-
-```
-
-```{r alpha diversity scatterplot, fig.width=10, fig.height=4}
-
-ggplot((dset.alpha %>%
-        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP, SEX, diversities_shannon) %>%
-        gather(Covariate, Value, -SEX, -diversities_shannon)), 
-       aes(x=diversities_shannon, y=Value)) +
-    facet_wrap(~Covariate, ncol=4, scales = "free") +
-    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
-    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
-    stat_poly_eq(formula = "y ~ x",
-                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
-                 label.y = 0,
-                 parse = TRUE,
-                 size = 3) +   
-    ylim(0, 200) +
-    ggplot2::theme_minimal()
-
-```
-
-Linear models
-
-```{r lm model}
-lm(MAP ~ diversities_shannon, data=dset.alpha) %>%
-    summary(.) %>% coef %>% kable(., digits=4)
-lm(MAP ~ BL_AGE + SEX + diversities_shannon, data=dset.alpha) %>%
-    summary(.) %>% coef %>% kable(., digits=4)
-lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT + diversities_shannon, data=dset.alpha) %>%
-    summary(.) %>% coef %>% kable(., digits=4)
-
-```
-
-# Beta diversity
-
-Beetadiversiteetti-matriisin lasku. Tiedosista comp_bc_dissimilarity_matrix.R ja käyttö comp_adonis_covariates.R.
-
-```{r vegdist}
-
-phf_profile.rel <- microbiome::transform(pseq_filtered, 'compositional')
-otu <- microbiome::abundances(phf_profile.rel)
-
-if (!file.exists("rds/bray.dist.m.rds")) {
-    bray.dist.m <- vegdist(t(otu), method="bray")
-    saveRDS(bray.dist.m, file = "rds/bray.dist.m.rds")
-} else {
-    bray.dist.m <- readRDS("rds/bray.dist.m.rds")
-}
-
-dist.matrix <- as.matrix(bray.dist.m)
-attr(dist.matrix, "method") <- "bray"
-
-```
-
-```{r beta hclust, fig.width = 5, fig.height = 4, dpi = 300}
-gplots::heatmap.2(dist.matrix,
-                  distfun = dist,
-                  trace = "none",
-                  hclustfun = hclust,
-                  key.title = NA,
-                  dendrogram = "column",
-                  density.info = "none",
-                  labRow = NA,
-                  labCol = NA)
-```
-
-Mallin lasku Adoniksella
-
-```{r adonis1}
-adonis(dist.matrix ~ MAP, 
-       data = meta(pseq_filtered),
-       permutations = 99)$aov.tab %>%
-                            as.data.frame %>%
-                            kable
-```
-
-```{r adonis2}
-adonis(dist.matrix ~ BL_AGE + SEX + MAP,
-       data = meta(pseq_filtered),
-       permutations = 99)$aov.tab %>%
-                            as.data.frame %>%
-                            kable
-```
-
-```{r adonis3}
-adonis(dist.matrix ~ MAP + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT,
-       data = meta(pseq_filtered),
-       permutations = 99)$aov.tab %>%
-                            as.data.frame %>%
-                            kable
-```
-
-# PCA
-
-```{r pca}
-pseq.clr <- microbiome::transform(pseq_filtered, "clr")
-X <- as.matrix(t(abundances(pseq.clr)))
-prcomp <- stats::prcomp(X)
-```
-
-```{r pca labels}
-percentage <- round(prcomp$sdev/sum(prcomp$sdev) * 100, 2)
-pca.labels <- paste0(colnames(prcomp$x), " (", paste0(as.character(percentage), "%", ")"))
-```
-
-```{r PCA plots categorical, fig.width = 10, fig.height = 10, dpi = 300}
-ggplot((cbind(meta(pseq_filtered), prcomp$x[,1:2]) %>%
-        select(PC1, PC2, BP_TREAT, CURR_SMOKE, DIAB, BL_USE_RX_C09, Q57X,
-               BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08, HYPERTENSION, SEX) %>%
-        gather(Covariate, Value, -PC1, -PC2)), 
-       aes(x=PC1, y=PC2)) +
-    facet_wrap(~Covariate, ncol=2, scales = "free") +
-    geom_point(aes(color = Value), size = 0.5) +
-    scale_color_brewer(palette="Dark2") +
-    xlab(pca.labels[1]) +
-    ylab(pca.labels[2]) +
-    ggplot2::theme_minimal()
-
-```
-
-```{r PCA plots continuous, fig.width = 10, fig.height = 10, dpi = 300}
-
-ggplot((cbind(meta(pseq_filtered), prcomp$x[,1:2]) %>%
-        select(PC1, PC2, BL_AGE, SYSTM, DIASM, BMI, MAP, PULSEPRESSURE) %>%
-        gather(Covariate, Value, -PC1, -PC2)), 
-       aes(x=PC1, y=PC2)) +
-    facet_wrap(~Covariate, ncol=2, scales = "free") +
-    geom_point(aes(color = Value), size = 0.5) +
-    scale_color_gradientn(colours = rainbow(12)) +
-    xlab(pca.labels[1]) +
-    ylab(pca.labels[2]) +
-    ggplot2::theme_minimal()
-
-```
-
-# Genus
-
-Aaron work-december-2018 prepare_maaslin_input.R
-
-```{r prepare maaslin}
-
-pheno <- t(meta(sample_data(pseq_filtered)))
-sample.ids <- colnames(pheno)
-phfg.pseq.rel <- microbiome::transform(pseq_filtered, "compositional")
-abud.rel <- abundances(phfg.pseq.rel)
-
-rownames(abud.rel) <- sapply(rownames(abud.rel),
-                             function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
-
-merged <- rbind(sample.ids, pheno, abud.rel)
-write.table(t(merged), file="merged_phfinrisk_genus.tsv", sep='\t', 
-            quote=FALSE, row.names=FALSE, col.names=TRUE)
-
-variables <- rownames(pheno)
-variables_wo_ctrls <- variables[!variables %in% c("BL_AGE","BMI","SEX")]
-data <- file("maaslin_config.config" , open='wt')
-pcl.contents <- c("Matrix: Metadata",
-                  paste(c("Read_PCL_Rows: BL_AGE,BMI,SEX",variables_wo_ctrls), collapse=","),
-                  "",
-                  "Matrix: Adundance",
-                  "Read_PCL_Rows: Archaea.Archaea-")
-writeLines(pcl.contents, con=data)
-close(data)
-```
-
-```{r run maaslin, eval = FALSE} 
-Maaslin("merged_phfinrisk_genus.tsv",
-        "maaslin",
-        strInputConfig = "maaslin_config.config",
-        strModelSelection="none",
-        dSignificanceLevel = 1.0, 
-        fAllvAll = TRUE,
-        strForcedPredictors = c("BL_AGE", "BMI", "SEX"))
-
-```
-
-Data framen luku *comp_maaslin_results_from_single_job.R* ja plottaus käyttäen *comp_maaslin_plots.R* ja *edgeR_plot_functions.R*.
-
-```{r Reading maaslins output}
-
-maas.results.df <- read.table("maaslin/merged_phfinrisk_genus.txt", header=TRUE, sep='\t')
-maas.results.df[,'Covariate'] <- maas.results.df[,'Variable']
-maas.results.df <- merge(maas.results.df,
-                         names.dset[,c('Covariate','Category', 'Name')],
-                         by=c('Covariate'))
-
-maas.results.df[,'Maaslin.Q.value'] <- maas.results.df[,'Q.value']
-maas.results.df[,'Qval'] <- p.adjust(maas.results.df[,'P.value'], method="BH")
-maas.results.df[,'beta'] <- maas.results.df[,'Coefficient']
-maas.results.df.signf <- dplyr::filter(maas.results.df, Qval < 0.05)
-
-```
-
-```{r scatter plots}
-
-sysm.signf.features <- maas.results.df.signf %>%
-    dplyr::filter(Covariate == "SYSTM") %>%
-    pull(Feature) %>%
-    as.character() %>%
-    gsub("_", ".", .)
-df.merged <- t(merged) %>% as.data.frame
-```
-
-```{r maaslin scatter plot, fig.width = 10, fig.height = 20, dpi = 300}
-
-#df.merged %>% pull("Rikenella.Bacteria") %>% as.character I%>% as.numeric %>% sqrt %>% asin
-
-ggplot(df.merged %>%
-       select(sysm.signf.features, SYSTM, SEX) %>%
-       gather(Covariate, Value, sysm.signf.features) %>%
-       mutate(SYSTM = as.numeric(as.character(SYSTM)),
-              Value = as.numeric(as.character(Value))),
-       aes(x = asin(sqrt(Value)), y = SYSTM)) +
-    facet_wrap(~Covariate, ncol=2, scales = "free") +
-    geom_point(aes(color = SEX), size = 0.5) +
-    scale_y_continuous(breaks=seq(0,200,40)) +
-    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
-    stat_poly_eq(formula = "y ~ x",
-                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
-                 label.y = 0,
-                 parse = TRUE,
-                 size = 3) +   
-    ggplot2::theme_minimal()
-
-```
-
-```{r maaslin plots, fig.width = 10, fig.height = 4, dpi = 300}
-source("edgeR_plot_functions.R")
-
-plot_maaslin_heatmap_final(maas.results.df.signf,
-                      xlab.text="Genera",
-                      beta.maxabs=max(abs(maas.results.df.signf[,'beta'])))
-
-```
-
-```{r scp html}
+```{r scp html, cache = FALSE}
 system('scp rrbiome.html utu:articletwo')

### 2018-12-31 rrbiome.Rmd
Better subgrouping real

--- 2018-12-31 rrbiome.Rmd
@@
         stats::na.omit() %>%
-        dplyr::select(-MEN, -DNA_OK) %>%
+        dplyr::select(-MEN, -DNA_OK, -BP_TREAT) %>%
         tibble::remove_rownames() %>%
@@
 
-Phenotypical data between genus and species all.equal `r
-isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data
-frames are `r dim(dset.genus)` and `r dim(dset.species)`.
+Phenotypical data between genus and species all.equal `r isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data
+frames are (`r dim(dset.genus)`) and (`r dim(dset.species)`).
 
@@
 
-Number of participants in each combination of categorical variables
+Number of participants in each combination of all categorical variables
 
@@
 dset.species %>%
-    group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION, Q57X,
+    group_by(SEX, CURR_SMOKE,  HYPERTENSION, Q57X,
              BL_USE_RX_C09, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
@@
 
-Number of particiapants in each combination of subset of categorical variables
+Number of particiapants in each combination after combining RR
+medication and exercice to binomial variables.
 
 ```{r numbers by group reduced}
-dset.species %>% group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION) %>%
-    select(SEX, BP_TREAT, CURR_SMOKE, HYPERTENSION) %>%
+dset.species %>%
+    mutate(ANYDRUG = ifelse(BL_USE_RX_C09  == 1 | BL_USE_RX_C03 == 1 |
+                            BL_USE_RX_C07 == 1  | BL_USE_RX_C08 == 1, 1, 0),
+           EXERCICE = ifelse(Q57X == 1, 0, 1)) %>%
+    group_by(SEX, CURR_SMOKE,  HYPERTENSION, ANYDRUG, EXERCICE) %>%
+    select(SEX, CURR_SMOKE, HYPERTENSION, EXERCICE) %>%
     summarise(n = n()) %>%
@@
     kable
+
+dset.species %>% pull(Q57X) %>% summary
 ```
@@
     summary(.) %>% coef %>% kable(., digits=4)
-lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT + diversities_shannon, data=dset.alpha) %>%
+lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + diversities_shannon, data=dset.alpha) %>%
     summary(.) %>% coef %>% kable(., digits=4)
@@
 
-Aaron work-december-2018 prepare_maaslin_input.R
+Aaron work-december-2018 *prepare_maaslin_input.R*
 
@@
 
-```{r scp html}
+```{r scp html, cache = FALSE}
 system('scp rrbiome.html utu:articletwo')

### 2018-12-31 rrbiome.Rmd
Redefined HYPERTENSION and added chi squared

@@
         dplyr::select(rowname, included) %>% 
-        dplyr::mutate(PULSEPRESSURE = SYSTM - DIASM,
-                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | BP_TREAT == 1, 1, 0)),
+        dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
+                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
+                      ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
+                      PULSEPRESSURE = SYSTM - DIASM,
+                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
                       SEX = factor(ifelse(MEN == "Female", 1, 0)),
@@
         stats::na.omit() %>%
-        dplyr::select(-MEN, -DNA_OK, -BP_TREAT) %>%
+        dplyr::select(-MEN, -DNA_OK) %>%
         tibble::remove_rownames() %>%
@@
 
-Phenotypical data between genus and species all.equal `r isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data
+Phenotypical data between genus and species all.equal `r 
+isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data
 frames are (`r dim(dset.genus)`) and (`r dim(dset.species)`).
@@
              BL_USE_RX_C09, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
-    select(-SYSTM, -DIASM, -BL_AGE, -BMI, -PULSEPRESSURE, -MAP) %>%
     summarise(n = n()) %>%
@@
 dset.species %>%
-    mutate(ANYDRUG = ifelse(BL_USE_RX_C09  == 1 | BL_USE_RX_C03 == 1 |
-                            BL_USE_RX_C07 == 1  | BL_USE_RX_C08 == 1, 1, 0),
-           EXERCICE = ifelse(Q57X == 1, 0, 1)) %>%
-    group_by(SEX, CURR_SMOKE,  HYPERTENSION, ANYDRUG, EXERCICE) %>%
-    select(SEX, CURR_SMOKE, HYPERTENSION, EXERCICE) %>%
+     group_by(SEX, CURR_SMOKE,  HYPERTENSION, ANYDRUG, ANYEXERCICE) %>%
     summarise(n = n()) %>%
@@
     kable
-
-dset.species %>% pull(Q57X) %>% summary
 ```
 
-# Alpha diversity
+# Self reported vs. registered medication
+
+Contigency table for RR mediaction. P-value for chi-squared test is
+`r chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value %>% format.pval`.
+
+```{r self vs registered medication}
+table(dset.species$BP_TREAT, dset.species$ANYDRUG)
+```
 

### 2018-12-31 rrbiome.Rmd
Minor

--- 2018-12-31 rrbiome.Rmd
@@
 
-Phenotypical data between genus and species all.equal `r isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data frames are (`r dim(dset.genus)`) and (`r dim(dset.species)`).
+Phenotypical data between genus and species all.equal `r 
+isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data
+frames are (`r dim(dset.genus)`) and (`r dim(dset.species)`).
 
@@
 dset.species %>%
-    group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION, Q57X,
+    group_by(SEX, CURR_SMOKE,  HYPERTENSION, Q57X,
              BL_USE_RX_C09, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
-    select(-SYSTM, -DIASM, -BL_AGE, -BMI, -PULSEPRESSURE, -MAP) %>%
     summarise(n = n()) %>%
@@
 
-Number of particiapants in each combination of subset of categorical variables
+Number of particiapants in each combination after combining RR
+medication and exercice to binomial variables.
 
 ```{r numbers by group reduced}
-dset.species %>% group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION) %>%
-    select(SEX, BP_TREAT, CURR_SMOKE, HYPERTENSION) %>%
+dset.species %>%
+     group_by(SEX, CURR_SMOKE,  HYPERTENSION, ANYDRUG, ANYEXERCICE) %>%
     summarise(n = n()) %>%
     arrange(desc(n)) %>%
+    kable
+```
+
+# Self reported vs. registered medication
+
+Contigency table for RR mediaction. P-value for chi-squared test is
+`r chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value %>% format.pval`.
+
+```{r self vs registered medication}
+table(dset.species$BP_TREAT, dset.species$ANYDRUG)  %>%
     kable

### 2018-12-31 rrbiome.Rmd
Minor

--- 2018-12-31 rrbiome.Rmd
@@
 ```{r self vs registered medication}
-table(dset.species$BP_TREAT, dset.species$ANYDRUG)
+table(dset.species$BP_TREAT, dset.species$ANYDRUG) %>% kable
 ```
@@
 
-```
-
-```{r maaslin plots, fig.width = 10, fig.height = 4, dpi = 300}
-source("edgeR_plot_functions.R")
-
-plot_maaslin_heatmap_final(maas.results.df.signf,
-                      xlab.text="Genera",
-                      beta.maxabs=max(abs(maas.results.df.signf[,'beta'])))
-
-```
-
 ```{r scp html, cache = FALSE}

### 2018-12-31 rrbiome.Rmd
Minor
Table labels

@@
 
-Contigency table for RR mediaction. P-value for chi-squared test is
-`r chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value %>% format.pval`.
+Contigency table for RR mediaction.
 
 ```{r self vs registered medication}
-table(dset.species$BP_TREAT, dset.species$ANYDRUG) %>% kable
+tbl <- table(dset.species$BP_TREAT, dset.species$ANYDRUG)
+colnames(tbl) <- c("ANYDRUG 0", "ANYDRUG 1")
+rownames(tbl) <- c("BP_TREAT 0", "BP_TREAT 1")
+tbl %>% kable(.)
 ```
+
+\xi^2-test for medication `r format.pval(chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value)`.
 

### 2018-12-31 rrbiome.Rmd
ALPHA-MAP plot

--- 2018-12-31 rrbiome.Rmd
@@
 library(knitr)
+library(kableExtra)
 library(tidyr)
@@
         tibble::column_to_rownames(var = "rowname")
+
+meta.merge.alphadiversity <- function(pseq, index = "shannon") {
+    alphadiversity  <- microbiome::global(pseq, index = index)
+    base::merge(meta(pseq), alphadiversity, by=0, all=TRUE) %>%
+        dplyr::rename(Sample_ID = Row.names) %>%
+            dplyr::mutate(BL_AGE_GROUP = group_age(dset.alpha$BL_AGE))
+}
 ```
@@
 ```{r updating phyloseq objects}
-pseq.genus <- sample_data(dset.genus)
-pseq.species <- sample_data(dset.species)
+sample_data(pseq.genus) <- sample_data(dset.genus)
+sample_data(pseq.species) <- sample_data(dset.species)
 ```
@@
 
-Contigency table for RR mediaction. P-value for chi-squared test is
-`r chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value %>% format.pval`.
+Contigency table for RR mediaction.
 
 ```{r self vs registered medication}
-table(dset.species$BP_TREAT, dset.species$ANYDRUG) %>% kable
+tbl <- table(dset.species$BP_TREAT, dset.species$ANYDRUG)
+colnames(tbl) <- c("ANYDRUG 0", "ANYDRUG 1")
+rownames(tbl) <- c("BP_TREAT 0", "BP_TREAT 1")
+tbl %>%
+    kable(.) %>%
+    kable_styling(full_width = F) %>%
+    column_spec(1, bold = T)
 ```
+
+$\chi^2$-test for medication `r format.pval(chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value)`.
+
+# Alpha diversity
 
@@
 ```{r alpha diversity}
+dset.alpha <- meta.merge.alphadiversity(pseq.species)
+```
 
-alphadiversity <- microbiome::global(pseq_filtered, index="shannon")
-dset.alpha <- merge(dset, alphadiversity, by=0, all=TRUE)
-
+```{r alpha-MAP scatter plots, fig.width=10, fig.height=4}
+ggplot((dset.alpha, aes(x=diversities_shannon, y=MAP)) +
+    facet_wrap(BL_AGE_GROUP, ncol=4, scales = "free") +
+    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
+    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
+    stat_poly_eq(formula = "y ~ x",
+                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
+                 label.y = 0,
+                 parse = TRUE,
+                 size = 3) +   
+    ylim(0, 200) +
+    ggplot2::theme_minimal() 
 ```
@@
 ```{r alpha diversity scatterplot, fig.width=10, fig.height=4}
-
 ggplot((dset.alpha %>%
@@
     ggplot2::theme_minimal() 
-
 ```
@@
     ggplot2::theme_minimal() 
-
-```{r scp html, cache = FALSE}
-system('scp rrbiome.html utu:articletwo')
 ```

### 2018-12-31 rrbiome.Rmd
eval FALSE

@@
 
-```{r alpha diversity scatterplot, fig.width=10, fig.height=4}
+```{r alpha diversity scatterplot, fig.width=10, fig.height=4, eval = FALSE}
 ggplot((dset.alpha %>%
@@
 
-```{r lm model}
+```{r lm model, eval = FALSE}
 lm(MAP ~ diversities_shannon, data=dset.alpha) %>%
@@
 
-```{r vegdist}
+```{r vegdist, eval = FALSE}
 
@@
 
-```{r beta hclust, fig.width = 5, fig.height = 4, dpi = 300}
+```{r beta hclust, fig.width = 5, fig.height = 4, dpi = 300, eval = FALSE}
 gplots::heatmap.2(dist.matrix,
@@
 
-```{r adonis1}
+```{r adonis1, eval = FALSE}
 adonis(dist.matrix ~ MAP, 
@@
 
-```{r adonis2}
+```{r adonis2, eval = FALSE}
 adonis(dist.matrix ~ BL_AGE + SEX + MAP,
@@
 
-```{r adonis3}
+```{r adonis3, eval = FALSE}
 adonis(dist.matrix ~ MAP + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT,
@@
 
-```{r pca}
+```{r pca, eval = FALSE}
 pseq.clr <- microbiome::transform(pseq_filtered, "clr")
@@
 
-```{r pca labels}
+```{r pca labels, eval = FALSE}
 percentage <- round(prcomp$sdev/sum(prcomp$sdev) * 100, 2)
@@
 
-```{r PCA plots categorical, fig.width = 10, fig.height = 10, dpi = 300}
+```{r PCA plots categorical, fig.width = 10, fig.height = 10, dpi = 300, eval = FALSE}
 ggplot((cbind(meta(pseq_filtered), prcomp$x[,1:2]) %>%
@@
 
-```{r PCA plots continuous, fig.width = 10, fig.height = 10, dpi = 300}
+```{r PCA plots continuous, fig.width = 10, fig.height = 10, dpi = 300,  eval = FALSE}
 
@@
 
-```{r prepare maaslin}
+```{r prepare maaslin, eval = FALSE}
 
@@
 
-```{r Reading maaslins output}
+```{r Reading maaslins output, eval = FALSE}
 
@@
 
-```{r scatter plots}
+```{r scatter plots, eval = FALSE}
 
@@
 
-```{r maaslin scatter plot, fig.width = 10, fig.height = 20, dpi = 300}
+```{r maaslin scatter plot, fig.width = 10, fig.height = 20, dpi = 300, eval = FALSE}
 

### 2018-12-31 rrbiome.Rmd
MAP ALPHA

@@
 ```{r alpha-MAP scatter plots, fig.width=10, fig.height=4}
-ggplot((dset.alpha, aes(x=diversities_shannon, y=MAP)) +
-    facet_wrap(BL_AGE_GROUP, ncol=4, scales = "free") +
-    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
+ggplot(dset.alpha %>% mutate(SEX = factor(ifelse(SEX == 0, "Male", "Female"))),
+       aes(x=diversities_shannon, y=MAP)) +
+    facet_wrap(~BL_AGE_GROUP + SEX, ncol=2, scales = "free") +
+    geom_point(size=0.1, shape=1, aes(colour = ANYDRUG)) +
     geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +

### 2018-12-31 bin/reports.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ${HOME}/phd/research/articletwo/ || exit 1
+
+rsync -avu utu:articletwo/*.{html,pdf} reports/ #&& open reports/rrbiome.html

### 2018-12-31 rrbiome.Rmd
Minor

--- 2018-12-31 rrbiome.Rmd
@@
 ```{r alpha-MAP scatter plots, fig.width=10, fig.height=4}
-ggplot((dset.alpha, aes(x=diversities_shannon, y=MAP)) +
-    facet_wrap(BL_AGE_GROUP, ncol=4, scales = "free") +
-    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
+ggplot((dset.alpha %>% mutate(SEX = factor(ifelse(SEX == 0, "Male", "Female")))),
+       aes(x=diversities_shannon, y=MAP)) +
+    facet_wrap(~BL_AGE_GROUP + SEX, ncol=2, scales = "free") +
+    geom_point(size=0.1, shape=1, aes(colour = ANYDRUG)) +
     geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +

### 2018-12-31 rrbiome.Rmd
Minor

--- 2018-12-31 rrbiome.Rmd
@@
 ```{r alpha-MAP scatter plots, fig.width=10, fig.height=4}
-ggplot((dset.alpha, aes(x=diversities_shannon, y=MAP)) +
-    facet_wrap(BL_AGE_GROUP, ncol=4, scales = "free") +
-    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
+ggplot(dset.alpha %>% mutate(SEX = ifelse(SEX == 0, "Male", "Female")), aes(x=diversities_shannon, y=MAP)) +
+    facet_wrap(~BL_AGE_GROUP + SEX, ncol=2, scales = "free") +
+    geom_point(size=0.1, shape=1, aes(colour = ANYDRUG)) +
     geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +

### 2018-12-31 rrbiome.Rmd
Bug

@@
         dplyr::rename(Sample_ID = Row.names) %>%
-            dplyr::mutate(BL_AGE_GROUP = group_age(dset.alpha$BL_AGE))
+            dplyr::mutate(BL_AGE_GROUP = group_age(BL_AGE))
 }

### 2018-12-31 rrbiome.Rmd
Colors

@@
 
-```{r alpha-MAP scatter plots, fig.width=10, fig.height=4}
+```{r alpha-MAP scatter plots, fig.width=10, fig.height=15}
 ggplot(dset.alpha %>% mutate(SEX = ifelse(SEX == 0, "Male", "Female")), aes(x=diversities_shannon, y=MAP)) +
@@
     geom_point(size=0.1, shape=1, aes(colour = ANYDRUG)) +
+    scale_colour_manual(values = c("red","blue", "green")) +
     geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +

### 2018-12-31 rrbiome.Rmd
alpha diversity

@@
 dset.species %>%
-    group_by(SEX, CURR_SMOKE,  HYPERTENSION, Q57X,
+    group_by(SEX, CURR_SMOKE,  HYPERTENSION, Q57X, DIAB,
              BL_USE_RX_C09, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
@@
 dset.species %>%
-     group_by(SEX, CURR_SMOKE,  HYPERTENSION, ANYDRUG, ANYEXERCICE) %>%
+     group_by(SEX, CURR_SMOKE,  HYPERTENSION, DIAB, ANYDRUG, ANYEXERCICE) %>%
     summarise(n = n()) %>%
@@
 
+Scatterplots for alpha diversity by age group and sex
+
 ```{r alpha-MAP scatter plots, fig.width=10, fig.height=15}
@@
 
-```{r alpha diversity scatterplot, fig.width=10, fig.height=4, eval = FALSE}
-ggplot((dset.alpha %>%
-        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP, SEX, diversities_shannon) %>%
-        gather(Covariate, Value, -SEX, -diversities_shannon)), 
-       aes(x=diversities_shannon, y=Value)) +
-    facet_wrap(~Covariate, ncol=4, scales = "free") +
-    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
-    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
-    stat_poly_eq(formula = "y ~ x",
-                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
-                 label.y = 0,
-                 parse = TRUE,
-                 size = 3) +   
-    ylim(0, 200) +
-    ggplot2::theme_minimal() 
-```
-
 Linear models
 
-```{r lm model, eval = FALSE}
-lm(MAP ~ diversities_shannon, data=dset.alpha) %>%
-    summary(.) %>% coef %>% kable(., digits=4)
-lm(MAP ~ BL_AGE + SEX + diversities_shannon, data=dset.alpha) %>%
-    summary(.) %>% coef %>% kable(., digits=4)
-lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + diversities_shannon, data=dset.alpha) %>%
-    summary(.) %>% coef %>% kable(., digits=4)
+```{r lm model}
+calculateglm <- function(dset,
+                         responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
+                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
+                         min_n_for_continuous = 10,
+                         modelstring = "%s ~ AGE + sex") {
+    glmlist <- lapply(responses, function(response) {
+        fo.family <- ifelse(length(unique(pull(dset, response)))>min_n_for_continuous, stats::gaussian, stats::binomial)
+        fo <- sprintf(modelstring, response)
+        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
+            broom::tidy() %>%
+                dplyr::filter(grepl("diversities_", term)) %>%
+                dplyr::mutate(response = response, fo = fo) %>%
+                dplyr::select(-term)
+    })
+    data.table::rbindlist(test, id = "model_name") %>%
+        as.data.frame %>%
+        dplyr::mutate(conf.low = estimate - qnorm(1- 0.05/2) * std.error,
+                      conf.high = estimate + qnorm(1- 0.05/2) * std.error)
+}
+
+calculateglm(as.data.frame(dset.alpha), modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon") %>%
+    select(response, estimate, conf.low, conf.high, p.value) %>%
+    kable
+
+#lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon, data = dset.alpha) %>% summary()
+
+###%>% kable(., digits=4)
 

### 2018-12-31 rrbiome.Rmd
minor

@@
     glmlist <- lapply(responses, function(response) {
-        fo.family <- ifelse(length(unique(pull(dset, response)))>min_n_for_continuous, stats::gaussian, stats::binomial)
+        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
         fo <- sprintf(modelstring, response)
@@
     })
-    data.table::rbindlist(test, id = "model_name") %>%
+    data.table::rbindlist(glmlist, id = "model_name") %>%
         as.data.frame %>%

### 2018-12-31 rrbiome.Rmd
Minor

--- 2018-12-31 rrbiome.Rmd
@@
 library(knitr)
-library(kableExtra)
 library(tidyr)
@@
 tbl %>%
-    kable(.) %>%
-    kable_styling(full_width = F) %>%
-    column_spec(1, bold = T)
+    kable(.)
 ```
@@
 
-```{r lm model}
+```{r function for glm calculation, include = FALSE}
 calculateglm <- function(dset,
@@
 }
+```
 
+```{r calculate alpha models}
 calculateglm(as.data.frame(dset.alpha), modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon") %>%
-    select(response, estimate, conf.low, conf.high, p.value) %>%
-    kable
-
-#lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon, data = dset.alpha) %>% summary()
-
-###%>% kable(., digits=4)
-
+    select(response, estimate, std.error, conf.low, conf.high, p.value) %>%
+    kable(.)
 ```
@@
 
-Beetadiversiteetti-matriisin lasku. Tiedosista comp_bc_dissimilarity_matrix.R ja käyttö comp_adonis_covariates.R.
+Beetadiversiteetti-matriisin lasku. Tiedosista *comp_bc_dissimilarity_matrix.R* ja käyttö *comp_adonis_covariates.R*.
 

### 2018-12-31 rrbiome.Rmd
Minor

@@
 
-Scatterplots for alpha diversity by age group and sex
+Scatterplots for alpha diversity by age group and sex.
 
@@
 ```
-
-Linear models
 
@@
 
-```{r calculate alpha models}
-calculateglm(as.data.frame(dset.alpha), modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon") %>%
+Calculating linear models for alpha diversity.
+
+```{r calculate full alpha model}
+calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon") %>%
+    select(response, estimate, std.error, conf.low, conf.high, p.value) %>%
+    kable(.)
+```
+
+Calculating linear model using binomial variable for exercice and medication
+
+```{r calculate full alpha model}
+calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + ANYEXERCICE + DIAB + ANYDRUG  + diversities_shannon") %>%
     select(response, estimate, std.error, conf.low, conf.high, p.value) %>%

### 2018-12-31 rrbiome.Rmd
Minor

@@
 
-```{r calculate full alpha model}
+```{r calculate binomial alpha model}
 calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + ANYEXERCICE + DIAB + ANYDRUG  + diversities_shannon") %>%

### 2018-12-31 rrbiome.Rmd
Minimal model

@@
 
-Calculating linear model using binomial variable for exercice and medication
+Calculating linear model using binomial variable for exercice and medication.
 
@@
 calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + ANYEXERCICE + DIAB + ANYDRUG  + diversities_shannon") %>%
+    select(response, estimate, std.error, conf.low, conf.high, p.value) %>%
+    kable(.)
+```
+
+Calculating linear model using minimal variables.
+
+```{r calculate minimal alpha model}
+calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + diversities_shannon") %>%
     select(response, estimate, std.error, conf.low, conf.high, p.value) %>%

### 2018-12-31 rrbiome.Rmd
Beta diversity

@@
 
-```{r vegdist, eval = FALSE}
-
-phf_profile.rel <- microbiome::transform(pseq_filtered, 'compositional')
-otu <- microbiome::abundances(phf_profile.rel)
-
-if (!file.exists("rds/bray.dist.m.rds")) {
-    bray.dist.m <- vegdist(t(otu), method="bray")
-    saveRDS(bray.dist.m, file = "rds/bray.dist.m.rds")
-} else {
-    bray.dist.m <- readRDS("rds/bray.dist.m.rds")
+```{r vegdist}
+calculate.beta.matrix <- function(pseq) {
+    compositional.profile <- microbiome::transform(pseq, 'compositional')
+    otu <- microbiome::abundances(compositional.profile)
+    vegan::vegdist(t(otu), method="bray")
 }
 
+if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds") {
+    bray.dist.m.species <- calculate.beta.matrix(pseq.species)
+    saveRDS(bray.dist.m.species, file = "rds/bray.dist.m.species.rds")
+    bray.dist.m.genus <- calculate.beta.matrix(pseq.genus)
+    saveRDS(bray.dist.m.genus, file = "rds/bray.dist.m.genus.rds")
+} else {
+    bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
+    bray.dist.m.genus <- readRDS("rds/bray.dist.m.genus.rds")
+}
+}
+```
+
+```{r temp}
 dist.matrix <- as.matrix(bray.dist.m)
 attr(dist.matrix, "method") <- "bray"
-
 ```

### 2018-12-31 rrbiome.Rmd
Beta bug

--- 2018-12-31 rrbiome.Rmd
@@
 
-```{r vegdist, eval = FALSE}
+```{r matrix calculation}
+if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds") {
+    bray.dist.m.species <- calculate.beta.matrix(pseq.species)
+    saveRDS(bray.dist.m.species, file = "rds/bray.dist.m.species.rds")
+    bray.dist.m.genus <- calculate.beta.matrix(pseq.genus)
+    saveRDS(bray.dist.m.genus, file = "rds/bray.dist.m.genus.rds")
+} else {
+    bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
+    bray.dist.m.genus <- readRDS("rds/bray.dist.m.genus.rds")
+}
+```
 
-phf_profile.rel <- microbiome::transform(pseq_filtered, 'compositional')
-otu <- microbiome::abundances(phf_profile.rel)
-
-if (!file.exists("rds/bray.dist.m.rds")) {
-    bray.dist.m <- vegdist(t(otu), method="bray")
-    saveRDS(bray.dist.m, file = "rds/bray.dist.m.rds")
-} else {
-    bray.dist.m <- readRDS("rds/bray.dist.m.rds")
-}
-
+```{r temp}
 dist.matrix <- as.matrix(bray.dist.m)
 attr(dist.matrix, "method") <- "bray"
-
 ```

### 2018-12-31 rrbiome.Rmd
Missing bracket

@@
 ```{r matrix calculation}
-if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds") {
+if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds")) {
     bray.dist.m.species <- calculate.beta.matrix(pseq.species)

### 2018-12-31 rrbiome.Rmd
Better matrix

@@
             dplyr::mutate(BL_AGE_GROUP = group_age(BL_AGE))
+}
+
+calculate.beta.matrix <- function(pseq) {
+    compositional.profile <- microbiome::transform(pseq, 'compositional')
+    otu <- microbiome::abundances(compositional.profile)
+    bray.dist.m <- vegan::vegdist(t(otu), method="bray")
+    dist.matrix <- as.matrix(bray.dist.m)
+    attr(dist.matrix, "method") <- "bray"
+    dist.matrix
 }
@@
 }
-```
-
-```{r temp}
-dist.matrix <- as.matrix(bray.dist.m)
-attr(dist.matrix, "method") <- "bray"
 ```

### 2019-01-01 rrbiome.Rmd
Adonis, bin

@@
                          min_n_for_continuous = 10,
-                         modelstring = "%s ~ AGE + sex") {
+                         modelstring = "%s ~ BL_AGE + SEX") {
     glmlist <- lapply(responses, function(response) {
@@
 
-```{r matrix calculation}
+```{r matrix calculation, cache = FALSE}
 if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds")) {
@@
 
-```{r beta hclust, fig.width = 5, fig.height = 4, dpi = 300, eval = FALSE}
-gplots::heatmap.2(dist.matrix,
-                  distfun = dist,
-                  trace = "none",
-                  hclustfun = hclust,
-                  key.title = NA,
-                  dendrogram = "column",
-                  density.info = "none",
-                  labRow = NA,
-                  labCol = NA)
+Mallin lasku Adoniksella
+
+```{r function for glm calculation, include = FALSE}
+calculateadonis <- function(dset,
+                            matrix,
+                            responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
+                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
+                            covariates = "BL_AGE + SEX",
+                            modelstring = "matrix ~ %s + %s",
+                            npermutations = 99) {
+    glmlist <- lapply(responses, function(response) {
+        fo <- as.formula(sprintf(modelstring, response, covariates))
+        ad <- adonis(fo, data = dset, permutations = npermutations)
+        ad$aov.tab %>% as.data.frame
+    }) 
+}
 ```
 
-Mallin lasku Adoniksella
+ ```{r adonis calculation, cache = FALSE}
+if (!file.exists("rds/adonis.species.rds")) {
+    adonis.species <- calculateadonis(dset = meta(pseq.species),
+                    matrix = bray.dist.m.species,
+                    covariates = "BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
+                    npermutations = 99)
+     saveRDS(adonis.species, file = "rds/adonis.species.rds")
+} else {
+     adonis.species  <- readRDS("rds/adonis.species.rds")
+}
+```
 
-```{r adonis1, eval = FALSE}
-adonis(dist.matrix ~ MAP, 
-       data = meta(pseq_filtered),
-       permutations = 99)$aov.tab %>%
+```{r adonis3}
+
+ret
+
+adonis(bray.dist.m.species ~ MAP + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09,
+       data = meta(pseq.species),
+       permutations = 2)$aov.tab %>%
                             as.data.frame %>%
                             kable
-```
 
-```{r adonis2, eval = FALSE}
-adonis(dist.matrix ~ BL_AGE + SEX + MAP,
-       data = meta(pseq_filtered),
-       permutations = 99)$aov.tab %>%
-                            as.data.frame %>%
-                            kable
-```
-
-```{r adonis3, eval = FALSE}
-adonis(dist.matrix ~ MAP + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT,
-       data = meta(pseq_filtered),
-       permutations = 99)$aov.tab %>%
-                            as.data.frame %>%
-                            kable
+ glmlist <- lapply(responses, function(response) {
+        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
+        fo <- sprintf(modelstring, response)
+        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
+            broom::tidy() %>%
+                dplyr::filter(grepl("diversities_", term)) %>%
+                dplyr::mutate(response = response, fo = fo) %>%
+                dplyr::select(-term)
+    })
 ```

### 2019-01-01 bin/reports.sh

@@ -1,5 +1,13 @@
 #!/bin/bash
 
-cd ${HOME}/phd/research/articletwo/ || exit 1
+cd "${HOME}/phd/research/articletwo/reports" || exit 1
+HOSTPATH="utu:articletwo"
+FILE="rrbiome.html"
+BACKUP="rrbiome_$(date +%Y%m%d).html"
 
-rsync -avu utu:articletwo/*.{html,pdf} reports/ #&& open reports/rrbiome.html
+RET=$(rsync -aEim $HOSTPATH/$FILE $FILE)
+
+if [ $? -eq 0 -a -n "${RET}" ]; then
+	cp $FILE $BACKUP
+	open $FILE
+fi

### 2019-01-01 rrbiome.Rmd
rename

@@
 
-```{r function for glm calculation, include = FALSE}
+```{r function for adnois calculation, include = FALSE}
 calculateadonis <- function(dset,

### 2019-01-01 rrbiome.Rmd
mclapply

--- 2019-01-01 rrbiome.Rmd
@@
 
-```{r function for glm calculation, include = FALSE}
+```{r function for adnoiscalculation, include = FALSE}
 calculateadonis <- function(dset,
@@
                             npermutations = 99) {
-    glmlist <- lapply(responses, function(response) {
+    glmlist <- mclapply(responses, function(response) {
         fo <- as.formula(sprintf(modelstring, response, covariates))
@@
         ad$aov.tab %>% as.data.frame
-    }) 
+    }, mc.cores = lenfth(responses)) 
 }

### 2019-01-01 rrbiome.Rmd
Bug

--- 2019-01-01 rrbiome.Rmd
@@
 library(Maaslin)
+library(parallel)
 ```
@@
                       MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
- #       dplyr::filter(DNA_OK == 1) %>%
          stats::na.omit() %>%
@@
         ad$aov.tab %>% as.data.frame
-    }) 
+    }, mc.cores = length(responses)) 
 }
@@
 
-```{r adonis3}
+```{r adonis3, eval = FALSE}
 
 ret
+
+ adonis.species <- calculateadonis(dset = meta(pseq.species),
+                    matrix = bray.dist.m.species,
+                    covariates = "BL_AGE + SEX",
+                    npermutations = 92)
 

### 2019-01-01 rrbiome.Rmd
Cleaning adonis output

@@
                             npermutations = 99) {
-    glmlist <- mclapply(responses, function(response) {
+    ret  <- mclapply(responses, function(response) {
         fo <- as.formula(sprintf(modelstring, response, covariates))
         ad <- adonis(fo, data = dset, permutations = npermutations)
-        ad$aov.tab %>% as.data.frame
+        ad$aov.tab %>% as.data.frame %>%
+            tibble::rownames_to_column(var = "response") %>%
+            dplyr::mutate(fo = fo)
     }, mc.cores = length(responses))
+    data.table::rbindlist(ret, id = "model_name") 
 }
@@
 ```{r adonis3, eval = FALSE}
+adonis.species %>%
+    broom::tidy()
 
-ret
+ret <- list(model_1 = adonis.species$model_1 %>% tibble::rownames_to_column(var = "response"),
+     model_2 =  adonis.species$model_2 %>% tibble::rownames_to_column(var = "response"))
 
- adonis.species <- calculateadonis(dset = meta(pseq.species),
-                    matrix = bray.dist.m.species,
-                    covariates = "BL_AGE + SEX",
-                    npermutations = 92)
+data.table::rbindlist(ret, id = "model_name") %>%
+    dplyr::filter(response %in% c("MAP", "SYSTM"))
 
-adonis(bray.dist.m.species ~ MAP + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09,
-       data = meta(pseq.species),
-       permutations = 2)$aov.tab %>%
-                            as.data.frame %>%
-                            kable
-
- glmlist <- lapply(responses, function(response) {
-        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
-        fo <- sprintf(modelstring, response)
-        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
-            broom::tidy() %>%
-                dplyr::filter(grepl("diversities_", term)) %>%
-                dplyr::mutate(response = response, fo = fo) %>%
-                dplyr::select(-term)
     })

### 2019-01-01 rrbiome.Rmd
Minor

@@
         ad <- adonis(fo, data = dset, permutations = npermutations)
-        ad$aov.tab %>% as.data.frame %>%
-            tibble::rownames_to_column(var = "response") %>%
-            dplyr::mutate(fo = fo)
+        ad$aov.tab %>%
+            as.data.frame %>%
+            tibble::rownames_to_column(var = "term") %>%
+            dplyr::mutate(response = response, fo = fo)
     }, mc.cores = length(responses))

### 2019-01-01 rrbiome.Rmd
Fixing adonis implementation

--- 2019-01-01 rrbiome.Rmd
@@
     ret  <- mclapply(responses, function(response) {
-        fo <- as.formula(sprintf(modelstring, response, covariates))
-        ad <- adonis(fo, data = dset, permutations = npermutations)
-        ad$aov.tab %>% as.data.frame %>%
-            tibble::rownames_to_column(var = "response") %>%
-            dplyr::mutate(fo = fo)
+        fo <- sprintf(modelstring, response, covariates)
+        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
+        ad$aov.tab %>%
+            as.data.frame %>%
+            tibble::rownames_to_column(var = "term") %>%
+            dplyr::mutate(response = response, fo = fo)
       }, mc.cores = length(responses))
-    data.table::rbindlist(ret, id = "model_name") 
+    data.table::rbindlist(ret, id = "model_name") %>%
+        dplyr::filter(term %in% responses) %>%
+            dplyr::select(-term)
 }
@@
 ```{r adonis3, eval = FALSE}
-adonis.species %>%
-    broom::tidy()
+ adonis.species <- calculateadonis(dset = meta(pseq.species),
+                    matrix = bray.dist.m.species,
+                    covariates = "BL_AGE + SEX",
+                    npermutations = 2)
+
+ adonis.species
 
@@
 data.table::rbindlist(ret, id = "model_name") %>%
-    dplyr::filter(response %in% c("MAP", "SYSTM"))
+    dplyr::filter(response %in% list(a = "MAP", b = "SYSTM"))
 

### 2019-01-01 rrbiome.Rmd
Create rds

@@
                       file.path = 'cache/', fig.path = 'cache/')
+
+dir.create("rds/", showWarnings = FALSE)
 ```
@@
 ```{r adonis3, eval = FALSE}
- adonis.species <- calculateadonis(dset = meta(pseq.species),
-                    matrix = bray.dist.m.species,
-                    covariates = "BL_AGE + SEX",
-                    npermutations = 2)
-
- adonis.species
-
-ret <- list(model_1 = adonis.species$model_1 %>% tibble::rownames_to_column(var = "response"),
-     model_2 =  adonis.species$model_2 %>% tibble::rownames_to_column(var = "response"))
-
-data.table::rbindlist(ret, id = "model_name") %>%
-    dplyr::filter(response %in% list(a = "MAP", b = "SYSTM"))
-
-    })
+ adonis.species %>%
+    select(response, SumsOfSqs, MeanSqs, F.Model, R2, "Pr(>F)") %>%
+    kable(.)
 ```

### 2019-01-02 rrbiome.Rmd
Minor reordering and cleaning

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, results='asis', cache=TRUE, warning=FALSE)
+knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis', cache=TRUE, warning=FALSE)
 knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
@@
 
-Importing libraries
+Importing libraries:
 
@@
 
-```{r functions, include = FALSE}
-print.dset <- function(dset) {
-    dset %>%
-        gather(Covariate, Value) %>%
-        group_by(Covariate) %>%
-        summarise_all(., .funs = funs(mean(as.numeric(.), na.rm = TRUE),
-                                         median(as.numeric(.), na.rm = TRUE),
-                                         sd(as.numeric(.), na.rm = TRUE),
-                                         min(as.numeric(.), na.rm = TRUE),
-                                         max(as.numeric(.), na.rm = TRUE),
-                                         NAs = sum(is.na(.)),
-                                         n = n(),
-                                         n_distinct)) %>%
-            mutate_if(is.numeric, round, 2) %>%
-            kable
-}
+Function definitions
 
+```{r functions}
 filter.phenotype.data <- function(pseq,
@@
 
+Sample data has dimensions (`r dim(sample_data(pseq.species)`).
+
+# Self reported vs. registered medication
+
+Contigency table for RR mediaction.
+
+```{r self vs registered medication}
+tbl <- table(dset.species$BP_TREAT, dset.species$ANYDRUG)
+colnames(tbl) <- c("ANYDRUG 0", "ANYDRUG 1")
+rownames(tbl) <- c("BP_TREAT 0", "BP_TREAT 1")
+tbl %>%
+    kable(.)
+```
+
+$\chi^2$-test for medication `r format.pval(chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value)`.
+
+# Self reported vs. registered medication
+
+Contigency table for RR mediaction.
+
+```{r self vs registered medication}
+tbl <- table(dset.species$BP_TREAT, dset.species$ANYDRUG)
+colnames(tbl) <- c("ANYDRUG 0", "ANYDRUG 1")
+rownames(tbl) <- c("BP_TREAT 0", "BP_TREAT 1")
+tbl %>%
+    kable(.)
+```
+
+$\chi^2$-test for medication `r format.pval(chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value)`.
+
 # How large subgroups are?
@@
 ```
-
-# Self reported vs. registered medication
-
-Contigency table for RR mediaction.
-
-```{r self vs registered medication}
-tbl <- table(dset.species$BP_TREAT, dset.species$ANYDRUG)
-colnames(tbl) <- c("ANYDRUG 0", "ANYDRUG 1")
-rownames(tbl) <- c("BP_TREAT 0", "BP_TREAT 1")
-tbl %>%
-    kable(.)
-```
-
-$\chi^2$-test for medication `r format.pval(chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value)`.
 
@@
 
-```{r adonis3, eval = FALSE}
+```{r adonis3}
  adonis.species %>%

### 2019-01-02 rrbiome.Rmd
Removing duplication
Missing bracket

@@
 
-Function definitions
+# Function definitions
 
@@
 }
+
+calculateadonis <- function(dset,
+                            matrix,
+                            responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
+                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
+                            covariates = "BL_AGE + SEX",
+                            modelstring = "matrix ~ %s + %s",
+                            npermutations = 99) {
+    ret  <- mclapply(responses, function(response) {
+        fo <- sprintf(modelstring, response, covariates)
+        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
+        ad$aov.tab %>%
+            as.data.frame %>%
+            tibble::rownames_to_column(var = "term") %>%
+            dplyr::mutate(response = response, fo = fo)
+      }, mc.cores = length(responses))
+    data.table::rbindlist(ret, id = "model_name") %>%
+        dplyr::filter(term %in% responses) %>%
+            dplyr::select(-term)
+}
+
+calculateglm <- function(dset,
+                         responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
+                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
+                         min_n_for_continuous = 10,
+                         modelstring = "%s ~ BL_AGE + SEX") {
+    glmlist <- lapply(responses, function(response) {
+        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
+        fo <- sprintf(modelstring, response)
+        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
+            broom::tidy() %>%
+                dplyr::filter(grepl("diversities_", term)) %>%
+                dplyr::mutate(response = response, fo = fo) %>%
+                dplyr::select(-term)
+    })
+    data.table::rbindlist(glmlist, id = "model_name") %>%
+        as.data.frame %>%
+        dplyr::mutate(conf.low = estimate - qnorm(1- 0.05/2) * std.error,
+                      conf.high = estimate + qnorm(1- 0.05/2) * std.error)
+}
 ```
+
+# Loading data
 
@@
 
-Sample data has dimensions (`r dim(sample_data(pseq.species)`).
-
-# Self reported vs. registered medication
-
-Contigency table for RR mediaction.
-
-```{r self vs registered medication}
-tbl <- table(dset.species$BP_TREAT, dset.species$ANYDRUG)
-colnames(tbl) <- c("ANYDRUG 0", "ANYDRUG 1")
-rownames(tbl) <- c("BP_TREAT 0", "BP_TREAT 1")
-tbl %>%
-    kable(.)
-```
-
-$\chi^2$-test for medication `r format.pval(chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value)`.
+Sample data has dimensions (`r dim(sample_data(pseq.species))`).
 
@@
 
-```{r function for glm calculation, include = FALSE}
-calculateglm <- function(dset,
-                         responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
-                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
-                         min_n_for_continuous = 10,
-                         modelstring = "%s ~ BL_AGE + SEX") {
-    glmlist <- lapply(responses, function(response) {
-        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
-        fo <- sprintf(modelstring, response)
-        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
-            broom::tidy() %>%
-                dplyr::filter(grepl("diversities_", term)) %>%
-                dplyr::mutate(response = response, fo = fo) %>%
-                dplyr::select(-term)
-    })
-    data.table::rbindlist(glmlist, id = "model_name") %>%
-        as.data.frame %>%
-        dplyr::mutate(conf.low = estimate - qnorm(1- 0.05/2) * std.error,
-                      conf.high = estimate + qnorm(1- 0.05/2) * std.error)
-}
-```
-
 Calculating linear models for alpha diversity.
@@
 
-Beetadiversiteetti-matriisin lasku. Tiedosista *comp_bc_dissimilarity_matrix.R* ja käyttö *comp_adonis_covariates.R*.
+Calculating the distance matrix from data after compositional
+transformation using vegdist and bray method.
 
@@
 
-Mallin lasku Adoniksella
-
-```{r function for adnoiscalculation, include = FALSE}
-calculateadonis <- function(dset,
-                            matrix,
-                            responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
-                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
-                            covariates = "BL_AGE + SEX",
-                            modelstring = "matrix ~ %s + %s",
-                            npermutations = 99) {
-    ret  <- mclapply(responses, function(response) {
-        fo <- sprintf(modelstring, response, covariates)
-        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
-        ad$aov.tab %>%
-            as.data.frame %>%
-            tibble::rownames_to_column(var = "term") %>%
-            dplyr::mutate(response = response, fo = fo)
-      }, mc.cores = length(responses))
-    data.table::rbindlist(ret, id = "model_name") %>%
-        dplyr::filter(term %in% responses) %>%
-            dplyr::select(-term)
-}
-```
+Calculating beta diversity for blood pressure variables using adonis
 
@@
 ```
+
+Printing the results for adonis run for blood pressure variables in
+full model
 

### 2019-01-02 rrbiome.Rmd
pca

@@
 }
+calculateglm <- function(dset,
+                         responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
+                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
+                         min_n_for_continuous = 10,
+                         modelstring = "%s ~ BL_AGE + SEX") {
+    glmlist <- lapply(responses, function(response) {
+        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
+        fo <- sprintf(modelstring, response)
+        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
+            broom::tidy() %>%
+                dplyr::filter(grepl("diversities_", term)) %>%
+                dplyr::mutate(response = response, fo = fo) %>%
+                dplyr::select(-term)
+    })
+    data.table::rbindlist(glmlist, id = "model_name") %>%
+        as.data.frame %>%
+        dplyr::mutate(conf.low = estimate - qnorm(1- 0.05/2) * std.error,
+                      conf.high = estimate + qnorm(1- 0.05/2) * std.error)
+}
 
@@
             dplyr::select(-term)
-}
-
-calculateglm <- function(dset,
-                         responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
-                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
-                         min_n_for_continuous = 10,
-                         modelstring = "%s ~ BL_AGE + SEX") {
-    glmlist <- lapply(responses, function(response) {
-        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
-        fo <- sprintf(modelstring, response)
-        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
-            broom::tidy() %>%
-                dplyr::filter(grepl("diversities_", term)) %>%
-                dplyr::mutate(response = response, fo = fo) %>%
-                dplyr::select(-term)
-    })
-    data.table::rbindlist(glmlist, id = "model_name") %>%
-        as.data.frame %>%
-        dplyr::mutate(conf.low = estimate - qnorm(1- 0.05/2) * std.error,
-                      conf.high = estimate + qnorm(1- 0.05/2) * std.error)
 }
@@
 
-```{r pca, eval = FALSE}
-pseq.clr <- microbiome::transform(pseq_filtered, "clr")
-X <- as.matrix(t(abundances(pseq.clr)))
-prcomp <- stats::prcomp(X)
+We calculate primary component analysis using centered log-ratio
+transformation for the abunances.
+
+```{r pca}
+if (!file.exists("rds/pca.prcomp.rds")) {
+    pca.abundances  <- as.matrix(t(microbiome::abundances(microbiome::transform(pseq.species, "clr"))))
+    pca.prcomp <- stats::prcomp(pca.abundances)
+    saveRDS(pca.prcomp, file = "rds/pca.prcomp.rds")
+} else {
+     pca.prcomp <- readRDS("rds/pca.prcomp.rds")
+}
 ```
 
-```{r pca labels, eval = FALSE}
-percentage <- round(prcomp$sdev/sum(prcomp$sdev) * 100, 2)
-pca.labels <- paste0(colnames(prcomp$x), " (", paste0(as.character(percentage), "%", ")"))
+```{r pca percentage}
+(pca.prcomp$sdev / sum(pca.prcomp$sdev)) %>% head %>% kable
+```
+
+Labels for PCA plot containing the
+
+```{r pca labels}
+pca.percentage <- round(pca.prcomp$sdev / sum(pca.prcomp$sdev) * 100, 2)
+pca.labels <- paste0(colnames(pca.prcomp$x), " (", paste0(as.character(pca.percentage), "%", ")"))
 ```

### 2019-01-02 rrbiome.Rmd
PCA ready?

@@
 
+Proportion of variance for main PC axis
+
 ```{r pca percentage}
-(pca.prcomp$sdev / sum(pca.prcomp$sdev)) %>% head %>% kable
+pca.eigs <- pca.prcomp$sdev^2
+pca.percentage <- 100 * eigs /sum(eigs)
+pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)
+pca.percentage %>% head %>% kable
 ```
 
-Labels for PCA plot containing the
+PCA plot for hypertension
 
-```{r pca labels}
-pca.percentage <- round(pca.prcomp$sdev / sum(pca.prcomp$sdev) * 100, 2)
-pca.labels <- paste0(colnames(pca.prcomp$x), " (", paste0(as.character(pca.percentage), "%", ")"))
-```
-
-```{r PCA plots categorical, fig.width = 10, fig.height = 10, dpi = 300, eval = FALSE}
-ggplot((cbind(meta(pseq_filtered), prcomp$x[,1:2]) %>%
-        select(PC1, PC2, BP_TREAT, CURR_SMOKE, DIAB, BL_USE_RX_C09, Q57X,
-               BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08, HYPERTENSION, SEX) %>%
-        gather(Covariate, Value, -PC1, -PC2)), 
-       aes(x=PC1, y=PC2)) +
-    facet_wrap(~Covariate, ncol=2, scales = "free") +
-    geom_point(aes(color = Value), size = 0.5) +
-    scale_color_brewer(palette="Dark2") +
+```{r PCA plots continuous, fig.width = 10, fig.height = 4, dpi = 300}
+ggplot(cbind(meta(pseq.species), pca.prcomp$x[,1:2]), aes(x=PC1, y=PC2)) +
+    geom_point(aes(color = HYPERTENSION), size = 0.2) +
     xlab(pca.labels[1]) +
     ylab(pca.labels[2]) +
+    scale_color_brewer(palette="Dark2") +
     ggplot2::theme_minimal() 
-
-```
-
-```{r PCA plots continuous, fig.width = 10, fig.height = 10, dpi = 300,  eval = FALSE}
-
-ggplot((cbind(meta(pseq_filtered), prcomp$x[,1:2]) %>%
-        select(PC1, PC2, BL_AGE, SYSTM, DIASM, BMI, MAP, PULSEPRESSURE) %>%
-        gather(Covariate, Value, -PC1, -PC2)), 
-       aes(x=PC1, y=PC2)) +
-    facet_wrap(~Covariate, ncol=2, scales = "free") +
-    geom_point(aes(color = Value), size = 0.5) +
-    scale_color_gradientn(colours = rainbow(12)) +
-    xlab(pca.labels[1]) +
-    ylab(pca.labels[2]) +
-    ggplot2::theme_minimal()
-
 ```

### 2019-01-02 rrbiome.Rmd
renamed var

@@
 pca.eigs <- pca.prcomp$sdev^2
-pca.percentage <- 100 * eigs /sum(eigs)
+pca.percentage <- 100 * pca.eigs /sum(pca.eigs)
 pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)

### 2019-01-02 rrbiome.Rmd
minor

@@
 
-```{r libraries}
+```{r libraries, cache = FALSE}
 library(dplyr)
@@
 ```{r updating phyloseq objects}
-sample_data(pseq.genus) <- sample_data(dset.genus)
-sample_data(pseq.species) <- sample_data(dset.species)
+phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(dset.genus)
+phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(dset.species)
 ```

### 2019-01-02 rrbiome.Rmd
Colnames

@@
 pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)
-pca.percentage %>% head %>% kable
+setNames(pca.percentage, pca.prcomp$x %>% colnames) %>% head(10) %>% t %>% kable
 ```
@@
 
-```{r PCA plots continuous, fig.width = 10, fig.height = 4, dpi = 300}
+```{r PCA plots continuous, fig.width = 10, fig.height = 6, dpi = 300}
 ggplot(cbind(meta(pseq.species), pca.prcomp$x[,1:2]), aes(x=PC1, y=PC2)) +

### 2019-01-02 rrbiome.Rmd
Maaslin

@@
 
-# Genus
+# Connection between genus and blood pressure
 
@@
 
-```{r prepare maaslin, eval = FALSE}
+```{r prepare maaslin}
 
-pheno <- t(meta(sample_data(pseq_filtered)))
+prepare.maaslin <- function (pseq,
+                             tsvfile = "merged_phfinrisk_genus.tsv",
+                             conffile = "maaslin_config.config",
+                             includedvars = c("BL_AGE", "BMI", "SEX")) {
+    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
     sample.ids <- colnames(pheno)
-phfg.pseq.rel <- microbiome::transform(pseq_filtered, "compositional")
+    phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
     abud.rel <- abundances(phfg.pseq.rel)
-
-rownames(abud.rel) <- sapply(rownames(abud.rel),
-                             function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
-
+    rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
     merged <- rbind(sample.ids, pheno, abud.rel)
-write.table(t(merged), file="merged_phfinrisk_genus.tsv", sep='\t', 
-            quote=FALSE, row.names=FALSE, col.names=TRUE)
-
-variables <- rownames(pheno)
-variables_wo_ctrls <- variables[!variables %in% c("BL_AGE","BMI","SEX")]
-data <- file("maaslin_config.config" , open='wt')
+    write.table(t(merged), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
+    data <- file(conffile , open='wt')
     pcl.contents <- c("Matrix: Metadata",
-                  paste(c("Read_PCL_Rows: BL_AGE,BMI,SEX",variables_wo_ctrls), collapse=","),
+                      paste(c("Read_PCL_Rows: ", paste(includedvars, collapse=",")), collapse = ""),
                       "",
@@
     close(data)
+}
+
 ```
 
-```{r run maaslin, eval = FALSE} 
+```{r run maaslin} 
+maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
+prepare.maaslin(pseq.genus,
+                 tsvfile = "merged_phfinrisk_genus.tsv",
+                conffile = "maaslin_config.config",
+                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars))
+
 Maaslin("merged_phfinrisk_genus.tsv",
@@
         fAllvAll = TRUE,
-        strForcedPredictors = c("BL_AGE", "BMI", "SEX"))
+        strForcedPredictors = maaslin.force.vars)
 

### 2019-01-02 rrbiome.Rmd
White space
Functional

--- 2019-01-02 rrbiome.Rmd
@@
                      Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
+names.dset <- bind_rows(names.dset,
+                        data.frame(Covariate = c("MAP", "PULSEPRESSURE", "HYPERTENSION"),
+                                   Category = rep("Physical", 3),
+                                   Name =c( "Mean artrerial pressure", "Pulse pressure", "Hypertension")))
 ```
@@
 
-```{r prepare maaslin, eval = FALSE}
+```{r prepare maaslin}
 
-pheno <- t(meta(sample_data(pseq_filtered)))
+get.merged.dataset <- function(pseq) {
+    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
     sample.ids <- colnames(pheno)
-phfg.pseq.rel <- microbiome::transform(pseq_filtered, "compositional")
+    phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
     abud.rel <- abundances(phfg.pseq.rel)
+    rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
+    t(rbind(sample.ids, pheno, abud.rel))
+}
 
-rownames(abud.rel) <- sapply(rownames(abud.rel),
-                             function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
-
-merged <- rbind(sample.ids, pheno, abud.rel)
-write.table(t(merged), file="merged_phfinrisk_genus.tsv", sep='\t', 
-            quote=FALSE, row.names=FALSE, col.names=TRUE)
-
-variables <- rownames(pheno)
-variables_wo_ctrls <- variables[!variables %in% c("BL_AGE","BMI","SEX")]
-data <- file("maaslin_config.config" , open='wt')
+prepare.maaslin <- function (pseq,
+                             tsvfile = "merged_phfinrisk_genus.tsv",
+                             conffile = "maaslin_config.config",
+                             includedvars = c("BL_AGE", "BMI", "SEX")) {
+    write.table(get.merged.dataset(pseq), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
+    data <- file(conffile , open='wt')
     pcl.contents <- c("Matrix: Metadata",
-                  paste(c("Read_PCL_Rows: BL_AGE,BMI,SEX",variables_wo_ctrls), collapse=","),
+                      paste(c("Read_PCL_Rows: ", paste(includedvars, collapse=",")), collapse = ""),
                       "",
@@
     close(data)
+}
+
+ merge.pheno.abu <- function(pseq) {
+     pheno <- microbiome::meta(phyloseq::sample_data(pseq))
+     sample.ids <- rownames(pheno)
+     phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
+     abud.rel <- abundances(phfg.pseq.rel)
+     rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
+     cbind(pheno, t(abud.rel))
+}
 ```
 
-```{r run maaslin, eval = FALSE} 
-Maaslin("merged_phfinrisk_genus.tsv",
+Running maaslin for blood pressure variables while rest of the
+phenotype variables are forced in the model
+
+```{r run maaslin} 
+maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
+
+prepare.maaslin(pseq.genus,
+                 tsvfile = "maaslin_merged_phfinrisk_genus.tsv",
+                conffile = "maaslin_config.config",
+                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars))
+
+Maaslin("maaslin_merged_phfinrisk_genus.tsv",
         "maaslin",
@@
         fAllvAll = TRUE,
-        strForcedPredictors = c("BL_AGE", "BMI", "SEX"))
-
+        strForcedPredictors = maaslin.force.vars)
 ```
@@
 
-```{r Reading maaslins output, eval = FALSE}
-
+```{r Reading maaslins output}
 maas.results.df <- read.table("maaslin/merged_phfinrisk_genus.txt", header=TRUE, sep='\t')
 maas.results.df[,'Covariate'] <- maas.results.df[,'Variable']
-maas.results.df <- merge(maas.results.df,
-                         names.dset[,c('Covariate','Category', 'Name')],
-                         by=c('Covariate'))
-
+maas.results.df <- merge(maas.results.df, names.dset[,c('Covariate','Category', 'Name')], by=c('Covariate'))
 maas.results.df[,'Maaslin.Q.value'] <- maas.results.df[,'Q.value']
@@
 
-```{r scatter plots, eval = FALSE}
-
+```{r scatter plots}
 sysm.signf.features <- maas.results.df.signf %>%
-    dplyr::filter(Covariate == "SYSTM") %>%
     pull(Feature) %>%
@@
     gsub("_", ".", .)
-df.merged <- t(merged) %>% as.data.frame
+
+df.merged <-  merge.pheno.abu(pseq.genus)
 ```
 
-```{r maaslin scatter plot, fig.width = 10, fig.height = 20, dpi = 300, eval = FALSE}
-
-#df.merged %>% pull("Rikenella.Bacteria") %>% as.character I%>% as.numeric %>% sqrt %>% asin
-
-ggplot(df.merged %>%
-       select(sysm.signf.features, SYSTM, SEX) %>%
-       gather(Covariate, Value, sysm.signf.features) %>%
-       mutate(SYSTM = as.numeric(as.character(SYSTM)),
-              Value = as.numeric(as.character(Value))),
-       aes(x = asin(sqrt(Value)), y = SYSTM)) +
+```{r maaslin scatter plot, fig.width = 10, fig.height = 20, dpi = 300}
+ggplot(df.merged %>% gather(Covariate, relabundance, sysm.signf.features), aes(x = asin(sqrt(relabundance)), y = SYSTM)) +
     facet_wrap(~Covariate, ncol=2, scales = "free") +
-    geom_point(aes(color = SEX), size = 0.5) +
+    geom_point(aes(color = SEX), size = 0.2) +
     scale_y_continuous(breaks=seq(0,200,40)) +
@@
 ```
+
+```{r maaslin plot}
+source("edgeR_plot_functions.R")
+
+plot_maaslin_heatmap_final(maas.results.df.signf,
+                      xlab.text="Genera",
+                      beta.maxabs=max(abs(maas.results.df.signf[,'beta'])),
+                      order.by.category=FALSE)
+```

### 2019-01-02 rrbiome.Rmd
Minor

@@
 
-```{r run maaslin} 
+```{r run maaslin, echo = FALSE, warnings = FALSE, message=FALSE} 
 maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
@@
 
-Data framen luku *comp_maaslin_results_from_single_job.R* ja plottaus käyttäen *comp_maaslin_plots.R* ja *edgeR_plot_functions.R*.
+Reading and filtering the data Maaslin provided
 
 ```{r Reading maaslins output}
-maas.results.df <- read.table("maaslin/merged_phfinrisk_genus.txt", header=TRUE, sep='\t')
-maas.results.df[,'Covariate'] <- maas.results.df[,'Variable']
-maas.results.df <- merge(maas.results.df, names.dset[,c('Covariate','Category', 'Name')], by=c('Covariate'))
-maas.results.df[,'Maaslin.Q.value'] <- maas.results.df[,'Q.value']
-maas.results.df[,'Qval'] <- p.adjust(maas.results.df[,'P.value'], method="BH")
-maas.results.df[,'beta'] <- maas.results.df[,'Coefficient']
-maas.results.df.signf <- dplyr::filter(maas.results.df, Qval < 0.05)
+maas.results.df.signf <- read.table("maaslin/merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
+    dplyr::mutate(Qval = p.adjust(maas.results.df[,'P.value'], method="BH"),
+           Beta = Coefficient) %>%
+   dplyr::filter(Qval < 0.05)
 
-```
-
-```{r scatter plots}
-sysm.signf.features <- maas.results.df.signf %>%
+maas.signf.features <- maas.results.df.signf %>%
     pull(Feature) %>%
@@
 
+Table for significant relations between blood pressure variables and genera
+
+```{r maaslin significant table}
+maas.results.df.signf %>%
+    select(Name, Feature, Beta, Qval) %>%
+    kable           
+```
+
+A scatter plot for the significant genera against blood
+pressuer. Compositional data is transformed using 
+$\textrm{arcsin}(\textrm{sqrt} (\cdot))$.
+
 ```{r maaslin scatter plot, fig.width = 10, fig.height = 20, dpi = 300}
-ggplot(df.merged %>% gather(Covariate, relabundance, sysm.signf.features), aes(x = asin(sqrt(relabundance)), y = SYSTM)) +
+ggplot(df.merged %>% gather(Covariate, relabundance, maas.signf.features), aes(x = asin(sqrt(relabundance)), y = SYSTM)) +
     facet_wrap(~Covariate, ncol=2, scales = "free") +

### 2019-01-02 rrbiome.Rmd
Maaslin

@@
 ```{r prepare maaslin}
-
-get.merged.dataset <- function(pseq) {
+prepare.maaslin <- function (pseq,
+                             tsvfile = "merged_phfinrisk_genus.tsv",
+                             conffile = "maaslin_config.config",
+                             includedvars = c("BL_AGE", "BMI", "SEX")) {
     pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
@@
     rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
-    t(rbind(sample.ids, pheno, abud.rel))
-}
-
-prepare.maaslin <- function (pseq,
-                             tsvfile = "merged_phfinrisk_genus.tsv",
-                             conffile = "maaslin_config.config",
-                             includedvars = c("BL_AGE", "BMI", "SEX")) {
-    write.table(get.merged.dataset(pseq), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
+    write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=TRUE, row.names=TRUE, col.names=NA)
     data <- file(conffile , open='wt')
@@
     merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
-    dplyr::mutate(Qval = p.adjust(maas.results.df[,'P.value'], method="BH"),
+    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
            Beta = Coefficient) %>%
@@
 
-```{r maaslin plot}
+```{r maaslin plot, message = FALSE, echo = FALSE}
 source("edgeR_plot_functions.R")
@@
 ```
+
+# Lactobacillus
+
+```{r lactobacillus}
+
+taxa <- subset_taxa(pseq.species, Genus=="Lactobacillus") %>% taxa()
+
+```

### 2019-01-02 rrbiome.Rmd
Bug

@@
     rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
-    write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=TRUE, row.names=TRUE, col.names=NA)
+    write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
     data <- file(conffile , open='wt')
@@
 
-```{r lactobacillus}
+```{r lactobacillus, eval = FALSE}
+pseq.lacto  <- subset_taxa(pseq.species, Genus=="Lactobacillus")
+pseq.lacto.comp  <- microbiome::transform(pseq.lacto, "compositional")
+abud.rel.lacto <- abundances(pseq.lacto.comp)
 
-taxa <- subset_taxa(pseq.species, Genus=="Lactobacillus") %>% taxa()
+abud.rel.lacto %>%
+    gather(Key, Value) %>%
+    group_by(Subject, Type) %>%
+     summarise_each(funs(mean= mean(., na.rm=TRUE)))
 

### 2019-01-02 rrbiome.Rmd
Lactobacillus

@@
 ```{r Reading maaslins output}
-maas.results.df.signf <- read.table("maaslin/merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
+maas.results.df.signf <- read.table("maaslin/maaslin_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
     merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
@@
 
-```{r lactobacillus, eval = FALSE}
+```{r lactobacillus}
 pseq.lacto  <- subset_taxa(pseq.species, Genus=="Lactobacillus")
@@
 abud.rel.lacto %>%
-    gather(Key, Value) %>%
-    group_by(Subject, Type) %>%
-     summarise_each(funs(mean= mean(., na.rm=TRUE)))
-
+    t %>%
+    as.data.frame %>%
+    #dplyr::select("Lactobacillus_paracasei (BacteriaPlasmid)", "Lactobacillus_selangorensis (Bacteria)") %>%
+    gather(key, value) %>%
+    group_by(key) %>%
+    summarise_all(funs(mean= mean(value),
+                       min = min(value),
+                       max = max(value),
+                       sd = sd(value),
+                       n = n(),
+                       nonzeros = sum(abs(value) > 1e-12),
+                       zeros = n - nonzeros)) %>%
+    arrange(zeros) %>%
+    kable
 ```

### 2019-01-02 rrbiome.Rmd
Minor typo

@@
                       xlab.text="Genera",
-                      beta.maxabs=max(abs(maas.results.df.signf[,'beta'])),
+                      beta.maxabs=max(abs(maas.results.df.signf[,'Beta'])),
                       order.by.category=FALSE)

### 2019-01-02 rrbiome.Rmd
Minor

--- 2019-01-02 rrbiome.Rmd
@@
     dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-           Beta = Coefficient) %>%
+           beta = Coefficient) %>%
    dplyr::filter(Qval < 0.05)
@@
 maas.results.df.signf %>%
-    select(Name, Feature, Beta, Qval) %>%
+    select(Name, Feature, beta, Qval) %>%
     kable           

### 2019-01-02 rrbiome.Rmd
lacto maaslin

@@
                              conffile = "maaslin_config.config",
-                             includedvars = c("BL_AGE", "BMI", "SEX")) {
+                             includedvars = c("BL_AGE", "BMI", "SEX"),
+                             readpclrows = "Archaea.Archaea-") {
     pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
@@
     abud.rel <- abundances(phfg.pseq.rel)
-    rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
+    rownames(abud.rel) <- sapply(rownames(abud.rel),
+                                 function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
     write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
@@
                       "Matrix: Adundance",
-                      "Read_PCL_Rows: Archaea.Archaea-")
+                      paste("Read_PCL_Rows:", readpclrows))
     writeLines(pcl.contents, con=data)
@@
 ```
+
+```{r lactobacillus scatterplot}
+df.merged.lacto <-  merge.pheno.abu(pseq.lacto)
+
+colnames(df.merged.lacto)
+
+prepare.maaslin(pseq.lacto,
+                 tsvfile = "maaslin_lacto_merged_phfinrisk_genus.tsv",
+                conffile = "maaslin_lacto_config.config",
+                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars),
+                readpclrows = "-MAP")
+
+Maaslin("maaslin_lacto_merged_phfinrisk_genus.tsv",
+        "maaslin_lacto",
+        strInputConfig = "maaslin_lacto_config.config",
+        strModelSelection="none",
+        dSignificanceLevel = 1.0, 
+        fAllvAll = TRUE,
+        strForcedPredictors = maaslin.force.vars)
+
+```

### 2019-01-02 rrbiome.Rmd
Final

@@
 
-```{r lactobacillus scatterplot}
-df.merged.lacto <-  merge.pheno.abu(pseq.lacto)
-
-colnames(df.merged.lacto)
+```{r lactobacillus scatterplot, eval = FALSE, include = FALSE}
+maaslin.lacto.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "DIAB")
 
@@
                 conffile = "maaslin_lacto_config.config",
-                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars),
-                readpclrows = "-MAP")
+                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.lacto.force.vars),
+                readpclrows = "Lactobacillus_acetotolerans.Bacteria-")
 
@@
         fAllvAll = TRUE,
-        strForcedPredictors = maaslin.force.vars)
+        strForcedPredictors = maaslin.lacto.force.vars)
+```
+
+```{r Reading maaslin lacto output, eval = FALSE, include = FALSE}
+maas.lacto.results.df.signf <- read.table("maaslin_lacto/maaslin_lacto_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
+    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
+                  beta = Coefficient)
+
+                                        #%>%
+   dplyr::filter(Qval < 0.05)
+
+maas.lacto.signf.features <- maas.results.df.signf %>%
+    pull(Feature) %>%
+    as.character() %>%
+    gsub("_", ".", .)
+
+df.lacto.merged <-  merge.pheno.abu(pseq.lacto)
+
+maas.lacto.results.df.signf %>% select(Feature, Qval, P.value)
 
 ```
+
+```{r maaslin lacto significant table, eval = FALSE, include = FALSE}
+maas.lacto.results.df.signf %>%
+    select(Name, Feature, beta, Qval) %>%
+    kable           
+```

### 2019-01-03 rrbiome.Rmd
Minor

@@
     dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  beta = Coefficient)
-
-                                        #%>%
+                  beta = Coefficient) %>%
    dplyr::filter(Qval < 0.05)
@@
 df.lacto.merged <-  merge.pheno.abu(pseq.lacto)
-
-maas.lacto.results.df.signf %>% select(Feature, Qval, P.value)
-
 ```

### 2019-01-03 rrbiome.Rmd
Minor

@@
             dplyr::select(-term)
+}
+
+prepare.maaslin <- function (pseq,
+                             tsvfile = "merged_phfinrisk_genus.tsv",
+                             conffile = "maaslin_config.config",
+                             includedvars = c("BL_AGE", "BMI", "SEX"),
+                             readpclrows = "Archaea.Archaea-") {
+    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
+    sample.ids <- colnames(pheno)
+    phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
+    abud.rel <- abundances(phfg.pseq.rel)
+    rownames(abud.rel) <- sapply(rownames(abud.rel),
+                                 function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
+    write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
+    data <- file(conffile , open='wt')
+    pcl.contents <- c("Matrix: Metadata",
+                      paste(c("Read_PCL_Rows: ", paste(includedvars, collapse=",")), collapse = ""),
+                      "",
+                      "Matrix: Adundance",
+                      paste("Read_PCL_Rows:", readpclrows))
+    writeLines(pcl.contents, con=data)
+    close(data)
+}
+
+ merge.pheno.abu <- function(pseq) {
+     pheno <- microbiome::meta(phyloseq::sample_data(pseq))
+     sample.ids <- rownames(pheno)
+     phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
+     abud.rel <- abundances(phfg.pseq.rel)
+     rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
+     cbind(pheno, t(abud.rel))
 }
@@
 
-Aaron work-december-2018 *prepare_maaslin_input.R*
-
-```{r prepare maaslin}
-prepare.maaslin <- function (pseq,
-                             tsvfile = "merged_phfinrisk_genus.tsv",
-                             conffile = "maaslin_config.config",
-                             includedvars = c("BL_AGE", "BMI", "SEX"),
-                             readpclrows = "Archaea.Archaea-") {
-    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
-    sample.ids <- colnames(pheno)
-    phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
-    abud.rel <- abundances(phfg.pseq.rel)
-    rownames(abud.rel) <- sapply(rownames(abud.rel),
-                                 function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
-    write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
-    data <- file(conffile , open='wt')
-    pcl.contents <- c("Matrix: Metadata",
-                      paste(c("Read_PCL_Rows: ", paste(includedvars, collapse=",")), collapse = ""),
-                      "",
-                      "Matrix: Adundance",
-                      paste("Read_PCL_Rows:", readpclrows))
-    writeLines(pcl.contents, con=data)
-    close(data)
-}
-
- merge.pheno.abu <- function(pseq) {
-     pheno <- microbiome::meta(phyloseq::sample_data(pseq))
-     sample.ids <- rownames(pheno)
-     phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
-     abud.rel <- abundances(phfg.pseq.rel)
-     rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
-     cbind(pheno, t(abud.rel))
-}
-```
-
 Running maaslin for blood pressure variables while rest of the
@@
 
-```{r run maaslin, echo = FALSE, warnings = FALSE, message=FALSE} 
+```{r run maaslin, results = 'hide'} 
 maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
+
+#pseq.core <- microbiome::core(pseq.genus, detection = 0.001, pervalence = 0.01)
 

### 2019-01-03 rrbiome.Rmd
Better scatter plot

@@
 ```{r maaslin scatter plot, fig.width = 10, fig.height = 20, dpi = 300}
-ggplot(df.merged %>% gather(Covariate, relabundance, maas.signf.features), aes(x = asin(sqrt(relabundance)), y = SYSTM)) +
+ggplot(df.merged %>% gather(Covariate, relabundance, maas.signf.features), aes(x = asin(sqrt(relabundance)), y = SYSTM, color = SEX)) +
     facet_wrap(~Covariate, ncol=2, scales = "free") +
-    geom_point(aes(color = SEX), size = 0.2) +
+    geom_point(size = 0.2) +
     scale_y_continuous(breaks=seq(0,200,40)) +
-    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
+    geom_smooth(method=lm, se=TRUE, size=0.5) +
     stat_poly_eq(formula = "y ~ x",
                  aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
-                 label.y = 0,
+                 label.y.npc = "bottom",
                  parse = TRUE,

### 2019-01-03 rrbiome.Rmd
Minor

@@
                  label.y.npc = "bottom",
+                 label.x.npc = "right",
                  parse = TRUE,

### 2019-01-03 rrbiome.Rmd
More permutations for adonis

@@
                     covariates = "BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
-                    npermutations = 99)
+                    npermutations = 500)
      saveRDS(adonis.species, file = "rds/adonis.species.rds")

### 2019-01-03 rrbiome.Rmd
Adonis fix?

--- 2019-01-03 rrbiome.Rmd
@@
                                           model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
-                            covariates = "BL_AGE + SEX",
-                            modelstring = "matrix ~ %s + %s",
-                            npermutations = 99) {
-    ret  <- mclapply(responses, function(response) {
-        fo <- sprintf(modelstring, response, covariates)
-        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
+                            covariates = c("BL_AGE", "SEX"),
+                            npermutations = 99,
+                            maxcores = 100) {
+    mclapply(responses, function(response) {
+        fo <- paste(response, covariates, sep = " + ", collapse = " + ")
+        ad <- adonis(as.formula(paste(matrix, fo, sep = "~")), data = dset, permutations = npermutations)
         ad$aov.tab %>%
@@
             tibble::rownames_to_column(var = "term") %>%
-            dplyr::mutate(response = response, fo = fo)
-      }, mc.cores = length(responses))
-    data.table::rbindlist(ret, id = "model_name") %>%
-        dplyr::filter(term %in% responses) %>%
-            dplyr::select(-term)
+            dplyr::mutate(response = response)
+      }, mc.cores = min(maxcores, length(responses)))
 }
@@
                     matrix = bray.dist.m.species,
-                    covariates = "BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
-                    npermutations = 99)
+                        covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
+                    npermutations = 500)
     saveRDS(adonis.species, file = "rds/adonis.species.rds")
+
 } else {
@@
 adonis.species %>%
+    dplyr::filter(term %in% responses) %>%
+    dplyr::select(-term) %>%
     select(response, SumsOfSqs, MeanSqs, F.Model, R2, "Pr(>F)") %>%

### 2019-01-03 rrbiome.Rmd
Full re run for adonis

@@
     mclapply(responses, function(response) {
-        fo <- paste(response, covariates, sep = " + ", collapse = " + ")
-        ad <- adonis(as.formula(paste(matrix, fo, sep = "~")), data = dset, permutations = npermutations)
+	fo <- sprintf("matrix ~ %s + %s", response, paste(covariates, collapse = " + "))
+        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
         ad$aov.tab %>%

### 2019-01-03 rrbiomefunctions.R
Functionq

@@ -1,0 +1,117 @@
+# Helper functions for articletwo
+
+replace.brackets <- function (genus) gsub('(.*) \\(+(.+)\\)', '\\1.\\2', genus)
+
+filter.phenotype.data <- function(pseq,
+                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
+                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
+                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
+    meta(pseq) %>%
+        tibble::rownames_to_column(var = "rowname") %>%
+        dplyr::select(rowname, included) %>%
+        dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
+                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
+                      ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
+                      PULSEPRESSURE = SYSTM - DIASM,
+                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
+                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
+                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
+         stats::na.omit() %>%
+        dplyr::select(-MEN, -DNA_OK) %>%
+        tibble::remove_rownames() %>%
+        tibble::column_to_rownames(var = "rowname")
+
+meta.merge.alphadiversity <- function(pseq, index = "shannon") {
+    alphadiversity  <- microbiome::global(pseq, index = index)
+    base::merge(meta(pseq), alphadiversity, by=0, all=TRUE) %>%
+        dplyr::rename(Sample_ID = Row.names) %>%
+            dplyr::mutate(BL_AGE_GROUP = group_age(BL_AGE))
+}
+
+calculate.beta.matrix <- function(pseq) {
+    compositional.profile <- microbiome::transform(pseq, 'compositional')
+    otu <- microbiome::abundances(compositional.profile)
+    bray.dist.m <- vegan::vegdist(t(otu), method="bray")
+    dist.matrix <- as.matrix(bray.dist.m)
[…]

### 2019-01-03 articletwo.Rmd
Debuging adonis

--- 2019-01-03 rrbiome.Rmd
@@
 library(parallel)
+library(officer)
+library(tableone)
 ```
@@
 
+```{r Importing functions}
+source("rrbiomefunctions.R", echo = TRUE)
 ```{r functions}
-filter.phenotype.data <- function(pseq,
-                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
-                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
-                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
-    meta(pseq) %>%
-        tibble::rownames_to_column(var = "rowname") %>%
-        dplyr::select(rowname, included) %>% 
-        dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
-                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
-                      ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
-                      PULSEPRESSURE = SYSTM - DIASM,
-                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
-                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
-                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
-         stats::na.omit() %>%
-        dplyr::select(-MEN, -DNA_OK) %>%
-        tibble::remove_rownames() %>%
-        tibble::column_to_rownames(var = "rowname")
-
-meta.merge.alphadiversity <- function(pseq, index = "shannon") {
-    alphadiversity  <- microbiome::global(pseq, index = index)
-    base::merge(meta(pseq), alphadiversity, by=0, all=TRUE) %>%
-        dplyr::rename(Sample_ID = Row.names) %>%
-            dplyr::mutate(BL_AGE_GROUP = group_age(BL_AGE))
-}
-
-calculate.beta.matrix <- function(pseq) {
-    compositional.profile <- microbiome::transform(pseq, 'compositional')
-    otu <- microbiome::abundances(compositional.profile)
-    bray.dist.m <- vegan::vegdist(t(otu), method="bray")
-    dist.matrix <- as.matrix(bray.dist.m)
-    attr(dist.matrix, "method") <- "bray"
-    dist.matrix
-}
-calculateglm <- function(dset,
-                         responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
-                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
-                         min_n_for_continuous = 10,
-                         modelstring = "%s ~ BL_AGE + SEX") {
-    glmlist <- lapply(responses, function(response) {
-        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
-        fo <- sprintf(modelstring, response)
-        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
-            broom::tidy() %>%
-                dplyr::filter(grepl("diversities_", term)) %>%
-                dplyr::mutate(response = response, fo = fo) %>%
-                dplyr::select(-term)
-    })
-    data.table::rbindlist(glmlist, id = "model_name") %>%
-        as.data.frame %>%
-        dplyr::mutate(conf.low = estimate - qnorm(1- 0.05/2) * std.error,
-                      conf.high = estimate + qnorm(1- 0.05/2) * std.error)
-}
-
-calculateadonis <- function(dset,
-                            matrix,
-                            responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
-                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
-                            covariates = c("BL_AGE", "SEX"),
-                            npermutations = 99,
-                            maxcores = 100) {
-    mclapply(responses, function(response) {
-	fo <- sprintf("matrix ~ %s + %s", response, paste(covariates, collapse = " + "))
-        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
-        ad$aov.tab %>%
-            as.data.frame %>%
-            tibble::rownames_to_column(var = "term") %>%
-            dplyr::mutate(response = response)
-      }, mc.cores = min(maxcores, length(responses)))
-}
-
-prepare.maaslin <- function (pseq,
-                             tsvfile = "merged_phfinrisk_genus.tsv",
-                             conffile = "maaslin_config.config",
-                             includedvars = c("BL_AGE", "BMI", "SEX"),
-                             readpclrows = "Archaea.Archaea-") {
-    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
-    sample.ids <- colnames(pheno)
-    phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
-    abud.rel <- abundances(phfg.pseq.rel)
-    rownames(abud.rel) <- sapply(rownames(abud.rel),
-                                 function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
-    write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
-    data <- file(conffile , open='wt')
-    pcl.contents <- c("Matrix: Metadata",
-                      paste(c("Read_PCL_Rows: ", paste(includedvars, collapse=",")), collapse = ""),
-                      "",
-                      "Matrix: Adundance",
-                      paste("Read_PCL_Rows:", readpclrows))
-    writeLines(pcl.contents, con=data)
-    close(data)
-}
-
- merge.pheno.abu <- function(pseq) {
-     pheno <- microbiome::meta(phyloseq::sample_data(pseq))
-     sample.ids <- rownames(pheno)
-     phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
-     abud.rel <- abundances(phfg.pseq.rel)
-     rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
-     cbind(pheno, t(abud.rel))
-}
 ```
@@
 Sample data has dimensions (`r dim(sample_data(pseq.species))`).
-
-# Self reported vs. registered medication
-
-Contigency table for RR mediaction.
-
-```{r self vs registered medication}
-tbl <- table(dset.species$BP_TREAT, dset.species$ANYDRUG)
-colnames(tbl) <- c("ANYDRUG 0", "ANYDRUG 1")
-rownames(tbl) <- c("BP_TREAT 0", "BP_TREAT 1")
-tbl %>%
-    kable(.)
-```
-
-$\chi^2$-test for medication `r format.pval(chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value)`.
-
-# How large subgroups are?
-
-Number of participants in each combination of all categorical variables
-
-```{r numbers by group}
-dset.species %>%
-    group_by(SEX, CURR_SMOKE,  HYPERTENSION, Q57X, DIAB,
-             BL_USE_RX_C09, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
-    summarise(n = n()) %>%
-    arrange(desc(n)) %>%
-    kable            
-```
-
-Number of particiapants in each combination after combining RR
-medication and exercice to binomial variables.
-
-```{r numbers by group reduced}
-dset.species %>%
-     group_by(SEX, CURR_SMOKE,  HYPERTENSION, DIAB, ANYDRUG, ANYEXERCICE) %>%
-    summarise(n = n()) %>%
-    arrange(desc(n)) %>%
-    kable
-```
-
-# Alpha diversity
-
-Dset with alpha diversity
-
-```{r alpha diversity}
-dset.alpha <- meta.merge.alphadiversity(pseq.species)
-```
-
-Scatterplots for alpha diversity by age group and sex.
-
-```{r alpha-MAP scatter plots, fig.width=10, fig.height=15}
-ggplot(dset.alpha %>% mutate(SEX = ifelse(SEX == 0, "Male", "Female")), aes(x=diversities_shannon, y=MAP)) +
-    facet_wrap(~BL_AGE_GROUP + SEX, ncol=2, scales = "free") +
-    geom_point(size=0.1, shape=1, aes(colour = ANYDRUG)) +
-    scale_colour_manual(values = c("red","blue", "green")) +
-    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
-    stat_poly_eq(formula = "y ~ x",
-                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
-                 label.y = 0,
-                 parse = TRUE,
-                 size = 3) +   
-    ylim(0, 200) +
-    ggplot2::theme_minimal() 
-```
-
-Calculating linear models for alpha diversity.
-
-```{r calculate full alpha model}
-calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon") %>%
-    select(response, estimate, std.error, conf.low, conf.high, p.value) %>%
-    kable(.)
-```
-
-Calculating linear model using binomial variable for exercice and medication.
-
-```{r calculate binomial alpha model}
-calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + ANYEXERCICE + DIAB + ANYDRUG  + diversities_shannon") %>%
-    select(response, estimate, std.error, conf.low, conf.high, p.value) %>%
-    kable(.)
-```
-
-Calculating linear model using minimal variables.
-
-```{r calculate minimal alpha model}
-calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + diversities_shannon") %>%
-    select(response, estimate, std.error, conf.low, conf.high, p.value) %>%
-    kable(.)
-```
-
-# Beta diversity
-
-Calculating the distance matrix from data after compositional
-transformation using vegdist and bray method.
-
-```{r matrix calculation, cache = FALSE}
-if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds")) {
-    bray.dist.m.species <- calculate.beta.matrix(pseq.species)
-    saveRDS(bray.dist.m.species, file = "rds/bray.dist.m.species.rds")
-    bray.dist.m.genus <- calculate.beta.matrix(pseq.genus)
-    saveRDS(bray.dist.m.genus, file = "rds/bray.dist.m.genus.rds")
-} else {
-    bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
-    bray.dist.m.genus <- readRDS("rds/bray.dist.m.genus.rds")
-}
-```
-
-Calculating beta diversity for blood pressure variables using adonis
-
- ```{r adonis calculation, cache = FALSE}
-if (!file.exists("rds/adonis.species.rds")) {
-    adonis.species <- calculateadonis(dset = meta(pseq.species),
-                    matrix = bray.dist.m.species,
-                        covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                    npermutations = 500)
-    saveRDS(adonis.species, file = "rds/adonis.species.rds")
-
-} else {
-     adonis.species  <- readRDS("rds/adonis.species.rds")
-}
-```
-
-Printing the results for adonis run for blood pressure variables in
-full model
-
-```{r adonis3}
-adonis.species %>%
-    dplyr::filter(term %in% responses) %>%
-    dplyr::select(-term) %>%
-    select(response, SumsOfSqs, MeanSqs, F.Model, R2, "Pr(>F)") %>%
-    kable(.)
-```
-
-# PCA
-
-We calculate primary component analysis using centered log-ratio
-transformation for the abunances.
-
-```{r pca}
-if (!file.exists("rds/pca.prcomp.rds")) {
-    pca.abundances  <- as.matrix(t(microbiome::abundances(microbiome::transform(pseq.species, "clr"))))
-    pca.prcomp <- stats::prcomp(pca.abundances)
-    saveRDS(pca.prcomp, file = "rds/pca.prcomp.rds")
-} else {
-     pca.prcomp <- readRDS("rds/pca.prcomp.rds")
-}
-```
-
-Proportion of variance for main PC axis
-
-```{r pca percentage}
-pca.eigs <- pca.prcomp$sdev^2
-pca.percentage <- 100 * pca.eigs /sum(pca.eigs)
-pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)
-setNames(pca.percentage, pca.prcomp$x %>% colnames) %>% head(10) %>% t %>% kable
-```
-
-PCA plot for hypertension
-
-```{r PCA plots continuous, fig.width = 10, fig.height = 6, dpi = 300}
-ggplot(cbind(meta(pseq.species), pca.prcomp$x[,1:2]), aes(x=PC1, y=PC2)) +
-    geom_point(aes(color = HYPERTENSION), size = 0.2) +
-    xlab(pca.labels[1]) +
-    ylab(pca.labels[2]) +
-    scale_color_brewer(palette="Dark2") +
-    ggplot2::theme_minimal() 
-```
-
-# Connection between genus and blood pressure
-
-Running maaslin for blood pressure variables while rest of the
-phenotype variables are forced in the model
-
-```{r run maaslin, results = 'hide'} 
-maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
-
-#pseq.core <- microbiome::core(pseq.genus, detection = 0.001, pervalence = 0.01)
-
-prepare.maaslin(pseq.genus,
-                 tsvfile = "maaslin_merged_phfinrisk_genus.tsv",
-                conffile = "maaslin_config.config",
-                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars))
-
-Maaslin("maaslin_merged_phfinrisk_genus.tsv",
-        "maaslin",
-        strInputConfig = "maaslin_config.config",
-        strModelSelection="none",
-        dSignificanceLevel = 1.0, 
-        fAllvAll = TRUE,
-        strForcedPredictors = maaslin.force.vars)
-```
-
-Reading and filtering the data Maaslin provided
-
-```{r Reading maaslins output}
-maas.results.df.signf <- read.table("maaslin/maaslin_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
-    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-           beta = Coefficient) %>%
-   dplyr::filter(Qval < 0.05)
-
-maas.signf.features <- maas.results.df.signf %>%
-    pull(Feature) %>%
-    as.character() %>%
-    gsub("_", ".", .)
-
-df.merged <-  merge.pheno.abu(pseq.genus)
-```
-
-Table for significant relations between blood pressure variables and genera
-
-```{r maaslin significant table}
-maas.results.df.signf %>%
-    select(Name, Feature, beta, Qval) %>%
-    kable           
-```
-
-A scatter plot for the significant genera against blood
-pressuer. Compositional data is transformed using 
-$\textrm{arcsin}(\textrm{sqrt} (\cdot))$.
-
-```{r maaslin scatter plot, fig.width = 10, fig.height = 20, dpi = 300}
-ggplot(df.merged %>% gather(Covariate, relabundance, maas.signf.features), aes(x = asin(sqrt(relabundance)), y = SYSTM, color = SEX)) +
-    facet_wrap(~Covariate, ncol=2, scales = "free") +
-    geom_point(size = 0.2) +
-    scale_y_continuous(breaks=seq(0,200,40)) +
-    geom_smooth(method=lm, se=TRUE, size=0.5) +
-    stat_poly_eq(formula = "y ~ x",
-                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
-                 label.y.npc = "bottom",
-                 label.x.npc = "right",
-                 parse = TRUE,
-                 size = 3) +   
-    ggplot2::theme_minimal() 
-```
-
-```{r maaslin plot, message = FALSE, echo = FALSE}
-source("edgeR_plot_functions.R")
-
-plot_maaslin_heatmap_final(maas.results.df.signf,
-                      xlab.text="Genera",
-                      beta.maxabs=max(abs(maas.results.df.signf[,'beta'])),
-                      order.by.category=FALSE)
-```
-
-# Lactobacillus
-
-```{r lactobacillus}
-pseq.lacto  <- subset_taxa(pseq.species, Genus=="Lactobacillus")
-pseq.lacto.comp  <- microbiome::transform(pseq.lacto, "compositional")
-abud.rel.lacto <- abundances(pseq.lacto.comp)
-
-abud.rel.lacto %>%
-    t %>%
-    as.data.frame %>%
-    #dplyr::select("Lactobacillus_paracasei (BacteriaPlasmid)", "Lactobacillus_selangorensis (Bacteria)") %>%
-    gather(key, value) %>%
-    group_by(key) %>%
-    summarise_all(funs(mean= mean(value),
-                       min = min(value),
-                       max = max(value),
-                       sd = sd(value),
-                       n = n(),
-                       nonzeros = sum(abs(value) > 1e-12),
-                       zeros = n - nonzeros)) %>%
-    arrange(zeros) %>%
-    kable
-```
-
-```{r lactobacillus scatterplot, eval = FALSE, include = FALSE}
-maaslin.lacto.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "DIAB")
-
-prepare.maaslin(pseq.lacto,
-                 tsvfile = "maaslin_lacto_merged_phfinrisk_genus.tsv",
-                conffile = "maaslin_lacto_config.config",
-                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.lacto.force.vars),
-                readpclrows = "Lactobacillus_acetotolerans.Bacteria-")
-
-Maaslin("maaslin_lacto_merged_phfinrisk_genus.tsv",
-        "maaslin_lacto",
-        strInputConfig = "maaslin_lacto_config.config",
-        strModelSelection="none",
-        dSignificanceLevel = 1.0, 
-        fAllvAll = TRUE,
-        strForcedPredictors = maaslin.lacto.force.vars)
-```
-
-```{r Reading maaslin lacto output, eval = FALSE, include = FALSE}
-maas.lacto.results.df.signf <- read.table("maaslin_lacto/maaslin_lacto_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
-    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  beta = Coefficient) %>%
-   dplyr::filter(Qval < 0.05)
-
-maas.lacto.signf.features <- maas.results.df.signf %>%
-    pull(Feature) %>%
-    as.character() %>%
-    gsub("_", ".", .)
-
-df.lacto.merged <-  merge.pheno.abu(pseq.lacto)
-```
-
-```{r maaslin lacto significant table, eval = FALSE, include = FALSE}
-maas.lacto.results.df.signf %>%
-    select(Name, Feature, beta, Qval) %>%
-    kable           
-```

### 2019-01-03 rrbiome.Rmd
Debuging adonis

@@
 # Function definitions
+
+```{r mini functions}
+replace.brackets <- function (genus) gsub('(.*) \\(+(.+)\\)', '\\1.\\2', genus)
+```
 
@@
                             maxcores = 100) {
-    mclapply(responses, function(response) {
+    lapply(responses, function(response) {
         fo <- sprintf("matrix ~ %s + %s", response, paste(covariates, collapse = " + "))
-        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
+        ad <- adonis(formula = as.formula(fo), data = dset, permutations = npermutations)
         ad$aov.tab %>%
@@
             dplyr::mutate(response = response)
-      }, mc.cores = min(maxcores, length(responses)))
+      })#, mc.cores = min(maxcores, length(responses)))
 }
@@
                              includedvars = c("BL_AGE", "BMI", "SEX"),
-                             readpclrows = "Archaea.Archaea-") {
-    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
+                             readpclrows = "Archaea.Archaea-",
+                             core = FALSE,
+                             core.detection = 0.001,
+                             core.prevalence = 0.01) {
+    pseq.comp <- microbiome::transform(pseq, "compositional")
+    if (core) {
+         pseq.comp  <- microbiome::core(pseq.comp, detection = core.detection, prevalence = core.prevalence)
+    }
+    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq.comp)))
     sample.ids <- colnames(pheno)
-    phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
-    abud.rel <- abundances(phfg.pseq.rel)
-    rownames(abud.rel) <- sapply(rownames(abud.rel),
-                                 function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
+    abud.rel <- abundances(pseq.comp)
+    rownames(abud.rel) <- sapply(rownames(abud.rel), replace.brackets)
     write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
@@
 
- merge.pheno.abu <- function(pseq) {
-     pheno <- microbiome::meta(phyloseq::sample_data(pseq))
+ merge.pheno.abu <- function(pseq,
+                             core = FALSE,
+                             core.detection = 0.001,
+                             core.prevalence = 0.01) {
+     pseq.comp <- microbiome::transform(pseq, "compositional")
+     if (core) {
+         pseq.comp  <- microbiome::core(pseq.comp, detection = core.detection, prevalence = core.prevalence)
+    }
+     pheno <- microbiome::meta(phyloseq::sample_data(pseq.comp))
      sample.ids <- rownames(pheno)
-     phfg.pseq.rel <- microbiome::transform(pseq, "compositional")
-     abud.rel <- abundances(phfg.pseq.rel)
-     rownames(abud.rel) <- sapply(rownames(abud.rel), function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))
+     abud.rel <- abundances(pseq.comp)
+     rownames(abud.rel) <- sapply(rownames(abud.rel), replace.brackets)
      cbind(pheno, t(abud.rel))
 }
+
 ```
@@
 if (!file.exists("rds/adonis.species.rds")) {
+
     adonis.species <- calculateadonis(dset = meta(pseq.species),
@@
                         covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                    npermutations = 500)
+                    npermutations = 1)
+
     saveRDS(adonis.species, file = "rds/adonis.species.rds")
@@
 } else {
-     pca.prcomp <- readRDS("rds/pca.prcomp.rds")
+5     pca.prcomp <- readRDS("rds/pca.prcomp.rds")
 }
@@
 
-#pseq.core <- microbiome::core(pseq.genus, detection = 0.001, pervalence = 0.01)
+prepare.maaslin(pseq.genus,
+                 tsvfile = "maaslin_core_merged_phfinrisk_genus.tsv",
+                conffile = "maaslin_core_config.config",
+                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars),
+                core = TRUE,
+                readpclrows = "Halapricum.Archaea-")
 
-prepare.maaslin(pseq.genus,
-                 tsvfile = "maaslin_merged_phfinrisk_genus.tsv",
-                conffile = "maaslin_config.config",
-                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars))
-
-Maaslin("maaslin_merged_phfinrisk_genus.tsv",
-        "maaslin",
-        strInputConfig = "maaslin_config.config",
+Maaslin("maaslin_core_merged_phfinrisk_genus.tsv",
+        "maaslin_core",
+        strInputConfig = "maaslin_core_config.config",
         strModelSelection="none",
@@
 ```{r Reading maaslins output}
-maas.results.df.signf <- read.table("maaslin/maaslin_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
+maas.results.df.signf <- read.table("maaslin_core/maaslin_core_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
     merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
@@
 
-df.merged <-  merge.pheno.abu(pseq.genus)
+df.merged <-  merge.pheno.abu(pseq.genus, core = TRUE)
 ```
@@
 
-plot_maaslin_heatmap_final(maas.results.df.signf,
+plot_maaslin_heatmap_final(maas.results.df.signf %>% dplyr::filter(Variable != "HYPERTENSION"),
                       xlab.text="Genera",

### 2019-01-03 rrbiome.Rmd

@@
                             maxcores = 100) {
-    lapply(responses, function(response) {
+    mclapply(responses, function(response) {
 	fo <- sprintf("matrix ~ %s + %s", response, paste(covariates, collapse = " + "))
-        ad <- adonis(formula = as.formula(fo), data = dset, permutations = npermutations)
+        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
         ad$aov.tab %>%
@@
             dplyr::mutate(response = response)
-      })#, mc.cores = min(maxcores, length(responses)))
+      }, mc.cores = min(maxcores, length(responses)))
 }

### 2019-01-03 articletwo.Rmd
officer

@@
 
+```{r Remove temporary data, include = FALSE, cache = FALSE, eval = FALSE}
+unlink("rds", recursive=TRUE)
+unlink("cache", recursive=TRUE)
+```
+
+```{r Initialize docx, include = FALSE, cache = FALSE}
+doc <- read_docx(path = "style/articlestyle.docx") %>%
+  body_remove()
+
+dir.create("rds", showWarnings = FALSE)
+dir.create("cache", showWarnings = FALSE)
+
+text.bold <- fp_text(font.size = 12, bold = TRUE, font.family = "Arial")
+text.normal <- fp_text(font.size = 12, font.family = "Arial")
+```
+
 # Function definitions
@@
 
-```{r variables}
+```{r variables, warning = FALSE}
 data("phfinrisk_metadatadesc", package="finriskmetagcommon")
@@
 
-Filtering clinical data
-
-```{r filtering}
-dset.genus <- filter.phenotype.data(pseq.genus)
-dset.species <- filter.phenotype.data(pseq.species)
-```
-
-Phenotypical data between genus and species all.equal `r 
-isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data
-frames are (`r dim(dset.genus)`) and (`r dim(dset.species)`).
-
 Updating phyloseq objects
@@
 ```{r updating phyloseq objects}
-phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(dset.genus)
-phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(dset.species)
+phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(filter.phenotype.data(pseq.genus))
+phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(filter.phenotype.data(pseq.species))
 ```
@@
 Sample data has dimensions (`r dim(sample_data(pseq.species))`).
+
+```{r officer init}
+
+```

### 2019-01-04 rrbiome.Rmd
Biome

@@
 ```{r calculate full alpha model}
-calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon") %>%
-    select(response, estimate, std.error, conf.low, conf.high, p.value) %>%
-    kable(.)
-```
-
-Calculating linear model using binomial variable for exercice and medication.
-
-```{r calculate binomial alpha model}
-calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + ANYEXERCICE + DIAB + ANYDRUG  + diversities_shannon") %>%
-    select(response, estimate, std.error, conf.low, conf.high, p.value) %>%
-    kable(.)
-```
-
-Calculating linear model using minimal variables.
-
-```{r calculate minimal alpha model}
-calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + diversities_shannon") %>%
+alphaglm <- calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon")
+alphaglm %>%
     select(response, estimate, std.error, conf.low, conf.high, p.value) %>%
@@
 adonis.species %>%
-    dplyr::filter(term %in% responses) %>%
-    dplyr::select(-term) %>%
+    data.table::rbindlist(ret, id = "model_name") %>%
+#    dplyr::filter(term %in% c("SY) %>%
+ #   dplyr::select(-term) %>%
     select(response, SumsOfSqs, MeanSqs, F.Model, R2, "Pr(>F)") %>%
@@
 } else {
-5     pca.prcomp <- readRDS("rds/pca.prcomp.rds")
+    pca.prcomp <- readRDS("rds/pca.prcomp.rds")
 }

### 2019-01-04 rrbiome.Rmd
Typo

@@
 adonis.species %>%
-    data.table::rbindlist(ret, id = "model_name") %>%
+    data.table::rbindlist(id = "model_name") %>%
 #    dplyr::filter(term %in% c("SY) %>%

### 2019-01-04 articletwo.Rmd
Functioning

@@
 library(officer)
+library(flextable)
 library(tableone)
@@
 ```{r Importing functions}
+source("aaro/edgeR_plot_functions.R", echo = TRUE)
+source("aaro/comp_fig3_plots.R", echo = TRUE)
 source("rrbiomefunctions.R", echo = TRUE)
-```{r functions}
+
+writetable <- function(doc, tbl,  head, foot) {
+    doc <- doc %>% 
+        body_add_fpar(fpar(ftext("Table . ", prop = text.bold), ftext(head, prop = text.normal)), style = "Table Caption") %>%
+        body_add_flextable(tbl, align = "left") %>%
+        body_add_par(foot, style = "footnote text") %>%
+        body_add_break(pos = "after")
+}
+
+writeggplot <- function(doc, ggplot, header, footer, width = 7, height = 7) {
+    doc <- doc %>% 
+        body_add_fpar(fpar(ftext("Figure . ", prop = text.bold), ftext(header, prop = text.normal)), style = "Image Caption") %>%
+        body_add_gg(value = ggplot, width = width, height = height) %>%
+        body_add_par(footer, style = "footnote text") %>%
+        body_add_break(pos = "after") 
+}
+
 ```
@@
 names.dset <- bind_rows(names.dset,
-                        data.frame(Covariate = c("MAP", "PULSEPRESSURE", "HYPERTENSION"),
-                                   Category = rep("Physical", 3),
-                                   Name =c( "Mean artrerial pressure", "Pulse pressure", "Hypertension")))
+                        data.frame(Covariate = c("MAP", "PULSEPRESSURE", "HYPERTENSION", "SEX"),
+                                   Category = rep("Physical", 4),
+                                   Name =c( "Mean artrerial pressure", "Pulse pressure", "Hypertension", "Female")))
 ```
@@
 
-```{r officer init}
+```{r Characteristics}
+tableone.names <- list("BL_AGE" = "Age, y (SD)",
+              "SEX" = "Female, N (%)",
+              "BMI" = "BMI, kg/m² (SD)",
+              "SYSTM" = "Systolic blood pressure, mmHg (SD)",
+              "DIASM" = "Diastolic blood pressure, mmHg (SD)",
+              "PULSEPRESSURE" = "Pulse pressure, mmHg (SD)",
+              "MAP" = "Mean arterial pressure, mmHg (SD)",
+              "HYPERTENSION" = "Hypertension, N (%)",
+              "CURR_SMOKE" = "Current smoker, N (%)",
+              "DIAB" = "Diabetes mellitus, N (%)",
+              "Q57X" = "Exercice, N (%)",
+              "ANYDRUG" = "Antihypertensive medication, N (%)",
+              "BL_USE_RX_C03" = "  Diuretics, N (%)",
+              "BL_USE_RX_C07" = "  Beta blockers, N (%)",
+              "BL_USE_RX_C08" = "  Calcium channel blocker, N (%)",
+              "BL_USE_RX_C09" = "  Agents acting on the renin-angiotensin system, N (%)")
+tableone.factors <- c("SEX", "HYPERTENSION",  "ANYDRUG", "CURR_SMOKE", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
+
+characteristics <- function(dset, names, factors) {
+    title <- "Characteristics"
+    overall <- paste0("Cases, n=", dim(dset)[1])
+    tableobject <- tableone::CreateTableOne(data = dset, vars = names, factorVars = factors)
+    tablecsv <- print(tableobject,
+                      exact = "stage",
+                      quote = FALSE,
+                      noSpaces = TRUE,
+                      printToggle = FALSE,
+                      digits = 1,
+                      pDigits = 3,
+                      contDigits=1)
+}
+
+data <- characteristics(meta(pseq.species), names(tableone.names), tableone.factors) %>%
+    as.data.frame %>%
+    tibble::rownames_to_column(var = "rowname") %>%
+    dplyr::filter(row_number() > 1) %>%
+    format(justify = "left", trim = TRUE) %>%
+    rowwise() %>%
+    mutate(id = gsub("([A-Za-z_0-9]+).*", "\\1", rowname)) %>%
+    mutate(present = id %in%  names(tableone.names)) %>%
+    mutate(rowname = ifelse(present == TRUE, tableone.names[[id]], rowname)) %>%
+    select(rowname, Overall)
+
+tbl1  <- flextable(data = data) %>%
+    flextable::width(j = 1, width = 3) %>%
+    flextable::width(j = 2, width = 1.2)
+
+doc <- writetable(doc, tbl1, "Continuous variables are presented as mean (standard deviation).
+Categorical variables reported as absolute and relative frequencies.
+BMI indicates body mass index.", "Characteristics of the study sample.")
+```
+```{r alpha-beta picture}
+dset.alpha <- meta.merge.alphadiversity(pseq.species)
+alphaglm <- calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon")
+alpha.diversity <- alphaglm %>% select(response, estimate, p.value)
+
+adonis.species  <- readRDS("rds/adonis.species.rds")
+beta.diversity <- adonis.species %>%
+    data.table::rbindlist(id = "model_name") %>%
+    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
+    dplyr::select(response, R2)
+
+diversity <- merge(alpha.diversity, beta.diversity, by = "response") %>%
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate")
+```
+
+```{r pca}
+pca.prcomp <- readRDS("rds-20190102/pca.prcomp.rds")
+pca.eigs <- pca.prcomp$sdev^2
+pca.percentage <- 100 * pca.eigs /sum(pca.eigs)
+pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)
+setNames(pca.percentage, pca.prcomp$x %>% colnames) %>% head(10) %>% t %>% kable
+pcaplot <- ggplot(cbind(meta(pseq.species), pca.prcomp$x[,1:2]), aes(x=PC1, y=PC2)) +
+    geom_point(aes(color = HYPERTENSION), size = 0.2) +
+    xlab(pca.labels[1]) +
+    ylab(pca.labels[2]) +
+    scale_color_brewer(palette="Dark2") +
+    ggplot2::theme_minimal()
+
+writeggplot(doc, pcaplot2, "PCA", "footer", width = 7, height = 7)
+```
+
+```{r maaslin}
+maas.results.df.signf <- read.table("maaslin_core/maaslin_core_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
+    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
+           beta = Coefficient) %>%
+   dplyr::filter(Qval < 0.05)
+
+maaslinplot <- plot_maaslin_heatmap_final(maas.results.df.signf %>% dplyr::filter(Variable != "HYPERTENSION"),
+                      xlab.text="Genera",
+                      beta.maxabs=max(abs(maas.results.df.signf[,'beta'])),
+                      order.by.category=FALSE)
+
+writeggplot(doc, maaslinplot, "maaslin", "footer", width = 7, height = 3)
 
 ```
+
+```{r}
+``
+
+```{r Write docx to file, cache = FALSE, results = FALSE, include = FALSE}
+print(doc, target = "articletwo.docx")
+```

### 2019-01-04 aaro/comp_fig3_plots.R

@@ -1,156 +1,14 @@
-# Plots for Figure 3
-
-library(biomformat)
-library(microbiome)
-library(phyloseq)
-library(data.table)
-library(RColorBrewer)
-library(dplyr)
-packageVersion('phyloseq')
-library(finriskmetagcommon)
-library(ggplot2)
-library(gridExtra)
-library(cowplot)
-library(ggpubr)
-library(grid)
-library(gtable)
 library(scales)
-# http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/81-ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page/<Paste>
+library(gridExtra)
 
-source('code/phyloseq/plot_catcol_palette.R')
-
-pltdir <- 'docs/assets/figure3/'
-if (!dir.exists(pltdir)) {
-  dir.create(pltdir)
-}
-
-### Preliminaries: load and subset data
-
-maindir <- "/csc/fr_metagenome/microbiome_scratch/scratch/"
-datadir <- file.path(maindir, "data_private/")
-
-
-# alpha diversity linear models
-
-
-shannon.df <- readRDS(file.path(datadir, 'lm_stats_dichodiversities_shannon.spt0.2018-10-22.RDs'))
-#isimpsn.df <- readRDS(file.path(datadir, 'lm_stats.diversities_inverse_simpson.Rds'))
-#observd.df <- readRDS(file.path(datadir, 'lm_stats.observed.Rds'))
-
-
-# PCA regression results
-
-
-## PERMANOVA results
-# (preprocessing lifted from comp_adonis_plots.R)
-ident <- "20181030_nperm50"
-
-adon.res <- readRDS(paste0("/csc/fr_metagenome/microbiome_scratch/scratch/",
-                           "data_private/adonis.result.", ident, ".RDs"))
-
-
-adon.res.df <- adon.res[[1]] # only the variate of interest
-adon.res.all.df <- readRDS(paste0("/csc/fr_metagenome/microbiome_scratch/scratch/",
-                           "data_private/adonis.result.", ident, ".all.df.RDs"))# all output of aov.tab
-rm(adon.res)
-
-
-# PERMANOVA: merge covariate category into df
-
-if (!exists('phfinrisk_metadatadesc')) {
-  data("phfinrisk_metadatadesc", package="finriskmetagcommon")
-}
[…]

### 2019-01-04 articletwo.Rmd
PCA plot

@@
 library(flextable)
+library(rvg)
 library(tableone)
+library(scales)
+library(gridExtra)
 ```
@@
 source("aaro/edgeR_plot_functions.R", echo = TRUE)
-source("aaro/comp_fig3_plots.R", echo = TRUE)
 source("rrbiomefunctions.R", echo = TRUE)
@@
         body_add_gg(value = ggplot, width = width, height = height) %>%
+        body_add_par(footer, style = "footnote text") %>%
+        body_add_break(pos = "after") 
+}
+
+writeimage <- function(doc, filename, header, footer, width = 7, height = 7) {
+    doc <- doc %>% 
+        body_add_fpar(fpar(ftext("Figure . ", prop = text.bold), ftext(header, prop = text.normal)), style = "Image Caption") %>%
+        body_add_img(src = filename, width = width, height = height) %>% 
         body_add_par(footer, style = "footnote text") %>%
@@
 ```
-```{r alpha-beta picture}
+```{r alphabeta definitions}
 dset.alpha <- meta.merge.alphadiversity(pseq.species)
@@
 diversity <- merge(alpha.diversity, beta.diversity, by = "response") %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate")
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate") %>%
+    mutate(Qstar = ifelse(p.value < 0.05, '*', ' '))
+
+fig.height <- 4.0 + nrow(diversity) * 0.33
+fig.permar2.text <- "R2 of variable \nin Bray-Curtis distance" # TODO latexify expression?
+fig.textsize <- 15
+amaxb <- diversity %>% pull(estimate) %>% abs %>% max %>% signif(., digits = 2)
+```
+
+```{r alphabeta alpha}
+plot.alpha.diversity <- ggplot(diversity, aes(x=reorder(Name, R2),  y = 1, fill = estimate)) +
+    geom_bar(stat="identity", color="black") +
+    coord_flip() +
+    scale_fill_gradientn(colours = c("darkblue", "blue", "white", "red", "darkred"),
+                         values = rescale(c(-amaxb,-0.1, 0, 0.1, amaxb)),
+                         name = 'Regression coefficient \nin linear model \nfor Shannon index',
+                         limits = c(-1.05*amaxb, 1.05*amaxb),
+                         na.value = "black") +
+    theme_classic(20) +
+    geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=TRUE, color='black', size=20) +
+    scale_shape_manual(name="",
+                       values=c('*'='*', ' '=' '),
+                       labels=c("*"="significant at FDR 0.05", ' '=' '),
+                       breaks=c("*", ' ')) +
+    xlab("") +
+    ylab("") +
+    theme(legend.position="right",
+          legend.title=element_text(size = 20),
+          legend.direction="vertical",
+          legend.text=element_text(size = floor(0.75*20)),
+          legend.box.margin=margin(0,0,0,0,"pt"),
+          legend.margin=margin(0,0,0,0,"pt"),
+          legend.justification="left",
+          axis.line.y = element_line(linetype="blank"),
+          axis.line.x = element_line(linetype="blank"),
+          legend.title.align = 0,
+          legend.text.align = 0,
+          axis.text.y = element_text(color="black"),
+          axis.text.x = element_text(color="black"))+
+    scale_y_continuous(breaks=c(0.5), labels=c(expression(alpha)))
+
+plot.alpha.diversity.gtable <- ggplot_gtable(ggplot_build(plot.alpha.diversity))
+plot.alpha.diversity.legend <- legend_extractor(plot.alpha.diversity.gtable)
+plot.alpha.diversity.yaxis <- plot.alpha.diversity.gtable$grobs[[3]]
+
+plot.alpha.diversity <- plot.alpha.diversity +
+    theme(legend.position = "none",
+          axis.text.x = element_blank(),
+          axis.text.y=element_blank())
+
+plot.gtable.alpha.diversity.stripped <- ggplot_gtable(ggplot_build(plot.alpha.diversity))
+```
+
+```{r alphabeta beta}
+plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2)) +
+    geom_bar(stat="identity", color='black') +
+    coord_flip() +
+    theme_classic(fig.textsize) +
+    ylab(fig.permar2.text) +
+    theme(axis.text.y = element_text(color="black"),
+          axis.text.x = element_text(color="black"),
+          legend.position="right",
+          legend.direction="vertical",
+          legend.box.margin=margin(0,0,0,0,"pt"),
+          legend.margin=margin(0,0,0,0,"pt"),
+          legend.justification="left",
+          legend.title.align = 0,
+          legend.text.align = 0)
+
+plot.gtable.beta.diversity <- ggplot_gtable(ggplot_build(plot.beta.diversity))
+plot.xlab.beta.diversity <- xlabb_extractor(gtable.beta.diversity)
+
+plot.beta.diversity <- plot.beta.diversity  +
+    theme(axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.title.x  =  element_blank(),
+          axis.title.y = element_blank(),
+          legend.position = "none")
+plot.gtable.beta.diversity.stripped <- ggplot_gtable(ggplot_build(plot.beta.diversity))
+```
+
+```{r save grob}
+gs4 <- list(plot.alpha.diversity.yaxis,
+            plot.gtable.alpha.diversity.stripped$grobs[[6]],
+            plot.gtable.beta.diversity.stripped$grobs[[6]],
+            plot.alpha.diversity.legend,
+            plot.alpha.diversity.gtable$grobs[[7]],
+            plot.gtable.beta.diversity$grobs[[7]],
+           plot.xlab.beta.diversity)
+#           plot.gtable.beta.diversity.stripped[[3]])
+
+g4 <- arrangeGrob(grobs = gs4,
+            layout_matrix = rbind(
+                c(1,2,8,3,NA),
+                c(1,2,8,3,4),
+                c(1,2,8,3,NA),
+                c(1,2,8,3,NA),
+                c(NA,5,NA,6,NA),
+                c(NA,NA,NA,7,NA)),
+            widths=c(4,0.7,0.5,5,5),
+            heights=c(3,12,8,21,1,2))
+
+ggsave(file = "cache/alpha-beta.png", plot=g4, height=fig.height, width=11)
+writeimage(doc, "cache/alpha-beta.png", "Alpha-beta", "footer", width = 11, height = fig.height)
 ```
@@
 ```{r pca}
-pca.prcomp <- readRDS("rds-20190102/pca.prcomp.rds")
+pca.prcomp <- readRDS("rds/pca.prcomp.rds")
 pca.eigs <- pca.prcomp$sdev^2
@@
 pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)
-setNames(pca.percentage, pca.prcomp$x %>% colnames) %>% head(10) %>% t %>% kable
+
 pcaplot <- ggplot(cbind(meta(pseq.species), pca.prcomp$x[,1:2]), aes(x=PC1, y=PC2)) +
-    geom_point(aes(color = HYPERTENSION), size = 0.2) +
+    geom_point(aes(color = HYPERTENSION), size = 0.5) +
+    scale_color_manual(values=c("red", "blue"),
+                       name= "Hypertension",
+                       breaks=c("1", "0"),
+                       labels=c("Yes", "No")) +
     xlab(pca.labels[1]) +
     ylab(pca.labels[2]) +
-    scale_color_brewer(palette="Dark2") +
-    ggplot2::theme_minimal()
+    coord_fixed() +
+    scale_x_continuous(breaks=seq(-100, 100, 20)) +
+    scale_y_continuous(breaks=seq(-30, 30, 10)) +
+    theme_classic() +
+    theme(text = element_text(size=20),
+          legend.justification=c(1,0),
+          legend.position=c(1.03, 0.75),
+          legend.title=element_text(size = 20),
+          legend.direction="vertical",
+          legend.text=element_text(size = floor(0.75*20)),
+          legend.box.margin = margin(1, 1, 1, 1, "mm"),
+          legend.box.background = element_rect(colour = "black"))
 
-writeggplot(doc, pcaplot2, "PCA", "footer", width = 7, height = 7)
+ggsave(file = "cache/pca.png", plot = pcaplot, height = 5, width = 9, dpi = 300)
+
+writeimage(doc, "cache/pca.png", "Alpha-beta", "footer", width = 11, height = fig.height)
 ```
@@
 writeggplot(doc, maaslinplot, "maaslin", "footer", width = 7, height = 3)
-
 ```

### 2019-01-04 aaro/comp_fig3_plots.R

@@ -12,21 +12,18 @@ amaxb <- diversity %>% pull(estimate) %>% abs %>% max %>% signif(., digits = 2)
 
 legend_extractor <- function (tmp) {
   leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
-  legend <- tmp$grobs[[leg]]
-  legend
+  tmp$grobs[[leg]]
 }
 
 ylabl_extractor <- function(tmp) {
   yl <- which(grepl('axis.title.y.left', sapply(tmp$grobs, function(x) x$name)))
-  ylabb <- tmp$grobs[[yl]]
-  ylabb
+  tmp$grobs[[yl]]
 }
 
 
 xlabb_extractor <- function(tmp) {
   xl <- which(grepl('axis.title.x.bottom', sapply(tmp$grobs, function(x) x$name)))
-  xlabb <- tmp$grobs[[xl]]
-  xlabb
+  tmp$grobs[[xl]]
 }
 
 

### 2019-01-04 rrbiomefunctions.R

@@ -115,3 +115,19 @@ prepare.maaslin <- function (pseq,
      cbind(pheno, t(abud.rel))
 }
 
+
+legend_extractor <- function (tmp) {
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  tmp$grobs[[leg]]
+}
+
+ylabl_extractor <- function(tmp) {
+  yl <- which(grepl('axis.title.y.left', sapply(tmp$grobs, function(x) x$name)))
+  tmp$grobs[[yl]]
+}
+
+xlabb_extractor <- function(tmp) {
+  xl <- which(grepl('axis.title.x.bottom', sapply(tmp$grobs, function(x) x$name)))
+  tmp$grobs[[xl]]
+}
+

### 2019-01-04 aaro/comp_supp_pca_subnet_regres_results_heatmap.R
For fig 4

@@ -1,0 +1,214 @@
+library(biomformat)
+library(microbiome)
+library(phyloseq)
+library(data.table)
+library(RColorBrewer)
+library(dplyr)
+packageVersion('phyloseq')
+library(finriskmetagcommon)
+library(ggplot2)
+library(gridExtra)
+library(cowplot)
+library(ggpubr)
+library(grid)
+library(gtable)
+library(scales)
+
+source('code/phyloseq/generic_functions.R')
+source("code/phyloseq/taxa_names_underscored.R")
+source('code/phyloseq/netw_plot_funs.R')
+source('code/phyloseq/lm_comp_functions.R')
+
+source('code/phyloseq/netw_plot_funs.R')
+
+
+source('code/phyloseq/edgeR_functions.R')
+source('code/phyloseq/edgeR_plot_functions.R')
+
+maindir <- "/csc/fr_metagenome/microbiome_scratch/scratch/"
+datadir <- file.path(maindir, "data_private/")
+
+pltdir <- 'docs/assets/PC_subnet_pheno_association/'
+if (!dir.exists(pltdir)) {
+  dir.create(pltdir)
+}
+
+pseqdate <- "2018-12-21"
+phfg.pseq <- readRDS(paste0(datadir, "phfinrisk_genus_ok_drop50k_", pseqdate, ".RDs"))
+pseq.rel.core <- core(microbiome::transform(phfg.pseq, "compositional"), detection=0.1/100, prevalence=1/100)
+
+pseq.rel.core.rel.clr <- transform(transform(pseq.rel.core, "compositional"), "clr")
+
+
+# latest sample data
+phfg.pseq.all <- readRDS(paste0(datadir, "phfinrisk_genus_all_drop50k_", pseqdate, ".RDs"))
+
+data("phfinrisk_metadatadesc", package="finriskmetagcommon")
+names.df <- filter(phfinrisk_metadatadesc,
+                   Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115 == 0)
+phfg.pseq <- drop_unused_samplevars(phfg.pseq.all, names.df)
+
+# read subnet info + rel abundances of subnets:
+
+# from comp_fig2_plots_d50k.R
+n.components <- readRDS(paste0(datadir, "spieceasi_0_", pseqdate,
+                              "_components_fig2.RDs"))
+
+n.relabunds.core.genera <- abundances(pseq.rel.core)
[…]

### 2019-01-04 articletwo.Rmd
Friday

@@
 library(gridExtra)
+library(png)
 ```
@@
 
-writeimage <- function(doc, filename, header, footer, width = 7, height = 7) {
+writeimage <- function(doc, filename, header, footer, width = 7.5) {
+    img <- readPNG("cache/pca.png")
     doc <- doc %>% 
         body_add_fpar(fpar(ftext("Figure . ", prop = text.bold), ftext(header, prop = text.normal)), style = "Image Caption") %>%
-        body_add_img(src = filename, width = width, height = height) %>% 
+        body_add_img(src = filename, width = width, height = width*dim(img)[1]/dim(img)[2]) %>% 
         body_add_par(footer, style = "footnote text") %>%
@@
 ggsave(file = "cache/alpha-beta.png", plot=g4, height=fig.height, width=11)
-writeimage(doc, "cache/alpha-beta.png", "Alpha-beta", "footer", width = 11, height = fig.height)
+writeimage(doc, "cache/alpha-beta.png", "Alpha-beta", "footer")
 ```
@@
 
-writeimage(doc, "cache/pca.png", "Alpha-beta", "footer", width = 11, height = fig.height)
+writeimage(doc, "cache/pca.png", "Alpha-beta", "footer", width = 7)
 ```
@@
 
-```{r}
-``
+```{r PCA heatmap}
+pca.axis <- pca.prcomp$x %>% data.frame
+pca.dset <- base::merge(meta(pseq.species), pca.axis[,1:5], by=0, all=TRUE) %>%
+  dplyr::rename(Sample_ID = Row.names)
+
+pca.dset %>% head
+
+pca.ret <- lapply(paste0("PC", seq(1,5)), function (pc) {
+    calculateglm(pca.dset,
+                 modelstring = paste("%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  +", pc),
+                 filterstr = "PC") %>%
+        select(response, estimate, p.value) %>%
+        mutate(model_name = pc)
+})
+
+pca.ret  %>%
+    data.table::rbindlist() %>%
+    as.data.frame %>%
+    mutate(p.value = p.adjust(p.value, method="BH")) %>%
+    arrange(p.value) %>%
+    kable
+```
+
+```{r PCA loadings}
+
+pca.rotation.abs <- abs(pca.prcomp$rotation)
+pca.loading <- sweep(pca.rotation.abs, 2, colSums(pca.rotation.abs), "/") %>%
+    as.data.frame %>%
+    tibble::rownames_to_column(var = "term")
+
+pca.pc3 <- pca.loading %>%
+    select(term, PC3) %>%
+    arrange(-PC3) %>%
+    head(20)
+
+pca.loading %>% pull(PC3) %>% head(2700) %>% sum
+
+ggplot(data=pca.pc3, aes(x=reorder(term, -PC3), y=PC3)) +
+    geom_bar(stat="identity") +
+    theme_minimal() +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+dev.off()
+
+```
 

### 2019-01-04 rrbiome.Rmd
For fig 4
Friday

@@
 
-We calculate primary component analysis using centered log-ratio
+We calculate principal component analysis using centered log-ratio
 transformation for the abunances.
@@
 ```{r maaslin plot, message = FALSE, echo = FALSE}
-source("edgeR_plot_functions.R")
+source("aaro/edgeR_plot_functions.R")
 
@@
 ```{r lactobacillus scatterplot, eval = FALSE, include = FALSE}
-maaslin.lacto.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "DIAB")
+maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
 
@@
                 conffile = "maaslin_lacto_config.config",
-                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.lacto.force.vars),
-                readpclrows = "Lactobacillus_acetotolerans.Bacteria-")
+                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars),
+                readpclrows = "Lactobacillus_delbrueckii.Bacteria,Lactobacillus_ruminis.Bacteria,Lactobacillus_acidophilus.Bacteria,Lactobacillus_fermentum.Bacteria,Lactobacillus_paracasei.Bacteria,Lactobacillus_salivarius.Bacteria")
 
@@
         fAllvAll = TRUE,
-        strForcedPredictors = maaslin.lacto.force.vars)
+        strForcedPredictors = maaslin.force.vars)
 ```

### 2019-01-04 rrbiomefunctions.R

@@ -40,13 +40,14 @@ calculateglm <- function(dset,
                          responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
                                           model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
                          min_n_for_continuous = 10,
-                         modelstring = "%s ~ BL_AGE + SEX") {
+                         modelstring = "%s ~ BL_AGE + SEX",
+                         filterstr = "diversities_") {
     glmlist <- lapply(responses, function(response) {
         fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
         fo <- sprintf(modelstring, response)
         stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
             broom::tidy() %>%
-                dplyr::filter(grepl("diversities_", term)) %>%
+                dplyr::filter(grepl(filterstr, term)) %>%
                 dplyr::mutate(response = response, fo = fo) %>%
                 dplyr::select(-term)
     })

### 2019-01-23 articletwo.Rmd
Minor

@@
 text.normal <- fp_text(font.size = 12, font.family = "Arial")
+```
+
+# Sources
+
+### Officer functions
+
+```{r Officer importer, code=readLines("articleone-officer.R")}
+source("articletwo-officer.R")
 ```
@@
 Sample data has dimensions (`r dim(sample_data(pseq.species))`).
+
+## Characteristics of the study sample. (Table 1.)
+
+```{r Characteristics}
+tbl1 <- xtableformatter(characteristics(dset$fr02$data.full)) %>%
+    flextable::width(j = 1, width = 3) %>%
+    flextable::width(j = 2, width = 1.2) 
+```
+
+```{r Write characteristics}
+tbl1head <- "Characteristics of the study sample."
+tbl1foot <- "Continuous variables are presented as mean (standard deviation). Categorical variables reported as absolute and relative frequencies. BMI indicates body mass index."
+writetable(doc, tbl1, number = 1, tbl1head, tbl1foot)
+```
 

### 2019-01-23 articletwo-officer.R

@@ -1,0 +1,148 @@
+characteristics <- function(dset, type = "fr02") {
+  if (type == "fr02") {
+    factors <- c("HT", "HRX", "sex", "curr_smk", "curr_diab", "plate")
+    names <- list("Age, y (SD)" = "AGE",
+                  "Female, N (%)" = "sex",
+                  "BMI, kg/m² (SD)" = "BMI",
+                  "Systolic blood pressure, mmHg (SD)" = "SBP",
+                  "Diastolic blood pressure, mmHg (SD)" = "DBP",
+                  "Pulse pressure, mmHg (SD)" = "PP",
+                  "Mean arterial pressure, mmHg (SD)" = "MAP",
+                  "Hypertension, N (%)" = "HT",
+                  "Antihypertensive medication, N (%)" = "HRX",
+                  "Current smoker, N (%)" = "curr_smk",
+                  "Diabetes mellitus, N (%)" = "curr_diab")
+  } else {
+    factors <- c("HT8", "HRX8", "sex", "CURRSMK8", "curr_diab8", "plate")
+    names <- list("Age, y (SD)" = "AGE8",
+                  "Female, N (%)" = "sex",
+                  "BMI, kg/m² (SD)" = "BMI8",
+                  "Systolic blood pressure, mmHg (SD)" = "SBP8",
+                  "Diastolic blood pressure, mmHg (SD)" = "DBP8",
+                  "Pulse pressure, mmHg (SD)" = "PP8",
+                  "Mean arterial pressure, mmHg (SD)" = "MAP8",
+                  "Hypertension, N (%)" = "HT8",
+                  "Antihypertensive medication, N (%)" = "HRX8",
+                  "Current smoker, N (%)" = "CURRSMK8",
+                  "Diabetes mellitus, N (%)" = "curr_diab8")
+  }
+  tableobject <- tableone::CreateTableOne(data = dset, vars = unlist(names), factorVars = factors)
+  tablecsv <- print(tableobject,
+                    exact = "stage",
+                    quote = FALSE,
+                    noSpaces = TRUE,
+                    printToggle = FALSE,
+                    digits = 1,
+                    pDigits = 3,
+                    contDigits=1)
+  
+  title <- "Characteristics"
+  overall <- paste0("Cases, n=", dim(dset)[1])
+  
+  tablecsv <- tablecsv %>%
+    as.data.frame %>%
+    dplyr::filter(row_number() > 1) %>%
[…]

### 2019-01-23 articletwo.Rmd
Minor

@@
 
+### RR biome functions
+
+```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
+source("articletwo-rrbiome.R")
+```
+
+### EdgeR plot functions
+
+```{r EdgeR functions, code=readLines("aaro/edgeR_plot_functions.R")}
+source("aaro/edgeR_plot_functions.R")
+```
+
 # Function definitions
@@
 ```{r Importing functions}
-source("aaro/edgeR_plot_functions.R", echo = TRUE)
-source("rrbiomefunctions.R", echo = TRUE)
-
 writetable <- function(doc, tbl,  head, foot) {

### 2019-01-23 articletwo.Rmd
Minor

@@
 
-# Function definitions
-
-```{r Importing functions}
-writetable <- function(doc, tbl,  head, foot) {
-    doc <- doc %>% 
-        body_add_fpar(fpar(ftext("Table . ", prop = text.bold), ftext(head, prop = text.normal)), style = "Table Caption") %>%
-        body_add_flextable(tbl, align = "left") %>%
-        body_add_par(foot, style = "footnote text") %>%
-        body_add_break(pos = "after")
-}
-
-writeggplot <- function(doc, ggplot, header, footer, width = 7, height = 7) {
-    doc <- doc %>% 
-        body_add_fpar(fpar(ftext("Figure . ", prop = text.bold), ftext(header, prop = text.normal)), style = "Image Caption") %>%
-        body_add_gg(value = ggplot, width = width, height = height) %>%
-        body_add_par(footer, style = "footnote text") %>%
-        body_add_break(pos = "after") 
-}
-
-writeimage <- function(doc, filename, header, footer, width = 7.5) {
-    img <- readPNG("cache/pca.png")
-    doc <- doc %>% 
-        body_add_fpar(fpar(ftext("Figure . ", prop = text.bold), ftext(header, prop = text.normal)), style = "Image Caption") %>%
-        body_add_img(src = filename, width = width, height = width*dim(img)[1]/dim(img)[2]) %>% 
-        body_add_par(footer, style = "footnote text") %>%
-        body_add_break(pos = "after") 
-}
-
-```
-
 # Loading data
@@
                                    Category = rep("Physical", 4),
-                                   Name =c( "Mean artrerial pressure", "Pulse pressure", "Hypertension", "Female")))
+                                   Name =c( "Mean artrerial pressure", "Pulse pressure", "Hypertension", "Female"),
+                                   Desc = c("Mean arterail presure, 2./3.*DIASM + 1./3.*SYSTM)",
+                                            "Pulse pressure, SYSTM - DIASM",
+                                            "Hypertension, SYSTM >=140 or DIAS >= 90 or BP mediaction",
+                                            "Sex is female, True for female, False for male")))
 ```
@@
 ```{r Characteristics}
-tbl1 <- xtableformatter(characteristics(dset$fr02$data.full)) %>%
-    flextable::width(j = 1, width = 3) %>%
-    flextable::width(j = 2, width = 1.2) 
+tbl1 <- tableone(meta(pseq.species))
 ```
@@
 tbl1head <- "Characteristics of the study sample."
-tbl1foot <- "Continuous variables are presented as mean (standard deviation). Categorical variables reported as absolute and relative frequencies. BMI indicates body mass index."
+tbl1foot <- "Continuous variables are presented as mean (standard deviation).
+Categorical variables reported as absolute and relative frequencies.
+BP indicates blood pressure, BMI body mass index, RAS renin-angiotensin system."
 writetable(doc, tbl1, number = 1, tbl1head, tbl1foot)
+print(doc, target = "articletwo.docx")
 ```
 
-```{r Characteristics}
-tableone.names <- list("BL_AGE" = "Age, y (SD)",
-              "SEX" = "Female, N (%)",
-              "BMI" = "BMI, kg/m² (SD)",
-              "SYSTM" = "Systolic blood pressure, mmHg (SD)",
-              "DIASM" = "Diastolic blood pressure, mmHg (SD)",
-              "PULSEPRESSURE" = "Pulse pressure, mmHg (SD)",
-              "MAP" = "Mean arterial pressure, mmHg (SD)",
-              "HYPERTENSION" = "Hypertension, N (%)",
-              "CURR_SMOKE" = "Current smoker, N (%)",
-              "DIAB" = "Diabetes mellitus, N (%)",
-              "Q57X" = "Exercice, N (%)",
-              "ANYDRUG" = "Antihypertensive medication, N (%)",
-              "BL_USE_RX_C03" = "  Diuretics, N (%)",
-              "BL_USE_RX_C07" = "  Beta blockers, N (%)",
-              "BL_USE_RX_C08" = "  Calcium channel blocker, N (%)",
-              "BL_USE_RX_C09" = "  Agents acting on the renin-angiotensin system, N (%)")
-tableone.factors <- c("SEX", "HYPERTENSION",  "ANYDRUG", "CURR_SMOKE", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
-
-characteristics <- function(dset, names, factors) {
-    title <- "Characteristics"
-    overall <- paste0("Cases, n=", dim(dset)[1])
-    tableobject <- tableone::CreateTableOne(data = dset, vars = names, factorVars = factors)
-    tablecsv <- print(tableobject,
-                      exact = "stage",
-                      quote = FALSE,
-                      noSpaces = TRUE,
-                      printToggle = FALSE,
-                      digits = 1,
-                      pDigits = 3,
-                      contDigits=1)
-}
-
-data <- characteristics(meta(pseq.species), names(tableone.names), tableone.factors) %>%
-    as.data.frame %>%
-    tibble::rownames_to_column(var = "rowname") %>%
-    dplyr::filter(row_number() > 1) %>%
-    format(justify = "left", trim = TRUE) %>%
-    rowwise() %>%
-    mutate(id = gsub("([A-Za-z_0-9]+).*", "\\1", rowname)) %>%
-    mutate(present = id %in%  names(tableone.names)) %>%
-    mutate(rowname = ifelse(present == TRUE, tableone.names[[id]], rowname)) %>%
-    select(rowname, Overall)
-
-tbl1  <- flextable(data = data) %>%
-    flextable::width(j = 1, width = 3) %>%
-    flextable::width(j = 2, width = 1.2)
-
-doc <- writetable(doc, tbl1, "Continuous variables are presented as mean (standard deviation).
-Categorical variables reported as absolute and relative frequencies.
-BMI indicates body mass index.", "Characteristics of the study sample.")
-```
 ```{r alphabeta definitions}

### 2019-01-23 articletwo-officer.R

@@ -1,51 +1,65 @@
-characteristics <- function(dset, type = "fr02") {
-  if (type == "fr02") {
-    factors <- c("HT", "HRX", "sex", "curr_smk", "curr_diab", "plate")
-    names <- list("Age, y (SD)" = "AGE",
-                  "Female, N (%)" = "sex",
-                  "BMI, kg/m² (SD)" = "BMI",
-                  "Systolic blood pressure, mmHg (SD)" = "SBP",
-                  "Diastolic blood pressure, mmHg (SD)" = "DBP",
-                  "Pulse pressure, mmHg (SD)" = "PP",
-                  "Mean arterial pressure, mmHg (SD)" = "MAP",
-                  "Hypertension, N (%)" = "HT",
-                  "Antihypertensive medication, N (%)" = "HRX",
-                  "Current smoker, N (%)" = "curr_smk",
-                  "Diabetes mellitus, N (%)" = "curr_diab")
-  } else {
-    factors <- c("HT8", "HRX8", "sex", "CURRSMK8", "curr_diab8", "plate")
-    names <- list("Age, y (SD)" = "AGE8",
-                  "Female, N (%)" = "sex",
-                  "BMI, kg/m² (SD)" = "BMI8",
-                  "Systolic blood pressure, mmHg (SD)" = "SBP8",
-                  "Diastolic blood pressure, mmHg (SD)" = "DBP8",
-                  "Pulse pressure, mmHg (SD)" = "PP8",
-                  "Mean arterial pressure, mmHg (SD)" = "MAP8",
-                  "Hypertension, N (%)" = "HT8",
-                  "Antihypertensive medication, N (%)" = "HRX8",
-                  "Current smoker, N (%)" = "CURRSMK8",
-                  "Diabetes mellitus, N (%)" = "curr_diab8")
-  }
-  tableobject <- tableone::CreateTableOne(data = dset, vars = unlist(names), factorVars = factors)
-  tablecsv <- print(tableobject,
-                    exact = "stage",
-                    quote = FALSE,
-                    noSpaces = TRUE,
-                    printToggle = FALSE,
-                    digits = 1,
-                    pDigits = 3,
-                    contDigits=1)
-  
-  title <- "Characteristics"
-  overall <- paste0("Cases, n=", dim(dset)[1])
-  
-  tablecsv <- tablecsv %>%
-    as.data.frame %>%
-    dplyr::filter(row_number() > 1) %>%
-    dplyr::mutate(!!title := names(names), !!overall := Overall) %>%
[…]

### 2019-01-23 articletwo-officer.R
Minor

@@ -22,7 +22,7 @@ characteristics <- function(dset, names, factors) {
         select(rowname, Overall)
 }
 
-tableone <- function(dset, ) {
+tableone <- function(dset) {
     tableone.names <- list("BL_AGE" = "Age, y (SD)",
                            "SEX" = "Female, N (%)",
                            "BMI" = "BMI, kg/m² (SD)",

### 2019-01-23 articletwo.Rmd
Minor

@@
 
-Importing libraries:
+# Running at
+
+Calculations ran at
+
+```{r Running at}
+if (!exists('now')) {
+    now <- format(Sys.time(), '%Y%m%d-%H%M%S')
+}
+now
+```
+
+# Importing libraries:
 
@@
   body_remove()
-
 dir.create("rds", showWarnings = FALSE)
 dir.create("cache", showWarnings = FALSE)
-
-text.bold <- fp_text(font.size = 12, bold = TRUE, font.family = "Arial")
-text.normal <- fp_text(font.size = 12, font.family = "Arial")
 ```
@@
 writetable(doc, tbl1, number = 1, tbl1head, tbl1foot)
-print(doc, target = "articletwo.docx")
 ```
+
+## Alpha & Beta plot (Figure 1.)
 
@@
 adonis.species  <- readRDS("rds/adonis.species.rds")
+
 beta.diversity <- adonis.species %>%
     data.table::rbindlist(id = "model_name") %>%
-    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
+    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"))
+
+                                        #%>%
     dplyr::select(response, R2)
@@
 
+## Principal component analysis (Figure 2.)
+
 ```{r pca}
@@
 ```{r Write docx to file, cache = FALSE, results = FALSE, include = FALSE}
-print(doc, target = "articletwo.docx")
+print(doc, target = paste0("report/articletwo-", now, ".docx"))
 ```

### 2019-01-23 rrbiome.Rmd
Minor

@@
 
-# Function definitions
+## RR biome functions
 
-```{r mini functions}
-replace.brackets <- function (genus) gsub('(.*) \\(+(.+)\\)', '\\1.\\2', genus)
-```
-
-```{r functions}
-filter.phenotype.data <- function(pseq,
-                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
-                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
-                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
-    meta(pseq) %>%
-        tibble::rownames_to_column(var = "rowname") %>%
-        dplyr::select(rowname, included) %>% 
-        dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
-                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
-                      ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
-                      PULSEPRESSURE = SYSTM - DIASM,
-                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
-                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
-                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
-         stats::na.omit() %>%
-        dplyr::select(-MEN, -DNA_OK) %>%
-        tibble::remove_rownames() %>%
-        tibble::column_to_rownames(var = "rowname")
-
-meta.merge.alphadiversity <- function(pseq, index = "shannon") {
-    alphadiversity  <- microbiome::global(pseq, index = index)
-    base::merge(meta(pseq), alphadiversity, by=0, all=TRUE) %>%
-        dplyr::rename(Sample_ID = Row.names) %>%
-            dplyr::mutate(BL_AGE_GROUP = group_age(BL_AGE))
-}
-
-calculate.beta.matrix <- function(pseq) {
-    compositional.profile <- microbiome::transform(pseq, 'compositional')
-    otu <- microbiome::abundances(compositional.profile)
-    bray.dist.m <- vegan::vegdist(t(otu), method="bray")
-    dist.matrix <- as.matrix(bray.dist.m)
-    attr(dist.matrix, "method") <- "bray"
-    dist.matrix
-}
-calculateglm <- function(dset,
-                         responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
-                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
-                         min_n_for_continuous = 10,
-                         modelstring = "%s ~ BL_AGE + SEX") {
-    glmlist <- lapply(responses, function(response) {
-        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
-        fo <- sprintf(modelstring, response)
-        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
-            broom::tidy() %>%
-                dplyr::filter(grepl("diversities_", term)) %>%
-                dplyr::mutate(response = response, fo = fo) %>%
-                dplyr::select(-term)
-    })
-    data.table::rbindlist(glmlist, id = "model_name") %>%
-        as.data.frame %>%
-        dplyr::mutate(conf.low = estimate - qnorm(1- 0.05/2) * std.error,
-                      conf.high = estimate + qnorm(1- 0.05/2) * std.error)
-}
-
-calculateadonis <- function(dset,
-                            matrix,
-                            responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
-                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
-                            covariates = c("BL_AGE", "SEX"),
-                            npermutations = 99,
-                            maxcores = 100) {
-    mclapply(responses, function(response) {
-	fo <- sprintf("matrix ~ %s + %s", response, paste(covariates, collapse = " + "))
-        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
-        ad$aov.tab %>%
-            as.data.frame %>%
-            tibble::rownames_to_column(var = "term") %>%
-            dplyr::mutate(response = response)
-      }, mc.cores = min(maxcores, length(responses)))
-}
-
-prepare.maaslin <- function (pseq,
-                             tsvfile = "merged_phfinrisk_genus.tsv",
-                             conffile = "maaslin_config.config",
-                             includedvars = c("BL_AGE", "BMI", "SEX"),
-                             readpclrows = "Archaea.Archaea-",
-                             core = FALSE,
-                             core.detection = 0.001,
-                             core.prevalence = 0.01) {
-    pseq.comp <- microbiome::transform(pseq, "compositional")
-    if (core) {
-         pseq.comp  <- microbiome::core(pseq.comp, detection = core.detection, prevalence = core.prevalence)
-    }
-    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq.comp)))
-    sample.ids <- colnames(pheno)
-    abud.rel <- abundances(pseq.comp)
-    rownames(abud.rel) <- sapply(rownames(abud.rel), replace.brackets)
-    write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
-    data <- file(conffile , open='wt')
-    pcl.contents <- c("Matrix: Metadata",
-                      paste(c("Read_PCL_Rows: ", paste(includedvars, collapse=",")), collapse = ""),
-                      "",
-                      "Matrix: Adundance",
-                      paste("Read_PCL_Rows:", readpclrows))
-    writeLines(pcl.contents, con=data)
-    close(data)
-}
-
- merge.pheno.abu <- function(pseq,
-                             core = FALSE,
-                             core.detection = 0.001,
-                             core.prevalence = 0.01) {
-     pseq.comp <- microbiome::transform(pseq, "compositional")
-     if (core) {
-         pseq.comp  <- microbiome::core(pseq.comp, detection = core.detection, prevalence = core.prevalence)
-    }
-     pheno <- microbiome::meta(phyloseq::sample_data(pseq.comp))
-     sample.ids <- rownames(pheno)
-     abud.rel <- abundances(pseq.comp)
-     rownames(abud.rel) <- sapply(rownames(abud.rel), replace.brackets)
-     cbind(pheno, t(abud.rel))
-}
-
+```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
+source("articletwo-rrbiome.R")
 ```

### 2019-01-23 articletwo.Rmd
Minor

@@
 now
+```
+
+Session info
+
+```{r session info}
+sessionInfo()
 ```
@@
 tbl1foot <- "Continuous variables are presented as mean (standard deviation).
+Exercise at free-timeis reported as not much (1), walk etc.
+(2), regular three hours or more or competitive sports (3).
 Categorical variables reported as absolute and relative frequencies.
@@
 ggsave(file = "cache/alpha-beta.png", plot=g4, height=fig.height, width=11)
-writeimage(doc, "cache/alpha-beta.png", "Alpha-beta", "footer")
+
 ```
@@
     geom_point(aes(color = HYPERTENSION), size = 0.5) +
-    scale_color_manual(values=c("red", "blue"),
+    scale_color_manual(values=c("blue", "red"),
                        name= "Hypertension",
@@
 
-writeimage(doc, "cache/pca.png", "Alpha-beta", "footer", width = 7)
+writeimage(doc, 2, "cache/pca.png", "Alpha-beta", "footer", width = 7)
 ```
+
+```{r Write figure 2}
+fig2head <- "Tenth of bacterial variation can be explained by two first principal axes"
+fig2foot <- sprintf("Principal component analysis for bacterial abundances at species level.
+Centered log ratio transform is performed for the compositional data.
+First two axes explain %.1f%% of the variation.
+Research participants with hypertension are denoted by red dot.", sum(pca.percentage[1:2]))
+writeimage(doc, 2, "cache/pca.png", fig2head, fig2foot)
+```
+
+## Direct associations between genus level and blood pressure variables (figure 3)
 
@@
 
-writeggplot(doc, maaslinplot, "maaslin", "footer", width = 7, height = 3)
+ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 7, dpi = 300)
 ```
+
+```{r Write figure 3}
+fig3head <- ""
+fig3foot <- ""
+writeimage(doc, 3, "cache/maaslin.png", fig3head, fig3foot)
+```
+
+## Figure 4
 
@@
 
-```{r PCA loadings}
+## Loadings for principal component axis 3 (Figure 5)
 
+```{r Main contributors of PC3}
 pca.rotation.abs <- abs(pca.prcomp$rotation)
@@
     tibble::rownames_to_column(var = "term")
+
+pca.prcomp$rotation %>% length
+
+pca.loading %>% pull(PC3) %>% sum
 
@@
 
-pca.loading %>% pull(PC3) %>% head(2700) %>% sum
-
-ggplot(data=pca.pc3, aes(x=reorder(term, -PC3), y=PC3)) +
+pca.pc3.gg <- ggplot(data=pca.pc3, aes(x=reorder(term, -PC3), y=PC3)) +
     geom_bar(stat="identity") +
@@
 
-dev.off()
+ggsave(file = "cache/pca.pc3.png", plot = pca.pc3.gg, height = 3, width = 7, dpi = 300)
+```
 
+```{r Write figure 4}
+fig4head <- ""
+fig4foot <- ""
+writeimage(doc, 4, "cache/pca.pc3.png", fig4head, fig4foot)
 ```

### 2019-01-23 rrbiome.Rmd
Minor

@@
 if (!file.exists("rds/adonis.species.rds")) {
-
     adonis.species <- calculateadonis(dset = meta(pseq.species),
@@
                                       covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                    npermutations = 1)
-
+                                      npermutations = 99)
     saveRDS(adonis.species, file = "rds/adonis.species.rds")
-
     } else {

### 2019-01-23 articletwo-rrbiome.R

@@ -64,14 +64,14 @@ calculateadonis <- function(dset,
                             covariates = c("BL_AGE", "SEX"),
                             npermutations = 99,
                             maxcores = 100) {
-    lapply(responses, function(response) {
+    mclapply(responses, function(response) {
         fo <- sprintf("matrix ~ %s + %s", response, paste(covariates, collapse = " + "))
         ad <- adonis(formula = as.formula(fo), data = dset, permutations = npermutations)
         ad$aov.tab %>%
             as.data.frame %>%
             tibble::rownames_to_column(var = "term") %>%
             dplyr::mutate(response = response)
-      })#, mc.cores = min(maxcores, length(responses)))
+      }, mc.cores = min(maxcores, length(responses)))
 }
 
 prepare.maaslin <- function (pseq,

### 2019-01-23 articletwo.Rmd
Figure 4

@@
 ```{r Main contributors of PC3}
-pca.rotation.abs <- abs(pca.prcomp$rotation)
-pca.loading <- sweep(pca.rotation.abs, 2, colSums(pca.rotation.abs), "/") %>%
+pca.loading <- pca.prcomp$rotation^2 %>%
     as.data.frame %>%
-    tibble::rownames_to_column(var = "term")
+    tibble::rownames_to_column(var = "term") %>%
+    mutate(name = rename.genus(term))
 
-pca.prcomp$rotation %>% length
-
-pca.loading %>% pull(PC3) %>% sum
-
-pca.pc3 <- pca.loading %>%
-    select(term, PC3) %>%
+pca.loading.pc3.top20 <- pca.loading %>%
+    select(name, PC3) %>%
     arrange(-PC3) %>%
@@
 
-pca.pc3.gg <- ggplot(data=pca.pc3, aes(x=reorder(term, -PC3), y=PC3)) +
-    geom_bar(stat="identity") +
-    theme_minimal() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-ggsave(file = "cache/pca.pc3.png", plot = pca.pc3.gg, height = 3, width = 7, dpi = 300)
+pca.pc3.gg <- ggplot(data=pca.loading.pc3.top20, aes(x=reorder(name, -PC3), y=100*PC3)) +
+    geom_bar(stat="identity", width=0.5, position = position_dodge(preserve = "single")) +
+    theme_classic(20) +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.3, hjust = 1)) +
+    xlab("") +
+    ylab("Contribution (%)") +
+#    coord_fixed(ratio = 1) +
+    theme(axis.title.x=element_blank(),
+          text = element_text(size=20),
+          legend.justification=c(1,0),
+          legend.position=c(1.03, 0.75),
+          legend.title=element_text(size = 20),
+          legend.direction="vertical",
+          legend.text=element_text(size = floor(0.75*20)),
+          legend.box.margin = margin(1, 1, 1, 1, "mm"),
+          legend.box.background = element_rect(colour = "black"),
+          aspect.ratio = 0.7)
+   ggsave(file = "cache/pca.pc3.png", plot = pca.pc3.gg, height = 8, width = 8, dpi = 300) 
 ```
@@
 ```{r Write figure 4}
-fig4head <- ""
-fig4foot <- ""
+fig4head <- "Bacteria causing dysentery and diarrhea contribute to third principal component"
+fig4foot <- sprintf("Top 20 bacterial contributors to third principal coordinate at species level.
+Centered log ratio transform is performed for the compositional data.
+Third principal component explain %.1f%% of the total variation.
+Plasmid DNA is denoted using star.", pca.percentage[3])
 writeimage(doc, 4, "cache/pca.pc3.png", fig4head, fig4foot)

### 2019-01-23 articletwo-rrbiome.R

@@ -2,6 +2,12 @@
 
 replace.brackets <- function (genus) gsub('(.*) \\(+(.+)\\)', '\\1.\\2', genus)
 
+rename.genus  <- function (genus, markword = "Plasmid", mark = "*") {
+    name <- gsub('(.*) \\(.*\\)', '\\1', genus)
+    star <- ifelse(grepl(markword, genus), mark, "")
+    paste0(gsub("_", " ", name), star)
+}
+
 filter.phenotype.data <- function(pseq,
                                   included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
                                                "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",

### 2019-01-23 articletwo.Rmd
Bugs at atlas

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis', cache=TRUE, warning=FALSE)
+knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis', cache=FALSE, warning=FALSE)
 knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
@@
 
-```{r Officer importer, code=readLines("articleone-officer.R")}
+```{r Officer importer, code=readLines("articletwo-officer.R")}
 source("articletwo-officer.R")
@@
     data.table::rbindlist(id = "model_name") %>%
-    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"))
-
-                                        #%>%
+    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
     dplyr::select(response, R2)
@@
 plot.gtable.beta.diversity <- ggplot_gtable(ggplot_build(plot.beta.diversity))
-plot.xlab.beta.diversity <- xlabb_extractor(gtable.beta.diversity)
+plot.xlab.beta.diversity <- xlabb_extractor(plot.gtable.beta.diversity)
 
@@
 ```{r Main contributors of PC3}
-pca.loading <- pca.prcomp$rotation^2 %>%
+pca.rotation.abs <- abs(pca.prcomp$rotation)
+pca.loading <- sweep(pca.rotation.abs, 2, colSums(pca.rotation.abs), "/") %>%
 	    as.data.frame %>%
@@
 
-```{r Write docx to file, cache = FALSE, results = FALSE, include = FALSE}
+```{r Write docx to file}
 print(doc, target = paste0("report/articletwo-", now, ".docx"))

### 2019-01-23 articletwo-officer.R

@@ -17,8 +17,8 @@ characteristics <- function(dset, names, factors) {
         format(justify = "left", trim = TRUE) %>%
         rowwise() %>%
         mutate(id = gsub("([A-Za-z_0-9]+).*", "\\1", rowname)) %>%
-        mutate(present = id %in%  names(tableone.names)) %>%
-        mutate(rowname = ifelse(present == TRUE, tableone.names[[id]], rowname)) %>%
+        mutate(present = id %in%  names(names)) %>%
+        mutate(rowname = ifelse(present == TRUE, names[[id]], rowname)) %>%
         select(rowname, Overall)
 }
 
@@ -145,6 +145,9 @@ riskmodel.joincolumns <- function(rset, riskmodels, idforn = "fwdbonf_riskclass"
 }
 
 writetable <- function(doc, tbl, number = 0, head, foot) {
+	text.bold <- fp_text(font.size = 12, bold = TRUE, font.family = "Arial")
+	text.normal <- fp_text(font.size = 12, font.family = "Arial")
+
   doc <- doc %>%
     body_add_fpar(fpar(ftext(paste0("Table ", number, ". "), prop = text.bold), ftext(head, prop = text.normal)), style = "Table Caption") %>%
     body_add_flextable(tbl, align = "left") %>%
@@ -153,6 +156,9 @@ writetable <- function(doc, tbl, number = 0, head, foot) {
 }
 
 writeimage <- function(doc, number = 0, filename, header, footer, width = 7) {
+text.bold <- fp_text(font.size = 12, bold = TRUE, font.family = "Arial")
+text.normal <- fp_text(font.size = 12, font.family = "Arial")
+
   img <- readPNG(filename)
   doc <- doc %>%
     body_add_fpar(fpar(ftext(paste0("Figure ", number, ". "), prop = text.bold), ftext(header, prop = text.normal)), style = "Image Caption") %>%

### 2019-01-23 articletwo.Rmd
Minoq

--- 2019-01-23 articletwo.Rmd
@@
 
-ggsave(file = "cache/alpha-beta.png", plot=g4, height=fig.height, width=11)
+ggsave(file = "cache/alpha-beta.png", plot=g4, height=10, width=11)
 
@@
     ylab("Contribution (%)") +
-#    coord_fixed(ratio = 1) +
+    scale_y_continuous(limits=c(0, 2.7), oob = rescale_none, expand = c(0,0)) +
     theme(axis.title.x=element_blank(),

### 2019-01-25 articletwo.Rmd
Minor

@@
 
-```{r session info}
-sessionInfo()
+```{r Officer importer, code = sessionInfo()}
 ```
@@
     mutate(Qstar = ifelse(p.value < 0.05, '*', ' '))
-
-fig.height <- 4.0 + nrow(diversity) * 0.33
-fig.permar2.text <- "R2 of variable \nin Bray-Curtis distance" # TODO latexify expression?
-fig.textsize <- 15
-amaxb <- diversity %>% pull(estimate) %>% abs %>% max %>% signif(., digits = 2)
 ```
@@
 ```{r alphabeta alpha}
+plot.alpha <- function(diversity,  width = .4, posdodge = .25, aspectratio = 2, colorlimit = 0.25) {
     plot.alpha.diversity <- ggplot(diversity, aes(x=reorder(Name, R2),  y = 1, fill = estimate)) +
@@
         scale_fill_gradientn(colours = c("darkblue", "blue", "white", "red", "darkred"),
-                         values = rescale(c(-amaxb,-0.1, 0, 0.1, amaxb)),
+                             values = rescale(c(-amaxb,-0.1, 0, 0.1, colorlimit)),
                              name = 'Regression coefficient \nin linear model \nfor Shannon index',
@@
         theme(legend.position="right",
-          legend.title=element_text(size = 20),
+              legend.title=element_text(size = 14),
               legend.direction="vertical",
@@
     plot.gtable.alpha.diversity.stripped <- ggplot_gtable(ggplot_build(plot.alpha.diversity))
+    return(list(plot = plot.gtable.alpha.diversity.stripped$grobs[[6]],
+                yaxis = plot.alpha.diversity.yaxis,
+                xaxis = plot.alpha.diversity.gtable$grobs[[7]],
+                legend = plot.alpha.diversity.legend))
+}
+
 ```
@@
 ```{r alphabeta beta}
+plot.beta <- function(diversity, fig.permar2.text = "R2 of variable \nin Bray-Curtis distance", fig.textsize = 15) {
     plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2)) +
-    geom_bar(stat="identity", color='black') +
+        geom_bar(stat="identity", color="black") +
         coord_flip() +
@@
     plot.gtable.beta.diversity.stripped <- ggplot_gtable(ggplot_build(plot.beta.diversity))
+
+      return(list(plot =plot.gtable.beta.diversity.stripped$grobs[[6]],
+                xaxis = plot.gtable.beta.diversity$grobs[[7]]))
+}
 ```
@@
 ```{r save grob}
-gs4 <- list(plot.alpha.diversity.yaxis,
-            plot.gtable.alpha.diversity.stripped$grobs[[6]],
-            plot.gtable.beta.diversity.stripped$grobs[[6]],
-            plot.alpha.diversity.legend,
-            plot.alpha.diversity.gtable$grobs[[7]],
-            plot.gtable.beta.diversity$grobs[[7]],
+alpha.diversity <- plot.alpha(diversity)
+beta.diversity <- plot.beta(diversity)
+gs4 <- list(alpha.diversity$yaxis,
+            alpha.diversity$plot,
+            beta.diversity$plot,  
+            alpha.diversity$legend,
+            alpha.diversity$xaxis,
+            beta.diversity$xaxis,
            plot.xlab.beta.diversity)
-#           plot.gtable.beta.diversity.stripped[[3]])
 
@@
             layout_matrix = rbind(
-                c(1,2,8,3,NA),
-                c(1,2,8,3,4),
-                c(1,2,8,3,NA),
-                c(1,2,8,3,NA),
-                c(NA,5,NA,6,NA),
-                c(NA,NA,NA,7,NA)),
-            widths=c(4,0.7,0.5,5,5),
-            heights=c(3,12,8,21,1,2))
-
-ggsave(file = "cache/alpha-beta.png", plot=g4, height=10, width=11)
-
+                c(1,    2,    3, 4),
+                c(NA, 5,    6, NA),
+                c(NA, NA, 7, NA)),
+            widths=c(4, 0.7, 5, 5),
+            heights=c(10, 1, 2))
+ggsave(file = "cache/alpha-beta.png", plot=g4, height=3, width=9)
 ```

### 2019-01-25 rrbiome.Rmd
Minor

--- 2019-01-23 rrbiome.Rmd
@@
  if (!file.exists("rds/adonis.species.rds")) {
-
-    adonis.species <- calculateadonis(dset = meta(pseq.species),
+     adonis.species = list()
+     adonis.species$min <- calculateadonis(dset = meta(pseq.species),
+                                           matrix = bray.dist.m.species,
+                                           covariates = c("BL_AGE", "SEX"),
+                                           npermutations = 99)
+     adonis.species$max <- calculateadonis(dset = meta(pseq.species),
                                            matrix = bray.dist.m.species,
                                            covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                    npermutations = 1)
+                                           npermutations = 99)
 
      saveRDS(adonis.species, file = "rds/adonis.species.rds")
-
     } else {

### 2019-01-25 articletwo.Rmd
Minor

@@
 
+adonis.species$min
+
 beta.diversity <- adonis.species %>%
@@
               legend.text=element_text(size = floor(0.75*20)),
-              legend.box.margin=margin(0,0,0,0,"pt"),
+              legend.box.margin=margin(t = 1, unit = "cm"),
               legend.margin=margin(0,0,0,0,"pt"),
@@
                   layout_matrix = rbind(
-                c(1,    2,    3, 4),
-                c(NA, 5,    6, NA),
-                c(NA, NA, 7, NA)),
-            widths=c(4, 0.7, 5, 5),
-            heights=c(10, 1, 2))
+                      c(1,    2,    3, NA, NA),
+                      c(1,    2,    3, NA, 4),
+                      c(NA, 5,    6, NA, NA),
+                      c(NA, NA, 7, NA, NA)),
+                  widths=c(4, 0.7, 5, 0.7, 3),
+            heights=c(3, 7, 1, 2))
 ggsave(file = "cache/alpha-beta.png", plot=g4, height=3, width=9)
+
 ```

### 2019-01-25 rrbiome.Rmd
Minor

--- 2019-01-23 rrbiome.Rmd
@@
  if (!file.exists("rds/adonis.species.rds")) {
-
-    adonis.species <- calculateadonis(dset = meta(pseq.species),
+     adonis.species = list()
+     adonis.species$min <- calculateadonis(dset = meta(pseq.species),
+                                           matrix = bray.dist.m.species,
+                                           covariates = c("BL_AGE", "SEX"),
+                                           npermutations = 99)
+     adonis.species$max <- calculateadonis(dset = meta(pseq.species),
                                            matrix = bray.dist.m.species,
                                            covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                    npermutations = 1)
-
+                                           npermutations = 99)
      saveRDS(adonis.species, file = "rds/adonis.species.rds")
-
     } else {

### 2019-01-25 articletwo.Rmd
Minor

@@
         scale_fill_gradientn(colours = c("darkblue", "blue", "white", "red", "darkred"),
-                             values = rescale(c(-amaxb,-0.1, 0, 0.1, colorlimit)),
+                             values = rescale(c(-colorlimit, 0, colorlimit)),
                              name = 'Regression coefficient \nin linear model \nfor Shannon index',
-                             limits = c(-1.05*amaxb, 1.05*amaxb),
+                             limits = c(-colorlimit, colorlimit),
                              na.value = "black") +
@@
               legend.text=element_text(size = floor(0.75*20)),
-              legend.box.margin=margin(t = 1, unit = "cm"),
+              legend.box.margin=margin(t = 0, unit = "cm"),
               legend.margin=margin(0,0,0,0,"pt"),
@@
                   layout_matrix = rbind(
+                      c(4,    2,    3, NA, NA),
                       c(1,    2,    3, NA, NA),
-                      c(1,    2,    3, NA, 4),
                       c(NA, 5,    6, NA, NA),
                       c(NA, NA, 7, NA, NA)),
-                  widths=c(4, 0.7, 5, 0.7, 3),
+                  widths=c(4, 0.7, 5, 0.7, 4),
             heights=c(3, 7, 1, 2))

### 2019-01-25 rrbiome.Rmd
Minor

--- 2019-01-23 rrbiome.Rmd
@@
 
-```{r libraries, cache = FALSE}
+```{r libraries}
 library(dplyr)
@@
 
-```{r matrix calculation, cache = FALSE}
+```{r matrix calculation}
 if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds")) {
@@
     saveRDS(bray.dist.m.species, file = "rds/bray.dist.m.species.rds")
-    bray.dist.m.genus <- calculate.beta.matrix(pseq.genus)
-    saveRDS(bray.dist.m.genus, file = "rds/bray.dist.m.genus.rds")
 } else {
     bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
-    bray.dist.m.genus <- readRDS("rds/bray.dist.m.genus.rds")
 }
@@
 
- ```{r adonis calculation, cache = FALSE}
+ ```{r adonis calculation}
  if (!file.exists("rds/adonis.species.rds")) {
-    adonis.species <- calculateadonis(dset = meta(pseq.species),
+
+     adonis.species = list()
+     adonis.species$min <- calculateadonis(dset = meta(pseq.species),
+                                           matrix = bray.dist.m.species,
+                                           covariates = c("BL_AGE", "SEX"),
+                                           npermutations = 99)
+
+     adonis.species$max <- calculateadonis(dset = meta(pseq.species),
                                            matrix = bray.dist.m.species,

### 2019-01-25 articletwo.Rmd
New file

--- 2019-01-25 articletwo.Rmd
@@
 
+### Plot functions
+
+```{r EdgeR functions, code=readLines("aaro/edgeR_plot_functions.R")}
+source("articletwo-ggplot.R")
+```
+
 # Loading data
@@
     mutate(Qstar = ifelse(p.value < 0.05, '*', ' '))
-```
-
-```{r alphabeta alpha}
-plot.alpha <- function(diversity,  width = .4, posdodge = .25, aspectratio = 2, colorlimit = 0.25) {
-    plot.alpha.diversity <- ggplot(diversity, aes(x=reorder(Name, R2),  y = 1, fill = estimate)) +
-        geom_bar(stat="identity", color="black") +
-        coord_flip() +
-        scale_fill_gradientn(colours = c("darkblue", "blue", "white", "red", "darkred"),
-                             values = rescale(c(-amaxb,-0.1, 0, 0.1, colorlimit)),
-                             name = 'Regression coefficient \nin linear model \nfor Shannon index',
-                             limits = c(-1.05*amaxb, 1.05*amaxb),
-                             na.value = "black") +
-        theme_classic(20) +
-        geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=TRUE, color='black', size=20) +
-        scale_shape_manual(name="",
-                           values=c('*'='*', ' '=' '),
-                           labels=c("*"="significant at FDR 0.05", ' '=' '),
-                           breaks=c("*", ' ')) +
-        xlab("") +
-        ylab("") +
-        theme(legend.position="right",
-              legend.title=element_text(size = 14),
-              legend.direction="vertical",
-              legend.text=element_text(size = floor(0.75*20)),
-              legend.box.margin=margin(0,0,0,0,"pt"),
-              legend.margin=margin(0,0,0,0,"pt"),
-              legend.justification="left",
-              axis.line.y = element_line(linetype="blank"),
-              axis.line.x = element_line(linetype="blank"),
-              legend.title.align = 0,
-              legend.text.align = 0,
-              axis.text.y = element_text(color="black"),
-              axis.text.x = element_text(color="black"))+
-        scale_y_continuous(breaks=c(0.5), labels=c(expression(alpha)))
-
-    plot.alpha.diversity.gtable <- ggplot_gtable(ggplot_build(plot.alpha.diversity))
-    plot.alpha.diversity.legend <- legend_extractor(plot.alpha.diversity.gtable)
-    plot.alpha.diversity.yaxis <- plot.alpha.diversity.gtable$grobs[[3]]
-
-    plot.alpha.diversity <- plot.alpha.diversity +
-        theme(legend.position = "none",
-              axis.text.x = element_blank(),
-              axis.text.y=element_blank())
-
-    plot.gtable.alpha.diversity.stripped <- ggplot_gtable(ggplot_build(plot.alpha.diversity))
-    return(list(plot = plot.gtable.alpha.diversity.stripped$grobs[[6]],
-                yaxis = plot.alpha.diversity.yaxis,
-                xaxis = plot.alpha.diversity.gtable$grobs[[7]],
-                legend = plot.alpha.diversity.legend))
-}
-
-```
-
-```{r alphabeta beta}
-plot.beta <- function(diversity, fig.permar2.text = "R2 of variable \nin Bray-Curtis distance", fig.textsize = 15) {
-    plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2)) +
-        geom_bar(stat="identity", color="black") +
-        coord_flip() +
-        theme_classic(fig.textsize) +
-        ylab(fig.permar2.text) +
-        theme(axis.text.y = element_text(color="black"),
-              axis.text.x = element_text(color="black"),
-              legend.position="right",
-              legend.direction="vertical",
-              legend.box.margin=margin(0,0,0,0,"pt"),
-              legend.margin=margin(0,0,0,0,"pt"),
-              legend.justification="left",
-              legend.title.align = 0,
-              legend.text.align = 0)
-
-    plot.gtable.beta.diversity <- ggplot_gtable(ggplot_build(plot.beta.diversity))
-    plot.xlab.beta.diversity <- xlabb_extractor(plot.gtable.beta.diversity)
-
-    plot.beta.diversity <- plot.beta.diversity  +
-        theme(axis.text.x = element_blank(),
-              axis.text.y = element_blank(),
-              axis.title.x  =  element_blank(),
-              axis.title.y = element_blank(),
-              legend.position = "none")
-    plot.gtable.beta.diversity.stripped <- ggplot_gtable(ggplot_build(plot.beta.diversity))
-
-      return(list(plot =plot.gtable.beta.diversity.stripped$grobs[[6]],
-                xaxis = plot.gtable.beta.diversity$grobs[[7]]))
-}
 ```
@@
                   layout_matrix = rbind(
-                c(1,    2,    3, 4),
-                c(NA, 5,    6, NA),
-                c(NA, NA, 7, NA)),
-            widths=c(4, 0.7, 5, 5),
-            heights=c(10, 1, 2))
+                      c(1,    2,    3, NA, 4),
+                      c(1,    2,    3, NA, NA),
+                      c(NA, 5,    6, NA, NA),
+                      c(NA, NA, 7, NA, NA)),
+                  widths=c(4, 0.7, 5, 0.7, 4),
+            heights=c(3, 7, 1, 2))
 ggsave(file = "cache/alpha-beta.png", plot=g4, height=3, width=9)
+
 ```

### 2019-01-25 articletwo-ggplot.R

@@ -1,0 +1,78 @@
+plot.alpha <- function(diversity,  colorlimit = 0.25) {
+    plot.alpha.diversity <- ggplot(diversity, aes(x=reorder(Name, R2),  y = 1, fill = estimate)) +
+        geom_bar(stat="identity", color="black") +
+        coord_flip() +
+        scale_fill_gradientn(colours = c("darkblue", "white", "darkred"),
+                             values = rescale(c(-colorlimit, 0, colorlimit)),
+                             name = 'Regression coefficient \nin linear model \nfor Shannon index',
+                             limits = c(-colorlimit, colorlimit),
+                             na.value = "black") +
+        theme_classic(20) +
+        geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=TRUE, color='black', size=20) +
+        scale_shape_manual(name="",
+                           values=c('*'='*', ' '=' '),
+                           labels=c("*"="significant at FDR 0.05", ' '=' '),
+                           breaks=c("*", ' ')) +
+        xlab("") +
+        ylab("") +
+        theme(legend.position="right",
+              legend.title=element_text(size = 14),
+              legend.direction="vertical",
+              legend.text=element_text(size = floor(0.75*20)),
+              legend.box.margin=margin(t = 0, unit = "cm"),
+              legend.margin=margin(0,0,0,0,"pt"),
+              legend.justification="left",
+              axis.line.y = element_line(linetype="blank"),
+              axis.line.x = element_line(linetype="blank"),
+              legend.title.align = 0,
+              legend.text.align = 0,
+              axis.text.y = element_text(color="black"),
+              axis.text.x = element_text(color="black"))+
+        scale_y_continuous(breaks=c(0.5), labels=c(expression(alpha)))
+
+    plot.alpha.diversity.gtable <- ggplot_gtable(ggplot_build(plot.alpha.diversity))
+    plot.alpha.diversity.legend <- legend_extractor(plot.alpha.diversity.gtable)
+    plot.alpha.diversity.yaxis <- plot.alpha.diversity.gtable$grobs[[3]]
+
+    plot.alpha.diversity <- plot.alpha.diversity +
+        theme(legend.position = "none",
+              axis.text.x = element_blank(),
+              axis.text.y=element_blank())
+
+    plot.gtable.alpha.diversity.stripped <- ggplot_gtable(ggplot_build(plot.alpha.diversity))
[…]

### 2019-01-25 rrbiome.Rmd
Minor

@@
 ```{r adonis3}
-adonis.species %>%
+adonis.species$max %>%
     data.table::rbindlist(id = "model_name") %>%

### 2019-01-25 articletwo.Rmd
Minor

@@
 library(scales)
+library(ggplot2)
 library(gridExtra)
@@
 
-```{r EdgeR functions, code=readLines("aaro/edgeR_plot_functions.R")}
+```{r EdgeR functions, code=readLines("articletwo-ggplot.R")}
 source("articletwo-ggplot.R")
@@
 adonis.species  <- readRDS("rds/adonis.species.rds")
-
 adonis.species$min
@@
             beta.diversity$xaxis,
-           plot.xlab.beta.diversity)
+            plot.xlab.beta.diversity,
+            text_grob("A", vjust = "top", hjust= "left", size = 20))
 
@@
                   layout_matrix = rbind(
-                      c(1,    2,    3, NA, 4),
-                      c(1,    2,    3, NA, NA),
-                      c(NA, 5,    6, NA, NA),
-                      c(NA, NA, 7, NA, NA)),
-                  widths=c(4, 0.7, 5, 0.7, 4),
-            heights=c(3, 7, 1, 2))
+                      c(8,    NA,    NA,    NA, NA, 4),
+                      c(NA, 1,    2,    3, NA, 4),
+                      c(NA, 1,    2,    3, NA, NA),
+                      c(NA, NA, 5,    6, NA, NA),
+                      c(NA, NA, NA, 7, NA, NA)),
+                  widths=c(0.7, 4, 0.7, 5, 0.7, 4),
+                  heights=c(1, 1, 6, 1, 2))
 ggsave(file = "cache/alpha-beta.png", plot=g4, height=3, width=9)

### 2019-01-25 articletwo.Rmd
Minor

@@
 dset.alpha <- meta.merge.alphadiversity(pseq.species)
-alphaglm <- calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon")
-alpha.diversity <- alphaglm %>% select(response, estimate, p.value)
+alphadiversity <- list(
+    min = calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX  + diversities_shannon"),
+    max = calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon"))
+betadiversity  <- readRDS("rds/adonis.species.rds")
+diversities <- mergediversities(alphadiversity, betadiversity)
 
-adonis.species  <- readRDS("rds/adonis.species.rds")
-adonis.species$min
+betadiversity$min$model_1 %>% select(term, R2)
+betadiversity$max$model_1 %>% select(term, R2)
 
-beta.diversity <- adonis.species %>%
-    data.table::rbindlist(id = "model_name") %>%
-    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
-    dplyr::select(response, R2)
-
-diversity <- merge(alpha.diversity, beta.diversity, by = "response") %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate") %>%
-    mutate(Qstar = ifelse(p.value < 0.05, '*', ' '))
 ```
 
+Max values for alpha diversity `r diversities$max %>% pull(estimate)
+								  %>% abs %>% max %>% round(., 2)` and
+								  `r diversities$mmin %>%
+								  pull(estimate) %>% abs %>% max %>%
+								  round(., 2)`.
+
 ```{r save grob}
-alpha.diversity <- plot.alpha(diversity)
-beta.diversity <- plot.beta(diversity)
-gs4 <- list(alpha.diversity$yaxis,
-            alpha.diversity$plot,
-            beta.diversity$plot,  
-            alpha.diversity$legend,
-            alpha.diversity$xaxis,
-            beta.diversity$xaxis,
-            plot.xlab.beta.diversity,
-            text_grob("A", vjust = "top", hjust= "left", size = 20))
+alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 1.3)
+alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 1.3)
+beta.diversity.min <- plot.beta(diversities$min)
+beta.diversity.max <- plot.beta(diversities$max)
+
+gs4 <- list(alpha.diversity.min$yaxis,
+            alpha.diversity.min$plot,
+            beta.diversity.min$plot,  
+            alpha.diversity.min$legend,
+            alpha.diversity.min$xaxis,
+            beta.diversity.min$xaxis,
+            beta.diversity.min$xlab,
+            text_grob("A.", size = 18),
+            alpha.diversity.max$yaxis,
+            alpha.diversity.max$plot,
+            beta.diversity.max$plot,  
+            alpha.diversity.max$xaxis,
+            beta.diversity.max$xaxis,
+            beta.diversity.max$xlab,
+            text_grob("B.", size = 18))
 
@@
                   layout_matrix = rbind(
-                      c(8,    NA,    NA,    NA, NA, 4),
-                      c(NA, 1,    2,    3, NA, 4),
-                      c(NA, 1,    2,    3, NA, NA),
-                      c(NA, NA, 5,    6, NA, NA),
-                      c(NA, NA, NA, 7, NA, NA)),
-                  widths=c(0.7, 4, 0.7, 5, 0.7, 4),
-                  heights=c(1, 1, 6, 1, 2))
-ggsave(file = "cache/alpha-beta.png", plot=g4, height=3, width=9)
+                      c(8,    NA,    NA, NA,   NA, NA, NA),
+                      c(NA, 1,    2,    NA, 3, NA, NA),
+                      c(NA, 1,    2,    NA, 3, NA, 4),
+                      c(NA, NA, 5,    NA,6, NA, 4),
+                      c(NA, NA, NA, NA,7, NA, 4),
+                      c(15,    NA,    NA, NA,   NA, NA, 4),
+                      c(NA, 9,    10,  NA,  11, NA, 4),
+                      c(NA, 9,    10,  NA,  11, NA, NA),
+                      c(NA, NA, 12,   NA, 13, NA, NA),
+                      c(NA, NA, NA, NA,14, NA, NA)),
+                  widths=c(0.7, 4, 0.7, 0.1, 5, 0.7, 4),
+                  heights=rep(c(1, 1, 6, 1, 2), 2))
+ggsave(file = "cache/alpha-beta.png", plot=g4, height=6, width=9)
+```
 
+```{r Write figure 2}
+fig1head <- "alpha beta figure"
+fig1foot <- ""
+writeimage(doc, 1, "cache/alpha-beta.png", fig1head, fig1foot)
 ```
@@
 ggsave(file = "cache/pca.png", plot = pcaplot, height = 5, width = 9, dpi = 300)
-
-writeimage(doc, 2, "cache/pca.png", "Alpha-beta", "footer", width = 7)
 ```

### 2019-01-25 articletwo-ggplot.R

@@ -2,16 +2,16 @@ plot.alpha <- function(diversity,  colorlimit = 0.25) {
     plot.alpha.diversity <- ggplot(diversity, aes(x=reorder(Name, R2),  y = 1, fill = estimate)) +
         geom_bar(stat="identity", color="black") +
         coord_flip() +
-        scale_fill_gradientn(colours = c("darkblue", "white", "darkred"),
-                             values = rescale(c(-colorlimit, 0, colorlimit)),
+        scale_fill_gradientn(colours = c("blue", "white"),
+                             breaks = c(-colorlimit, 0),
                              name = 'Regression coefficient \nin linear model \nfor Shannon index',
-                             limits = c(-colorlimit, colorlimit),
+                             limits = c(-colorlimit, 0),
                              na.value = "black") +
         theme_classic(20) +
         geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=TRUE, color='black', size=20) +
         scale_shape_manual(name="",
                            values=c('*'='*', ' '=' '),
-                           labels=c("*"="significant at FDR 0.05", ' '=' '),
+                           labels=c("*"="significant at\nFDR 0.05", ' '=' '),
                            breaks=c("*", ' ')) +
         xlab("") +
         ylab("") +
@@ -74,5 +74,6 @@ plot.beta <- function(diversity, fig.permar2.text = "R2 of variable \nin Bray-Cu
     plot.gtable.beta.diversity.stripped <- ggplot_gtable(ggplot_build(plot.beta.diversity))
 
       return(list(plot =plot.gtable.beta.diversity.stripped$grobs[[6]],
-                xaxis = plot.gtable.beta.diversity$grobs[[7]]))
+                  xaxis = plot.gtable.beta.diversity$grobs[[7]],
+                  xlab = plot.xlab.beta.diversity))
 }

### 2019-01-25 articletwo-rrbiome.R

@@ -138,3 +138,21 @@ xlabb_extractor <- function(tmp) {
   tmp$grobs[[xl]]
 }
 
+vectortolist <- function(c) {
+  l <- as.list(c)
+  names(l) <- l
+  l
+}
+
+mergediversities <- function(alphadiversity, betadiversity) {
+    lapply(vectortolist(names(alphadiversity)), function(x) {
+        alpha.select <- alphadiversity[[x]] %>% select(estimate, p.value, response)
+        beta.rbind <- betadiversity[[x]] %>%
+            data.table::rbindlist(id = "model_name") %>%
+            dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
+            dplyr::select(response, R2)
+        merge( alpha.select , beta.rbind, by = "response") %>%
+            merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate") %>%
+            mutate(Qstar = ifelse(p.value < 0.05, '*', ' '))
+    })
+}

### 2019-01-25 articletwo.Rmd
Minor

@@
 ```{r variables, warning = FALSE}
-data("phfinrisk_metadatadesc", package="finriskmetagcommon")
-names.dset <- filter(phfinrisk_metadatadesc,
-                     Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
-names.dset <- bind_rows(names.dset,
-                        data.frame(Covariate = c("MAP", "PULSEPRESSURE", "HYPERTENSION", "SEX"),
-                                   Category = rep("Physical", 4),
-                                   Name =c( "Mean artrerial pressure", "Pulse pressure", "Hypertension", "Female"),
-                                   Desc = c("Mean arterail presure, 2./3.*DIASM + 1./3.*SYSTM)",
-                                            "Pulse pressure, SYSTM - DIASM",
-                                            "Hypertension, SYSTM >=140 or DIAS >= 90 or BP mediaction",
-                                            "Sex is female, True for female, False for male")))
+names.dset <- getdescriptions()
 ```
@@
 Sample data has dimensions (`r dim(sample_data(pseq.species))`).
+
+## Models
+
+```{r matrix calculation}
+if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds")) {
+    bray.dist.m.species <- calculate.beta.matrix(pseq.species)
+    saveRDS(bray.dist.m.species, file = "rds/bray.dist.m.species.rds")
+} else {
+    bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
+}
+```
+
+ ```{r adonis calculation}
+ if (!file.exists("rds/adonis.species.rds")) {
+     adonis.species = list()
+     adonis.species$min <- calculateadonis(dset = meta(pseq.species),
+                                           matrix = bray.dist.m.species,
+                                           covariates = c("BL_AGE", "SEX"),
+                                           npermutations = 99)
+     adonis.species$max <- calculateadonis(dset = meta(pseq.species),
+                                           matrix = bray.dist.m.species,
+                                           covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
+                                                          "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
+                                           npermutations = 99)
+     saveRDS(adonis.species, file = "rds/adonis.species.rds")
+    } else {
+     adonis.species  <- readRDS("rds/adonis.species.rds")
+}
+```
+
+```{r pca}
+if (!file.exists("rds/pca.prcomp.rds")) {
+    pca.abundances  <- as.matrix(t(microbiome::abundances(microbiome::transform(pseq.species, "clr"))))
+    pca.prcomp <- stats::prcomp(pca.abundances)
+    saveRDS(pca.prcomp, file = "rds/pca.prcomp.rds")
+} else {
+    pca.prcomp <- readRDS("rds/pca.prcomp.rds")
+}
+```
+
+```{r run maaslin, results = 'hide'} 
+maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
+
+prepare.maaslin(pseq.genus,
+                 tsvfile = "maaslin_core_merged_phfinrisk_genus.tsv",
+                conffile = "maaslin_core_config.config",
+                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars),
+                core = TRUE,
+                readpclrows = "Halapricum.Archaea-")
+
+Maaslin("maaslin_core_merged_phfinrisk_genus.tsv",
+        "maaslin_core",
+        strInputConfig = "maaslin_core_config.config",
+        strModelSelection="none",
+        dSignificanceLevel = 1.0, 
+        fAllvAll = TRUE,
+        strForcedPredictors = maaslin.force.vars)
+```
+
+```{r lactobacillus scatterplot,  results = 'hide'}
+pseq.lacto  <- subset_taxa(pseq.species, Genus=="Lactobacillus")
+maaslin.force.vars.lacto <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
+
+prepare.maaslin(pseq.lacto,
+                 tsvfile = "maaslin_lacto_merged_phfinrisk_genus.tsv",
+                conffile = "maaslin_lacto_config.config",
+                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars.lacto),
+                readpclrows = "Lactobacillus_delbrueckii.Bacteria,Lactobacillus_ruminis.Bacteria,Lactobacillus_acidophilus.Bacteria,Lactobacillus_fermentum.Bacteria,Lactobacillus_paracasei.Bacteria,Lactobacillus_salivarius.Bacteria")
+
+Maaslin("maaslin_lacto_merged_phfinrisk_genus.tsv",
+        "maaslin_lacto",
+        strInputConfig = "maaslin_lacto_config.config",
+        strModelSelection="none",
+        dSignificanceLevel = 1.0, 
+        fAllvAll = TRUE,
+        strForcedPredictors = maaslin.force.vars.lacto)
+```
 
@@
     max = calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon"))
-betadiversity  <- readRDS("rds/adonis.species.rds")
-diversities <- mergediversities(alphadiversity, betadiversity)
+diversities <- mergediversities(alphadiversity, adonis.species)
 
@@
 ```{r pca}
-pca.prcomp <- readRDS("rds/pca.prcomp.rds")
 pca.eigs <- pca.prcomp$sdev^2
@@
 
+### Lactobacillus
+
+```{r Reading maaslin lacto output}
+maas.lacto.results.df.signf <- read.table("maaslin_lacto/maaslin_lacto_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
+    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
+                  beta = Coefficient) %>%
+   dplyr::filter(Qval < 0.05)
+
+maas.lacto.signf.features <- maas.results.df.signf %>%
+    pull(Feature) %>%
+    as.character() %>%
+    gsub("_", ".", .)
+
+df.lacto.merged <-  merge.pheno.abu(pseq.lacto)
+```
+
 ```{r Write docx to file}

### 2019-01-25 articletwo-rrbiome.R

@@ -156,3 +156,17 @@ mergediversities <- function(alphadiversity, betadiversity) {
             mutate(Qstar = ifelse(p.value < 0.05, '*', ' '))
     })
 }
+
+getdescriptions <- function() {
+    data("phfinrisk_metadatadesc", package="finriskmetagcommon")
+    names.dset <- filter(phfinrisk_metadatadesc,
+                         Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
+    bind_rows(names.dset,
+              data.frame(Covariate = c("MAP", "PULSEPRESSURE", "HYPERTENSION", "SEX"),
+                         Category = rep("Physical", 4),
+                         Name =c( "Mean artrerial pressure", "Pulse pressure", "Hypertension", "Female"),
+                         Desc = c("Mean arterail presure, 2./3.*DIASM + 1./3.*SYSTM)",
+                                  "Pulse pressure, SYSTM - DIASM",
+                                  "Hypertension, SYSTM >=140 or DIAS >= 90 or BP mediaction",
+                                  "Sex is female, True for female, False for male")))
+}

### 2019-01-25 articletwo.Rmd
Minor

@@
 
-```{r EdgeR functions, code=readLines("articletwo-ggplot.R")}
+```{r Plot functions, code=readLines("articletwo-ggplot.R")}
 source("articletwo-ggplot.R")
@@
 
-```{r pca}
+```{r pca calculate}
 if (!file.exists("rds/pca.prcomp.rds")) {
@@
 
-```{r pca}
+```{r pca plot}
 pca.eigs <- pca.prcomp$sdev^2

### 2019-01-25 articletwo.Rmd
Minor

@@
 diversities <- mergediversities(alphadiversity, adonis.species)
-
-betadiversity$min$model_1 %>% select(term, R2)
-betadiversity$max$model_1 %>% select(term, R2)
-
 ```
@@
 
-```{r Write figure 2}
+```{r Write figure 1}
 fig1head <- "alpha beta figure"

### 2019-01-25 articletwo.Rmd
Minor

@@
 
+unlink("maaslin_core", recursive = TRUE)
+
 prepare.maaslin(pseq.genus,
@@
 maaslin.force.vars.lacto <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
+
+unlink("maaslin_lacto", recursive = TRUE)
 
@@
 ```
-
-Max values for alpha diversity `r diversities$max %>% pull(estimate)
-								  %>% abs %>% max %>% round(., 2)` and
-								  `r diversities$mmin %>%
-								  pull(estimate) %>% abs %>% max %>%
-								  round(., 2)`.
 
@@
     dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  beta = Coefficient) %>%
-   dplyr::filter(Qval < 0.05)
+                  beta = Coefficient) #%>%
 
-maas.lacto.signf.features <- maas.results.df.signf %>%
-    pull(Feature) %>%
-    as.character() %>%
-    gsub("_", ".", .)
-
-df.lacto.merged <-  merge.pheno.abu(pseq.lacto)
+colnames(maas.lacto.results.df.signf)
+maas.lacto.results.df.signf %>%
+    select(Variable, Feature, P.value, Q.value) %>%
+    arrange(Q.value) %>%
+    kable
 ```

### 2019-01-25 articletwo.Rmd
Minor

@@
 
-```{r run maaslin, results = 'hide'} 
-maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
+```{r maaslin runs, results = 'hide'} 
+maaslin <- list()
 
-unlink("maaslin_core", recursive = TRUE)
+maalin$genus <- maaslinwrapper(pseq = pseq.genus,
+                looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
+                forced = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03",
+                           "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
+                taxa = "Halapricum.Archaea-")
 
-prepare.maaslin(pseq.genus,
-                 tsvfile = "maaslin_core_merged_phfinrisk_genus.tsv",
-                conffile = "maaslin_core_config.config",
-                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars),
-                core = TRUE,
-                readpclrows = "Halapricum.Archaea-")
+lactotaxa <- paste0(c("Lactobacillus_delbrueckii.Bacteria", "Lactobacillus_ruminis.Bacteria", "Lactobacillus_acidophilus.Bacteria",
+                    "Lactobacillus_fermentum.Bacteria", "Lactobacillus_paracasei.Bacteria", "Lactobacillus_salivarius.Bacteria"), collapse = ",")
 
-Maaslin("maaslin_core_merged_phfinrisk_genus.tsv",
-        "maaslin_core",
-        strInputConfig = "maaslin_core_config.config",
-        strModelSelection="none",
-        dSignificanceLevel = 1.0, 
-        fAllvAll = TRUE,
-        strForcedPredictors = maaslin.force.vars)
-```
+maalin$lacto.min <- maaslinwrapper(pseq = (subset_taxa(pseq.species, Genus=="Lactobacillus")),
+                looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
+                forced = c("BL_AGE", "SEX"),
+                taxa = lactotaxa)
 
-```{r lactobacillus scatterplot,  results = 'hide'}
-pseq.lacto  <- subset_taxa(pseq.species, Genus=="Lactobacillus")
-maaslin.force.vars.lacto <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
-
-unlink("maaslin_lacto", recursive = TRUE)
-
-prepare.maaslin(pseq.lacto,
-                 tsvfile = "maaslin_lacto_merged_phfinrisk_genus.tsv",
-                conffile = "maaslin_lacto_config.config",
-                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars.lacto),
-                readpclrows = "Lactobacillus_delbrueckii.Bacteria,Lactobacillus_ruminis.Bacteria,Lactobacillus_acidophilus.Bacteria,Lactobacillus_fermentum.Bacteria,Lactobacillus_paracasei.Bacteria,Lactobacillus_salivarius.Bacteria")
-
-Maaslin("maaslin_lacto_merged_phfinrisk_genus.tsv",
-        "maaslin_lacto",
-        strInputConfig = "maaslin_lacto_config.config",
-        strModelSelection="none",
-        dSignificanceLevel = 1.0, 
-        fAllvAll = TRUE,
-        strForcedPredictors = maaslin.force.vars.lacto)
+maalin$lacto.max  <- maaslinwrapper(pseq = (subset_taxa(pseq.species, Genus=="Lactobacillus")),
+                looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
+                forced =c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03",
+                          "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
+                taxa = lactotaxa)
 ```
@@
 ```{r maaslin}
-maas.results.df.signf <- read.table("maaslin_core/maaslin_core_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
+maas.results.df.signf <- maaslin$genus %>%
     merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
@@
 
-```{r Reading maaslin lacto output}
-maas.lacto.results.df.signf <- read.table("maaslin_lacto/maaslin_lacto_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
+Lacto min
+
+```{r Reading maaslin lacto min output}
+maas.lacto.results.df.signf <- maaslin$lacto.min %>%
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
+    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
+                  beta = Coefficient) #%>%
+
+colnames(maas.lacto.results.df.signf)
+maas.lacto.results.df.signf %>%
+    select(Variable, Feature, P.value, Q.value) %>%
+    arrange(Q.value) %>%
+    kable
+```
+
+Lacto max
+
+```{r Reading maaslin lacto max output}
+maas.lacto.results.df.signf <- maaslin$lacto.max %>%
     merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%

### 2019-01-25 articletwo-rrbiome.R

@@ -164,9 +164,31 @@ getdescriptions <- function() {
     bind_rows(names.dset,
               data.frame(Covariate = c("MAP", "PULSEPRESSURE", "HYPERTENSION", "SEX"),
                          Category = rep("Physical", 4),
-                         Name =c( "Mean artrerial pressure", "Pulse pressure", "Hypertension", "Female"),
-                         Desc = c("Mean arterail presure, 2./3.*DIASM + 1./3.*SYSTM)",
+                         Name =c( "Mean arterial pressure", "Pulse pressure", "Hypertension", "Female"),
+                         Desc = c("Mean arterial presure, 2./3.*DIASM + 1./3.*SYSTM)",
                                   "Pulse pressure, SYSTM - DIASM",
                                   "Hypertension, SYSTM >=140 or DIAS >= 90 or BP mediaction",
                                   "Sex is female, True for female, False for male")))
+}
+
+maaslinwrapper <- function(pseq, looped, forced, taxa)  {
+    tempdir <- sprintf("%s/maaslin-%s", Sys.getenv("HOME"), format(Sys.time(), '%s'))
+    dir.create(tempdir)
+    cwd <- getwd()
+    setwd(tempdir)
+    prepare.maaslin(pseq,
+                    tsvfile = "maaslin.tsv",
+                    conffile = "maaslin.config",
+                    includedvars = c(looped, forced),
+                    readpclrows = taxa)
+    Maaslin("maaslin.tsv",
+            "maaslin",
+            strInputConfig = "maaslin.config",
+            strModelSelection="none",
+            dSignificanceLevel = 1.0,
+            fAllvAll = TRUE,
+            strForcedPredictors = forced)
+    ret <- read.table("maaslin/maaslin.txt", header=TRUE, sep='\t')
+    setwd(cwd)
+    ret
 }

### 2019-01-25 articletwo.Rmd
Minor

@@
                                            covariates = c("BL_AGE", "SEX"),
-                                           npermutations = 99)
+                                           npermutations = 1000)
      adonis.species$max <- calculateadonis(dset = meta(pseq.species),
@@
                                                           "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                                           npermutations = 99)
+                                           npermutations = 1000)
      saveRDS(adonis.species, file = "rds/adonis.species.rds")

### 2019-01-25 articletwo.Rmd
Minor

@@
 
-maalin$genus <- maaslinwrapper(pseq = pseq.genus,
+maaslin$genus <- maaslinwrapper(pseq = pseq.genus,
                 looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
@@
 
-maalin$lacto.min <- maaslinwrapper(pseq = (subset_taxa(pseq.species, Genus=="Lactobacillus")),
+maaslin$lacto.min <- maaslinwrapper(pseq = (subset_taxa(pseq.species, Genus=="Lactobacillus")),
                 looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
@@
 
-maalin$lacto.max  <- maaslinwrapper(pseq = (subset_taxa(pseq.species, Genus=="Lactobacillus")),
+maaslin
+
+maaslin$lacto.max  <- maaslinwrapper(pseq = (subset_taxa(pseq.species, Genus=="Lactobacillus")),
                 looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),

### 2019-01-25 articletwo-rrbiome.R

@@ -171,8 +171,8 @@ getdescriptions <- function() {
                                   "Sex is female, True for female, False for male")))
 }
 
-maaslinwrapper <- function(pseq, looped, forced, taxa)  {
-    tempdir <- sprintf("%s/maaslin-%s", Sys.getenv("HOME"), format(Sys.time(), '%s'))
+maaslinwrapper <- function(pseq, looped, forced, taxa, tempstr = "%s/temp/maaslin-%s")  {
+    tempdir <- sprintf(tempstr, Sys.getenv("HOME"), format(Sys.time(), '%s'))
     dir.create(tempdir)
     cwd <- getwd()
     setwd(tempdir)

### 2019-01-25 articletwo.Rmd
Minor

@@
 ```{r Running at}
-if (!exists('now')) {
-    now <- format(Sys.time(), '%Y%m%d-%H%M%S')
-}
-now
-```
-
-Session info
-
-```{r Officer importer, code = sessionInfo()}
+now <- ifelse(exists('now'),now, format(Sys.time(), '%Y%m%d-%H%M%S'))
 ```
@@
 library(png)
+library(pander)
+```
+
+Session info
+
+```{r Session info}
+pander(sessionInfo(), compact = TRUE)
 ```
@@
                 taxa = lactotaxa)
-
-maaslin
 
@@
 
-colnames(maas.lacto.results.df.signf)
 maas.lacto.results.df.signf %>%

### 2019-02-05 articletwo.Rmd
Minor

@@
 
-# Running at
+# Command line arguments
 
@@
 ```{r Running at}
-now <- ifelse(exists('now'),now, format(Sys.time(), '%Y%m%d-%H%M%S'))
+args <- R.utils::commandArgs(trailingOnly = TRUE, asValues = TRUE, unique = TRUE)
+now <- ifelse("time" %in% names(args),
+              args$time,
+              format(Sys.time(), '%Y%m%d-%H%M%S'))
+
+if("clean" %in% names(args)) {
+    unlink("rds", recursive=TRUE)
+    unlink("cache", recursive=TRUE)
+}
 ```
@@
 pander(sessionInfo(), compact = TRUE)
-```
-
-```{r Remove temporary data, include = FALSE, cache = FALSE, eval = FALSE}
-unlink("rds", recursive=TRUE)
-unlink("cache", recursive=TRUE)
 ```

### 2019-02-05 articletwo-officer.R

@@ -1,7 +1,7 @@
 characteristics <- function(dset, names, factors) {
     title <- "Characteristics"
     overall <- paste0("Cases, n=", dim(dset)[1])
-    tableobject <- tableone::CreateTableOne(data = dset, vars = names, factorVars = factors)
+    tableobject <- tableone::CreateTableOne(data = dset, vars = names(names), factorVars = factors)
     tablecsv <- print(tableobject,
                       exact = "stage",
                       quote = FALSE,
@@ -42,7 +42,7 @@ tableone <- function(dset) {
     tableone.factors <- c("SEX", "HYPERTENSION",  "ANYDRUG", "CURR_SMOKE", "DIAB",
                           "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
 
-    data <- characteristics(dset, names(tableone.names), tableone.factors)
+    data <- characteristics(dset, tableone.names, tableone.factors)
 
     flextable(data = data) %>%
         set_header_labels(rowname = "Characteristics",

### 2019-02-05 articletwo.Rmd
Minor

--- 2019-01-25 articletwo.Rmd
@@
 
-```{r Running at}
-now <- ifelse(exists('now'),now, format(Sys.time(), '%Y%m%d-%H%M%S'))
+```{r Command line arguments}
+now <- format(Sys.time(), '%Y%m%d-%H%M%S')
+
+if (exists(args)) {
+    if ("time" %in% names(args) now <- args$time,
+
+    if("clean" %in% names(args)) {
+        unlink("rds", recursive=TRUE)
+        unlink("cache", recursive=TRUE)
+    }
+}
 ```

### 2019-02-05 articletwo-officer.R
Minor

@@ -1,7 +1,7 @@
-characteristics <- function(dset, names, factors) {
+characteristics <- function(dset, tableone.names, tableone.factors, extras = list()) {
     title <- "Characteristics"
     overall <- paste0("Cases, n=", dim(dset)[1])
-    tableobject <- tableone::CreateTableOne(data = dset, vars = names(names), factorVars = factors)
+    tableobject <- tableone::CreateTableOne(data = dset, vars = names(tableone.names), factorVars = tableone.factors)
     tablecsv <- print(tableobject,
                       exact = "stage",
                       quote = FALSE,
@@ -10,19 +10,21 @@ characteristics <- function(dset, names, factors) {
                       digits = 1,
                       pDigits = 3,
                       contDigits=1)
+    tableone.fullnames <- c(tableone.names, extras)
     tablecsv %>%
         as.data.frame %>%
         tibble::rownames_to_column(var = "rowname") %>%
         dplyr::filter(row_number() > 1) %>%
         format(justify = "left", trim = TRUE) %>%
         rowwise() %>%
-        mutate(id = gsub("([A-Za-z_0-9]+).*", "\\1", rowname)) %>%
-        mutate(present = id %in%  names(names)) %>%
-        mutate(rowname = ifelse(present == TRUE, names[[id]], rowname)) %>%
+        mutate(id = gsub("^ *([A-Za-z_0-9]+).*", "\\1", rowname)) %>%
+        mutate(present = id %in%  names(tableone.fullnames)) %>%
+        mutate(rowname = ifelse(present == TRUE, tableone.fullnames[[id]], rowname)) %>%
         select(rowname, Overall)
 }
 
 tableone <- function(dset) {
+
     tableone.names <- list("BL_AGE" = "Age, y (SD)",
                            "SEX" = "Female, N (%)",
                            "BMI" = "BMI, kg/m² (SD)",
@@ -39,10 +41,12 @@ tableone <- function(dset) {
                            "BL_USE_RX_C07" = "  Beta blockers, N (%)",
                            "BL_USE_RX_C08" = "  Calcium channel blockers, N (%)",
                            "BL_USE_RX_C09" = "  Agents acting on the RAS, N (%)")
+    extras <- list("1"  =  "  Light",
+                           "2"  =  "  Moderate",
+                           "3"  =  "  Heavy")
[…]

### 2019-02-05 articletwo.Rmd
Minor
Figure 4

@@
 
-if (exists(args)) {
-    if ("time" %in% names(args) now <- args$time,
+if (exists("args")) {
+    if ("time" %in% names(args)) now <- args$time
 
@@
 maaslinplot <- plot_maaslin_heatmap_final(maas.results.df.signf %>% dplyr::filter(Variable != "HYPERTENSION"),
-                      xlab.text="Genera",
                       beta.maxabs=max(abs(maas.results.df.signf[,'beta'])),
@@
 ## Figure 4
+
+```{r PC plot}
+pca.prcomp.df <- pca.prcomp$x[,1:5] %>% data.frame
+#pca.prcomp$sdev[1:20]^2/sum(pca.prcomp$sdev^2)
+#cumsum(pca.prcomp$sdev[1:20]^2/sum(pca.prcomp$sdev^2))
+
+X <- base::merge(meta(pseq.species), pca.prcomp.df, by=0, all=TRUE) %>%
+    dplyr::rename(Sample_ID = Row.names)
+
+X.res <- lapply(vectortolist(paste0("PC", seq(5))), function(axis) {
+    modelstring <- paste0("%s ~  BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + ", axis)
+    calculateglm(X, modelstring = modelstring, filterstr = "PC")
+})
+
+Z.df <-  do.call(rbind.data.frame, X.res) %>%
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate") %>%
+    dplyr::mutate(Qval = p.adjust(p.value, method="BH"),
+                  beta = estimate,
+                  Feature = term)
+
+p2 <- plot_maaslin_heatmap_final(Z.df , fname="Feature", pname="Name",
+                                  xlab.text = "PC", signif="Qval", cluster=FALSE,
+                                  v.title.name = "Effect size (PC)",
+                                  ystr="reorder(Name, Qval)", sign.name="Significance.PC",
+                                  no.sign.legend=TRUE)
+
+ggsave(file = "cache/pca.p2.png", plot = p2, height = 8, width = 8, dpi = 300)
+
+```
+
+## Figure 5
 
@@
 
-```{r Write figure 4}
+```{r Write figure 5}
 fig4head <- "Bacteria causing dysentery and diarrhea contribute to third principal component"
@@
 Plasmid DNA is denoted using star.", pca.percentage[3])
-writeimage(doc, 4, "cache/pca.pc3.png", fig4head, fig4foot)
+writeimage(doc, 5, "cache/pca.pc3.png", fig4head, fig4foot)
 ```

### 2019-02-05 articletwo-rrbiome.R

@@ -54,8 +54,7 @@ calculateglm <- function(dset,
         stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
             broom::tidy() %>%
                 dplyr::filter(grepl(filterstr, term)) %>%
-                dplyr::mutate(response = response, fo = fo) %>%
-                dplyr::select(-term)
+                dplyr::mutate(response = response, fo = fo)
     })
     data.table::rbindlist(glmlist, id = "model_name") %>%
         as.data.frame %>%

### 2019-02-11 articletwo.Rmd
Minor

@@
 ggsave(file = "cache/pca.p2.png", plot = p2, height = 8, width = 8, dpi = 300) 
+```
 
+```{r Write figure 5}
+fig4head <- "P2"
+fig4foot <- ""
+writeimage(doc, 4, "cache/pca.p2.png", fig4head, fig4foot)
 ```
@@
 ```{r Write figure 5}
-fig4head <- "Bacteria causing dysentery and diarrhea contribute to third principal component"
-fig4foot <- sprintf("Top 20 bacterial contributors to third principal coordinate at species level.
+fig5head <- "Bacteria causing dysentery and diarrhea contribute to third principal component"
+fig5foot <- sprintf("Top 20 bacterial contributors to third principal coordinate at species level.
 Centered log ratio transform is performed for the compositional data.
@@
 Plasmid DNA is denoted using star.", pca.percentage[3])
-writeimage(doc, 5, "cache/pca.pc3.png", fig4head, fig4foot)
+writeimage(doc, 5, "cache/pca.pc3.png", fig5head, fig5foot)
 ```

### 2019-02-26 articletwo.Rmd
Minor

@@
 ```{r matrix calculation}
-if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds")) {
+if (!file.exists("rds/bray.dist.m.species.rds")) {
     bray.dist.m.species <- calculate.beta.matrix(pseq.species)
@@
 
-```{r Write figure 5}
+```{r Write figure 4}
 fig4head <- "P2"
@@
 ```
+
+```{r save session}
+save.session(paste0("cache/session-", now, ".Rda"))
+```

### 2019-02-27 articletwo.Rmd
Minor
Core

@@
 
+```{r core taxa}
+pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
+pseq.species.lactobacillus <- subset_taxa(pseq.species, Genus=="Lactobacillus")
+pseq.species.lactobacillus.core <- core(pseq.species.lactobacillus, detection = 0.001, prevalence = 0.01)
+```
+
+At genus level core has number of taxa `r ntaxa(pseq.genus.core)` (`r
+ntaxa(pseq.genus)`). For lactobacilli core has number of taxa `r
+ntaxa(pseq.species.lactobacillus.core)` (`r ntaxa(pseq.species.lactobacillus)`).
+
 ```{r maaslin runs, results = 'hide'} 
@@
 
-maaslin$genus <- maaslinwrapper(pseq = pseq.genus,
+maaslin$genus <- maaslinwrapper(pseq = pseq.genus.core,
                 looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
@@
 
-lactotaxa <- paste0(c("Lactobacillus_delbrueckii.Bacteria", "Lactobacillus_ruminis.Bacteria", "Lactobacillus_acidophilus.Bacteria",
-                    "Lactobacillus_fermentum.Bacteria", "Lactobacillus_paracasei.Bacteria", "Lactobacillus_salivarius.Bacteria"), collapse = ",")
+lactotaxa <-c("Lactobacillus_delbrueckii.Bacteria", "Lactobacillus_ruminis.Bacteria", "Lactobacillus_acidophilus.Bacteria",
+                    "Lactobacillus_fermentum.Bacteria", "Lactobacillus_paracasei.Bacteria", "Lactobacillus_salivarius.Bacteria")
 
-maaslin$lacto.min <- maaslinwrapper(pseq = (subset_taxa(pseq.species, Genus=="Lactobacillus")),
+maaslin$lacto.min <- maaslinwrapper(pseq = pseq.species.lactobacillus.core,
                 looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
                 forced = c("BL_AGE", "SEX"),
-                taxa = lactotaxa)
+                taxa =  paste0(lactotaxa,  collapse = ","))
 
-maaslin$lacto.max  <- maaslinwrapper(pseq = (subset_taxa(pseq.species, Genus=="Lactobacillus")),
+maaslin$lacto.max  <- maaslinwrapper(pseq = pseq.species.lactobacillus.core,
                 looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
@@
                           "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                taxa = lactotaxa)
+                taxa = paste0(lactotaxa collapse = ","))
 ```
@@
 ```{r save session}
-save.session(paste0("cache/session-", now, ".Rda"))
+save.image(file = paste0("cache/session-", now, ".Rdata"))
 ```

### 2019-02-27 articletwo.Rmd
Minor

@@
                           "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                taxa = paste0(lactotaxa collapse = ","))
+                taxa = paste0(lactotaxa, collapse = ","))
 ```
@@
 ```{r save session}
-save.image(file = paste0("cache/session-", now, ".Rdata"))
+save.image(file = paste0("rds/session-", now, ".Rdata"))
 ```

### 2019-03-22 articletwo.Rmd
Minor

@@
 
-maas.lacto.results.df.signf %>%
-    select(Variable, Feature, P.value, Q.value) %>%
-    arrange(Q.value) %>%
-    kable
+```
+
+```{r lactobacillus table}
+allruns.unamb <- turkumetabolites::unambiguous.lmrank(dset$fr02, linear.lmrank)
+allruns.sup <- turkumetabolites::pub.lmrank(allruns.unamb,  p = 0.05/330)
+
+maaslin.lacto.max.table <- maaslin$lacto.max %>%
+    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
+                  beta_ci =sprintf("%.1E", Coefficient, NA),
+                  p.value = sprintf("%.2f", p.adjust(P.value, method="BH"))) %>%
+    merge(., names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
+    select(Feature, Name, beta_ci, p.value) %>%
+    gather(variable, value, -Feature, -Name) %>%
+    unite(temp, Name, variable) %>%
+    spread(temp, value) %>%
+    mutate(Feature = rename.genus(gsub("_Bacteria", "", Feature))) %>%
+    arrange(Feature)
+
+names(maaslin.lacto.max.table) <- gsub(" ", "_", names(maaslin.lacto.max.table ))
+
+typology.tbl2 <- data.frame(col_keys = c("Feature",
+                                         "Mean_arterial_pressure_beta_ci",
+                                         "Mean_arterial_pressure_p.value",
+                                         "Systolic_BP_beta_ci",
+                                         "Systolic_BP_p.value",
+                                         "Diastolic_BP_beta_ci",
+                                         "Diastolic_BP_p.value",
+                                         "Pulse_pressure_beta_ci",
+                                         "Pulse_pressure_p.value",
+                                         "Hypertension_beta_ci",
+                                         "Hypertension_p.value"),
+                            what = c( "",
+                                     rep("MAP", 2), 
+                                     rep("Systolic BP", 2), 
+                                     rep("Diastolic BP", 2), 
+                                     rep("Pulse pressure", 2), 
+                                     rep("Hypertension", 2)),
+                            measure = c("Lactobacillus", rep(c("β±SE", "p"), 5)),
+                            stringsAsFactors = FALSE)
+
+tbl2 <- flextable(data = maaslin.lacto.max.table) %>%
+    set_header_df(mapping = typology.tbl2, key = "col_keys") %>%
+        flextable::theme_booktabs()
+writetable(doc, tbl2, number = 2, tbl2head, tbl2foot)
+print(doc, target = paste0("report/articletwo-test.docx"))
+```
+
+```{r write table 2}
+tbl2head <- "Lactobacilli"
+tbl2foot <- ""
+writetable(doc, tbl2, number = 2, tbl2head, tbl2foot)
 ```
@@
 ```{r save session}
-save.image(file = paste0("rds/session-", now, ".Rdata"))
+save.image(file = paste0("session/session-", now, ".Rdata"))
 ```

### 2019-03-22 articletwo.Rmd
Lacto coefficient to show

@@
 dir.create("rds/", showWarnings = FALSE)
+dir.create("session/", showWarnings = FALSE)
 ```
@@
                 taxa = paste0(lactotaxa, collapse = ","))
+
 ```
@@
 ```{r Reading maaslin lacto min output}
-maas.lacto.results.df.signf <- maaslin$lacto.min %>%
+maaslin$lacto.min %>%
     merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
     dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  beta = Coefficient) #%>%
-
-maas.lacto.results.df.signf %>%
-    select(Variable, Feature, P.value, Q.value) %>%
+                  beta = Coefficient) %>% 
+    select(Variable, Feature, Coefficient, P.value, Q.value) %>%
     arrange(Q.value) %>%
+    mutate_if(is.numeric, funs(ifelse(. < 0.01, sprintf("%.1e", .), round(., 2)))) %>%
     kable
@@
 ```{r Reading maaslin lacto max output}
-maas.lacto.results.df.signf <- maaslin$lacto.max %>%
+maaslin$lacto.max %>%
     merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
     dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  beta = Coefficient) #%>%
-
+                  beta = Coefficient) %>%
+    select(Variable, Feature, Coefficient, P.value, Q.value) %>%
+    arrange(Q.value) %>%
+    mutate_if(is.numeric, funs(ifelse(. < 0.01, sprintf("%.1e", .), round(., 2)))) %>%
+    kable
 ```

### 2019-04-24 articletwo.Rmd
Minor

@@
     }
+    if ("tags" %in% names(args)) 
+        rtags('~/phd/research/articletwo/', recursive = FALSE, pattern = '\\.[RrSs](rw|md)?$',ofile = '~/phd/research/articletwo/TAGS', verbose = TRUE, append = FALSE)
 }
@@
 
+}
+```
 # Importing libraries:
@@
 
+```{r test scaling}
+meta(pseq.genus) %>% summarize(mean(MAP), sd(MAP), mean(SYSTM), sd(SYSTM), mean(DIASM), sd(DIASM), mean(PULSEPRESSURE), sd(PULSEPRESSURE)) %>% kable
+```
+
 Sample data has dimensions (`r dim(sample_data(pseq.species))`).
@@
 tbl1foot <- "Continuous variables are presented as mean (standard deviation).
-Exercise at free-timeis reported as not much (1), walk etc.
-(2), regular three hours or more or competitive sports (3).
 Categorical variables reported as absolute and relative frequencies.
-BP indicates blood pressure, BMI body mass index, RAS renin-angiotensin system."
+BP indicates blood pressure, BMI, body mass index, RAS, renin-angiotensin system."
 writetable(doc, tbl1, number = 1, tbl1head, tbl1foot)
@@
 
+Minimum coefficient for alpha diversity is `r rbind(diversities$min,
+diversities$max) %>% pull(estimate) %>% min`.
+
 ```{r save grob}
@@
 alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 1.3)
+
 beta.diversity.min <- plot.beta(diversities$min)
@@
                       c(NA, 1,    2,    NA, 3, NA, 4),
-                      c(NA, NA, 5,    NA,6, NA, 4),
-                      c(NA, NA, NA, NA,7, NA, 4),
-                      c(15,    NA,    NA, NA,   NA, NA, 4),
-                      c(NA, 9,    10,  NA,  11, NA, 4),
+                      c(NA, NA, 5,    NA,6, NA, NA),
+                      c(NA, NA, NA, NA,7, NA, NA),
+                      c(15,    NA,    NA, NA,   NA, NA, NA),
+                      c(NA, 9,    10,  NA,  11, NA, NA),
                       c(NA, 9,    10,  NA,  11, NA, NA),
@@
                       c(NA, NA, NA, NA,14, NA, NA)),
-                  widths=c(0.7, 4, 0.7, 0.1, 5, 0.7, 4),
+                  widths=c(0.7, 4, 0.7, 0.1, 5, 0.7, 3),
                   heights=rep(c(1, 1, 6, 1, 2), 2))
+
 ggsave(file = "cache/alpha-beta.png", plot=g4, height=6, width=9)
@@
 ```{r Write figure 1}
-fig1head <- "alpha beta figure"
-fig1foot <- ""
+fig1head <- "Associations between BP variables and microbial diversity"
+fig1foot <- "Results in panel A are calculated using minimal model using only age and sex for covariates.
+Panel B has the full model where covariates are age, sex, BMI, smoker, exercise, and four antihypertensive medications.
+The blue tinted box on the left represents four linear and one logistic model where blood pressure variables are dependant variables and Shannon's index is included in covariates.
+The gray tinted bars represent beta diversity i.e.
+analysis of variance for Bray-Curtis distance matrix where blood pressure variables are included in covariates.
+Result marked with asterisk are significant at FDR 0.05.
+Permutational analysis for beta diverity has 1000 permutations (p=0.01)."
 writeimage(doc, 1, "cache/alpha-beta.png", fig1head, fig1foot)

### 2019-04-24 articletwo-ggplot.R

@@ -8,7 +8,7 @@ plot.alpha <- function(diversity,  colorlimit = 0.25) {
                              limits = c(-colorlimit, 0),
                              na.value = "black") +
         theme_classic(20) +
-        geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=TRUE, color='black', size=20) +
+        geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=FALSE, color='black', size=20) +
         scale_shape_manual(name="",
                            values=c('*'='*', ' '=' '),
                            labels=c("*"="significant at\nFDR 0.05", ' '=' '),

### 2019-04-24 articletwo-officer.R

@@ -35,12 +35,12 @@ tableone <- function(dset) {
                            "HYPERTENSION" = "Hypertension, N (%)",
                            "CURR_SMOKE" = "Current smoker, N (%)",
                            "DIAB" = "Diabetes mellitus, N (%)",
-                           "Q57X" = "Exercice, N (%)",
+                           "Q57X" = "Exercise, N (%)",
                            "ANYDRUG" = "Antihypertensive medication, N (%)",
                            "BL_USE_RX_C03" = "  Diuretics, N (%)",
                            "BL_USE_RX_C07" = "  Beta blockers, N (%)",
                            "BL_USE_RX_C08" = "  Calcium channel blockers, N (%)",
-                           "BL_USE_RX_C09" = "  Agents acting on the RAS, N (%)")
+                           "BL_USE_RX_C09" = "  RAS blockers, N (%)")
     extras <- list("1"  =  "  Light",
                            "2"  =  "  Moderate",
                            "3"  =  "  Heavy")

### 2019-04-24 articletwo-rrbiome.R

@@ -8,13 +8,18 @@ rename.genus  <- function (genus, markword = "Plasmid", mark = "*") {
     paste0(gsub("_", " ", name), star)
 }
 
+myscale <- function(x){
+  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
+}
+
 filter.phenotype.data <- function(pseq,
                                   included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
                                                "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
-                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
+                                               "BL_USE_RX_C07", "BL_USE_RX_C08")) {
     meta(pseq) %>%
         tibble::rownames_to_column(var = "rowname") %>%
         dplyr::select(rowname, included) %>%
+        stats::na.omit() %>%
         dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
                                        BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
                       ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
@@ -22,10 +27,18 @@ filter.phenotype.data <- function(pseq,
                       HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
                       SEX = factor(ifelse(MEN == "Female", 1, 0)),
                       MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
-         stats::na.omit() %>%
+        dplyr::mutate(oSYSTM = SYSTM,
+                      oDIASM = DIASM,
+                      oPULSEPRESSURE = PULSEPRESSURE,
+                      oMAP = MAP,
+                      SYSTM = myscale(SYSTM),
+                      DIASM = myscale(DIASM),
+                      PULSEPRESSURE = myscale(PULSEPRESSURE),
+                      MAP = myscale(MAP)) %>%
         dplyr::select(-MEN, -DNA_OK) %>%
         tibble::remove_rownames() %>%
         tibble::column_to_rownames(var = "rowname")
+}
 
 meta.merge.alphadiversity <- function(pseq, index = "shannon") {
     alphadiversity  <- microbiome::global(pseq, index = index)

### 2019-04-25 articletwo.Rmd
Minor

@@
     if ("tags" %in% names(args)) 
-        rtags('~/phd/research/articletwo/', recursive = FALSE, pattern = '\\.[RrSs](rw|md)?$',ofile = '~/phd/research/articletwo/TAGS', verbose = TRUE, append = FALSE)
+        rtags('~/phd/research/articletwo/', recursive = FALSE, pattern = '\\.[RrSs](rw|md)?$',
+              ofile = '~/phd/research/articletwo/TAGS', verbose = TRUE, append = FALSE)
 }
@@
 
-```{r test scaling}
-meta(pseq.genus) %>% summarize(mean(MAP), sd(MAP), mean(SYSTM), sd(SYSTM), mean(DIASM), sd(DIASM), mean(PULSEPRESSURE), sd(PULSEPRESSURE)) %>% kable
-```
-
 Sample data has dimensions (`r dim(sample_data(pseq.species))`).
@@
 
+```{r maalin manual, eval = FALSE}
+test.abud <- microbiome::abundances(microbiome::transform(pseq.genus.core, "compositional")) %>%
+    t %>%
+    as.data.frame
+df <- test.abud %>% mutate_all(funs(asin(sqrt(.))))
+```
+
 ```{r maaslin runs, results = 'hide'} 
 maaslin <- list()
+lm (Sutterella_Bacteria ~ MAP + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + 
+                           BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09, )
 
@@
 ```{r Write figure 2}
-fig2head <- "Tenth of bacterial variation can be explained by two first principal axes"
+fig2head <- "Principal component analysis for bacterial abundances"
 fig2foot <- sprintf("Principal component analysis for bacterial abundances at species level.
@@
 First two axes explain %.1f%% of the variation.
-Research participants with hypertension are denoted by red dot.", sum(pca.percentage[1:2]))
+Centered log-ratio (CLR) tranformation is used for microbiome abundances.", sum(pca.percentage[1:2]))
 writeimage(doc, 2, "cache/pca.png", fig2head, fig2foot)
@@
     dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
+                  Feature = gsub("_Bacteria", "", Feature),
                   beta = Coefficient) %>%
-   dplyr::filter(Qval < 0.05)
+    dplyr::filter(Qval < 0.05) %>%
+    arrange(Feature)
 
-maaslinplot <- plot_maaslin_heatmap_final(maas.results.df.signf %>% dplyr::filter(Variable != "HYPERTENSION"),
+maaslinplot <- plot_maaslin_heatmap_final(maas.results.df.signf %>% dplyr::filter(Variable!="HYPERTENSION"),
                       beta.maxabs=max(abs(maas.results.df.signf[,'beta'])),
@@
 writeimage(doc, 3, "cache/maaslin.png", fig3head, fig3foot)
-```
-
-## Figure 4
-
-```{r PC plot}
-pca.prcomp.df <- pca.prcomp$x[,1:5] %>% data.frame
-#pca.prcomp$sdev[1:20]^2/sum(pca.prcomp$sdev^2)
-#cumsum(pca.prcomp$sdev[1:20]^2/sum(pca.prcomp$sdev^2))
-
-X <- base::merge(meta(pseq.species), pca.prcomp.df, by=0, all=TRUE) %>%
-    dplyr::rename(Sample_ID = Row.names)
-
-X.res <- lapply(vectortolist(paste0("PC", seq(5))), function(axis) {
-    modelstring <- paste0("%s ~  BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + ", axis)
-    calculateglm(X, modelstring = modelstring, filterstr = "PC")
-})
-
-Z.df <-  do.call(rbind.data.frame, X.res) %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate") %>%
-    dplyr::mutate(Qval = p.adjust(p.value, method="BH"),
-                  beta = estimate,
-                  Feature = term)
-
-p2 <- plot_maaslin_heatmap_final(Z.df , fname="Feature", pname="Name",
-                                  xlab.text = "PC", signif="Qval", cluster=FALSE,
-                                  v.title.name = "Effect size (PC)",
-                                  ystr="reorder(Name, Qval)", sign.name="Significance.PC",
-                                  no.sign.legend=TRUE)
-
-ggsave(file = "cache/pca.p2.png", plot = p2, height = 8, width = 8, dpi = 300) 
-```
-
-```{r Write figure 4}
-fig4head <- "P2"
-fig4foot <- ""
-writeimage(doc, 4, "cache/pca.p2.png", fig4head, fig4foot)
-```
-
-## Figure 5
-
-```{r PCA heatmap}
-pca.axis <- pca.prcomp$x %>% data.frame
-pca.dset <- base::merge(meta(pseq.species), pca.axis[,1:5], by=0, all=TRUE) %>%
-  dplyr::rename(Sample_ID = Row.names)
-
-pca.dset %>% head
-
-pca.ret <- lapply(paste0("PC", seq(1,5)), function (pc) {
-    calculateglm(pca.dset,
-                 modelstring = paste("%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  +", pc),
-                 filterstr = "PC") %>%
-        select(response, estimate, p.value) %>%
-        mutate(model_name = pc)
-})
-
-pca.ret  %>%
-    data.table::rbindlist() %>%
-    as.data.frame %>%
-    mutate(p.value = p.adjust(p.value, method="BH")) %>%
-    arrange(p.value) %>%
-    kable
-```
-
-## Loadings for principal component axis 3 (Figure 5)
-
-```{r Main contributors of PC3}
-pca.rotation.abs <- abs(pca.prcomp$rotation)
-pca.loading <- sweep(pca.rotation.abs, 2, colSums(pca.rotation.abs), "/") %>%
-	    as.data.frame %>%
-    tibble::rownames_to_column(var = "term") %>%
-    mutate(name = rename.genus(term))
-
-pca.loading.pc3.top20 <- pca.loading %>%
-    select(name, PC3) %>%
-    arrange(-PC3) %>%
-    head(20)
-
-pca.pc3.gg <- ggplot(data=pca.loading.pc3.top20, aes(x=reorder(name, -PC3), y=100*PC3)) +
-    geom_bar(stat="identity", width=0.5, position = position_dodge(preserve = "single")) +
-    theme_classic(20) +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.3, hjust = 1)) +
-    xlab("") +
-    ylab("Contribution (%)") +
-    scale_y_continuous(limits=c(0, 2.7), oob = rescale_none, expand = c(0,0)) +
-    theme(axis.title.x=element_blank(),
-          text = element_text(size=20),
-          legend.justification=c(1,0),
-          legend.position=c(1.03, 0.75),
-          legend.title=element_text(size = 20),
-          legend.direction="vertical",
-          legend.text=element_text(size = floor(0.75*20)),
-          legend.box.margin = margin(1, 1, 1, 1, "mm"),
-          legend.box.background = element_rect(colour = "black"),
-          aspect.ratio = 0.7)
-   ggsave(file = "cache/pca.pc3.png", plot = pca.pc3.gg, height = 8, width = 8, dpi = 300) 
-```
-
-```{r Write figure 5}
-fig5head <- "Bacteria causing dysentery and diarrhea contribute to third principal component"
-fig5foot <- sprintf("Top 20 bacterial contributors to third principal coordinate at species level.
-Centered log ratio transform is performed for the compositional data.
-Third principal component explain %.1f%% of the total variation.
-Plasmid DNA is denoted using star.", pca.percentage[3])
-writeimage(doc, 5, "cache/pca.pc3.png", fig5head, fig5foot)
 ```
@@
 ```{r lactobacillus table}
-allruns.unamb <- turkumetabolites::unambiguous.lmrank(dset$fr02, linear.lmrank)
-allruns.sup <- turkumetabolites::pub.lmrank(allruns.unamb,  p = 0.05/330)
-
 maaslin.lacto.max.table <- maaslin$lacto.max %>%

### 2019-04-25 aaro/edgeR_plot_functions.R

@@ -138,7 +138,7 @@ plot_edgeR_fc_heatmap_final <- function (tables.df, xlab.text = "Phyla", logFC.m
   p
 }
 
-plot_maaslin_heatmap_final <- function (tables.df, xlab.text = "Phyla", beta.maxabs=NULL, text.size=10,
+plot_maaslin_heatmap_final <- function (tables.df, beta.maxabs=NULL, text.size=10,
                                         cluster=TRUE, signif=NULL, signif.th=0.05,
                                         fname='Feature', pname='Name',vname='beta',
                                         qvalues=NULL, colours=NULL, breaks=NULL, names.df=NULL,
@@ -229,7 +229,7 @@ plot_maaslin_heatmap_final <- function (tables.df, xlab.text = "Phyla", beta.max
   }
 
   vname2plot <- ifelse(discrete.colors, vname_discrete, vname)
-  p <- ggplot(tables.df, aes_string(x = paste0("reorder(", fname, " , as.numeric(",fname,"))"),
+  p <- ggplot(tables.df, aes_string(x = paste0("reorder(", fname, " , ",fname,")"),
                                     y = ystr,
                                     fill=vname2plot)) + geom_tile(colour="white", size=0.25)
 
@@ -263,7 +263,7 @@ plot_maaslin_heatmap_final <- function (tables.df, xlab.text = "Phyla", beta.max
   }
 
   p <- p + coord_fixed()
-  p <- p + xlab(xlab.text) + ylab("")
+  p <- p + xlab("") + ylab("")
   p <- p + theme_classic(text.size)
   p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust=0.5, size=floor(0.7*text.size)))
   p <- p + theme(axis.text.y = element_text(size=floor(0.7*text.size)))

### 2019-04-25 articletwo-officer.R

@@ -28,10 +28,10 @@ tableone <- function(dset) {
     tableone.names <- list("BL_AGE" = "Age, y (SD)",
                            "SEX" = "Female, N (%)",
                            "BMI" = "BMI, kg/m² (SD)",
-                           "SYSTM" = "Systolic BP, mmHg (SD)",
-                           "DIASM" = "Diastolic BP, mmHg (SD)",
-                           "PULSEPRESSURE" = "Pulse pressure, mmHg (SD)",
-                           "MAP" = "Mean arterial pressure, mmHg (SD)",
+                           "oSYSTM" = "Systolic BP, mmHg (SD)",
+                           "oDIASM" = "Diastolic BP, mmHg (SD)",
+                           "oPULSEPRESSURE" = "Pulse pressure, mmHg (SD)",
+                           "oMAP" = "Mean arterial pressure, mmHg (SD)",
                            "HYPERTENSION" = "Hypertension, N (%)",
                            "CURR_SMOKE" = "Current smoker, N (%)",
                            "DIAB" = "Diabetes mellitus, N (%)",

### 2019-04-25 articletwo.Rmd
Bug
Minor

@@
 maaslin <- list()
-lm (Sutterella_Bacteria ~ MAP + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + 
-                           BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09, )
 
@@
 
+```{r salt}
+lacto.meta <- meta(pseq.genus) %>%
+    tibble::rownames_to_column("id")
+
+lacto.abund <- subset_taxa(pseq.genus, Genus=="Lactobacillus") %>%
+    abundances(., transform = "clr") %>%
+    t %>%
+    as.data.frame %>%
+    tibble::rownames_to_column("id") %>%
+    setNames(tolower(replace.brackets(names(.))))
+
+lacto.full <- full_join(lacto.meta, lacto.abund, by = "id") %>%
+    dplyr::mutate(fake_salt = relevel(factor(dplyr::ntile(BL_AGE, 4)), ref='1'),
+                  lacto_groups = relevel(factor(dplyr::ntile(lactobacillus.bacteria, 4)), ref='1'))
+
+lacto.mean <- lacto.full %>%
+    group_by(fake_salt) %>%
+    summarize(estimate = t.test(lactobacillus.bacteria)$"estimate",
+              qi_low = t.test(lactobacillus.bacteria)$"conf.int"[1],
+              qi_high = t.test(lactobacillus.bacteria)$"conf.int"[2],
+              group = "mean")
+
+lacto.odds <- lacto.full %>%
+    group_by(lacto_groups) %>%
+    do(oddsmodel = glm(HYPERTENSION ~ lactobacillus.bacteria, family=binomial(link="logit"), data = .)) %>%
+    do(data.frame(
+        fake_salt = .$lacto_groups,
+        term = names(coef(.$oddsmodel)),
+        exp(cbind(coef(.$oddsmodel), confint(.$oddsmodel))))) %>%
+    dplyr::mutate(group = "odds") %>%
+    dplyr::rename(estimate = V1, qi_low = X2.5.., qi_high = X97.5..) %>%
+    filter(!grepl("Intercept", term)) %>%
+    select(-term)
+```
+
+```{r forest plot lacto, eval = FALSE}
+
+myforestplot <- function(data) {
+    ggplot(data = data,
+           aes(x = fake_salt, y = estimate, ymin = qi_low, ymax = qi_high)) +
+        facet_wrap(~group,strip.position="top",nrow=1,scales = "free_y") +
+        xlab('Group') +
+        ylab("Risk Ratio (95% Confidence Interval)")+
+        geom_pointrange()+
+        geom_hline(aes(fill=fake_salt), yintercept=1, linetype=2)+
+        geom_errorbar(aes(ymin=qi_low, ymax=qi_high),width=0.2,cex=1)+ 
+        theme(plot.title=element_text(size=16,face="bold"),
+              axis.text.x=element_text(face="bold"),
+              axis.title=element_text(size=12,face="bold"),
+              strip.background = element_blank(),
+              strip.placement = "outside") + 
+        ggplot2::theme_minimal() 
+}
+
+gsalt <- grid.arrange(myforestplot(lacto.mean),  myforestplot(lacto.odds), ncol=2)
+ggsave(file = "cache/salt-forest.png", plot=gsalt, height=6, width=9)
+```
+
 Lacto min
@@
         flextable::theme_booktabs()
-writetable(doc, tbl2, number = 2, tbl2head, tbl2foot)
-print(doc, target = paste0("report/articletwo-test.docx"))
 ```

### 2019-04-25 articletwo-rrbiome.R

@@ -204,3 +204,7 @@ maaslinwrapper <- function(pseq, looped, forced, taxa, tempstr = "%s/temp/maasli
     setwd(cwd)
     ret
 }
+
+mygrep <- function(word, list, ignorecase = TRUE) {
+    list[grepl(word, list, ignore.case = ignorecase)]
+}

### 2019-04-25 articletwo.Rmd
Forest plot getting ready

--- 2019-04-25 articletwo.Rmd
@@
 
+```{r salt}
+lacto.meta <- meta(pseq.genus) %>%
+    tibble::rownames_to_column("id")
+
+lacto.abund <- subset_taxa(pseq.genus, Genus=="Lactobacillus") %>%
+    abundances(., transform = "clr") %>%
+    t %>%
+    as.data.frame %>%
+    tibble::rownames_to_column("id") %>%
+    setNames(tolower(replace.brackets(names(.))))
+
+lacto.full <- full_join(lacto.meta, lacto.abund, by = "id") %>%
+    dplyr::mutate(fake_salt = relevel(factor(dplyr::ntile(BL_AGE, 4)), ref='1'),
+                  lacto_groups = relevel(factor(dplyr::ntile(lactobacillus.bacteria, 4)), ref='1'))
+
+lacto.mean <- lacto.full %>%
+    group_by(fake_salt) %>%
+    summarize(estimate = t.test(lactobacillus.bacteria)$"estimate",
+              conf.low = t.test(lactobacillus.bacteria)$"conf.int"[1],
+              conf.high = t.test(lactobacillus.bacteria)$"conf.int"[2]) %>%
+    dplyr::rename(term = fake_salt)
+
+lacto.odds <- glm(HYPERTENSION ~ lacto_groups, family = binomial(link=logit), data=lacto.full) %>%
+    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
+    filter(!grepl("Intercept", term)) %>%
+    select(term, estimate, conf.low, conf.high) %>%
+    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA)
+
+lacto.mean
+lacto.odds
+```
+
+```{r forest plot lacto, eval = FALSE}
+
+myforestplot <- function(data, xlab = "", ylab = "", ylimits = c(0,10), interceptone = FALSE) {
+    ggplot(data = data,
+           aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+        geom_point() +
+        { if (interceptone) geom_hline(aes(fill=fake_salt), yintercept=1, linetype=2) } +
+        geom_errorbar(width=0.1,cex=1)+ 
+        ggplot2::theme_minimal()  +
+        theme(plot.title=element_text(size=16),
+              axis.title=element_text(size=12),
+              strip.background = element_blank()) +
+        scale_y_continuous(limits = ylimits) +
+        theme(panel.border = element_blank(),
+              panel.grid.major = element_blank(),
+              panel.grid.minor = element_blank(),
+              axis.line = element_line(colour = "black")) +
+        scale_x_discrete(labels=c("Q1", "Q2", "Q3", "Q4")) +
+        xlab(xlab) +
+        ylab(ylab)
+}
+
+gsalt <- grid.arrange(
+    myforestplot(lacto.mean,
+                 xlab = "Dieritary salt",
+                 ylab = "Lactobacillus abundance",
+                 ylimits = c(2,3)),
+    myforestplot(lacto.odds,
+                 xlab = "Lactobacillus abundance",
+                 ylab = "Odds of hypertension",
+                 ylimits = c(0.5, 1.5),
+                 interceptone = TRUE),
+    ncol=2)
+ggsave(file = "cache/salt-forest.png", plot=gsalt, height=6, width=9)
+
+```
+
 Lacto min

### 2019-04-29 articletwo.Rmd
Salt model

--- 2019-04-25 articletwo.Rmd
@@
 library(phyloseq)
+library(nortest)
 library(microbiome)
@@
 library(pander)
+library(ggpubr)
+library(broom)
+library(ggfortify)
 ```
@@
 
+```{r lacto data frame}
+lacto.abund <- subset_taxa(pseq.genus, Genus == "Lactobacillus") %>%
+    abundances(., transform = "clr") %>%
+    t %>%
+    as.data.frame %>%
+    tibble::rownames_to_column("id") %>%
+    setNames(tolower(replace.brackets(names(.))))
+
+lacto.full <- meta(pseq.genus) %>%
+    tibble::rownames_to_column("id") %>%
+    full_join(., lacto.abund, by = "id") %>%
+    dplyr::mutate(fake_salt = relevel(factor(dplyr::ntile(BL_AGE, 4)), ref='1'),
+                  lacto_groups = relevel(factor(dplyr::ntile(lactobacillus.bacteria, 4)), ref='1'))
+```
+
+```{r lacto adjusted means}
+lacto.lm.salt <- lm(lactobacillus.bacteria ~ fake_salt + BL_AGE + SEX, data=lacto.full)
+
+lacto.mean.unadjusted <- data.frame(
+    fake_salt = as.factor(seq(4)),
+    SEX = as.factor(0),
+    BL_AGE = mean(lacto.full %>% pull(BL_AGE)))
+
+lacto.mean.adjusted <- predict(lacto.lm.salt, lacto.mean.unadjusted, interval = "confidence") %>%
+    as.data.frame %>%
+    dplyr::rename(estimate = fit,
+           conf.low = lwr,
+           conf.high = upr) %>%
+    tibble::rownames_to_column("term")
+```
+
+```{r lacto odds for hypertension}
+lacto.odds <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
+                  family = binomial(link=logit),
+                  data=lacto.full) %>%
+    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
+    filter(!grepl("Intercept", term)) %>%
+    select(term, estimate, conf.low, conf.high) %>%
+    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
+    dplyr::filter(grepl("lacto", term))
+```
+
+```{r lacto lm residuals}
+ggplot2::autoplot(lm(lactobacillus.bacteria ~ fake_salt + BL_AGE + SEX, data=lacto.full))
+```
+
+```{r lacto logistic residuals}
+ggplot2::autoplot(glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX, family = binomial(link=logit), data=lacto.full))
+```
+
+```{r lacto anova}
+lacto.anova <- aov(lactobacillus.bacteria ~ fake_salt, data = lacto.full)
+summary(lacto.anova)
+TukeyHSD(lacto.anova)$fake_salt %>% kable
+```
+
+```{r lactoto test normality}
+grid.arrange(ggqqplot(lacto.full$lactobacillus.bacteria, title = "Lactobacillus"),
+             ggqqplot(lacto.full$BL_AGE, title = "BL_AGE"), ncol=2)
+```
+
+```{r forest plot lacto}
+myforestplot <- function(data, xlab = "", ylab = "", ylimits = c(0,10), interceptone = FALSE, subtitle = "") {
+    ggplot(data = data,
+           aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+        geom_point() +
+        { if (interceptone) geom_hline(aes(fill=fake_salt), yintercept=1, linetype=2) } +
+        geom_errorbar(width=0.1)+ 
+        ggplot2::theme_minimal()  +
+        theme(plot.title=element_text(size=16),
+              axis.title=element_text(size=12),
+              strip.background = element_blank()) +
+        theme(panel.border = element_blank(),
+              panel.grid.major = element_blank(),
+              panel.grid.minor = element_blank(),
+              axis.ticks.x = element_line(),
+              axis.ticks.y = element_line(),
+              axis.line = element_line(colour = "black")) +
+        scale_x_discrete(labels=c("Q1", "Q2", "Q3", "Q4")) +
+        scale_y_continuous(limits = ylimits, breaks = pretty(ylimits, 4)) + 
+        xlab(xlab) +
+        ylab(ylab) +
+        labs(subtitle = subtitle)
+}
+
+gsalt <- grid.arrange(
+    myforestplot(lacto.mean.adjusted,
+                 xlab = "Dietary salt",
+                 ylab = "Lactobacillus abundance",
+                 subtitle = "A.",
+                 ylimits = c(2.5,3)),
+    myforestplot(lacto.odds,
+                 xlab = "Lactobacillus abundance",
+                 ylab = "Odds of hypertension",
+                 subtitle = "B.",
+                 ylimits = c(0.5, 1.5),
+                 interceptone = TRUE),
+    ncol=2)
+ggsave(file = "cache/salt-forest.png", plot=gsalt, height=4, width=6, dpi = 300)
+```
+
+```{r Write figure 3}
+fig3head <- "Associations between dietary salt, lactobacillus abundance, and odds for hypertensoin"
+fig3foot <- "In panel A we see lactobacillus abundances in different dietary salt groups.
+The mean value is adjusted using age and sex.
+In panel B we split paticipants in four groups by lactobacillus abundances and calulate odds ratio for hypertension.
+Age and sex are included in logistic model.
+Both panel show  95 % CI for respective variables."
+writeimage(doc, 3, "cache/salt-forest.png", fig3head, fig3foot)
+```
+
 Lacto min

### 2019-04-29 articletwo.Rmd
Species level lactobacillus methods removed

@@
 
-Lacto min
-
-```{r Reading maaslin lacto min output}
-maaslin$lacto.min %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
-    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  beta = Coefficient) %>% 
-    select(Variable, Feature, Coefficient, P.value, Q.value) %>%
-    arrange(Q.value) %>%
-    mutate_if(is.numeric, funs(ifelse(. < 0.01, sprintf("%.1e", .), round(., 2)))) %>%
-    kable
-```
-
-Lacto max
-
-```{r Reading maaslin lacto max output}
-maaslin$lacto.max %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
-    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  beta = Coefficient) %>%
-    select(Variable, Feature, Coefficient, P.value, Q.value) %>%
-    arrange(Q.value) %>%
-    mutate_if(is.numeric, funs(ifelse(. < 0.01, sprintf("%.1e", .), round(., 2)))) %>%
-    kable
-```
-
-```{r lactobacillus table}
-maaslin.lacto.max.table <- maaslin$lacto.max %>%
-    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  beta_ci =sprintf("%.1E", Coefficient, NA),
-                  p.value = sprintf("%.2f", p.adjust(P.value, method="BH"))) %>%
-    merge(., names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
-    select(Feature, Name, beta_ci, p.value) %>%
-    gather(variable, value, -Feature, -Name) %>%
-    unite(temp, Name, variable) %>%
-    spread(temp, value) %>%
-    mutate(Feature = rename.genus(gsub("_Bacteria", "", Feature))) %>%
-    arrange(Feature)
-
-names(maaslin.lacto.max.table) <- gsub(" ", "_", names(maaslin.lacto.max.table ))
-
-typology.tbl2 <- data.frame(col_keys = c("Feature",
-                                         "Mean_arterial_pressure_beta_ci",
-                                         "Mean_arterial_pressure_p.value",
-                                         "Systolic_BP_beta_ci",
-                                         "Systolic_BP_p.value",
-                                         "Diastolic_BP_beta_ci",
-                                         "Diastolic_BP_p.value",
-                                         "Pulse_pressure_beta_ci",
-                                         "Pulse_pressure_p.value",
-                                         "Hypertension_beta_ci",
-                                         "Hypertension_p.value"),
-                            what = c( "",
-                                     rep("MAP", 2), 
-                                     rep("Systolic BP", 2), 
-                                     rep("Diastolic BP", 2), 
-                                     rep("Pulse pressure", 2), 
-                                     rep("Hypertension", 2)),
-                            measure = c("Lactobacillus", rep(c("β±SE", "p"), 5)),
-                            stringsAsFactors = FALSE)
-
-tbl2 <- flextable(data = maaslin.lacto.max.table) %>%
-    set_header_df(mapping = typology.tbl2, key = "col_keys") %>%
-        flextable::theme_booktabs()
-```
-
-```{r write table 2}
-tbl2head <- "Lactobacilli"
-tbl2foot <- ""
-writetable(doc, tbl2, number = 2, tbl2head, tbl2foot)
-```
-
 ```{r Write docx to file}

### 2019-04-29 articletwo.Rmd
Lactobacillus species level calulations removed

@@
 pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
-pseq.species.lactobacillus <- subset_taxa(pseq.species, Genus=="Lactobacillus")
-pseq.species.lactobacillus.core <- core(pseq.species.lactobacillus, detection = 0.001, prevalence = 0.01)
-```
-
-At genus level core has number of taxa `r ntaxa(pseq.genus.core)` (`r
-ntaxa(pseq.genus)`). For lactobacilli core has number of taxa `r
-ntaxa(pseq.species.lactobacillus.core)` (`r ntaxa(pseq.species.lactobacillus)`).
-
-```{r maalin manual, eval = FALSE}
-test.abud <- microbiome::abundances(microbiome::transform(pseq.genus.core, "compositional")) %>%
-    t %>%
-    as.data.frame
-df <- test.abud %>% mutate_all(funs(asin(sqrt(.))))
 ```
@@
 maaslin <- list()
-
 maaslin$genus <- maaslinwrapper(pseq = pseq.genus.core,
@@
                 taxa = "Halapricum.Archaea-")
-
-lactotaxa <-c("Lactobacillus_delbrueckii.Bacteria", "Lactobacillus_ruminis.Bacteria", "Lactobacillus_acidophilus.Bacteria",
-                    "Lactobacillus_fermentum.Bacteria", "Lactobacillus_paracasei.Bacteria", "Lactobacillus_salivarius.Bacteria")
-
-maaslin$lacto.min <- maaslinwrapper(pseq = pseq.species.lactobacillus.core,
-                looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
-                forced = c("BL_AGE", "SEX"),
-                taxa =  paste0(lactotaxa,  collapse = ","))
-
-maaslin$lacto.max  <- maaslinwrapper(pseq = pseq.species.lactobacillus.core,
-                looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
-                forced =c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03",
-                          "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                taxa = paste0(lactotaxa, collapse = ","))
-
 ```

### 2019-04-29 articletwo.Rmd
Ready?

@@
                   beta = Coefficient) %>%
-    dplyr::filter(Qval < 0.05) %>%
+    dplyr::filter(Qval < 0.05, Variable != "HYPERTENSION") %>%
     arrange(Feature)
 
-maaslinplot <- plot_maaslin_heatmap_final(maas.results.df.signf %>% dplyr::filter(Variable!="HYPERTENSION"),
-                      beta.maxabs=max(abs(maas.results.df.signf[,'beta'])),
-                      order.by.category=FALSE)
+my.plot.order <- function(x) {
+    y <- seq(length(x))
+    y[x == "Mean arterial pressure"] = 4
+    y[x == "Systolic BP"] = 3
+    y[x == "Diastolic BP"] = 2
+    y[x == "Pulse pressure"] = 1
+    y
+}
+
+maas.all.pairs <- expand.grid(Feature = maas.results.df.signf %>% pull(Feature),
+                         Name = maas.results.df.signf %>% pull(Name))
+maas.tables.df <- merge(maas.results.df.signf,
+                   all.pairs,
+                   all.y=TRUE,
+                   by.y=c('Feature', 'Name'),
+                   by.x=c('Feature', 'Name'))
+
+maaslinplot <- ggplot(maas.tables.df, aes(x = Feature, y = reorder(Name, my.plot.order(Name)), fill = beta)) +
+    geom_tile(aes(fill = 1)) +
+    geom_tile(colour="white", size=0.25) +
+    scale_fill_distiller(palette = "RdBu", name = "Effect size",
+                         limits=c(-0.003, 0.003), breaks = c(-0.003, 0, 0.003), na.value="grey30") +
+    coord_fixed() +
+    xlab("") +
+    ylab("") +
+    theme_classic(10) +  
+    theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust=0.5, size=7),
+          axis.text.y = element_text(size=7))
 
@@
 ```{r Write figure 3}
-fig3head <- ""
-fig3foot <- ""
+fig3head <- "Associations between genus level abundances and blood pressure variables"
+fig3foot <- "Only backteria with significant associations are shown.
+Blood pressure variables are normalized and abundances are arcsin-square root transformed.
+Dark gray color represents unsignificant associations."
 writeimage(doc, 3, "cache/maaslin.png", fig3head, fig3foot)

### 2019-04-29 articletwo.Rmd
Duplicated label

@@
 
-```{r Write figure 3}
+```{r Write figure 4}
 fig3head <- "Associations between genus level abundances and blood pressure variables"

### 2019-04-29 articletwo.Rmd
Renamed var

@@
 maas.tables.df <- merge(maas.results.df.signf,
-                   all.pairs,
+                        maas.all.pairs,
                         all.y=TRUE,

### 2019-04-29 articletwo.Rmd
Typos

@@
 
-}
-```
 # Importing libraries:
@@
 ```{r save grob}
-alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 1.3)
-alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 1.3)
+alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 0.5)
+alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 0.5)
 
@@
 Panel B has the full model where covariates are age, sex, BMI, smoker, exercise, and four antihypertensive medications.
-The blue tinted box on the left represents four linear and one logistic model where blood pressure variables are dependant variables and Shannon's index is included in covariates.
+The blue tinted box on the left represents four linear and one logistic model where blood pressure variables are dependent variables and Shannon's index is included in covariates.
 The gray tinted bars represent beta diversity i.e.
@@
 Result marked with asterisk are significant at FDR 0.05.
-Permutational analysis for beta diverity has 1000 permutations (p=0.01)."
+Permutational analysis for beta diversity has 1000 permutations (p=0.01)."
 writeimage(doc, 1, "cache/alpha-beta.png", fig1head, fig1foot)
@@
 First two axes explain %.1f%% of the variation.
-Centered log-ratio (CLR) tranformation is used for microbiome abundances.", sum(pca.percentage[1:2]))
+Centered log-ratio (CLR) transformation is used for microbiome abundances.", sum(pca.percentage[1:2]))
 writeimage(doc, 2, "cache/pca.png", fig2head, fig2foot)
@@
 
-```{r Write figure 4}
+```{r Write figure 3}
 fig3head <- "Associations between genus level abundances and blood pressure variables"
-fig3foot <- "Only backteria with significant associations are shown.
+fig3foot <- "Only bacteria with significant associations are shown.
 Blood pressure variables are normalized and abundances are arcsin-square root transformed.
-Dark gray color represents unsignificant associations."
+Dark gray color represents insignificant associations."
 writeimage(doc, 3, "cache/maaslin.png", fig3head, fig3foot)
@@
 
-```{r Write figure 3}
-fig3head <- "Associations between dietary salt, lactobacillus abundance, and odds for hypertensoin"
-fig3foot <- "In panel A we see lactobacillus abundances in different dietary salt groups.
+```{r Write figure 4}
+fig4head <- "Associations between dietary salt, lactobacillus abundance, and odds for hypertensoin"
+fig4foot <- "In panel A we see lactobacillus abundances in different dietary salt groups.
 The mean value is adjusted using age and sex.
-In panel B we split paticipants in four groups by lactobacillus abundances and calulate odds ratio for hypertension.
+In panel B we split participants in four groups by lactobacillus abundances and caclulate odds ratio for hypertension.
 Age and sex are included in logistic model.
-Both panel show  95 % CI for respective variables."
-writeimage(doc, 3, "cache/salt-forest.png", fig3head, fig3foot)
+A 95 % CI is drawn in both panels."
+writeimage(doc, 4, "cache/salt-forest.png", fig4head, fig4foot)
 ```

### 2019-04-29 articletwo.Rmd
Toggle sourced
Minor

@@
 
+<details>
+  <summary>Toggle imported core</summary>
 ### Officer functions
@@
 
+</details>
 # Loading data
@@
     theme_classic(10) +  
-    theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust=0.5, size=7),
-          axis.text.y = element_text(size=7))
+    theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust=0.5, size=10),
+          axis.text.y = element_text(size=10))
 

### 2019-04-29 articletwo.Rmd
Minor

@@
 ```{r Write figure 1}
-fig1head <- "Associations between BP variables and microbial diversity"
+fig1head <- "Associations between BP variables and microbial diversity."
 fig1foot <- "Results in panel A are calculated using minimal model using only age and sex for covariates.
@@
 ```{r Write figure 2}
-fig2head <- "Principal component analysis for bacterial abundances"
+fig2head <- "Principal component analysis for bacterial abundances."
 fig2foot <- sprintf("Principal component analysis for bacterial abundances at species level.
@@
                   beta = Coefficient) %>%
-    dplyr::filter(Qval < 0.05, Variable != "HYPERTENSION") %>%
+    dplyr::filter(Qval < 0.05) %>%
     arrange(Feature)
@@
     y <- seq(length(x))
-    y[x == "Mean arterial pressure"] = 4
-    y[x == "Systolic BP"] = 3
-    y[x == "Diastolic BP"] = 2
-    y[x == "Pulse pressure"] = 1
+    y[x == "Systolic BP"] = 4
+    y[x == "Diastolic BP"] = 3
+    y[x == "Pulse pressure"] = 2
+    y[x == "Mean arterial pressure"] = 1
+    y[x == "Hypertension"] = 0
     y
@@
 
-ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 7, dpi = 300)
+ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 6, dpi = 300)
 ```
 
+Max value of abs beta is `r maas.results.df.signf %>% pull(Coefficient) %>% abs %>% max`.
+
 ```{r Write figure 3}
-fig3head <- "Associations between genus level abundances and blood pressure variables"
+fig3head <- "Associations between genus level abundances and blood pressure variables."
 fig3foot <- "Only bacteria with significant associations are shown.
@@
 ```{r Write figure 4}
-fig4head <- "Associations between dietary salt, lactobacillus abundance, and odds for hypertensoin"
+fig4head <- "Associations between dietary salt, lactobacillus abundance, and odds for hypertension."
 fig4foot <- "In panel A we see lactobacillus abundances in different dietary salt groups.
 The mean value is adjusted using age and sex.
-In panel B we split participants in four groups by lactobacillus abundances and caclulate odds ratio for hypertension.
+In panel B we split participants in four groups by lactobacillus abundances and calculate odds ratio for hypertension.
 Age and sex are included in logistic model.

### 2019-05-02 articletwo.Rmd
Minor
Scatter plot

@@
 library(ggfortify)
+library(RColorBrewer)
 ```
@@
 
+### Officer functions
+
 <details>
-  <summary>Toggle imported core</summary>
-### Officer functions
+  <summary>Open/Close</summary>
 
@@
 
+</details>
+
 ### RR biome functions
+
+<details>
+  <summary>Open/Close</summary>
 
@@
 
+</details>
+
 ### EdgeR plot functions
+
+<details>
+  <summary>Open/Close</summary>
 
@@
 
+</details>
+
 ### Plot functions
+
+<details>
+  <summary>Open/Close</summary>
 
@@
 </details>
+
 # Loading data
@@
 ```{r alphabeta definitions}
-dset.alpha <- meta.merge.alphadiversity(pseq.species)
+dset.alpha <- meta.merge.alphadiversity(pseq.species) %>%
+    mutate(diversities_shannon = scale(diversities_shannon),
+           MAP = oMAP,
+           PULSEPRESSURE = oPULSEPRESSURE,
+           SYSTM = oSYSTM,
+           DIASM = oDIASM)
+
 alphadiversity <- list(
-    min = calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX  + diversities_shannon"),
-    max = calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + diversities_shannon"))
+    min = calculateglm(dset.alpha,
+                       modelstring = "%s ~ BL_AGE + SEX  + diversities_shannon"),
+    max = calculateglm(dset.alpha,
+                       modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + diversities_shannon"))
+
 diversities <- mergediversities(alphadiversity, adonis.species)
@@
 ```{r save grob}
-alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 0.5)
-alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 0.5)
+alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 0.6)
+alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 0.6)
 
@@
                         by.y=c('Feature', 'Name'),
-                        by.x=c('Feature', 'Name'))
+                        by.x=c('Feature', 'Name')) %>%
+    mutate(p.groups = cut(P.value*ifelse(Coefficient < 0, -1, 1),
+                          breaks = c(-0.05, -0.01, -0.001, 0, 0.001, 0.01, 0.05),
+                          labels = c("beta < 0, p < 0.05",
+                                     "beta < 0, p < 0.01",
+                                     "beta < 0, p < 0.001",
+                                     "beta > 0, p < 0.001",
+                                     "beta > 0, p < 0.01",
+                                     "beta > 0, p < 0.05")))
 
-maaslinplot <- ggplot(maas.tables.df, aes(x = Feature, y = reorder(Name, my.plot.order(Name)), fill = beta)) +
-    geom_tile(aes(fill = 1)) +
-    geom_tile(colour="white", size=0.25) +
-    scale_fill_distiller(palette = "RdBu", name = "Effect size",
-                         limits=c(-0.003, 0.003), breaks = c(-0.003, 0, 0.003), na.value="grey30") +
+mycolors <- c("beta < 0, p < 0.05" = "#B2182B",
+              "beta < 0, p < 0.01" = "#EF8A62",
+              "beta < 0, p < 0.001" = "#FDDBC7",
+              "beta > 0, p < 0.001" = "#D1E5F0",
+              "beta > 0, p < 0.01" = "#67A9CF",
+              "beta > 0, p < 0.05" = "#2166AC",
+              "NA" = "gray50")
+
+maaslinplot <- ggplot(maas.tables.df, aes(x = Feature,
+                                          y = reorder(Name, my.plot.order(Name)),
+                                          fill = p.groups)) +
+    geom_tile(color = "white", aes(fill = "NA")) +
+    geom_tile(color = "white") +
+    scale_fill_manual(values = mycolors, name = "P-value", na.translate = F) +
+    scale_x_discrete() +
     coord_fixed() +
@@
           axis.text.y = element_text(size=10))
-
 ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 6, dpi = 300)
@@
 Max value of abs beta is `r maas.results.df.signf %>% pull(Coefficient) %>% abs %>% max`.
+
+```{r scatter plots}
+scatter.signf.bacteria <- paste0(maas.results.df.signf %>% pull(Feature) %>% unique %>% tolower, ".bacteria")
+
+scatter.meta <- meta(pseq.genus) %>%
+    tibble::rownames_to_column("id") %>%
+      mutate(MAP = oMAP,
+           PULSEPRESSURE = oPULSEPRESSURE,
+           SYSTM = oSYSTM,
+           DIASM = oDIASM) %>%
+    gather("keym", "valuem", c(SYSTM, DIASM, MAP, PULSEPRESSURE))
+
+scatter.abund <- abundances(pseq.genus, transform = "compositional") %>%
+    t %>%
+    as.data.frame %>%
+    tibble::rownames_to_column("id") %>%
+    setNames(tolower(replace.brackets(names(.)))) %>%
+    select(id, scatter.signf.bacteria) %>%
+    gather(key, value, -id) %>%
+    mutate(value = asin(sqrt(value)))
+
+scatter.full <- full_join(scatter.meta, scatter.abund, by = "id")
+
+scatterplot <- ggplot(scatter.full, aes(y = valuem, x = value)) +
+    facet_wrap(~key + keym, ncol=4, scales = "fixed") +
+    geom_point(size=0.1, shape=1) +
+    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
+    stat_poly_eq(formula = "y ~ x",
+                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
+                 label.y =  -4,
+                 parse = TRUE,
+                 size = 2.5) +
+    ylab("asin(sqrt(abundance))") +
+    ggplot2::theme_minimal()
+
+ggsave(file = "cache/scatter.png", plot = scatterplot, height = 20, width = 10, dpi = 300)
+```
 

### 2019-05-02 articletwo-rrbiome.R

@@ -156,12 +156,13 @@ vectortolist <- function(c) {
   l
 }
 
-mergediversities <- function(alphadiversity, betadiversity) {
+mergediversities <- function(alphadiversity, betadiversity,
+                             responses = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) {
     lapply(vectortolist(names(alphadiversity)), function(x) {
         alpha.select <- alphadiversity[[x]] %>% select(estimate, p.value, response)
         beta.rbind <- betadiversity[[x]] %>%
             data.table::rbindlist(id = "model_name") %>%
-            dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
+            dplyr::filter(term %in% responses) %>%
             dplyr::select(response, R2)
         merge( alpha.select , beta.rbind, by = "response") %>%
             merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate") %>%

### 2019-05-06 articletwo.Rmd
Scatter plot

@@
               ofile = '~/phd/research/articletwo/TAGS', verbose = TRUE, append = FALSE)
+    if ("loadlast" %in% names(args) )
+        load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
 }
@@
 maaslin <- list()
+
+genus.core.bacteria <- pseq.genus.core %>%
+    abundances %>%
+    rownames %>%
+    replace.brackets %>%
+    mygrep(".Bacteria$", .)
+
 maaslin$genus <- maaslinwrapper(pseq = pseq.genus.core,
@@
                            "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                taxa = "Halapricum.Archaea-")
+                taxa = paste0(genus.core.bacteria, collapse = ","))
 ```
@@
            SYSTM = oSYSTM,
-           DIASM = oDIASM) %>%
-    gather("keym", "valuem", c(SYSTM, DIASM, MAP, PULSEPRESSURE))
+           DIASM = oDIASM)
 
-scatter.abund <- abundances(pseq.genus, transform = "compositional") %>%
+scatter.abund <- abundances(pseq.genus, transform = "clr") %>%
     t %>%
@@
     select(id, scatter.signf.bacteria) %>%
-    gather(key, value, -id) %>%
-    mutate(value = asin(sqrt(value)))
+    mutate_at(.vars = vars(scatter.signf.bacteria), .funs = funs(myscale(.)))
 
@@
 
-scatterplot <- ggplot(scatter.full, aes(y = valuem, x = value)) +
-    facet_wrap(~key + keym, ncol=4, scales = "fixed") +
-    geom_point(size=0.1, shape=1) +
-    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
-    stat_poly_eq(formula = "y ~ x",
-                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
-                 label.y =  -4,
-                 parse = TRUE,
-                 size = 2.5) +
-    ylab("asin(sqrt(abundance))") +
-    ggplot2::theme_minimal()
+scatter.combinations <- expand.grid(cc("SYSTM", "DIASM", "PULSEPRESSURE", "MAP"),
+                                    c2l(scatter.signf.bacteria)) %>%
+    setNames(c("rr", "abund"))
 
+scatter.plot <- function(df,
+                         maas.results,
+                         rr,
+                         abund,
+                         covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
+                                        "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08",
+                                        "BL_USE_RX_C09"),
+                         prediction = TRUE,
+                         plots = TRUE) {
+    fo <- sprintf("%s ~ %s + %s", rr, abund, paste0(covariates, collapse="+"))
+    model <- lm(as.formula(fo), data = df)
+
+    if (prediction == FALSE) return(model %>% summary)
+
+    df.predict <- data.frame(BPV = seq(-5,5,0.1)) %>%
+        mutate(SEX = as.factor(0),
+               BL_AGE = df %>% pull(BL_AGE) %>% mean,
+               BMI = df %>% pull(BMI) %>% mean,
+               CURR_SMOKE = as.factor(0),
+               Q57X = as.factor(1),
+               DIAB = 0,
+               BL_USE_RX_C03 = as.factor(0),
+               BL_USE_RX_C07 = as.factor(0),
+               BL_USE_RX_C08 = as.factor(0),
+               BL_USE_RX_C09 = as.factor(0),
+               !!abund := seq(-5,5,0.1)
+               )
+    prediction <- cbind(df.predict, predict(model, df.predict, interval = 'confidence'))
+
+    if (plots == FALSE) return(prediction)
+
+    feature <- firstup(gsub(".bacteria", "", replace.brackets(abund)))
+    qval.maaslin <- maas.results %>%
+        dplyr::filter(Feature == feature, Value == rr) %>%
+        nrow
+
+    if (qval.maaslin == 0) { return (ggplot() + ggplot2::theme_minimal()) }
+    pval <- pub.p(summary(model)$coef[2,4])
+    rsqr <- signif(summary(model)$adj.r.squared, 2)
+    slope <- signif(model$coef[[2]], 2)
+
+    ggplot(prediction, aes_string(x = abund, y = "fit")) +
+        geom_line() +
+        geom_point(size=0.1, shape=1, aes_string(x = abund, y = rr, color = "SEX"),
+                   data = df) +
+        labs(title = sprintf("Adj R2 = %s, Slope = %s, p = %s", rsqr, slope, pval)) +
+        xlab(feature) +
+        ylab(rr) +
+        ylim(0,200) +
+        ggplot2::theme_minimal() +
+        theme(plot.title = element_text(size=8)) +
+        theme(legend.position="none")
+}
+
+scatterplots <- lapply(data.frame(t(scatter.combinations)),
+                       function(x)
+                           scatter.plot(scatter.full, maas.results.df.signf, x$rr, x$abund))
+
+scatterplot <- do.call("grid.arrange", c(scatterplots, ncol=4))
 ggsave(file = "cache/scatter.png", plot = scatterplot, height = 20, width = 10, dpi = 300)

### 2019-05-06 articletwo-rrbiome.R

@@ -209,3 +209,25 @@ maaslinwrapper <- function(pseq, looped, forced, taxa, tempstr = "%s/temp/maasli
 mygrep <- function(word, list, ignorecase = TRUE) {
     list[grepl(word, list, ignore.case = ignorecase)]
 }
+
+cc <- function(...) {
+    c2l(c(...))
+}
+
+
+c2l <- function(c) {
+    l <- as.list(c)
+    names(l) <- c
+    l
+}
+
+pub.p <- function(p, Nfdr = FALSE) {
+    p <- as.numeric(p)
+    if (Nfdr) p <- p.adjust(p, method="BH", n = Nfdr)
+    ifelse(p < 0.01, ifelse(p<0.001, sprintf("%.2e", p), sprintf("%.3f", p)), sprintf("%.2f", p))
+}
+
+firstup <- function(x) {
+  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
+  x
+}

### 2019-05-06 articletwo.Rmd
Basic PCoA functionality

@@
 
-```{r pca calculate}
-if (!file.exists("rds/pca.prcomp.rds")) {
-    pca.abundances  <- as.matrix(t(microbiome::abundances(microbiome::transform(pseq.species, "clr"))))
-    pca.prcomp <- stats::prcomp(pca.abundances)
-    saveRDS(pca.prcomp, file = "rds/pca.prcomp.rds")
+```{r pcoa calculate}
+if (!file.exists("rds/pcoa.ordinate.rds")) {
+    pcoa.abundances <- microbiome::transform(pseq.species, 'compositional')
+    pcoa.ordinate <- ordinate(pcoa.abundances, method="PCoA", distance="bray")
+    saveRDS(pca.prcomp, file = "rds/pcoa.ordinate.rds")
 } else {
-    pca.prcomp <- readRDS("rds/pca.prcomp.rds")
+    pcoa.ordinate <- readRDS("rds/pcoa.ordinate.rds")
 }
@@
 
-## Principal component analysis (Figure 2.)
+## Principal coordinate analysis (Figure 2.)
 
-```{r pca plot}
-pca.eigs <- pca.prcomp$sdev^2
-pca.percentage <- 100 * pca.eigs /sum(pca.eigs)
-pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)
+```{r pcoa plot}
+pcoa.vectors <- pcoa.ordinate$vectors %>% as.data.frame
+eig.fracs <- pcoa.ordinate$values$Relative_eig
+pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2),
+                 meta(pseq.species), by=0, all = TRUE)
 
-pcaplot <- ggplot(cbind(meta(pseq.species), pca.prcomp$x[,1:2]), aes(x=PC1, y=PC2)) +
-    geom_point(aes(color = HYPERTENSION), size = 0.5) +
+colnames(pcoa.df)
+
+pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.2, color = HYPERTENSION, shape = ANYDRUG)) +
+    geom_point(size=0.4, alpha=0.4) +
+    xlab(paste("PCoA axis 1,", round(100*eig.fracs[1],1), "%")) +
+    ylab(paste("PCoA axis 2,", round(100*eig.fracs[2],1), "%")) +
     scale_color_manual(values=c("blue", "red"),
@@
                        labels=c("Yes", "No")) +
-    xlab(pca.labels[1]) +
-    ylab(pca.labels[2]) +
+    scale_shape_manual(values=c(4, 1),
+                       name = "BP-medication",
+                       breaks = c("1", "0"),
+                       labels = c("Yes", "No")) +
     coord_fixed() +
@@
     theme_classic() +
-    theme(text = element_text(size=20),
+    guides(colour = guide_legend(override.aes = list(size=6, shape = 15)),
+           shape = guide_legend(override.aes = list(size=6))) +
+    theme(text = element_text(size=14),
+          legend.title = element_text(size = 12),
+          legend.text=element_text(size=10),
           legend.justification=c(1,0),
-          legend.position=c(1.03, 0.75),
-          legend.title=element_text(size = 20),
-          legend.direction="vertical",
-          legend.text=element_text(size = floor(0.75*20)),
-          legend.box.margin = margin(1, 1, 1, 1, "mm"),
-          legend.box.background = element_rect(colour = "black"))
+          legend.position=c(1,0),
+          legend.margin=margin(t = 0, unit='cm'),
+          legend.background = element_rect(fill=alpha('gray', 0.1)))
 
-ggsave(file = "cache/pca.png", plot = pcaplot, height = 5, width = 9, dpi = 300)
+ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=5, width=7, dpi = 300)
+
 ```

### 2019-05-06 articletwo.Rmd
PCoA ready

@@
 pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2),
-                 meta(pseq.species), by=0, all = TRUE)
+                 meta(pseq.species), by=0, all = TRUE) %>%
+    mutate(STATETREATMENT = factor(paste0(HYPERTENSION, ANYDRUG),
+                                   levels = c("11", "10", "01", "00")))
 
-colnames(pcoa.df)
+pcoa.df %>% pull(STATETREATMENT)
 
-pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.2, color = HYPERTENSION, shape = ANYDRUG)) +
+pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.2, color = STATETREATMENT, shape = STATETREATMENT)) +
     geom_point(size=0.4, alpha=0.4) +
@@
     ylab(paste("PCoA axis 2,", round(100*eig.fracs[2],1), "%")) +
-    scale_color_manual(values=c("blue", "red"),
-                       name= "Hypertension",
-                       breaks=c("1", "0"),
-                       labels=c("Yes", "No")) +
-    scale_shape_manual(values=c(4, 1),
-                       name = "BP-medication",
-                       breaks = c("1", "0"),
-                       labels = c("Yes", "No")) +
+    scale_colour_manual(name = "State and treatment",
+                        labels = c("Hypertensive, medication",
+                                   "Hypertensive, no medication",
+                                   "Normotensive, no medication"),
+                        breaks=c("11", "10", "00"),
+                        values = c("red", "red", "blue", "blue")) +   
+    scale_shape_manual(name = "State and treatment",
+                       labels = c("Hypertensive, medication",
+                                  "Hypertensive, no medication",
+                                  "Normotensive, no medication"),
+                       breaks = c("11", "10", "00"),
+                       values = c(4, 1, 1))+
     coord_fixed() +
@@
     theme_classic() +
-    guides(colour = guide_legend(override.aes = list(size=6, shape = 15)),
-           shape = guide_legend(override.aes = list(size=6))) +
+    guides(colour = guide_legend(override.aes = list(size=4))) +
     theme(text = element_text(size=14),
-          legend.title = element_text(size = 12),
+          legend.title = element_blank(),
           legend.text=element_text(size=10),
@@
           legend.position=c(1,0),
-          legend.margin=margin(t = 0, unit='cm'),
+          legend.margin=margin(t = 1, unit='mm'),
           legend.background = element_rect(fill=alpha('gray', 0.1)))
 
-ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=5, width=7, dpi = 300)
-
+ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=5, width=5, dpi = 300)
 ```
@@
 ```{r Write figure 2}
-fig2head <- "Principal component analysis for bacterial abundances."
-fig2foot <- sprintf("Principal component analysis for bacterial abundances at species level.
-Centered log ratio transform is performed for the compositional data.
-First two axes explain %.1f%% of the variation.
-Centered log-ratio (CLR) transformation is used for microbiome abundances.", sum(pca.percentage[1:2]))
+fig2head <- "Principal coordinate analysis for bacterial abundances."
+fig2foot <- sprintf("Principal coordinate analysis for bacterial abundances at species level.
+Bray-Curtis distances are calculated for abundances after compositional transformation.
+First two axes explain %.1f%% of the variation.", round(100*sum(eig.fracs[1:2]),1))
 writeimage(doc, 2, "cache/pca.png", fig2head, fig2foot)

### 2019-05-06 articletwo.Rmd
Minor

@@
     geom_point(size=0.4, alpha=0.4) +
-    xlab(paste("PCoA axis 1,", round(100*eig.fracs[1],1), "%")) +
-    ylab(paste("PCoA axis 2,", round(100*eig.fracs[2],1), "%")) +
+    xlab(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1])) +
+    ylab(sprintf("PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2])) +
     scale_colour_manual(name = "State and treatment",

### 2019-05-07 articletwo.Rmd
Typo

@@
     pcoa.ordinate <- ordinate(pcoa.abundances, method="PCoA", distance="bray")
-    saveRDS(pca.prcomp, file = "rds/pcoa.ordinate.rds")
+    saveRDS(pcoa.ordinate, file = "rds/pcoa.ordinate.rds")
 } else {
@@
 First two axes explain %.1f%% of the variation.", round(100*sum(eig.fracs[1:2]),1))
-writeimage(doc, 2, "cache/pca.png", fig2head, fig2foot)
+writeimage(doc, 2, "cache/pcoa-species.png", fig2head, fig2foot)
 ```

### 2019-05-09 partialrda.Rmd
Partial RDA

---
title: "Partial RDA"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis', cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

dir.create("rds/", showWarnings = FALSE)
dir.create("session/", showWarnings = FALSE)
```

```{r libraries, cache = FALSE}
library(dplyr)
library(vegan)
library(tibble)
```

```{r data, eval = FALSE}
rda.abundances  <- as.matrix(t(microbiome::abundances(microbiome::transform(pseq.species, "clr"))))
permutest(rda.species, method="direct", perm=100)
rda.species <- rda(rda.abundances ~ MAP + Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
                   data = meta(pseq.species))
saveRDS(rda.species, file = "rda.specids.rds")
```

```{r load}
rda.species <- readRDS("rda.specids.rds")
```

```{r variance}
data.frame(Inertia = c(
               Conditioned = summary(rda.species)$partial.chi,
               Constrained = summary(rda.species)$constr.chi,
               Unconstrained = summary(rda.species)$unconst.chi)) %>%
    tibble::rownames_to_column("Variance") %>%
    mutate(Proportion = Inertia/sum(Inertia)) %>%
    mutate_at(2:2, funs(round(., 1))) %>%
    mutate_at(3:3, funs(round(., 5))) %>%
    kable
```

```{r direct}
permutest(rda.species, method="direct", perm=100)
```

```{r axis}
permutest(rda.species, method="axis", perm=100)
```

```{r term}
permutest(rda.species, method="term", perm=100)
```

```{r rda plot}
jpeg(file="cache/rda.jpg", res = 300, width = 20, height = 20, units = 'cm')
plot(rda.species)
dev.off()
```

```{r loading, eval = FALSE}
loadings <- as.data.frame(rda.species$CA$v.eig[,1:2])

rda.species %>% str

loadings %>% str
```

```{r save session}
save.image(file = paste0("session/rda.Rdata"))
```

### 2019-05-10 partialrda.Rmd
Library
Genus + core

@@
 library(vegan)
+library(broom)
 library(tibble)
+library(knitr)
+library(phyloseq)
+library(microbiome)
 ```
 
-```{r data, eval = FALSE}
-rda.abundances  <- as.matrix(t(microbiome::abundances(microbiome::transform(pseq.species, "clr"))))
-permutest(rda.species, method="direct", perm=100)
-rda.species <- rda(rda.abundances ~ MAP + Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
-                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
-                   data = meta(pseq.species))
-saveRDS(rda.species, file = "rda.specids.rds")
+### RR biome functions
+
+<details>
+  <summary>Open/Close</summary>
+
+```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
+source("articletwo-rrbiome.R")
 ```
 
-```{r load}
-rda.species <- readRDS("rda.specids.rds")
-```
+</details>
 
-```{r variance}
+```{r functions}
+myresults.table <- function(model) {
     data.frame(Inertia = c(
-               Conditioned = summary(rda.species)$partial.chi,
-               Constrained = summary(rda.species)$constr.chi,
-               Unconstrained = summary(rda.species)$unconst.chi)) %>%
+                   Conditioned = summary(model)$partial.chi,
+                   Constrained = summary(model)$constr.chi,
+                   Unconstrained = summary(model)$unconst.chi)) %>%
         tibble::rownames_to_column("Variance") %>%
@@
         kable
+}
+
+anotab <- function(x) {
+    EPS <- sqrt(.Machine$double.eps)
+    Pval <- (colSums(sweep(x$F.perm, 2, x$F.0 - EPS, ">=")) + 1)/(x$nperm + 1)
+    anotab <- data.frame(x$df, x$chi, c(x$F.0, NA), c(Pval, NA))
+    colnames(anotab) <- c("Df", "Inertia", "F", "Pr(>F)")
+    rownames(anotab) <- c(x$termlabels, "Residual")
+    class(anotab) <- c("anova", "data.frame")
+    anotab 
+}
 ```
 
-```{r direct}
-permutest(rda.species, method="direct", perm=100)
+Loading phyloseq object
+
+```{r data}
+pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
+pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
 ```
 
-```{r axis}
-permutest(rda.species, method="axis", perm=100)
+Updating phyloseq objects
+
+```{r updating phyloseq objects}
+phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(filter.phenotype.data(pseq.genus))
+phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(filter.phenotype.data(pseq.species))
 ```
 
-```{r term}
-permutest(rda.species, method="term", perm=100)
+```{r data}
+rda.abundances.species  <- microbiome::transform(pseq.species, "clr") %>%
+    microbiome::abundances() %>%
+    t() %>%
+    as.matrix()
+
+rda.abundances.genus  <- microbiome::transform(pseq.genus, "clr") %>%
+    microbiome::abundances() %>%
+    t() %>%
+    as.matrix()
+
+rda.abundances.genus.core  <- microbiome::transform(pseq.genus.core, "clr") %>%
+    microbiome::abundances() %>%
+    t() %>%
+    as.matrix()
 ```
 
-```{r rda plot}
-jpeg(file="cache/rda.jpg", res = 300, width = 20, height = 20, units = 'cm')
-plot(rda.species)
-dev.off()
+```{r models}
+rda.species <- rda(rda.abundances.species ~ MAP +
+                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
+                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
+                   data = meta(pseq.species))
+rda.genus <- rda(rda.abundances.genus ~ MAP +
+                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
+                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
+                 data = meta(pseq.genus))
+rda.genus.core <- rda(rda.abundances.genus.core ~ MAP +
+                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
+                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
+                   data = meta(pseq.genus.core))
+
 ```
 
-```{r loading, eval = FALSE}
-loadings <- as.data.frame(rda.species$CA$v.eig[,1:2])
+```{r myresults species}
+myresults.table(rda.species)
+```
 
-rda.species %>% str
+```{r myresults genus}
+myresults.table(rda.genus)
+```
 
-loadings %>% str
+```{r myresults genus core}
+myresults.table(rda.genus.core)
+```
+
+# ANOVAs
+
+Species
+
+```{r anova species}
+anova.species <- permutest(rda.species, method="reduced", perm=1)
+anova.species %>% anotab %>% kable
+```
+
+Genus
+
+```{r anova genus}
+anova.genus <- permutest(rda.genus, method="reduced", perm=1)
+anova.genus %>% anotab %>% kable
+```
+
+```{r anova genus core}
+anova.genus.core <- permutest(rda.genus.core, method="reduced", perm=1)
+anova.genus.core %>% anotab %>% kable
+```
+
+# Plots
+
+Species
+
+```{r rda plot species}
+plot(rda.species)
+```
+
+Genus
+
+```{r rda plot species}
+plot(rda.genus)
+```
+
+Genus core
+
+```{r rda plot species}
+plot(rda.genus.core)
 ```

### 2019-05-11 partialrda.Rmd
Duplicate
sex missing

@@
 ---
-title: "Partial RDA"
+title: "Randomforest"
 author: "Joonatan Palmu"
@@
 pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
-pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
 ```
@@
 phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(filter.phenotype.data(pseq.species))
+pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
 ```
 
-```{r data}
+```{r transformations}
 rda.abundances.species  <- microbiome::transform(pseq.species, "clr") %>%

### 2019-05-11 partialrda.Rmd
Rename

--- 2019-05-10 partialrda.Rmd
@@
 
-```{r data}
+```{r transformations}
 rda.abundances.species  <- microbiome::transform(pseq.species, "clr") %>%
@@
 
-```{r rda plot species}
+```{r rda plot genus}
 plot(rda.genus)
@@
 
-```{r rda plot species}
+```{r rda plot genus core}
 plot(rda.genus.core)

### 2019-05-13 partialrda.Rmd
1->100

--- 2019-05-11 partialrda.Rmd
@@
 ```{r anova species}
-anova.species <- permutest(rda.species, method="reduced", perm=1)
+anova.species <- permutest(rda.species, method="reduced", perm=100)
 anova.species %>% anotab %>% kable
@@
 ```{r anova genus}
-anova.genus <- permutest(rda.genus, method="reduced", perm=1)
+anova.genus <- permutest(rda.genus, method="reduced", perm=100)
 anova.genus %>% anotab %>% kable
@@
 ```{r anova genus core}
-anova.genus.core <- permutest(rda.genus.core, method="reduced", perm=1)
+anova.genus.core <- permutest(rda.genus.core, method="reduced", perm=100)
 anova.genus.core %>% anotab %>% kable

### 2019-05-13 partialrda.Rmd
Minor
Genus and core

--- 2019-05-11 partialrda.Rmd
@@
 library(microbiome)
+library(purrr)
+library(parallel)
 ```
@@
 </details>
-
-```{r functions}
-myresults.table <- function(model) {
-    data.frame(Inertia = c(
-                   Conditioned = summary(model)$partial.chi,
-                   Constrained = summary(model)$constr.chi,
-                   Unconstrained = summary(model)$unconst.chi)) %>%
-        tibble::rownames_to_column("Variance") %>%
-        mutate(Proportion = Inertia/sum(Inertia)) %>%
-        mutate_at(2:2, funs(round(., 1))) %>%
-        mutate_at(3:3, funs(round(., 5))) %>%
-        kable
-}
-
-anotab <- function(x) {
-    EPS <- sqrt(.Machine$double.eps)
-    Pval <- (colSums(sweep(x$F.perm, 2, x$F.0 - EPS, ">=")) + 1)/(x$nperm + 1)
-    anotab <- data.frame(x$df, x$chi, c(x$F.0, NA), c(Pval, NA))
-    colnames(anotab) <- c("Df", "Inertia", "F", "Pr(>F)")
-    rownames(anotab) <- c(x$termlabels, "Residual")
-    class(anotab) <- c("anova", "data.frame")
-    anotab 
-}
-```
 
@@
 
-```{r models}
-rda.species <- rda(rda.abundances.species ~ MAP +
-                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
-                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
-                   data = meta(pseq.species))
-rda.genus <- rda(rda.abundances.genus ~ MAP +
-                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
-                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
-                 data = meta(pseq.genus))
-rda.genus.core <- rda(rda.abundances.genus.core ~ MAP +
-                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
-                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
-                   data = meta(pseq.genus.core))
-
+```{r model function}
+modelfunction <- function(x, rda.abund, df.meta) {
+    covariates <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
+                   "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07",
+                   "BL_USE_RX_C08", "BL_USE_RX_C09")
+    fo <- sprintf("rda.abund ~ %s + Condition(%s)", x, paste0(covariates, collapse = "+"))
+    model <- rda(as.formula(fo), data = df.meta)
+    anova.term <- anova(model, by="term", perm=100)
+    data.frame(var = as.character(x),
+               beta = coef(model)[,1][[ifelse(x == "HYPERTENSION", "HYPERTENSION1", x)]],
+               RDA.Conditioned = summary(model)$partial.chi,
+               RDA.Constrained = summary(model)$constr.chi,
+               RDA.Unconstrained = summary(model)$unconst.chi,
+               p.anova = anova.term[["Pr(>F)"]][1],
+               fo = fo)
+}
 ```
 
-```{r myresults species}
-myresults.table(rda.species)
+```{r return parser}
+returnparser <- function(ret){
+    res <- do.call(rbind, ret) %>%
+        mutate(RDA.total =  RDA.Conditioned + RDA.Constrained + RDA.Unconstrained,
+               RDA.Conditioned.percent = RDA.Conditioned/RDA.total,
+               RDA.Constrained.percent = RDA.Constrained/RDA.total,
+               RDA.Unconstrained.percent = RDA.Unconstrained/RDA.total
+               )
+
+    res %>%
+        dplyr::select(var,
+                      beta,
+                      p.anova,
+                      RDA.Conditioned.percent,
+                      RDA.Constrained.percent,
+                      RDA.Unconstrained.percent) %>%
+        kable()
+}
 ```
 
-```{r myresults genus}
-myresults.table(rda.genus)
+## Species
+
+```{r models species}
+ret.species <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
+              modelfunction,
+              rda.abund = rda.abundances.species,
+              df.meta = meta(pseq.species),
+              mc.cores = 5)
+returnparser(ret.species)
 ```
 
-```{r myresults genus core}
-myresults.table(rda.genus.core)
+## Genus
+
+```{r models genus}
+ret.genus <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
+              modelfunction,
+              rda.abund = rda.abundances.genus,
+              df.meta = meta(pseq.genus),
+              mc.cores = 5)
+returnparser(ret.genus)
 ```
 
-# ANOVAs
+## Genus core
 
-Species
-
-```{r anova species}
-anova.species <- permutest(rda.species, method="reduced", perm=1)
-anova.species %>% anotab %>% kable
-```
-
-Genus
-
-```{r anova genus}
-anova.genus <- permutest(rda.genus, method="reduced", perm=1)
-anova.genus %>% anotab %>% kable
-```
-
-```{r anova genus core}
-anova.genus.core <- permutest(rda.genus.core, method="reduced", perm=1)
-anova.genus.core %>% anotab %>% kable
-```
-
-# Plots
-
-Species
-
-```{r rda plot species}
-plot(rda.species)
-```
-
-Genus
-
-```{r rda plot species}
-plot(rda.genus)
-```
-
-Genus core
-
-```{r rda plot species}
-plot(rda.genus.core)
+```{r models genus core}
+ret.genus.core <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
+              modelfunction,
+              rda.abund = rda.abundances.genus.core,
+              df.meta = meta(pseq.genus.core),
+              mc.cores = 5)
+returnparser(ret.genus.core)
 ```

### 2019-05-13 articletwo-rrbiome.R

@@ -34,7 +34,14 @@ filter.phenotype.data <- function(pseq,
                       SYSTM = myscale(SYSTM),
                       DIASM = myscale(DIASM),
                       PULSEPRESSURE = myscale(PULSEPRESSURE),
-                      MAP = myscale(MAP)) %>%
+                      MAP = myscale(MAP),
+                      DIAB = as.factor(DIAB),
+                      SEX = as.factor(SEX),
+                      Q57X = as.factor(Q57X),
+                      BL_USE_RX_C03 = as.factor(BL_USE_RX_C03),
+                      BL_USE_RX_C07 = as.factor(BL_USE_RX_C07),
+                      BL_USE_RX_C08 = as.factor(BL_USE_RX_C08),
+                      BL_USE_RX_C09 = as.factor(BL_USE_RX_C09)) %>%
         dplyr::select(-MEN, -DNA_OK) %>%
         tibble::remove_rownames() %>%
         tibble::column_to_rownames(var = "rowname")
@@ -215,9 +222,9 @@ cc <- function(...) {
 }
 
 
-c2l <- function(c) {
-    l <- as.list(c)
-    names(l) <- c
+c2l <- function(...) {
+    l <- as.list(c(...))
+    names(l) <- c(...)
     l
 }
 

### 2019-05-13 partialrda.Rmd
Example

@@
 ```
+## All models
 
@@
 
-## Species
+### Species
 
@@
 
-## Genus
+### Genus
 
@@
 
-## Genus core
+### Genus core
 
@@
 
+## Example model (species, systole)
+
+RDA model
+
+```{r rda}
+ex.rda <- rda(rda.abundances.species ~ SYSTM + Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X
+                                                        + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 +
+                                                        + BL_USE_RX_C08 + BL_USE_RX_C09),
+              data = meta(pseq.species))
+```
+
+RDA summary
+
+```{r rda summary}
+data.frame(Inertia = c(
+                   Conditioned = summary(ex.rda)$partial.chi,
+                   Constrained = summary(ex.rda)$constr.chi,
+                   Unconstrained = summary(ex.rda)$unconst.chi)) %>%
+        tibble::rownames_to_column("Variance") %>%
+        mutate(Proportion = Inertia/sum(Inertia)) %>%
+        mutate_at(2:2, funs(round(., 1))) %>%
+        mutate_at(3:3, funs(round(., 5))) %>%
+        kable
+```
+
+RDA r-squares
+
+```{r rda rsqr}
+RsquareAdj(ex.rda)$r.squared %>% kable
+```
+
+RDA coefficients
+
+```{r rda coef}
+coef(ex.rda) %>% kable
+```
+
+Anova
+
+```{r anova}
+anova.term <- anova(ex.rda, by="term", perm=100)
+anova.term %>% kable
+```
+
+Plot
+
+```{r rda plot}
+plot(ex.rda)
+```
+
 ```{r save session}

### 2019-05-14 partialrda.Rmd
perms

@@
 ---
-title: "Randomforest"
+title: "Redundancy analysis for hypertension"
 author: "Joonatan Palmu"
@@
     model <- rda(as.formula(fo), data = df.meta)
-    anova.term <- anova(model, by="term", perm=100)
+    anova.term <- anova(model, by="term", perm=1000)
     data.frame(var = as.character(x),
@@
 ```{r anova}
-anova.term <- anova(ex.rda, by="term", perm=100)
+anova.term <- anova(ex.rda, by="term", perm=1000)
 anova.term %>% kable

### 2019-05-14 partialrda.Rmd
Non partial RDA

--- 2019-05-13 partialrda.Rmd
@@
 ```{r model function}
-modelfunction <- function(x, rda.abund, df.meta) {
+modelfunction <- function(x, rda.abund, df.meta, modelstr = "rda.abund ~ %s + Condition(%s)") {
     covariates <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
@@
                    "BL_USE_RX_C08", "BL_USE_RX_C09")
-    fo <- sprintf("rda.abund ~ %s + Condition(%s)", x, paste0(covariates, collapse = "+"))
+    fo <- sprintf(modelstr, x, paste0(covariates, collapse = "+"))
     model <- rda(as.formula(fo), data = df.meta)
@@
 
+## Non-partial RDA species
+
+```{r models species}
+ret.species <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
+              modelfunction,
+              rda.abund = rda.abundances.species,
+              df.meta = meta(pseq.species),
+              modelstr = "rda.abund ~ %s + %s",
+              mc.cores = 5)
+returnparser(ret.species)
+```
+
 ```{r save session}

### 2019-05-14 articletwo.Rmd
Better salt analysis

@@
 library(microbiome)
-library(finriskmetagcommon)
 library(knitr)
@@
 library(reshape)
-library(ggpmisc)
 library(Maaslin)
@@
 library(flextable)
+library(xtable)
 library(rvg)
@@
 library(RColorBrewer)
+library(gvlma)
 ```
@@
 }
-```
-
-```{r core taxa}
-pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
 ```
@@
 
+Load genus data with dU-NA and combine it in a data frame with clr transformed lactobacillus abundances.
+
 ```{r lacto data frame}
-lacto.abund <- subset_taxa(pseq.genus, Genus == "Lactobacillus") %>%
+pseq.genus.salt <- import_salt_pseq("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+
+abund.genus.salt <- subset_taxa(pseq.genus.salt, Genus == "Lactobacillus") %>%
     abundances(., transform = "clr") %>%
@@
 
-lacto.full <- meta(pseq.genus) %>%
+df.genus.salt <- meta(pseq.genus.salt) %>%
     tibble::rownames_to_column("id") %>%
-    full_join(., lacto.abund, by = "id") %>%
-    dplyr::mutate(fake_salt = relevel(factor(dplyr::ntile(BL_AGE, 4)), ref='1'),
-                  lacto_groups = relevel(factor(dplyr::ntile(lactobacillus.bacteria, 4)), ref='1'))
+    full_join(., abund.genus.salt, by = "id") %>%
+    dplyr::mutate(salt_groups = relevel(factor(dplyr::ntile(dUNA, 3)), ref='1'),
+                  lacto_groups = relevel(factor(dplyr::ntile(lactobacillus.bacteria, 3)), ref='1'))
 ```
 
+### Daily urine natrium
+
+Associations between lactobacillus and dU-NA
+
+```{r abund salt lm continuous}
+lm.lacto.duna <- lm(lactobacillus.bacteria ~ dUNA + BL_AGE + SEX, data = df.genus.salt)
+lm.lacto.duna %>% tidy %>% kable
+```
+
+### Salt groups
+
+Association between lactobacillus and salt groups
+
+```{r abund salt lm group}
+lm.lacto.saltgroup <- lm(lactobacillus.bacteria ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
+lm.lacto.saltgroup %>% tidy %>% kable
+```
+
+Assesment for linear model asuumptions
+
+```{r abund salt lm group checks plot, fig.width=12, fig.height=8}
+par(mfrow=c(3,3))
+plot(gvlma::gvlma(lm.lacto.saltgroup), onepage = FALSE)
+```
+
+Anova between lactobacillus and salt groups
+
+```{r salt anova}
+aov.lacto.saltgroup <- aov(lactobacillus.bacteria ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
+aov.lacto.saltgroup %>% xtable %>% kable
+```
+
+Tukey's range test for lactobacillus and salt groups
+
+```{r salt HSD}
+hsd.lacto.saltgroup <- TukeyHSD(aov.lacto.saltgroup, "salt_groups")
+hsd.lacto.saltgroup$salt_groups %>% kable
+```
+
+### Adjusted means
+
+Calculate adjusted means for lactobacillus
+
 ```{r lacto adjusted means}
-lacto.lm.salt <- lm(lactobacillus.bacteria ~ fake_salt + BL_AGE + SEX, data=lacto.full)
+df.genus.salt.unadjusted <- with(df.genus.salt, data.frame(salt_groups = as.factor(seq(3)),
+                                                           SEX = as.factor(0),
+                                                           BL_AGE = mean(BL_AGE)))
 
-lacto.mean.unadjusted <- data.frame(
-    fake_salt = as.factor(seq(4)),
-    SEX = as.factor(0),
-    BL_AGE = mean(lacto.full %>% pull(BL_AGE)))
-
-lacto.mean.adjusted <- predict(lacto.lm.salt, lacto.mean.unadjusted, interval = "confidence") %>%
+adjmean.lacto.saltgroup <- predict(lm.lacto.saltgroup,
+                               df.genus.salt.unadjusted,
+                               interval = "confidence") %>%
     as.data.frame %>%
@@
 
-```{r lacto odds for hypertension}
-lacto.odds <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
+### Odds for hypertension
+
+Odds for hypertension
+
+```{r glm htn lacto}
+glm.hypertension.lactogroup <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
                   family = binomial(link=logit),
-                  data=lacto.full) %>%
-    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
+                  data = df.genus.salt) %>%
+    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
+glm.hypertension.lactogroup %>% kable
+```
+
+```{r odds htn lacto}
+odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
     filter(!grepl("Intercept", term)) %>%
@@
 
-```{r lacto lm residuals}
-ggplot2::autoplot(lm(lactobacillus.bacteria ~ fake_salt + BL_AGE + SEX, data=lacto.full))
-```
-
-```{r lacto logistic residuals}
-ggplot2::autoplot(glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX, family = binomial(link=logit), data=lacto.full))
-```
-
-```{r lacto anova}
-lacto.anova <- aov(lactobacillus.bacteria ~ fake_salt, data = lacto.full)
-summary(lacto.anova)
-TukeyHSD(lacto.anova)$fake_salt %>% kable
-```
-
-```{r lactoto test normality}
-grid.arrange(ggqqplot(lacto.full$lactobacillus.bacteria, title = "Lactobacillus"),
-             ggqqplot(lacto.full$BL_AGE, title = "BL_AGE"), ncol=2)
-```
+### Twin-forest plot
 
 ```{r forest plot lacto}
-myforestplot <- function(data, xlab = "", ylab = "", ylimits = c(0,10), interceptone = FALSE, subtitle = "") {
-    ggplot(data = data,
-           aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
-        geom_point() +
-        { if (interceptone) geom_hline(aes(fill=fake_salt), yintercept=1, linetype=2) } +
-        geom_errorbar(width=0.1)+ 
-        ggplot2::theme_minimal()  +
-        theme(plot.title=element_text(size=16),
-              axis.title=element_text(size=12),
-              strip.background = element_blank()) +
-        theme(panel.border = element_blank(),
-              panel.grid.major = element_blank(),
-              panel.grid.minor = element_blank(),
-              axis.ticks.x = element_line(),
-              axis.ticks.y = element_line(),
-              axis.line = element_line(colour = "black")) +
-        scale_x_discrete(labels=c("Q1", "Q2", "Q3", "Q4")) +
-        scale_y_continuous(limits = ylimits, breaks = pretty(ylimits, 4)) + 
-        xlab(xlab) +
-        ylab(ylab) +
-        labs(subtitle = subtitle)
-}
-
 gsalt <- grid.arrange(
-    myforestplot(lacto.mean.adjusted,
+    myforestplot(adjmean.lacto.saltgroup,
                  xlab = "Dietary salt",
@@
                  subtitle = "A.",
-                 ylimits = c(2.5,3)),
-    myforestplot(lacto.odds,
+                 my_y_scale = scale_y_continuous(limits = c(2.5,3), breaks = seq(2.5, 3.0, 0.1))),
+    myforestplot(odds.hypertension.lactogroup,
                  xlab = "Lactobacillus abundance",
@@
                  subtitle = "B.",
-                 ylimits = c(0.5, 1.5),
+                 my_y_scale = scale_y_log10(limits = c(0.5, 2), breaks = seq(0.5, 2.0, 0.5)),
                  interceptone = TRUE),
     ncol=2)
+
 ggsave(file = "cache/salt-forest.png", plot=gsalt, height=4, width=6, dpi = 300)

### 2019-05-14 articletwo-ggplot.R

@@ -77,3 +77,31 @@ plot.beta <- function(diversity, fig.permar2.text = "R2 of variable \nin Bray-Cu
                   xaxis = plot.gtable.beta.diversity$grobs[[7]],
                   xlab = plot.xlab.beta.diversity))
 }
+
+myforestplot <- function(data,
+                         xlab = "",
+                         ylab = "",
+                         interceptone = FALSE,
+                         subtitle = "",
+                         my_y_scale = scale_y_continuous()) {
+    ggplot(data = data,
+           aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+        geom_point() +
+        { if (interceptone) geom_hline(aes(fill=fake_salt), yintercept=1, linetype=2) } +
+        geom_errorbar(width=0.1)+ 
+        ggplot2::theme_minimal()  +
+        theme(plot.title=element_text(size=16),
+              axis.title=element_text(size=12),
+              strip.background = element_blank()) +
+        theme(panel.border = element_blank(),
+              panel.grid.major = element_blank(),
+              panel.grid.minor = element_blank(),
+              axis.ticks.x = element_line(),
+              axis.ticks.y = element_line(),
+              axis.line = element_line(colour = "black")) +
+        scale_x_discrete(labels=c("Q1", "Q2", "Q3", "Q4")) +
+        my_y_scale + 
+        xlab(xlab) +
+        ylab(ylab) +
+        labs(subtitle = subtitle)
+}

### 2019-05-14 articletwo-rrbiome.R

@@ -82,6 +82,25 @@ calculateglm <- function(dset,
                       conf.high = estimate + qnorm(1- 0.05/2) * std.error)
 }
 
+import_salt_pseq <- function(file = "data/phfinrisk_genus_all_drop50k_2018-11-16.RDs") {
+    pseq.salt <- readRDS(file)
+
+    pseq.salt.meta <- meta(pseq.salt) %>%
+        tibble::rownames_to_column(var = "rowname") %>%
+        mutate(SEX = factor(ifelse(MEN == "Female", 1, 0)),
+               ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
+                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
+               HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
+               dUNA = myscale(NA.)) %>%
+        select("rowname", "BL_AGE", "SEX", "dUNA", "HYPERTENSION") %>%
+        na.omit %>%
+        tibble::remove_rownames() %>%
+        tibble::column_to_rownames(var = "rowname")
+    
+    phyloseq::sample_data(pseq.salt) <- phyloseq::sample_data(pseq.salt.meta)
+    return(pseq.salt)
+}
+
 calculateadonis <- function(dset,
                             matrix,
                             responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",

### 2019-05-14 articletwo.Rmd
GVLMA

@@
 
+```{r abund salt lm continuous checks table}
+gvlma.table.plot(lm.lacto.duna)$table %>% kable
+```
+
+```{r abund salt lm continuous checks plot, fig.width=12, fig.height=8}
+gvlma.table.plot(lm.lacto.duna)$plot %>% plot
+```
+
 ### Salt groups
@@
 
+```{r abund salt lm group checks table}
+gvlma.table.plot(lm.lacto.saltgroup)$table %>% kable
+```
+
 ```{r abund salt lm group checks plot, fig.width=12, fig.height=8}
-par(mfrow=c(3,3))
-plot(gvlma::gvlma(lm.lacto.saltgroup), onepage = FALSE)
+gvlma.table.plot(lm.lacto.saltgroup)$plot %>% plot
 ```

### 2019-05-14 articletwo-ggplot.R

@@ -105,3 +105,14 @@ myforestplot <- function(data,
         ylab(ylab) +
         labs(subtitle = subtitle)
 }
+
+
+gvlma.table.plot <- function(model) {
+    { sink("/dev/null");
+        gvlma.model <- gvlma::gvlma(model)
+        gvlma.modek.summary <- summary(gvlma.model)
+        ; sink(); }
+    ret <- list("table" = gvlma.modek.summary,
+                 "plot" = gvlma.model)
+    return(ret)
+}

### 2019-05-14 articletwo.Rmd
Genus core missing, cleaner import function

@@
 ```{r data}
-pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
-pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
+pseq.genus <- import_filter_data("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
 ```
 
-Updating phyloseq objects
-
-```{r updating phyloseq objects}
-phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(filter.phenotype.data(pseq.genus))
-phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(filter.phenotype.data(pseq.species))
-```
-
-Sample data has dimensions (`r dim(sample_data(pseq.species))`).
+At species level meta has dimensions (`r dim(meta(pseq.species))`) and
+there ntaxa is `r ntaxa(pseq.species)`, and at genus level meta has
+dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is `r
+ntaxa(pseq.genus)`.
 
@@
 }
+```
+
+```{r core taxa}
+pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
 ```

### 2019-05-14 articletwo-rrbiome.R

@@ -82,6 +82,12 @@ calculateglm <- function(dset,
                       conf.high = estimate + qnorm(1- 0.05/2) * std.error)
 }
 
+import_filter_data <- function(file) {
+    pseq.full <- readRDS(file)
+    phyloseq::sample_data(pseq.full) <- phyloseq::sample_data(filter.phenotype.data(pseq.full))
+    return(pseq.full)
+}
+
 import_salt_pseq <- function(file = "data/phfinrisk_genus_all_drop50k_2018-11-16.RDs") {
     pseq.salt <- readRDS(file)
 

### 2019-05-15 partialrda.Rmd
Support for non-partial models

--- 2019-05-14 partialrda.Rmd
@@
 ```{r data}
-pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
-pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
-```
-
-Updating phyloseq objects
-
-```{r updating phyloseq objects}
-phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(filter.phenotype.data(pseq.genus))
-phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(filter.phenotype.data(pseq.species))
+pseq.genus <- import_filter_data("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
 pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
@@
 
-```{r transformations}
-rda.abundances.species  <- microbiome::transform(pseq.species, "clr") %>%
+```{r abundances to matrix function}
+transform.matrix <- function(pseq, transformation = "clr") {
+    microbiome::transform(pseq, transformation) %>%
         microbiome::abundances() %>%
@@
             as.matrix()
+ }
+```
 
-rda.abundances.genus  <- microbiome::transform(pseq.genus, "clr") %>%
-    microbiome::abundances() %>%
-    t() %>%
-    as.matrix()
-
-rda.abundances.genus.core  <- microbiome::transform(pseq.genus.core, "clr") %>%
-    microbiome::abundances() %>%
-    t() %>%
-    as.matrix()
+```{r transformations}
+rda.abundances.species  <- transform.matrix(pseq.species)
+rda.abundances.genus  <- transform.matrix(pseq.genus)
+rda.abundances.genus.core  <- transform.matrix(pseq.genus.core)
 ```
@@
 ```{r model function}
-modelfunction <- function(x, rda.abund, df.meta) {
+modelfunction <- function(x,
+                          rda.abund,
+                          df.meta,
+                          modelstr = "rda.abund ~ %s + Condition(%s)",
+                          permutations = 1000) {
     covariates <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
@@
                    "BL_USE_RX_C08", "BL_USE_RX_C09")
-    fo <- sprintf("rda.abund ~ %s + Condition(%s)", x, paste0(covariates, collapse = "+"))
+    fo <- sprintf(modelstr, x, paste0(covariates, collapse = "+"))
     model <- rda(as.formula(fo), data = df.meta)
-    anova.term <- anova(model, by="term", perm=1000)
+    anova.term <- anova(model, by="term", perm=permutations)
     data.frame(var = as.character(x),
                beta = coef(model)[,1][[ifelse(x == "HYPERTENSION", "HYPERTENSION1", x)]],
-               RDA.Conditioned = summary(model)$partial.chi,
-               RDA.Constrained = summary(model)$constr.chi,
-               RDA.Unconstrained = summary(model)$unconst.chi,
+               conditioned = ifelse("partial.chi" %in% names(summary(model)),
+                                        summary(model)$partial.chi,
+                                        0),
+               constrained = summary(model)$constr.chi,
+               unconstrained = summary(model)$unconst.chi,
                 p.anova = anova.term[["Pr(>F)"]][1],
@@
     res <- do.call(rbind, ret) %>%
-        mutate(RDA.total =  RDA.Conditioned + RDA.Constrained + RDA.Unconstrained,
-               RDA.Conditioned.percent = RDA.Conditioned/RDA.total,
-               RDA.Constrained.percent = RDA.Constrained/RDA.total,
-               RDA.Unconstrained.percent = RDA.Unconstrained/RDA.total
+        mutate(RDA.total =  conditioned + constrained + unconstrained,
+               conditioned = conditioned/RDA.total,
+               constrained = constrained/RDA.total,
+               unconstrained = unconstrained/RDA.total
                )
@@
                       p.anova,
-                      RDA.Conditioned.percent,
-                      RDA.Constrained.percent,
-                      RDA.Unconstrained.percent) %>%
+                      conditioned,
+                      constrained,
+                      unconstrained) %>%
         kable()
@@
 
+## Non-partial RDA species
+
+```{r species non condition covariatesmodels test, eval = FALSE}
+rda.subset <- rda.abundances.species[1:100, 1:100]
+meta.subset <- meta(pseq.species) %>%
+    tibble::rownames_to_column("id") %>%
+    dplyr::filter(id %in% rownames(rda.subset))
+
+ret.species <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
+              modelfunction,
+              rda.abund = rda.subset,
+              df.meta = meta.subset,
+              modelstr = "rda.abund ~ %s + %s",
+              mc.cores = 5)
+
+returnparser(ret.species)
+```
+
 ```{r save session}

### 2019-05-15 articletwo.Rmd
diab predictor missing as.factor

@@
                Q57X = as.factor(1),
-               DIAB = 0,
+               DIAB = as.factor(0),
                BL_USE_RX_C03 = as.factor(0),

### 2019-05-15 articletwo.Rmd
Bigger plots

@@
 
-```{r abund salt lm continuous checks plot, fig.width=12, fig.height=8}
+```{r abund salt lm continuous checks plot, fig.width=12, fig.height=14}
 gvlma.table.plot(lm.lacto.duna)$plot %>% plot
@@
 
-```{r abund salt lm group checks plot, fig.width=12, fig.height=8}
+```{r abund salt lm group checks plot, fig.width=12, fig.height=14}
 gvlma.table.plot(lm.lacto.saltgroup)$plot %>% plot

### 2019-05-15 articletwo.Rmd
More salt models

@@
 
+```{r check adonis, eval = FALSE}
+ad.min <- adonis(bray.dist.m.species ~ DIASM + BL_AGE + SEX,
+             data = meta(pseq.species),
+             permutations = 100)
+
+ad.max <- adonis(bray.dist.m.species ~ DIASM + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
+                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09,
+             data = meta(pseq.species),
+             permutations = 1000)
+
+```
+
 ```{r pcoa calculate}
@@
 
+```{r difference in beta diversity}
+merge(diversities$min %>% select(response, R2),
+      diversities$max %>% select(response, R2), by = "response") %>%
+    mutate(R2.diff = R2.y - R2.x) %>%
+    kable
+```
+
 ```{r save grob}
@@
 
-pcoa.df %>% pull(STATETREATMENT)
-
 pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.2, color = STATETREATMENT, shape = STATETREATMENT)) +
@@
     arrange(Feature)
+
+maas.results.df.signf %>% pull(Feature) %>% sort %>% unique %>% paste(., collapse=", ")
 
@@
 
+Characteristics in salt analysis
+
+```{r salt characteristics}
+
+pseq.genus.salt %>%
+    meta %>%
+    gather(key,value) %>%
+    group_by(key) %>%
+    summarise_all(funs(N = n(),
+                       Mean = mean(as.numeric(value), na.rm = TRUE),
+                       sd = sd(as.numeric(value), na.rm = TRUE),
+                       NAs = sum(is.na(value))))  %>%
+    kable()
+
+```
+
 ### Daily urine natrium
@@
 
+```{r equal variance}
+bartlett.test(lactobacillus.bacteria ~ salt_groups, data = df.genus.salt)
+```
+
+```{r one way test salt groups}
+oneway.test(lactobacillus.bacteria ~ salt_groups, data = df.genus.salt, var.equal = TRUE)
+```
+
 Anova between lactobacillus and salt groups
@@
                   conf.high = upr) %>%
-    tibble::rownames_to_column("term")
+    rownames::tibble_to_column("term")
 ```
@@
 
-```{r glm htn lacto}
+```{r glm htn lacto groups}
 glm.hypertension.lactogroup <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
@@
 glm.hypertension.lactogroup %>% kable
+```
+
+```{r glm htn lacto continuous}
+glm.hypertension.lactocontinuous <- glm(HYPERTENSION ~ lactobacillus.bacteria + BL_AGE + SEX,
+                  family = binomial(link=logit),
+                  data = df.genus.salt) %>%
+    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
+glm.hypertension.lactocontinuous %>% kable
+```
+
+Daily urine sodium and BP
+
+```{r dUNA RR}
+calculateglm(df.genus.salt, modelstring = "%s ~ dUNA + BL_AGE + SEX", filterstr = "dUNA") %>%
+    select(response, term, estimate, p.value) %>%
+    kable
+```
+
+Daily urine sodium and BP
+
+```{r lactobacillus RR}
+calculateglm(df.genus.salt,
+             modelstring = "%s ~ lactobacillus.bacteria + BL_AGE + SEX",
+             filterstr = "lactobacillus") %>%
+    select(response, term, estimate, p.value) %>%
+    kable
+```
+
+Assesment for linear model asumptions
+
+```{r table map lactobacillus}
+gvlma.table.plot(lm(log(MAP) ~ lactobacillus.bacteria + BL_AGE + SEX, data = df.genus.salt))$table %>% kable
+```
+
+```{r plot map lactobacillus, fig.width=12, fig.height=14}
+gvlma.table.plot(lm(MAP ~ lactobacillus.bacteria + BL_AGE + SEX, data = df.genus.salt))$plot %>% plot
 ```

### 2019-05-15 articletwo-rrbiome.R

@@ -97,8 +97,15 @@ import_salt_pseq <- function(file = "data/phfinrisk_genus_all_drop50k_2018-11-16
                ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
                                        BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
                HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
+               PULSEPRESSURE = SYSTM - DIASM,
+               MAP = 2./3.*DIASM + 1./3.*SYSTM,
+               SYSTM = myscale(SYSTM),
+               DIASM = myscale(DIASM),
+               PULSEPRESSURE = myscale(PULSEPRESSURE),
+               MAP = myscale(MAP),
                dUNA = myscale(NA.)) %>%
-        select("rowname", "BL_AGE", "SEX", "dUNA", "HYPERTENSION") %>%
+        select("rowname", "BL_AGE", "SEX", "dUNA", "HYPERTENSION", "SYSTM",
+               "DIASM", "PULSEPRESSURE", "MAP") %>%
         na.omit %>%
         tibble::remove_rownames() %>%
         tibble::column_to_rownames(var = "rowname")

### 2019-05-15 articletwo.Rmd
Typo

@@
                   conf.high = upr) %>%
-    rownames::tibble_to_column("term")
+    tibble::rownames_to_column("term")
 ```
@@
 ```{r table map lactobacillus}
-gvlma.table.plot(lm(log(MAP) ~ lactobacillus.bacteria + BL_AGE + SEX, data = df.genus.salt))$table %>% kable
+gvlma.table.plot(lm(MAP ~ lactobacillus.bacteria + BL_AGE + SEX, data = df.genus.salt))$table %>% kable
 ```

### 2019-05-15 comparepseqs.Rmd
Comparing different pseq object

---
title: "Compare pseqs"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis', cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

dir.create("rds/", showWarnings = FALSE)
dir.create("session/", showWarnings = FALSE)
```

```{r Command line arguments}
now <- format(Sys.time(), '%Y%m%d-%H%M%S')

if (exists("args")) {
    if ("time" %in% names(args)) now <- args$time

    if("clean" %in% names(args)) {
        unlink("rds", recursive=TRUE)
        unlink("cache", recursive=TRUE)
    }
    if ("loadlast" %in% names(args) )
        load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
}
```

# Importing libraries:

```{r libraries}
library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(parallel)
library(knitr)
```

### RR biome functions

<details>
  <summary>Open/Close</summary>

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```

</details>

# Loading data

 phyloseq-objektit: 1) SHOGUN 2) CENTRIFUGE 3) NCBI70 4) GTDB 5) 16S

```{r data}
pseq = list()
pseq$shogun <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
pseq$R16S <- import_filter_data("data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
pseq$gtdb <- import_filter_data("data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs")
pseq$ncbi <- import_filter_data("data/phfinrisk_custom70_species_all_2019-01-29.RDs")
#pseq$centrifuge <- import_filter_data("data/phfinrisk_centrifuge_species_all_2018-11-30.RDs")
```

```{r alpha diversities}
dset <- mclapply(pseq, function(x) meta.merge.alphadiversity(x) %>%
                                   mutate(diversities_shannon = scale(diversities_shannon)),
                 mc.cores = 4)

fo.alpha.min <- "%s ~ BL_AGE + SEX  + diversities_shannon"

alpha.min <- mclapply(dset, function(x) calculateglm(x, modelstring = fo.alpha.min),
                 mc.cores = 4)

fo.alpha.full <- "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + diversities_shannon"

alpha.full <- mclapply(dset, function(x) calculateglm(x, modelstring = fo.alpha.full),
                 mc.cores = 4)
```

```{r join alphas}
sfx <- names(dset)
res <- alpha.full[[1]]

for(i in head(seq_along(alpha.full), -1)) {
    res <- merge(res,
                 alpha.full[[i+1]],
                 all = TRUE,
                 suffixes = paste0("_", sfx[i:(i+1)]),
                 by = c("model_name", "term", "response"))
}

res %>%
    dplyr::select(response, starts_with("estimate_")) %>%
    kable
```

```{r save session}
save.image(file = paste0("session/comparepseqs.Rdata"))
```

### 2019-05-15 comparepseqs.Rmd
Bray matrix

@@
 
+```{r matrix calculation}
+if (!file.exists("rds/comparepseqs_bray_matrix.rds")) {
+    bray.dist.matrix <- mclapply(pseq, function(x) calculate.beta.matrix(x), mc.cores = 4)
+    saveRDS(bray.dist.matrix, file = "rds/comparepseqs_bray_matrix.rds")
+} else {
+    bray.dist.matrix  <- readRDS("rds/comparepseqs_bray_matrix.rds")
+}
+```
+
+ ```{r adonis calculation, eval = FALSE}
+ if (!file.exists("rds/adonis.species.rds")) {
+     adonis.species = list()
+     adonis.species$min <- calculateadonis(dset = meta(pseq.species),
+                                           matrix = bray.dist.m.species,
+                                           covariates = c("BL_AGE", "SEX"),
+                                           npermutations = 1000)
+     adonis.species$max <- calculateadonis(dset = meta(pseq.species),
+                                           matrix = bray.dist.m.species,
+                                           covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
+                                                          "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
+                                           npermutations = 1000)
+     saveRDS(adonis.species, file = "rds/adonis.species.rds")
+    } else {
+     adonis.species  <- readRDS("rds/adonis.species.rds")
+}
+```
+
 ```{r save session}

### 2019-05-15 comparepseqs.Rmd
Different rds location

@@
 
-dir.create("rds/", showWarnings = FALSE)
+dir.create("rds-comparepseqs/", showWarnings = FALSE)
 dir.create("session/", showWarnings = FALSE)
@@
     if("clean" %in% names(args)) {
-        unlink("rds", recursive=TRUE)
+        unlink("rds-comparepseqs", recursive=TRUE)
         unlink("cache", recursive=TRUE)
@@
 ```{r matrix calculation}
-if (!file.exists("rds/comparepseqs_bray_matrix.rds")) {
+if (!file.exists("rds-comparepseqs/comparepseqs_bray_matrix.rds")) {
     bray.dist.matrix <- mclapply(pseq, function(x) calculate.beta.matrix(x), mc.cores = 4)
-    saveRDS(bray.dist.matrix, file = "rds/comparepseqs_bray_matrix.rds")
+    saveRDS(bray.dist.matrix, file = "rds-comparepseqs/comparepseqs_bray_matrix.rds")
 } else {
-    bray.dist.matrix  <- readRDS("rds/comparepseqs_bray_matrix.rds")
+    bray.dist.matrix  <- readRDS("rds-comparepseqs/comparepseqs_bray_matrix.rds")
 }

### 2019-05-16 comparepseqs.Rmd
All five pseqs added when DIAB is excluded
Beta diversity

--- 2019-05-15 comparepseqs.Rmd
@@
 library(knitr)
+library(gridExtra)
 ```
@@
 source("articletwo-rrbiome.R")
+```
+
+</details>
+
+### Plot functions
+
+<details>
+  <summary>Open/Close</summary>
+
+```{r Plot functions, code=readLines("articletwo-ggplot.R")}
+source("articletwo-ggplot.R")
 ```
@@
 pseq$ncbi <- import_filter_data("data/phfinrisk_custom70_species_all_2019-01-29.RDs")
-#pseq$centrifuge <- import_filter_data("data/phfinrisk_centrifuge_species_all_2018-11-30.RDs")
+pseq$centrifuge <- import_filter_data("data/phfinrisk_centrifuge_species_all_2018-11-30.RDs",
+                                      excluded = c("DIAB"))
 ```
@@
                                              mutate(diversities_shannon = scale(diversities_shannon)),
-                 mc.cores = 4)
+                           mc.cores = length(pseq))
 
 fo.alpha.min <- "%s ~ BL_AGE + SEX  + diversities_shannon"
+fo.alpha.full <- "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + diversities_shannon" # MISSING DIAB
 
-alpha.min <- mclapply(dset, function(x) calculateglm(x, modelstring = fo.alpha.min),
-                 mc.cores = 4)
-
-fo.alpha.full <- "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + diversities_shannon"
-
-alpha.full <- mclapply(dset, function(x) calculateglm(x, modelstring = fo.alpha.full),
-                 mc.cores = 4)
+alphadiversity <-  mclapply(dset, function(x)
+    list(min =  calculateglm(x, modelstring = fo.alpha.min),
+         full = calculateglm(x, modelstring = fo.alpha.full)),
+    mc.cores = length(pseq))
 ```
 
-```{r join alphas}
+```{r dissimilarity matrix}
+if (!file.exists("rds-comparepseqs/comparepseqs_bray_matrix.rds")) {
+    bray.dist.matrix <- mclapply(pseq, function(x) calculate.beta.matrix(x), mc.cores = length(pseq))
+    saveRDS(bray.dist.matrix, file = "rds-comparepseqs/comparepseqs_bray_matrix.rds")
+} else {
+    bray.dist.matrix  <- readRDS("rds-comparepseqs/comparepseqs_bray_matrix.rds")
+}
+```
+
+ ```{r adonis calculation}
+ if (!file.exists("rds-comparepseqs/beta.diversity.rds")) {
+     covariates.min <- c("BL_AGE", "SEX")
+     covariates.full <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE",
+                          "Q57X", "BL_USE_RX_C03", "BL_USE_RX_C07",
+                          "BL_USE_RX_C08", "BL_USE_RX_C09")
+
+     betadiversity <- mclapply(c2l(names(pseq)), function(x)
+         list(min = calculateadonis(dset = meta(pseq[[x]]),
+                                    matrix = bray.dist.matrix[[x]],
+                                    covariates = covariates.min,
+                                    npermutations = 1000),
+              full = calculateadonis(dset = meta(pseq[[x]]),
+                                           matrix = bray.dist.matrix[[x]],
+                                           covariates = covariates.full
+                                           npermutations = 1000)
+         mc.cores = length(pseq))
+     saveRDS(betadiversity, file = "rds-comparepseqs/beta.diversity.rds")
+    } else {
+     betadiversity <- readRDS("rds-comparepseqs/beta.diversity.rds")
+}
+```
+
+```{r plot alpha diversity, eval = FALSE}
+beta.min <- lapply(c2l(names(pseq)), function(x) {
+    alpha.min[[x]] %>% mutate(R2 = 0) %>% select(model_name, R2, response)
+})
+
+beta.min[[1]] %>%
+         data.table::rbindlist(id = "model_name")
+
+#            dplyr::filter(term %in% responses) %>%
+            dplyr::select(response, R2)
+
+diversity.min <- mergediversities(alpha.min, beta.min)
+
+alpha.plot.min <- mclapply(alpha.min, function(x) { mergediversities(x, 
+                               },
+                               ## plot.alpha(x %>%
+                               ##                    mutate(Name = term,
+                               ##                           R2 = 0,
+                               ##                           Qstar = ifelse(p.value < 0.05, '*', ' ')),
+                               ##                    colorlimit = 0.6),
+                           mc.cores = length(pseq))
+
+gs4 <- c(list("yaxis" = alpha.plot.min[[1]]$yaxis,
+              "xaxis" = alpha.plot.min[[1]]$xaxis),
+              lapply(alpha.plot.min, function(x) x$plot))
+
+g4.firstline <- c(1, seq(3, length(alpha.plot.min)+2))
+g4.secondline <- c(NA, rep(2, length(alpha.plot.min)))
+
+g4 <- arrangeGrob(grobs = gs4, layout_matrix = rbind(g4.firstline, g4.secondline))
+
+ggsave(file = "cache/pseqdiff-alpha-min.png", plot=g4, height=6, width=9)
+
+```
+
+```{r join alphas, eval = FALSE}
 sfx <- names(dset)
-res <- alpha.full[[1]]
+res <- alpha.min[[1]]
 
-for(i in head(seq_along(alpha.full), -1)) {
+for(i in head(seq_along(alpha.min), -1)) {
     res <- merge(res,
@@
 
+res %>% colnames
+
 res %>%
-    dplyr::select(response, starts_with("estimate_")) %>%
-    kable
+    dplyr::select(response, starts_with("estimate_"), starts_with("p.value"))
+
 ```

### 2019-05-16 articletwo-ggplot.R

@@ -46,7 +46,9 @@ plot.alpha <- function(diversity,  colorlimit = 0.25) {
                 legend = plot.alpha.diversity.legend))
 }
 
-plot.beta <- function(diversity, fig.permar2.text = "R2 of variable \nin Bray-Curtis distance", fig.textsize = 15) {
+plot.beta <- function(diversity,
+                      fig.permar2.text = "R2 of variable \nin Bray-Curtis distance",
+                      fig.textsize = 15) {
     plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2)) +
         geom_bar(stat="identity", color="black") +
         coord_flip() +

### 2019-05-16 articletwo-rrbiome.R

@@ -13,12 +13,14 @@ myscale <- function(x){
 }
 
 filter.phenotype.data <- function(pseq,
-                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
-                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
-                                               "BL_USE_RX_C07", "BL_USE_RX_C08")) {
+                                  included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
+                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X",
+                                               "BL_USE_RX_C03",
+                                               "BL_USE_RX_C07", "BL_USE_RX_C08"),
+                                  excluded) {
     meta(pseq) %>%
         tibble::rownames_to_column(var = "rowname") %>%
-        dplyr::select(rowname, included) %>%
+        dplyr::select(rowname, included[!included %in% excluded]) %>%
         stats::na.omit() %>%
         dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
                                        BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
@@ -35,14 +37,14 @@ filter.phenotype.data <- function(pseq,
                       DIASM = myscale(DIASM),
                       PULSEPRESSURE = myscale(PULSEPRESSURE),
                       MAP = myscale(MAP),
-                      DIAB = as.factor(DIAB),
+#                      DIAB = ifelse("DIAB" %in% excluded, NA, as.factor(DIAB)),
                       SEX = as.factor(SEX),
                       Q57X = as.factor(Q57X),
                       BL_USE_RX_C03 = as.factor(BL_USE_RX_C03),
                       BL_USE_RX_C07 = as.factor(BL_USE_RX_C07),
                       BL_USE_RX_C08 = as.factor(BL_USE_RX_C08),
                       BL_USE_RX_C09 = as.factor(BL_USE_RX_C09)) %>%
-        dplyr::select(-MEN, -DNA_OK) %>%
+        dplyr::select(-MEN) %>%
         tibble::remove_rownames() %>%
         tibble::column_to_rownames(var = "rowname")
 }
[…]

### 2019-05-16 articletwo.Rmd
Cleaning salt results

--- 2019-05-15 articletwo.Rmd
@@
 
-### Lactobacillus
+## Lactobacillus (figure 4)
 
-Load genus data with dU-NA and combine it in a data frame with clr transformed lactobacillus abundances.
+The fourth and final step is to study how the amount of dietary salt
+affect the lactobacillus abundances and how these abundances affect
+the odds for hypertension. We load genus data with dU-NA and combine
+it in a data frame with CLR transformed lactobacillus abundances.
 
@@
 
-### Daily urine natrium
+Characteristics in salt analysis. Analyses are limited by the number
+of participants with collected dU-NA.
 
-Associations between lactobacillus and dU-NA
+```{r salt characteristics}
+pseq.genus.salt %>%
+    meta %>%
+    gather(key,value) %>%
+    group_by(key) %>%
+    summarise_all(funs(N = n(),
+                       Mean = mean(as.numeric(value), na.rm = TRUE),
+                       sd = sd(as.numeric(value), na.rm = TRUE),
+                       NAs = sum(is.na(value))))  %>%
+    mutate_if(is.numeric, round, 2) %>%
+    kable()
+```
+
+### How sodium affects lactobacillus abundances?
+
+Associations between continuous dU-NA and lactobacillus abundances.
 
@@
 lm.lacto.duna <- lm(lactobacillus.bacteria ~ dUNA + BL_AGE + SEX, data = df.genus.salt)
-lm.lacto.duna %>% tidy %>% kable
+lm.lacto.duna %>% tidy %>% mutate_if(is.numeric, round, 3) %>% kable
 ```
+
+Testing assumptions for linear regression analysis
 
@@
 
+Diagnostic plots for linear model
+
 ```{r abund salt lm continuous checks plot, fig.width=12, fig.height=14}
@@
 
-### Salt groups
+### Is there associations between salt groups and abudances?
 
-Association between lactobacillus and salt groups
+Associations between categorical dU-NA split in three equal size groups and lactobacillus abundances
 
@@
 lm.lacto.saltgroup <- lm(lactobacillus.bacteria ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
-lm.lacto.saltgroup %>% tidy %>% kable
+lm.lacto.saltgroup %>% tidy %>% mutate_if(is.numeric, round, 3) %>% kable
 ```
 
-Assesment for linear model asuumptions
+Testing assumptions for linear regression analysis
 
@@
 
+Diagnostic plots for linear model
+
 ```{r abund salt lm group checks plot, fig.width=12, fig.height=14}
@@
 
-Anova between lactobacillus and salt groups
+Bartlett's test for equality of variance gives p = `r
+bartlett.test(lactobacillus.bacteria ~ salt_groups, data =
+df.genus.salt)$p.value`.
+
+One-way test assuming normal distributions gives for equal mean p = `r
+oneway.test(lactobacillus.bacteria ~ salt_groups, data =
+df.genus.salt, var.equal = TRUE)$p.value`.
+
+Analysing also differences between salt groups using ANOVA.
 
@@
 
-Tukey's range test for lactobacillus and salt groups
+Tukey's range for pairwise test between salt groups
 
@@
 
-### Adjusted means
+### Adjusting means for forest plot
 
-Calculate adjusted means for lactobacillus
+Calculate adjusted means for lactobacillus assuming male sex and mean age.
 
@@
 
-### Odds for hypertension
+## Is there associations between lactobacillus abudance and hypertension?
 
-Odds for hypertension
+Studying associations between blood pressure variables and lactobacillus abundances
 
-```{r glm htn lacto}
-glm.hypertension.lactogroup <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
-                  family = binomial(link=logit),
-                  data = df.genus.salt) %>%
-    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
-glm.hypertension.lactogroup %>% kable
+```{r lactobacillus RR}
+calculateglm(df.genus.salt,
+             modelstring = "%s ~ lactobacillus.bacteria + BL_AGE + SEX",
+             filterstr = "lactobacillus") %>% 
+    select(response, term, estimate, std.error, p.value) %>%
+    mutate_if(is.numeric, round, 3) %>%
+    kable
+```
+
+Checking assumptions for linear regression for MAP
+
+```{r table map lactobacillus}
+gvlma.table.plot(lm(MAP ~ lactobacillus.bacteria + BL_AGE + SEX, data = df.genus.salt))$table %>% kable
+```
+
+Diagnostic plots for linear model
+
+```{r plot map lactobacillus, fig.width=12, fig.height=14}
+gvlma.table.plot(lm(MAP ~ lactobacillus.bacteria + BL_AGE + SEX, data = df.genus.salt))$plot %>% plot
 ```
@@
     dplyr::filter(grepl("lacto", term))
+odds.hypertension.lactogroup %>% kable
 ```
 
-### Twin-forest plot
+Odds for hypertension by categorical lactobacillus abundance.
+
+```{r glm htn lacto groups}
+glm.hypertension.lactogroup <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
+                  family = binomial(link=logit),
+                  data = df.genus.salt) %>%
+    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
+
+odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
+    filter(!grepl("Intercept", term)) %>%
+    select(term, estimate, conf.low, conf.high) %>%
+    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
+    dplyr::filter(grepl("lacto", term))
+
+glm.hypertension.lactogroup %>%
+    mutate_if(is.numeric, round, 3) %>%
+    kable
+```
+
+## Twin-forest plot
 

### 2019-05-16 articletwo.Rmd
Scatter plot

@@
 
-```{r Write characteristics}
+```{r Write characteristics, include=FALSE}
 tbl1head <- "Characteristics of the study sample."
@@
 
-```{r Write figure 1}
+```{r Write figure 1, include=FALSE}
 fig1head <- "Associations between BP variables and microbial diversity."
@@
 
-```{r Write figure 2}
+```{r Write figure 2, include=FALSE}
 fig2head <- "Principal coordinate analysis for bacterial abundances."
@@
                            scatter.plot(scatter.full, maas.results.df.signf, x$rr, x$abund))
+```
 
+```{r scatter plot at genus level for BP, fig.width=12, fig.height=20}
 scatterplot <- do.call("grid.arrange", c(scatterplots, ncol=4))
@@
 
-```{r Write figure 3}
+```{r Write figure 3, include=FALSE}
 fig3head <- "Associations between genus level abundances and blood pressure variables."
@@
 
-```{r Write figure 4}
+```{r Write figure 4, include=FALSE}
 fig4head <- "Associations between dietary salt, lactobacillus abundance, and odds for hypertension."
@@
 
-```{r Write docx to file}
+```{r Write docx to file, include=FALSE}
 print(doc, target = paste0("report/articletwo-", now, ".docx"))

### 2019-05-16 articletwo.Rmd
DIAB missing

@@
                   family = binomial(link=logit),
-                  data = df.genus.salt) %>%
-    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
+                  data = df.genus.salt)
 
 odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
+    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
     filter(!grepl("Intercept", term)) %>%
@@
 glm.hypertension.lactogroup %>%
+    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
     mutate_if(is.numeric, round, 3) %>%
     kable
+```
+
+```{r check odds model assumptions}
+g.odds.autoplot <- autoplot(glm.hypertension.lactogroup,
+                            which = 1:4,
+                            nrow = 1,
+                            ncol = 2,
+                            label.size = 3,
+                            size = 0.1) +
+    theme_minimal()
+ggsave(file = "cache/odds-autoplot.png", plot=g.odds.autoplot, height=10, width=6, dpi = 300)
+
 ```

### 2019-05-16 articletwo-rrbiome.R

@@ -37,7 +37,7 @@ filter.phenotype.data <- function(pseq,
                       DIASM = myscale(DIASM),
                       PULSEPRESSURE = myscale(PULSEPRESSURE),
                       MAP = myscale(MAP),
-#                      DIAB = ifelse("DIAB" %in% excluded, NA, as.factor(DIAB)),
+                      DIAB = ifelse("DIAB" %in% excluded, NA, as.factor(DIAB)),
                       SEX = as.factor(SEX),
                       Q57X = as.factor(Q57X),
                       BL_USE_RX_C03 = as.factor(BL_USE_RX_C03),

### 2019-05-16 articletwo.Rmd
Line change

@@
 ggsave(file = "cache/odds-autoplot.png", plot=g.odds.autoplot, height=10, width=6, dpi = 300)
-
 ```

### 2019-05-17 comparepseqs.Rmd
Missing commas

@@
 library(gridExtra)
+library(vegan)
 ```
@@
  ```{r adonis calculation}
+bray.dist.matrix %>% str
+
  if (!file.exists("rds-comparepseqs/beta.diversity.rds")) {
@@
                                            matrix = bray.dist.matrix[[x]],
-                                           covariates = covariates.full
-                                           npermutations = 1000)
-         mc.cores = length(pseq))
+                                           covariates = covariates.full,
+                                           npermutations = 1000),
+              mc.cores = length(pseq)))
      saveRDS(betadiversity, file = "rds-comparepseqs/beta.diversity.rds")
@@
 ```{r plot alpha diversity, eval = FALSE}
+
+betadiversity
 beta.min <- lapply(c2l(names(pseq)), function(x) {

### 2019-05-17 articletwo-rrbiome.R

@@ -273,3 +273,8 @@ firstup <- function(x) {
   substr(x, 1, 1) <- toupper(substr(x, 1, 1))
   x
 }
+
+mykable <- function(x, ...) {
+  capture.output(x <- print(x))
+  knitr::kable(x, ...)
+}

### 2019-05-17 comparepseqs.Rmd
Fixed adonis

--- 2019-05-16 comparepseqs.Rmd
@@
                                    covariates = covariates.min,
-                                    npermutations = 1000),
+                                   npermutations = 1),
              full = calculateadonis(dset = meta(pseq[[x]]),
                                     matrix = bray.dist.matrix[[x]],
-                                           covariates = covariates.full
-                                           npermutations = 1000)
+                                    covariates = covariates.full,
+                                    npermutations = 1)),
         mc.cores = length(pseq))
+
      saveRDS(betadiversity, file = "rds-comparepseqs/beta.diversity.rds")
@@
 ```{r plot alpha diversity, eval = FALSE}
+
+str(betadiversity)
+
+a <- mergediversities(alphadiversity[["ncbi"]], betadiversity[["ncbi"]])
+
+alphadiversity[["ncbi"]]
+betadiversity[["ncbi"]]
+
+names(betadiversity)
+betadiversity[[""]]
+
+ alpha.select <- alphadiversity[["R16S"]][["min"]] %>% select(estimate, p.value, response)
+
+beta.rbind <- betadiversity[["R16S"]][["min"]] %>%
+    data.table::rbindlist(id = "model_name")
+
+%>%
+            dplyr::filter(term %in% responses) %>%
+            dplyr::select(response, R2)
+
+        merge( alpha.select , beta.rbind, by = "response") %>%
+            merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate") %>%
+            mutate(Qstar = ifelse(p.value < 0.05, '*', ' '))
+    })
+
+vectortolist(names(alphadiversity[["R16S"]]))
+
+c2l(names(pseq))
 beta.min <- lapply(c2l(names(pseq)), function(x) {
-    alpha.min[[x]] %>% mutate(R2 = 0) %>% select(model_name, R2, response)
+    mergediversities(alphadiversity[[x]], betadiversity[[x]])
 })
@@
 
-alpha.plot.min <- mclapply(alpha.min, function(x) { mergediversities(x, 
-                               },
-                               ## plot.alpha(x %>%
-                               ##                    mutate(Name = term,
-                               ##                           R2 = 0,
-                               ##                           Qstar = ifelse(p.value < 0.05, '*', ' ')),
-                               ##                    colorlimit = 0.6),
+alpha.plot.min <- mclapply(alpha.min, function(x) { mergediversities(},
                                mc.cores = length(pseq))

### 2019-05-17 comparepseqs.Rmd
Perm 1000

@@
 if (!file.exists("rds-comparepseqs/beta.diversity.rds")) {
+
     covariates.min <- c("BL_AGE", "SEX")
@@
                                    covariates = covariates.min,
-                                   npermutations = 1),
+                                   npermutations = 1000),
              full = calculateadonis(dset = meta(pseq[[x]]),
@@
                                     covariates = covariates.full,
-                                    npermutations = 1)),
+                                    npermutations = 1000)),
         mc.cores = length(pseq))
@@
 ```{r plot alpha diversity, eval = FALSE}
-
-str(betadiversity)
 

### 2019-05-17 comparepseqs.Rmd
Less memory

@@
          full = calculateglm(x, modelstring = fo.alpha.full)),
-    mc.cores = length(pseq))
+    mc.cores = length(dset))
 ```
@@
 }
+rm(pseq)
 ```
@@
 
-    betadiversity <- mclapply(c2l(names(pseq)), function(x)
-        list(min = calculateadonis(dset = meta(pseq[[x]]),
+    betadiversity <- mclapply(c2l(names(dset)), function(x)
+        list(min = calculateadonis(dset = dset[[x]],
                                    matrix = bray.dist.matrix[[x]],
@@
                                    npermutations = 1000),
-             full = calculateadonis(dset = meta(pseq[[x]]),
+             full = calculateadonis(dset = dset[[x]],
                                     matrix = bray.dist.matrix[[x]],
@@
                                     npermutations = 1000)),
-        mc.cores = length(pseq))
+        mc.cores = length(dset))
 
@@
     }
+rm(bray.dist.matrix)
 ```

### 2019-05-20 comparepseqs.Rmd
Lapply to for

--- 2019-05-17 comparepseqs.Rmd
@@
 
+```{r packrat, include = FALSE, eval = FALSE}
+packrat::opts$ignored.packages(c("Maaslin", "finriskmetagcommon", "biomformat",
+                                 "microbiome", "phyloseq" ))
+packrat::snapshot()
+```
+
 ```{r Command line arguments}
@@
     if ("loadlast" %in% names(args) )
-        load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
+        load("session/comparepseqs.Rdata")
 }
@@
 # Importing libraries:
+
+```{r install, eval = FALSE, include = FALSE}
+source("https://bioconductor.org/biocLite.R")
+biocLite("BiocInstaller")
+library(BiocInstaller)
+source("http://www.bioconductor.org/biocLite.R")
+biocLite("microbiome")
+```
 
@@
 
-```{r data}
-pseq = list()
-pseq$shogun <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
-pseq$R16S <- import_filter_data("data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
-pseq$gtdb <- import_filter_data("data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs")
-pseq$ncbi <- import_filter_data("data/phfinrisk_custom70_species_all_2019-01-29.RDs")
-pseq$centrifuge <- import_filter_data("data/phfinrisk_centrifuge_species_all_2018-11-30.RDs",
-                                      excluded = c("DIAB"))
-```
+#         "data/phfinrisk_16S_otu_raw_2019-02-19.RDs",
+     #         "data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs",
+     #         "data/phfinrisk_custom70_species_all_2019-01-29.RDs",
 
-```{r alpha diversities}
-dset <- mclapply(pseq, function(x) meta.merge.alphadiversity(x) %>%
-                                             mutate(diversities_shannon = scale(diversities_shannon)),
-                           mc.cores = length(pseq))
-
+```{r constants}
 fo.alpha.min <- "%s ~ BL_AGE + SEX  + diversities_shannon"
 fo.alpha.full <- "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + diversities_shannon" # MISSING DIAB
-
-alphadiversity <-  mclapply(dset, function(x)
-    list(min =  calculateglm(x, modelstring = fo.alpha.min),
-         full = calculateglm(x, modelstring = fo.alpha.full)),
-    mc.cores = length(pseq))
-```
-
-```{r dissimilarity matrix}
-if (!file.exists("rds-comparepseqs/comparepseqs_bray_matrix.rds")) {
-    bray.dist.matrix <- mclapply(pseq, function(x) calculate.beta.matrix(x), mc.cores = length(pseq))
-    saveRDS(bray.dist.matrix, file = "rds-comparepseqs/comparepseqs_bray_matrix.rds")
-} else {
-    bray.dist.matrix  <- readRDS("rds-comparepseqs/comparepseqs_bray_matrix.rds")
-}
-```
-
- ```{r adonis calculation}
-if (!file.exists("rds-comparepseqs/beta.diversity.rds")) {
 
@@
                          "BL_USE_RX_C08", "BL_USE_RX_C09")
+```
 
-    betadiversity <- mclapply(c2l(names(pseq)), function(x)
-        list(min = calculateadonis(dset = meta(pseq[[x]]),
-                                   matrix = bray.dist.matrix[[x]],
+```{r rds files}
+rds <- list("shogun" = "data/phfinrisk_species_all_drop50k_2018-12-21.RDs",
+            "centrifuge" = "data/phfinrisk_centrifuge_species_all_2018-11-30.RDs")
+```
+
+```{r models}
+ret <- list()
+
+for (key in names(rds)) {
+    pseq <- import_filter_data(rds[[key]] ,excluded = c("DIAB"))
+    dset <- meta.merge.alphadiversity(x) %>%
+        mutate(diversities_shannon = scale(diversities_shannon))
+
+    alphadiversity <- list(min =  calculateglm(x, modelstring = fo.alpha.min),
+                           full = calculateglm(x, modelstring = fo.alpha.full))
+
+    saveRDS(alphadiversity,
+            file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
+
+    bray.dist.matrix <- calculate.beta.matrix(pseq)
+
+    betadiversity <- list(min = calculateadonis(dset = dset,
+                                                matrix = bray.dist.matrix,
                                                 covariates = covariates.min,
                                                 npermutations = 1000),
-             full = calculateadonis(dset = meta(pseq[[x]]),
-                                    matrix = bray.dist.matrix[[x]],
+                          full = calculateadonis(dset = dset,
+                                                 matrix = bray.dist.matrix,
                                                  covariates = covariates.full,
-                                    npermutations = 1000)),
-        mc.cores = length(pseq))
+                                                 npermutations = 1000))
 
-     saveRDS(betadiversity, file = "rds-comparepseqs/beta.diversity.rds")
-    } else {
-     betadiversity <- readRDS("rds-comparepseqs/beta.diversity.rds")
+    saveRDS(betadiversity,
+            file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
+
+    ret[[key]] = list("alphadiversity"  = alphadiversity,
+                      "betadiversity" = betadiversity)
 }
@@
 ```{r plot alpha diversity, eval = FALSE}
+
+alphadiversity[["ncbi"]]
+betadiversity[["ncbi"]]
+
+betadiversity
+
+ls()
 

### 2019-05-20 articletwo-rrbiome.R

@@ -213,7 +213,7 @@ mergediversities <- function(alphadiversity, betadiversity,
 }
 
 getdescriptions <- function() {
-    data("phfinrisk_metadatadesc", package="finriskmetagcommon")
+#    data("phfinrisk_metadatadesc", package="finriskmetagcommon")
     names.dset <- filter(phfinrisk_metadatadesc,
                          Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
     bind_rows(names.dset,
@@ -278,3 +278,6 @@ mykable <- function(x, ...) {
   capture.output(x <- print(x))
   knitr::kable(x, ...)
 }
+
+
+

### 2019-05-20 comparepseqs.Rmd
Minor

@@
 
-#         "data/phfinrisk_16S_otu_raw_2019-02-19.RDs",
-     #         "data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs",
-     #         "data/phfinrisk_custom70_species_all_2019-01-29.RDs",
-
 ```{r constants}
@@
 ```{r rds files}
-rds <- list("shogun" = "data/phfinrisk_species_all_drop50k_2018-12-21.RDs",
-            "centrifuge" = "data/phfinrisk_centrifuge_species_all_2018-11-30.RDs")
+rds <- list("R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs",
+            "shogun" = "data/phfinrisk_species_all_drop50k_2018-12-21.RDs",
+            "centrifuge" = "data/phfinrisk_centrifuge_species_all_2018-11-30.RDs",
+            "gtdb" = "data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs",
+            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs")
 ```

### 2019-05-20 comparepseqs.Rmd
x to dset
Packrat

@@
     pseq <- import_filter_data(rds[[key]] ,excluded = c("DIAB"))
-    dset <- meta.merge.alphadiversity(x) %>%
+    dset <- meta.merge.alphadiversity(dset) %>%
         mutate(diversities_shannon = scale(diversities_shannon))
 
-    alphadiversity <- list(min =  calculateglm(x, modelstring = fo.alpha.min),
-                           full = calculateglm(x, modelstring = fo.alpha.full))
+    alphadiversity <- list(min =  calculateglm(dset, modelstring = fo.alpha.min),
+                           full = calculateglm(dset, modelstring = fo.alpha.full))
 

### 2019-05-20 comparepseqs.Rmd
dset to pseq

@@
     pseq <- import_filter_data(rds[[key]] ,excluded = c("DIAB"))
-    dset <- meta.merge.alphadiversity(dset) %>%
+    dset <- meta.merge.alphadiversity(pseq) %>%
         mutate(diversities_shannon = scale(diversities_shannon))

### 2019-05-23 comparepseqs.Rmd
save bray also

@@
 ```{r rds files}
-rds <- list("R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs",
-            "shogun" = "data/phfinrisk_species_all_drop50k_2018-12-21.RDs",
+rds <- list("shogun" = "data/phfinrisk_species_all_drop50k_2018-12-21.RDs",
             "centrifuge" = "data/phfinrisk_centrifuge_species_all_2018-11-30.RDs",
             "gtdb" = "data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs",
-            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs")
+            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs",
+            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
 ```
@@
 
+    saveRDS(bray.dist.matrix,
+            file = paste0("rds-comparepseqs/bray-", key , ".rds"))
+
     betadiversity <- list(min = calculateadonis(dset = dset,
@@
 
+```{r plot, eval = FALSE}
+plotlist <- list()
+
+for (key in names(rds)) {
+    diversities <- mergediversities(ret[[key]][["alphadiversity"]],
+                                    ret[[key]][["betadiversity"]])
+
+    alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 0.6)
+    alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 0.6)
+
+    beta.diversity.min <- plot.beta(diversities$min)
+    beta.diversity.max <- plot.beta(diversities$max)
+
+    gs4 <- list(alpha.diversity.min$yaxis,
+                alpha.diversity.min$plot,
+                beta.diversity.min$plot,
+                alpha.diversity.min$legend,
+                alpha.diversity.min$xaxis,
+                beta.diversity.min$xaxis,
+                beta.diversity.min$xlab,
+                text_grob("A.", size = 18),
+                alpha.diversity.max$yaxis,
+                alpha.diversity.max$plot,
+                beta.diversity.max$plot,  
+                alpha.diversity.max$xaxis,
+                beta.diversity.max$xaxis,
+                beta.diversity.max$xlab,
+                text_grob("B.", size = 18))
+
+    g4 <- arrangeGrob(grobs = gs4,
+                      layout_matrix = rbind(
+                          c(8,    NA,    NA, NA,   NA, NA, NA),
+                          c(NA, 1,    2,    NA, 3, NA, NA),
+                          c(NA, 1,    2,    NA, 3, NA, 4),
+                          c(NA, NA, 5,    NA,6, NA, NA),
+                          c(NA, NA, NA, NA,7, NA, NA),
+                          c(15,    NA,    NA, NA,   NA, NA, NA),
+                          c(NA, 9,    10,  NA,  11, NA, NA),
+                          c(NA, 9,    10,  NA,  11, NA, NA),
+                          c(NA, NA, 12,   NA, 13, NA, NA),
+                          c(NA, NA, NA, NA,14, NA, NA)),
+                      widths=c(0.7, 4, 0.7, 0.1, 5, 0.7, 3),
+                      heights=rep(c(1, 1, 6, 1, 2), 2))
+
+    ggsave(file = paste0("cache/alpha-beta-", key , ".png"), plot=g4, height=6, width=9)
+    plotlist[[key]] = g4
+}
+```
+
 ```{r plot alpha diversity, eval = FALSE}
 
-alphadiversity[["ncbi"]]
-betadiversity[["ncbi"]]
-
-betadiversity
-
-ls()
-
-a <- mergediversities(alphadiversity[["ncbi"]], betadiversity[["ncbi"]])
-
-alphadiversity[["ncbi"]]
-betadiversity[["ncbi"]]
-
-names(betadiversity)
-betadiversity[[""]]
-
- alpha.select <- alphadiversity[["R16S"]][["min"]] %>% select(estimate, p.value, response)
-
-beta.rbind <- betadiversity[["R16S"]][["min"]] %>%
-    data.table::rbindlist(id = "model_name")
-
-%>%
-            dplyr::filter(term %in% responses) %>%
-            dplyr::select(response, R2)
-
-        merge( alpha.select , beta.rbind, by = "response") %>%
-            merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate") %>%
-            mutate(Qstar = ifelse(p.value < 0.05, '*', ' '))
-    })
-
-vectortolist(names(alphadiversity[["R16S"]]))
-
-c2l(names(pseq))
-beta.min <- lapply(c2l(names(pseq)), function(x) {
-    mergediversities(alphadiversity[[x]], betadiversity[[x]])
-})
-
-beta.min[[1]] %>%
-         data.table::rbindlist(id = "model_name")
-
-#            dplyr::filter(term %in% responses) %>%
-            dplyr::select(response, R2)
+alphadiversity <- readRDS("rds-comparepseqs/alpha-R16S.rds")
+alphadiversity
 

### 2019-05-23 comparepseqs.Rmd
Subset

@@
         load("session/comparepseqs.Rdata")
+
+    if ("subset" %in% names(args))
+        subset.run <- args$subset
 }
@@
             "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
+```
+
+```{r subset runs}
+if (exists(subset.run)) {
+    rds <- rds[[subset.run]]
+}
+
+print(rds)
+quit()
 ```
@@
 ```
-
-```{r save session}
-save.image(file = paste0("session/comparepseqs.Rdata"))
-```

### 2019-05-23 comparepseqs.Rmd
Minor
As numeric

@@
     if ("subset" %in% names(args))
-        subset.run <- args$subset
+        subset.run <- as.numeric(args$subset)
 }
@@
 ```{r subset runs}
-if (exists(subset.run)) {
+if (exists("subset.run")) {
+    message(paste0("Subset ", subset.run))
     rds <- rds[[subset.run]]
@@
 
-print(rds)
+message(rds)
 quit()

### 2019-05-24 comparepseqs.Rmd
Delete quit
list cast mistake
Beta saved over alpha

@@
 ```{r install, eval = FALSE, include = FALSE}
+install.packages("BiocManager")
+BiocManager::install("phyloseq")
+
 source("https://bioconductor.org/biocLite.R")
@@
 
+```{r rds subset, eval = FALSE}
+subset.run <- 3
+```
+
 ```{r rds files}
@@
     message(paste0("Subset ", subset.run))
-    rds <- rds[[subset.run]]
+    rds <- rds[subset.run]
 }
-
-message(rds)
-quit()
 ```
@@
 for (key in names(rds)) {
+    missing = list()
+
+    if (!file.exists(paste0("rds-comparepseqs/alpha-", key , ".rds"))) {
+        missing$alpha = TRUE
+    } else {
+        alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
+    }
+
+    if (!file.exists(paste0("rds-comparepseqs/bray-", key , ".rds"))) {
+        missing$bray = TRUE
+    } else {
+        bray.dist.matrix <- readRDS(file = paste0("rds-comparepseqs/bray-", key , ".rds"))
+    }
+
+    if (!file.exists(paste0("rds-comparepseqs/beta-", key , ".rds"))) {
+        missing$beta = TRUE
+    } else {
+        betadiversity <- readRDS(file = paste0("rds-comparepseqs/beta-", key , ".rds"))
+    }
+
+    if (length(missing) > 0) {
         pseq <- import_filter_data(rds[[key]] ,excluded = c("DIAB"))
@@
 
+        if ("alpha" %in% names(missing)) {
             alphadiversity <- list(min =  calculateglm(dset, modelstring = fo.alpha.min),
                                    full = calculateglm(dset, modelstring = fo.alpha.full))
+            saveRDS(alphadiversity, file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
+        }
 
-    saveRDS(alphadiversity,
-            file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
+        if ("bray" %in% names(missing)) {
+            bray.dist.matrix <- calculate.beta.matrix(pseq)
+            saveRDS(bray.dist.matrix, file = paste0("rds-comparepseqs/bray-", key , ".rds"))
+        }
 
-    bray.dist.matrix <- calculate.beta.matrix(pseq)
-
-    saveRDS(bray.dist.matrix,
-            file = paste0("rds-comparepseqs/bray-", key , ".rds"))
-
+        if ("beta" %in% names(missing)) {        
             betadiversity <- list(min = calculateadonis(dset = dset,
@@
                                                          npermutations = 1000))
-
-    saveRDS(betadiversity,
-            file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
+            saveRDS(betadiversity, file = paste0("rds-comparepseqs/beta-", key , ".rds"))
+        }
+    }
 
@@
 for (key in names(rds)) {
+    alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
+
     diversities <- mergediversities(ret[[key]][["alphadiversity"]],

### 2019-05-24 articletwo.Rmd
Missing commas
PCoA analysis and axis 1+3

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis', cache=FALSE, warning=FALSE)
+knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis',
+                      cache=FALSE, warning=FALSE)
 knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
@@
 
+## Firmicutes bacteroides ratio
+
+Firmicutes:Bacteroidetes ratio
+
+```{r firmicutes bacteroides, eval = FALSE}
+bfratio <- function (x) {
+    a <- transform(aggregate_taxa(x, level = "Phylum"), "compositional")
+    b <- abundances(a)["Bacteroidetes",]
+    f <- abundances(a)["Firmicutes",]  
+    #data.frame(Bacteroidetes = b, Firmicutes = f, Ratio = b/f)
+    b/f
+}
+
+packageVersion("microbiome")
+
+bfratio(pseq.species)
+
+phy3 <- aggregate_taxa(pseq.species, level = "Phylum")
+aggregate_taxa
+
+ transform(aggregate_taxa(x, level = "Phylum"), "compositional")
+sessionInfo()
+
+```
+
 ## Alpha & Beta plot (Figure 1.)
@@
     kable
+```
+
+Results in minimum model of alpha diversity
+
+```{r talbe alpha div min}
+alphadiversity$min %>%
+    select(response, estimate, p.value, conf.low, conf.high) %>%
+    mutate(mean_ci = sprintf("p=%s; 95%%-CI %.2f-%.2f",
+                              pub.p(p.value), conf.low, conf.high)) %>%
+    mutate_if(is.numeric, round, 3) %>%
+    arrange(p.value) %>%
+    kable
+```
+
+Results for full model of alpha diversity
+
+```{r talbe alpha div max}
+alphadiversity$max %>%
+    select(response, estimate, p.value, conf.low, conf.high) %>%
+    mutate(mean_ci = sprintf("p=%s; 95%%-CI %.2f-%.2f",
+                              pub.p(p.value), conf.low, conf.high)) %>%
+    mutate_if(is.numeric, round, 3) %>%
+    arrange(p.value) %>%
+    kable
+```
+
+Results for beta diversity
+
+```{r table beta diversity}
+adonis.species$max %>%
+    data.table::rbindlist(id = "model_name") %>%
+    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
+    dplyr::select(response, R2, "Pr(>F)") %>%
+    mutate(p.value = pub.p(`Pr(>F)`))
 ```
@@
 eig.fracs <- pcoa.ordinate$values$Relative_eig
-pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2),
+pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2, Axis.3),
                  meta(pseq.species), by=0, all = TRUE) %>%
     mutate(STATETREATMENT = factor(paste0(HYPERTENSION, ANYDRUG),
-                                   levels = c("11", "10", "01", "00")))
+                                   levels = c("11", "10", "01", "00")),
+           gMAP = cut(oMAP, seq(60, 160, 15)))
 
-pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.2, color = STATETREATMENT, shape = STATETREATMENT)) +
-    geom_point(size=0.4, alpha=0.4) +
+pcoa.df %>% pull(HYPERTENSION) %>% summary
+
+pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.3, color = HYPERTENSION)) +
+    geom_point(size=0.1, alpha=0.4) +
     xlab(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1])) +
-    ylab(sprintf("PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2])) +
+    ylab(sprintf("PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3])) +
     scale_colour_manual(name = "State and treatment",
-                        labels = c("Hypertensive, medication",
-                                   "Hypertensive, no medication",
-                                   "Normotensive, no medication"),
-                        breaks=c("11", "10", "00"),
-                        values = c("red", "red", "blue", "blue")) +   
-    scale_shape_manual(name = "State and treatment",
-                       labels = c("Hypertensive, medication",
-                                  "Hypertensive, no medication",
-                                  "Normotensive, no medication"),
-                       breaks = c("11", "10", "00"),
-                       values = c(4, 1, 1))+
+                        labels = c("Normotensive", "Hypertensive"),
+                        breaks=c(0, 1),
+                        values = c("blue", "red")) +   
     coord_fixed() +
-    scale_x_continuous(breaks=seq(-100, 100, 20)) +
-    scale_y_continuous(breaks=seq(-30, 30, 10)) +
+    scale_x_continuous(breaks=seq(-1, 1, .25)) +
+    scale_y_continuous(breaks=seq(-1, 1, .25)) +
     theme_classic() +
@@
 
-ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=5, width=5, dpi = 300)
+ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=4, width=5, units = "in", dpi = 300)
+```
+
+First three axis explain portion (`r eig.fracs[1:3]`) of
+variation. Testing associations to blood pressure variables on PCoA
+axis 1
+
+```{r pcoa axis analysis axis 1}
+calculateglm(pcoa.df,
+             modelstring = "%s ~ Axis.1 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
+                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
+             filterstr = "Axis") %>% 
+    select(response, term, estimate, std.error, p.value) %>%
+    mutate_if(is.numeric, round, 3) %>%
+    kable
+```
+
+Testing associations to blood pressure variables on PCoA axis 2
+
+```{r pcoa axis analysis axis 2}
+calculateglm(pcoa.df,
+             modelstring = "%s ~ Axis.2 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
+                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
+             filterstr = "Axis") %>% 
+    select(response, term, estimate, std.error, p.value) %>%
+    mutate_if(is.numeric, round, 3) %>%
+    kable
+```
+
+Testing associations to blood pressure variables on PCoA axis 3
+
+```{r pcoa axis analysis axis 3}
+calculateglm(pcoa.df,
+             modelstring = "%s ~ Axis.3 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
+                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
+             filterstr = "Axis") %>%
+    select(response, term, estimate, std.error, p.value) %>%
+    mutate_if(is.numeric, round, 3) %>%
+    kable
+```
+
 ```
@@
 Bray-Curtis distances are calculated for abundances after compositional transformation.
-First two axes explain %.1f%% of the variation.", round(100*sum(eig.fracs[1:2]),1))
+First two axes explain %.1f%% of the variation.", round(100*sum(eig.fracs[c(1,3)]),1))
 writeimage(doc, 2, "cache/pcoa-species.png", fig2head, fig2foot)
@@
 Max value of abs beta is `r maas.results.df.signf %>% pull(Coefficient) %>% abs %>% max`.
+
+```{r maaslin table}
+maas.results.df.signf %>%
+    select(Feature, Value, Coefficient, Qval) %>%
+    arrange(Qval) %>%
+    kable
+```
 
@@
     filter(!grepl("Intercept", term)) %>%
-    select(term, estimate, conf.low, conf.high) %>%
+    select(term, estimate, p.value, conf.low, conf.high) %>%
     add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%

### 2019-05-24 articletwo-ggplot.R

@@ -101,7 +101,7 @@ myforestplot <- function(data,
               axis.ticks.x = element_line(),
               axis.ticks.y = element_line(),
               axis.line = element_line(colour = "black")) +
-        scale_x_discrete(labels=c("Q1", "Q2", "Q3", "Q4")) +
+        scale_x_discrete(labels=c("T1", "T2", "T3")) +
         my_y_scale + 
         xlab(xlab) +
         ylab(ylab) +

### 2019-05-24 articletwo.Rmd
Removing BFratio

@@
 writetable(doc, tbl1, number = 1, tbl1head, tbl1foot)
-```
-
-## Firmicutes bacteroides ratio
-
-Firmicutes:Bacteroidetes ratio
-
-```{r firmicutes bacteroides, eval = FALSE}
-bfratio <- function (x) {
-    a <- transform(aggregate_taxa(x, level = "Phylum"), "compositional")
-    b <- abundances(a)["Bacteroidetes",]
-    f <- abundances(a)["Firmicutes",]  
-    #data.frame(Bacteroidetes = b, Firmicutes = f, Ratio = b/f)
-    b/f
-}
-
-packageVersion("microbiome")
-
-bfratio(pseq.species)
-
-phy3 <- aggregate_taxa(pseq.species, level = "Phylum")
-aggregate_taxa
-
- transform(aggregate_taxa(x, level = "Phylum"), "compositional")
-sessionInfo()
-
 ```

### 2019-05-24 articletwo-rrbiome.R
phfinrisk_metadatadesc

@@ -213,7 +213,7 @@ mergediversities <- function(alphadiversity, betadiversity,
 }
 
 getdescriptions <- function() {
-#    data("phfinrisk_metadatadesc", package="finriskmetagcommon")
+    data("phfinrisk_metadatadesc", package="finriskmetagcommon")
     names.dset <- filter(phfinrisk_metadatadesc,
                          Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
     bind_rows(names.dset,

### 2019-05-27 articletwo.Rmd
Fixing DIAB = 1 bug

@@
 ```{r maaslin runs, results = 'hide'} 
+if (!file.exists("rds/maaslin.rds")) {
     maaslin <- list()
-
     genus.core.bacteria <- pseq.genus.core %>%
@@
 
+    pseq.species %>% meta %>% pull(DIAB) %>% summary
+    pseq.genus.core %>% meta %>% pull(DIAB) %>% as.factor %>% summary
+
     maaslin$genus <- maaslinwrapper(pseq = pseq.genus.core,
-                looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
-                forced = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03",
+                                    looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE",
+                                               "HYPERTENSION"),
+                                    forced = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
+                                               "DIAB", "BL_USE_RX_C03",
                                                "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
                                     taxa = paste0(genus.core.bacteria, collapse = ","))
+    saveRDS(maaslin, file = "rds/maaslin.rds")
+} else {
+    maaslin <- readRDS("rds/maaslin.rds")
+}
 ```
@@
     dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
-    dplyr::select(response, R2, "Pr(>F)") %>%
+#    dplyr::select(response, R2, "Pr(>F)") %>%
     mutate(p.value = pub.p(`Pr(>F)`))
@@
 
+```{r adonis demo}
+data(dune)
+data(dune.env)
+adonis(dune ~ Management, data=dune.env, permutations=99)$aov.tab
+adonis(dune ~ Management, data=dune.env, permutations=99)$coefficients
+```
+
+```{r age sex bmi}
+lm(MAP ~ BL_AGE + SEX + BMI, data = dset.alpha) %>% summary
+```
+
 ```{r save session}

### 2019-05-27 articletwo-rrbiome.R

@@ -16,11 +16,10 @@ filter.phenotype.data <- function(pseq,
                                   included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
                                                "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X",
                                                "BL_USE_RX_C03",
-                                               "BL_USE_RX_C07", "BL_USE_RX_C08"),
-                                  excluded) {
+                                               "BL_USE_RX_C07", "BL_USE_RX_C08")) {
     meta(pseq) %>%
         tibble::rownames_to_column(var = "rowname") %>%
-        dplyr::select(rowname, included[!included %in% excluded]) %>%
+        dplyr::select(rowname, included) %>%
         stats::na.omit() %>%
         dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
                                        BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
@@ -37,13 +36,13 @@ filter.phenotype.data <- function(pseq,
                       DIASM = myscale(DIASM),
                       PULSEPRESSURE = myscale(PULSEPRESSURE),
                       MAP = myscale(MAP),
-                      DIAB = ifelse("DIAB" %in% excluded, NA, as.factor(DIAB)),
                       SEX = as.factor(SEX),
                       Q57X = as.factor(Q57X),
                       BL_USE_RX_C03 = as.factor(BL_USE_RX_C03),
                       BL_USE_RX_C07 = as.factor(BL_USE_RX_C07),
                       BL_USE_RX_C08 = as.factor(BL_USE_RX_C08),
                       BL_USE_RX_C09 = as.factor(BL_USE_RX_C09)) %>%
+        mutate_at(vars(one_of("DIAB")), as.factor) %>%
         dplyr::select(-MEN) %>%
         tibble::remove_rownames() %>%
         tibble::column_to_rownames(var = "rowname")
@@ -84,9 +83,9 @@ calculateglm <- function(dset,
                       conf.high = estimate + qnorm(1- 0.05/2) * std.error)
 }
 
-import_filter_data <- function(file, excluded = c()) {
+import_filter_data <- function(file) {
     pseq.full <- readRDS(file)
-    pseq.meta <- filter.phenotype.data(pseq.full, excluded = excluded)
[…]

### 2019-05-27 articletwo.Rmd
Fixing maaslin

@@
 if (!file.exists("rds/maaslin.rds")) {
+
     maaslin <- list()
@@
         mygrep(".Bacteria$", .)
-
-    pseq.species %>% meta %>% pull(DIAB) %>% summary
-    pseq.genus.core %>% meta %>% pull(DIAB) %>% as.factor %>% summary
-
     maaslin$genus <- maaslinwrapper(pseq = pseq.genus.core,
@@
                                     taxa = paste0(genus.core.bacteria, collapse = ","))
+
     saveRDS(maaslin, file = "rds/maaslin.rds")

### 2019-05-27 comparepseqs.Rmd
Wrong order

@@
 library(vegan)
+library(ggpubr)
 ```
@@
 # Loading data
+
+Loading descriptions for clinical data
+
+```{r variables, warning = FALSE}
+names.dset <- getdescriptions()
+```
 
@@
 }
+
+rds <- list("gtdb" = 1 , "centrifuge" = 2)
 ```
@@
     if (length(missing) > 0) {
-        pseq <- import_filter_data(rds[[key]] ,excluded = c("DIAB"))
+        pseq <- import_filter_data(rds[[key]])
         dset <- meta.merge.alphadiversity(pseq) %>%
@@
                                     ret[[key]][["betadiversity"]])
-
+diversities$min
     alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 0.6)
-    alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 0.6)
+    alpha.diversity.full <- plot.alpha(diversities$full, colorlimit = 0.6)
 
     beta.diversity.min <- plot.beta(diversities$min)
-    beta.diversity.max <- plot.beta(diversities$max)
+    beta.diversity.full <- plot.beta(diversities$full)
 
@@
                 beta.diversity.min$xlab,
-                text_grob("A.", size = 18),
-                alpha.diversity.max$yaxis,
-                alpha.diversity.max$plot,
-                beta.diversity.max$plot,  
-                alpha.diversity.max$xaxis,
-                beta.diversity.max$xaxis,
-                beta.diversity.max$xlab,
-                text_grob("B.", size = 18))
+                text_grob(paste("A.", key), size = 18, just = "left"),
+                alpha.diversity.full$yaxis,
+                alpha.diversity.full$plot,
+                beta.diversity.full$plot,  
+                alpha.diversity.full$xaxis,
+                beta.diversity.full$xaxis,
+                beta.diversity.full$xlab,
+                text_grob(paste("B.", key), size = 18, just = "left"))
 
@@
 ```{r plot alpha diversity, eval = FALSE}
-
-alphadiversity <- readRDS("rds-comparepseqs/alpha-R16S.rds")
-alphadiversity
-
-diversity.min <- mergediversities(alpha.min, beta.min)
-
-alpha.plot.min <- mclapply(alpha.min, function(x) { mergediversities(},
-                               mc.cores = length(pseq))
-
-gs4 <- c(list("yaxis" = alpha.plot.min[[1]]$yaxis,
-              "xaxis" = alpha.plot.min[[1]]$xaxis),
-              lapply(alpha.plot.min, function(x) x$plot))
-
-g4.firstline <- c(1, seq(3, length(alpha.plot.min)+2))
-g4.secondline <- c(NA, rep(2, length(alpha.plot.min)))
-
-g4 <- arrangeGrob(grobs = gs4, layout_matrix = rbind(g4.firstline, g4.secondline))
-
-ggsave(file = "cache/pseqdiff-alpha-min.png", plot=g4, height=6, width=9)
-
+g <- arrangeGrob(plotlist[[1]], plotlist[[2]], ncol=1)
+ggsave(file="cache/plotlist.pdf", g, height=6*length(plotlist), width=10)
 ```
 
-```{r join alphas, eval = FALSE}
-sfx <- names(dset)
-res <- alpha.min[[1]]
-
-for(i in head(seq_along(alpha.min), -1)) {
-    res <- merge(res,
-                 alpha.full[[i+1]],
-                 all = TRUE,
-                 suffixes = paste0("_", sfx[i:(i+1)]),
-                 by = c("model_name", "term", "response"))
-}
-
-res %>% colnames
-
-res %>%
-    dplyr::select(response, starts_with("estimate_"), starts_with("p.value"))
-
+```{r debug, eval = FALSE}
+ret[["centrifuge"]]$betadiversity$full$model_1
 ```

### 2019-05-27 articletwo-ggplot.R

@@ -48,12 +48,14 @@ plot.alpha <- function(diversity,  colorlimit = 0.25) {
 
 plot.beta <- function(diversity,
                       fig.permar2.text = "R2 of variable \nin Bray-Curtis distance",
-                      fig.textsize = 15) {
+                      fig.textsize = 15,
+                      ymax = 0.0016) {
     plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2)) +
         geom_bar(stat="identity", color="black") +
         coord_flip() +
         theme_classic(fig.textsize) +
         ylab(fig.permar2.text) +
+        scale_y_continuous(limits = c(0, ymax)) +
         theme(axis.text.y = element_text(color="black"),
               axis.text.x = element_text(color="black"),
               legend.position="right",
@@ -73,11 +75,12 @@ plot.beta <- function(diversity,
               axis.title.x  =  element_blank(),
               axis.title.y = element_blank(),
               legend.position = "none")
+    
     plot.gtable.beta.diversity.stripped <- ggplot_gtable(ggplot_build(plot.beta.diversity))
 
-      return(list(plot =plot.gtable.beta.diversity.stripped$grobs[[6]],
-                  xaxis = plot.gtable.beta.diversity$grobs[[7]],
-                  xlab = plot.xlab.beta.diversity))
+    return(list(plot = plot.gtable.beta.diversity.stripped$grobs[[6]],
+                xaxis = plot.gtable.beta.diversity$grobs[[7]],
+                xlab = plot.xlab.beta.diversity))
 }
 
 myforestplot <- function(data,

### 2019-05-27 articletwo.Rmd
Wrong order
Minor

@@
 
-```{r odds htn lacto}
-odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
-    filter(!grepl("Intercept", term)) %>%
-    select(term, estimate, p.value, conf.low, conf.high) %>%
-    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
-    dplyr::filter(grepl("lacto", term))
-odds.hypertension.lactogroup %>% kable
-```
-
 Odds for hypertension by categorical lactobacillus abundance.
@@
     kable
+```
+
+Odds for hypertension
+
+```{r odds htn lacto}
+odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
+    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
+    filter(!grepl("Intercept", term)) %>%
+    select(term, estimate, p.value, conf.low, conf.high) %>%
+    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
+    dplyr::filter(grepl("lacto", term))
+odds.hypertension.lactogroup %>% kable
 ```

### 2019-05-27 comparepseqs.Rmd
Minor

@@
 
-rds <- list("gtdb" = 1 , "centrifuge" = 2)
+rds <- list("gtdb" = 1 , "centrifuge" = 2, "ncbi" = 3)
 ```
@@
                 beta.diversity.min$xlab,
-                text_grob(paste("A.", key), size = 18, just = "left"),
+                text_grob(sprintf("A. %s (%s)", key, paste(covariates.min, collapse = ", ")),
+                          size = 16,
+                          just = "left"),
                 alpha.diversity.full$yaxis,
@@
                 beta.diversity.full$xlab,
-                text_grob(paste("B.", key), size = 18, just = "left"))
+                text_grob(sprintf("B. %s (%s)", key, paste(covariates.full %>%
+                                                           gsub("BL_USE_", "", .), collapse = ", ")),
+                          size = 16, just = "left"))
 
@@
 ```{r plot alpha diversity, eval = FALSE}
-g <- arrangeGrob(plotlist[[1]], plotlist[[2]], ncol=1)
-ggsave(file="cache/plotlist.pdf", g, height=6*length(plotlist), width=10)
+g <- arrangeGrob(grobs = plotlist, ncol=2)
+ggsave(file="cache/plotlist.pdf", g, height=6*length(plotlist)/1.5, width=10*2)
 ```

### 2019-05-27 articletwo.Rmd
Cleaning

@@
     dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
-#    dplyr::select(response, R2, "Pr(>F)") %>%
-    mutate(p.value = pub.p(`Pr(>F)`))
+    dplyr::select(response, R2, "Pr(>F)") %>%
+    mutate(p.value = pub.p(`Pr(>F)`)) %>%
+    kable
 ```
@@
 ## Direct associations between genus level and blood pressure variables (figure 3)
+
+For magnitude, age, sex, and BMI explain `r round(summary(lm(MAP ~ BL_AGE +
+SEX + BMI, data = meta(pseq.genus)))$adj.r.squared * 100, 1)`% of MAP.
 
@@
 
-```{r adonis demo}
-data(dune)
-data(dune.env)
-adonis(dune ~ Management, data=dune.env, permutations=99)$aov.tab
-adonis(dune ~ Management, data=dune.env, permutations=99)$coefficients
-```
-
-```{r age sex bmi}
-lm(MAP ~ BL_AGE + SEX + BMI, data = dset.alpha) %>% summary
-```
-
 ```{r save session}

### 2019-05-27 comparepseqs.Rmd
Print subset

@@
 library(ggpubr)
+library(broom)
+library(tidyr)
 ```
@@
 if (exists("subset.run")) {
-    message(paste0("Subset ", subset.run))
+    message(sprintf("Subset %s of %s", subset.run, rds[[subset.run]]))
     rds <- rds[subset.run]
@@
 for (key in names(rds)) {
-    alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
-
     diversities <- mergediversities(ret[[key]][["alphadiversity"]],
                                     ret[[key]][["betadiversity"]])
-diversities$min
-    alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 0.6)
-    alpha.diversity.full <- plot.alpha(diversities$full, colorlimit = 0.6)
+
+    alpha.diversity.min <- plot.alpha(diversities$min, alphabreaks = c(-0.15, -0.10, -0.05, 0))
+    alpha.diversity.full <- plot.alpha(diversities$full, alphabreaks = c(-0.15, -0.10, -0.05, -0))
 
@@
 ```{r debug, eval = FALSE}
-ret[["centrifuge"]]$betadiversity$full$model_1
+alphalist <- c()
+for (key in names(rds)) {
+    alphalist <- c(alphalist,
+              ret[[key]]$alphadiversity$min %>% pull(estimate),
+              ret[[key]]$alphadiversity$full %>% pull(estimate))
+}
+
+summary(alphalist)
 ```
+
+```{r alpha size, eval = FALSE}
+pseq <- import_filter_data(rds[["gtdb"]])
+
+dset <- meta.merge.alphadiversity(pseq) %>%
+    mutate(diversities_shannon = scale(diversities_shannon))
+
+dset %>%
+    select(MAP, diversities_shannon) %>%
+    gather(key, value) %>%
+    group_by(key) %>%
+    summarise_all(funs(N = n(),
+                       min = min(as.numeric(value), na.rm = TRUE),
+                       max = max(as.numeric(value), na.rm = TRUE),
+                       Mean = mean(as.numeric(value), na.rm = TRUE),
+                       SD = sd(as.numeric(value), na.rm = TRUE),
+                       NAs = sum(is.na(value))))  %>%
+    kable()
+
+lm(sprintf(fo.alpha.full, "SYSTM"), data = dset) %>% summary
+```
+
+```{r size}
+Reduce(function(...) merge(..., all=T), ret)
+```

### 2019-05-27 articletwo-ggplot.R

@@ -1,12 +1,20 @@
-plot.alpha <- function(diversity,  colorlimit = 0.25) {
-    plot.alpha.diversity <- ggplot(diversity, aes(x=reorder(Name, R2),  y = 1, fill = estimate)) +
+plot.alpha <- function(diversity,
+                       alphabreaks = c(-Inf, -1, -0.1, -0.01, 0)) {
+    diversity <- diversity %>%
+        mutate(estimate_fac =  cut(estimate, breaks = alphabreaks))
+    plot.alpha.diversity <- ggplot(diversity, aes(x=reorder(Name, R2),
+                                                  y = 1,
+                                                  fill = estimate_fac)) +
         geom_bar(stat="identity", color="black") +
         coord_flip() +
-        scale_fill_gradientn(colours = c("blue", "white"),
-                             breaks = c(-colorlimit, 0),
-                             name = 'Regression coefficient \nin linear model \nfor Shannon index',
-                             limits = c(-colorlimit, 0),
-                             na.value = "black") +
+        scale_fill_brewer(palette="Blues",
+                          name = 'Regression coefficient \nin linear model \nfor Shannon index') +
+        #scale_color_discrete_sequential(palette = "Blues", nmax = 6, order = 2:6) +
+#        scale_fill_gradientn(colours = c("blue", "white"),
+#                             breaks = c(-colorlimit, 0),
+#                             name = 'Regression coefficient \nin linear model \nfor Shannon index',
+#                             limits = c(-colorlimit, 0),
+#                             na.value = "black") +
         theme_classic(20) +
         geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=FALSE, color='black', size=20) +
         scale_shape_manual(name="",

### 2019-05-27 comparepseqs.Rmd
Remove debug
iDebug

@@
         if ("alpha" %in% names(missing)) {
+            message("Missing alpha")
             alphadiversity <- list(min =  calculateglm(dset, modelstring = fo.alpha.min),
@@
         if ("bray" %in% names(missing)) {
+            message("Missing Bray")
             bray.dist.matrix <- calculate.beta.matrix(pseq)
@@
         if ("beta" %in% names(missing)) {
+            message("Missing beta")
             betadiversity <- list(min = calculateadonis(dset = dset,
@@
 ```
-
-```{r size}
-Reduce(function(...) merge(..., all=T), ret)
-```

### 2019-05-28 comparepseqs.Rmd
DEbug line removed
Minor

@@
 ```{r rds subset, eval = FALSE}
-subset.run <- 3
+esubset.run <- 3
 ```
@@
             "gtdb" = "data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs",
-            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs",
-            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
+            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs")
+#            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
 ```
@@
 }
-
-rds <- list("gtdb" = 1 , "centrifuge" = 2, "ncbi" = 3)
 ```
@@
 ```{r plot, eval = FALSE}
+if (!exists("subset.run")) {
     plotlist <- list()
@@
 
-    alpha.diversity.min <- plot.alpha(diversities$min, alphabreaks = c(-0.15, -0.10, -0.05, 0))
-    alpha.diversity.full <- plot.alpha(diversities$full, alphabreaks = c(-0.15, -0.10, -0.05, -0))
+        alpha.diversity.min <- plot.alpha(diversities$min, alphabreaks = c(-0.15, -0.10, -0.05, -0.01, 0))
+        alpha.diversity.full <- plot.alpha(diversities$full, alphabreaks = c(-0.15, -0.10, -0.05, -0.01, -0))
 
@@
     }
-```
 
-```{r plot alpha diversity, eval = FALSE}
     g <- arrangeGrob(grobs = plotlist, ncol=2)
-ggsave(file="cache/plotlist.pdf", g, height=6*length(plotlist)/1.5, width=10*2)
+    ggsave(file="cache/plotlist.png", g, height=6*length(plotlist)/1.5, width=10*2)
+} else {
+    message("Skipping graphics")
+}
 ```
@@
 ```
+
+```{r debug alpha, eval = FALSE}
+alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alpha-R16S.rds"))
+alphadiversity
+
+```
+
+Full model R2
+
+```{r diff beta debug, eval = FALSE}
+betadf <- data.frame()
+for (key in names(rds)) {
+    moddf.full <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$full) %>%
+        dplyr::filter(response == term) %>%
+        mutate(model = key, covariates = "full") %>%
+        select(term, model, covariates, R2)
+    moddf.min <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$min) %>%
+        dplyr::filter(response == term) %>%
+        mutate(model = key, covariates = "min") %>%
+        select(term, model, covariates, R2)
+    betadf <- rbind(betadf, moddf.full, moddf.min)
+}
+
+betadf %>% arrange(term) %>% spread(model, R2) %>% kable
+```
+
+```{r debugging adonis, eval = FALSE}
+pseq <- import_filter_data(rds[["gtdb"]])
+
+dset <- meta.merge.alphadiversity(pseq) %>%
+    mutate(diversities_shannon = scale(diversities_shannon))
+
+bray.dist.matrix <- readRDS(file = paste0("rds-comparepseqs/bray-gtdb.rds"))
+
+dset.sample <- dset %>% sample_n(50)
+
+dset.ids <- dset.sample %>% pull(Sample_ID)
+
+betadiversity <- list(min = calculateadonis(dset = dset.sample,
+                                            matrix = bray.dist.matrix[dset.ids, dset.ids],
+                                            covariates = covariates.min,
+                                            npermutations = 10000),
+                      full = calculateadonis(dset = dset.sample,
+                                      matrix = bray.dist.matrix[dset.ids, dset.ids],
+                                      covariates = covariates.full,
+                                      npermutations = 2))
+
+betadiversity$full
+
+betadf <- data.frame()
+for (key in names(rds)) {
+    moddf.full <- Reduce(function(...) merge(..., all=T), betadiversity$full) %>%
+        dplyr::filter(response == term) %>%
+        mutate(model = key, covariates = "full") %>%
+        select(term, model, covariates, R2)
+    moddf.min <- Reduce(function(...) merge(..., all=T), betadiversity$min) %>%
+        dplyr::filter(response == term) %>%
+        mutate(model = key, covariates = "min") %>%
+        select(term, model, covariates, R2)
+    betadf <- rbind(betadf, moddf.full, moddf.min)
+}
+betadf %>% arrange(term) %>% spread(model, R2) %>% kable
+
+```

### 2019-05-28 articletwo-ggplot.R

@@ -8,7 +8,9 @@ plot.alpha <- function(diversity,
         geom_bar(stat="identity", color="black") +
         coord_flip() +
         scale_fill_brewer(palette="Blues",
-                          name = 'Regression coefficient \nin linear model \nfor Shannon index') +
+                          name = 'Regression coefficient \nin linear model \nfor Shannon index',
+                          drop=FALSE,
+                          direction = -1) +
         #scale_color_discrete_sequential(palette = "Blues", nmax = 6, order = 2:6) +
 #        scale_fill_gradientn(colours = c("blue", "white"),
 #                             breaks = c(-colorlimit, 0),

### 2019-05-28 articletwo-rrbiome.R

@@ -120,7 +120,7 @@ calculateadonis <- function(dset,
                             matrix,
                             responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
                                           model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
-                            covariates = c("BL_AGE", "SEX"),
+                            covariates = ".",
                             npermutations = 99,
                             maxcores = 100) {
     mclapply(responses, function(response) {

### 2019-05-28 comparepseqs.Rmd
Minor

@@
             "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs")
-#            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
+            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
 ```

### 2019-05-29 articletwo.Rmd
Minor

@@
 ```{r data}
-pseq.genus <- import_filter_data("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
-pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
+pseq.genus <- import_filter_data("data/phfinrisk_centrifuge_genus_all_drop50k_2018-11-30.RDs")
+pseq.species <- import_filter_data("data/phfinrisk_centrifuge_species_all_2018-11-30.RDs")
 ```

### 2019-05-29 comparepseqs.Rmd
Minor

@@
 
-```{r rds subset, eval = FALSE}
-esubset.run <- 3
-```
-
 ```{r rds files}
@@
             "gtdb" = "data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs",
-            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs")
+            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs",
             "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
@@
 
-```{r plot, eval = FALSE}
+```{r plot alpha-beta}
 if (!exists("subset.run")) {

### 2019-05-29 comparepseqs.Rmd
ggpairs

@@
 library(tidyr)
+library(GGally)
 ```
@@
 
+    if (!file.exists(paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))) {
+        missing$alphadistribution = TRUE
+    } else {
+        alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))
+    }
+
     if (!file.exists(paste0("rds-comparepseqs/alpha-", key , ".rds"))) {
@@
         dset <- meta.merge.alphadiversity(pseq) %>%
-            mutate(diversities_shannon = scale(diversities_shannon))
+            mutate(diversities_shannon = as.numeric(scale(diversities_shannon)))
+
+        if ("alphadistribution" %in% names(missing)) {
+            message("Missing alpha distribution")
+            alphadistribution <- dset %>%
+                mutate(!!key := diversities_shannon) %>%
+                select(Sample_ID, key)
+            saveRDS(alphadistribution, file = paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))
+        }
 
@@
     ret[[key]] = list("alphadiversity" = alphadiversity,
-                      "betadiversity" = betadiversity)
+                      "betadiversity" = betadiversity,
+                      "alphadistribution" = alphadistribution)
 }
@@
 ```{r debug alpha, eval = FALSE}
-alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alpha-R16S.rds"))
-alphadiversity
+scatterplot.df <- lapply(ret, function(x) x$alphadistribution) %>%
+    Reduce(function(x, y) merge(x, y, all=TRUE, by="Sample_ID"), .) %>%
+    select(-Sample_ID) %>%
+    na.omit()
+
+ggsave(file = paste0("cache/ggpairs-alphas.png"),
+       plot=ggpairs(scatterplot.df,
+                        lower = list(continuous = wrap("points",
+                                                       alpha = 0.3,
+                                                       size=0.1), 
+                                     combo = wrap("dot",
+                                                  alpha = 0.4,
+                                                  size=0.2))),
+       height=6,
+       width=9,
+       unit = "in",
+       dpi = 300)
 

### 2019-05-29 comparepseqs.Rmd
Allow set included variables for import

@@
             "gtdb" = "data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs",
-            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs",
-            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
+            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs")#,
+#            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
 ```
@@
     } else {
-        alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))
+        alphadistribution <- readRDS(file =
+                                         paste0("rds-comparepseqs/alphadistribution-",
+                                                key , ".rds"))
+    }
+
+    ret[[key]] = list(                      "alphadistribution" = alphadistribution)
     }
@@
     if (length(missing) > 0) {
-        pseq <- import_filter_data(rds[[key]])
+        pseq <- import_filter_data(rds[[key]],
+                                   included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
+                                                "CURR_SMOKE", "BL_USE_RX_C09",
+                                                "Q57X", "BL_USE_RX_C03",
+                                                "BL_USE_RX_C07", "BL_USE_RX_C08"))
         dset <- meta.merge.alphadiversity(pseq) %>%
@@
 
+scatterplot.df %>% head
+
+ lapply(ret, function(x) x$alphadistribution %>% head)
+
 ggsave(file = paste0("cache/ggpairs-alphas.png"),

### 2019-05-29 articletwo-rrbiome.R

@@ -83,9 +83,13 @@ calculateglm <- function(dset,
                       conf.high = estimate + qnorm(1- 0.05/2) * std.error)
 }
 
-import_filter_data <- function(file) {
+import_filter_data <- function(file,
+                               included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
+                                            "CURR_SMOKE", "DIAB", "BL_USE_RX_C09",
+                                            "Q57X", "BL_USE_RX_C03",
+                                            "BL_USE_RX_C07", "BL_USE_RX_C08")) {
     pseq.full <- readRDS(file)
-    pseq.meta <- filter.phenotype.data(pseq.full)
+    pseq.meta <- filter.phenotype.data(pseq.full, included = included)
     phyloseq::sample_data(pseq.full) <- phyloseq::sample_data(pseq.meta)
     return(pseq.full)
 }

### 2019-05-29 comparepseqs.Rmd
Typo, split model function

--- 2019-05-29 comparepseqs.Rmd
@@
     } else {
-        alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))
+        alphadistribution <- readRDS(file =
+                                         paste0("rds-comparepseqs/alphadistribution-",
+                                                key , ".rds"))
     }
@@
     }
+}
+```
 
+```{r calculate missing models}
 if (length(missing) > 0) {
-        pseq <- import_filter_data(rds[[key]])
+    for (key in names(rds)) {
+        pseq <- import_filter_data(rds[[key]],
+                                   included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
+                                                "CURR_SMOKE", "BL_USE_RX_C09",
+                                                "Q57X", "BL_USE_RX_C03",
+                                                "BL_USE_RX_C07", "BL_USE_RX_C08"))
         dset <- meta.merge.alphadiversity(pseq) %>%
@@
                 select(Sample_ID, key)
-            saveRDS(alphadistribution, file = paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))
+            saveRDS(alphadistribution,
+                    file = paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))
         }
@@
         }
-    }
 
@@
                           "alphadistribution" = alphadistribution)
+    }
 }

### 2019-05-29 comparepseqs.Rmd
Split not possible

--- 2019-05-29 comparepseqs.Rmd
@@
             "gtdb" = "data/phfinrisk_gtdb_species_all_drop50k_2019-01-14.RDs",
-            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs")#,
-#            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
+            "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs",
+            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
 ```
@@
                                                 key , ".rds"))
-    }
-
-    ret[[key]] = list(                      "alphadistribution" = alphadistribution)
     }

### 2019-05-31 comparepseqs.Rmd
Minor

@@
 # Importing libraries:
-
-```{r install, eval = FALSE, include = FALSE}
-install.packages("BiocManager")
-BiocManager::install("phyloseq")
-
-source("https://bioconductor.org/biocLite.R")
-biocLite("BiocInstaller")
-library(BiocInstaller)
-source("http://www.bioconductor.org/biocLite.R")
-biocLite("microbiome")
-```
 
@@
 
-```{r models}
+```{r alpha scatter, include = FALSE, eval = FALSE}
+rds = rds[c(1,2,4)]
+ret <- list()
+
+for (key in names(rds)) {
+    missing = list()
+
+    if (!file.exists(paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))) {
+        missing$alphadistribution = TRUE
+    } else {
+        alphadistribution <- readRDS(file =
+                                         paste0("rds-comparepseqs/alphadistribution-",
+                                                key , ".rds"))
+    }
+
+    ret[[key]] = list("alphadistribution" = alphadistribution)
+}
+```
+
+```{R models}
 ret <- list()
@@
                                                         covariates = covariates.min,
-                                                        npermutations = 1000),
+                                                        npermutations = 100),
                                   full = calculateadonis(dset = dset,
@@
                                                          covariates = covariates.full,
-                                                         npermutations = 1000))
+                                                         npermutations = 100))
             saveRDS(betadiversity, file = paste0("rds-comparepseqs/beta-", key , ".rds"))
@@
 ```{r plot alpha-beta}
+names(rds)
 if (!exists("subset.run")) {
@@
         ggsave(file = paste0("cache/alpha-beta-", key , ".png"), plot=g4, height=6, width=9)
+
         plotlist[[key]] = g4
@@
     ggsave(file="cache/plotlist.png", g, height=6*length(plotlist)/1.5, width=10*2)
+
 } else {
@@
 
-```{r debug alpha, eval = FALSE}
+# Alpha scatter
+
+```{r draw alpha distribution scatter plot, eval = FALSE}
 scatterplot.df <- lapply(ret, function(x) x$alphadistribution) %>%
@@
 
-scatterplot.df %>% head
-
- lapply(ret, function(x) x$alphadistribution %>% head)
-
-ggsave(file = paste0("cache/ggpairs-alphas.png"),
-       plot=ggpairs(scatterplot.df,
+galphascatter <- ggpairs(scatterplot.df,
                          lower = list(continuous = wrap("points",
@@
                                                    alpha = 0.4,
-                                                  size=0.2))),
+                                                   size=0.2))) +
+    theme(legend.position = "none", 
+          panel.grid.major = element_blank(), 
+          axis.ticks = element_blank(), 
+          panel.border = element_rect(linetype = "dashed", colour = "black", fill = NA))
+
+ggsave(file = paste0("cache/ggpairs-alphas.png"),
+       plot= galphascatter,
        height=6,

### 2019-05-31 comparepseqs.Rmd
Increase permutations

@@
                                                         covariates = covariates.min,
-                                                        npermutations = 100),
+                                                        npermutations = 1000),
                                   full = calculateadonis(dset = dset,
@@
                                                          covariates = covariates.full,
-                                                         npermutations = 100))
+                                                         npermutations = 1000))
             saveRDS(betadiversity, file = paste0("rds-comparepseqs/beta-", key , ".rds"))

### 2019-06-03 articletwo.Rmd
Pcoa loading

@@
 library(gvlma)
+library(purrr)
 ```
@@
 there ntaxa is `r ntaxa(pseq.species)`, and at genus level meta has
-dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is `r
-ntaxa(pseq.genus)`.
+dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is 
+`r ntaxa(pseq.genus)`.
 
@@
 
-pcoa.df %>% pull(HYPERTENSION) %>% summary
-
 pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.3, color = HYPERTENSION)) +
@@
 axis 1
+
+```{r top generas, eval = FALSE, include = FALSE}
+abund.genus <- abundances(pseq.genus, transform = "compositional") %>%
+    t %>%
+    as.data.frame %>%
+    tibble::rownames_to_column("id") %>%
+    setNames(tolower(replace.brackets(names(.))))
+
+abund.genus.bacteria <- colnames(abund.genus)[-1] %>% mygrep(".bacteria$", .)
+
+pcoacorabund <- function(names, pcoa) {
+    lapply(c2l(abund.genus.bacteria), function (name) 
+                     lapply(c2l(1,2,3), function(index) {
+                             model <- lm(abund.genus[,name] ~ pcoa[,index])
+                             summary(model)$r.squared
+                         }))
+}
+
+pcoacor.genus <- pcoacorabund(abund.genus.bacteria, pcoa.vectors)
+pcoacor.genus.df <- map_df(pcoacor.genus, ~as.data.frame(.x), .id="id")
+
+pcoacor.genus.df %>%
+    arrange(-X3) %>%
+    head(10) %>% kable
+```
 
@@
 ```
+
+```{r pcoa 3d}
+theta <- 130
+phi <- 10
+pcoa3d.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y = Axis.2, z=Axis.3, color = HYPERTENSION)) +
+  axes_3D(theta=theta, phi=phi) +
+  stat_3D(theta=theta, phi=phi, size = 0.01, alpha=0.4) +
+  labs_3D(theta=theta, phi=phi, 
+          hjust=c(1,0,0), vjust=c(1.5,1,-.2),
+          labs=c(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1]),
+                 sprintf("PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2]),
+                 sprintf("PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3]))) #+
+#  theme_void()
+
+ggsave(file = "cache/pcoa-species-3d.png", plot=pcoa3d.plot, height=4, width=7, units = "in", dpi = 300)
 
@@
     kable
+```
+
+Catabacter, Robinsoniella
+
+```{r compare maaslin to sun, eval = FALSE}
+maas.results.df.all <- maaslin$genus %>%
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
+    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
+                  Feature = gsub("_Bacteria", "", Feature),
+                  beta = Coefficient)
+
+maas.results.df.all %>% pull(Feature) %>% sort %>% unique %>% mygrep("Robin", .)
+
+genus.taxa <- pseq.genus.core %>% abundances %>% rownames
+
+genus.taxa %>% mygrep("Catabacter", .)
+
+maaslin.test <- maaslinwrapper(pseq = pseq.genus.core,
+                                    looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE",
+                                               "HYPERTENSION"),
+                                    forced = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
+                                               "DIAB", "BL_USE_RX_C03",
+                                               "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
+                               taxa = paste(c("Faecalicatena.Bacteria",
+                                        "Catabacter.Bacteria",
+                                        "Robinsoniella.Bacteria"), collapse = ", "))
+
 ```
@@
 
+```{r squared clinical, eval = FALSE, include = FALSE}
+mylmrsqr <- function(df, response, loopcovariates) {
+    model.list <- list()
+    for (i in seq(length(loopcovariates))) {
+        fo <- sprintf("%s ~ %s", response, paste(loopcovariates, collapse = " + "))
+        message(sprintf("formula : %s", fo))
+        model <- lm(as.formula(fo), data = df)
+        model$call <- as.formula(fo)
+        model.list[[i]] <- model
+        loopcovariates <- loopcovariates[-length(loopcovariates)]
+    }
+    return(model.list)
+}
+
+af.models <- mylmrsqr(df = meta(pseq.species),
+         response = "MAP",
+         loopcovariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
+                            "ANYDRUG"))
+
+af <- do.call(anova, af.models)
+
+af %>%
+    as.data.frame %>%
+        mutate(Rsqr =  `Sum of Sq`/sum(`Sum of Sq`, na.rm = T)*100)
+
+c(summary(lm(SYSTM ~ BL_AGE + SEX, meta(pseq.species)))$adj.r.squared,
+summary(lm(DIASM ~ BL_AGE + SEX, meta(pseq.species)))$adj.r.squared,
+summary(lm(PULSEPRESSURE ~ BL_AGE + SEX, meta(pseq.species)))$adj.r.squared,
+summary(lm(MAP ~ BL_AGE + SEX, meta(pseq.species)))$adj.r.squared) %>% mean
+
+c(summary(lm(SYSTM ~ BL_AGE + SEX + BMI, meta(pseq.species)))$adj.r.squared,
+summary(lm(DIASM ~ BL_AGE + SEX + BMI, meta(pseq.species)))$adj.r.squared,
+summary(lm(PULSEPRESSURE ~ BL_AGE + SEX + BMI, meta(pseq.species)))$adj.r.squared,
+summary(lm(MAP ~ BL_AGE + SEX + BMI, meta(pseq.species)))$adj.r.squared) %>% sort
+
+```
+
 ```{r Write docx to file, include=FALSE}

### 2019-06-03 articletwo.Rmd
Top genera

@@
 
-```{r top generas, eval = FALSE, include = FALSE}
+Genera with top R2 with PCoA axes
+
+```{r top generas}
 abund.genus <- abundances(pseq.genus, transform = "compositional") %>%
@@
 
-pcoacorabund <- function(names, pcoa) {
-    lapply(c2l(abund.genus.bacteria), function (name) 
-                     lapply(c2l(1,2,3), function(index) {
-                             model <- lm(abund.genus[,name] ~ pcoa[,index])
+abund.df <- merge(pcoa.vectors %>%
+                  select("Axis.1", "Axis.2", "Axis.3") %>%
+                  tibble::rownames_to_column("id"),
+                  abund.genus %>% select(id, abund.genus.bacteria),
+                  by = "id") %>%
+    merge(.,  meta(pseq.genus) %>% tibble::rownames_to_column("id"), by = "id")
+
+pcoacorabund <- function(df,
+                         genera.names,
+                         axis.names = c("Axis.1", "Axis.2", "Axis.3")) {
+    lapply(c2l(genera.names), function (taxa) 
+        lapply(c2l(axis.names), function(axis) {
+            fo <- sprintf("%s ~ %s", axis, taxa)
+            model <- lm(as.formula(fo), data = df)
             summary(model)$r.squared
@@
 
-pcoacor.genus <- pcoacorabund(abund.genus.bacteria, pcoa.vectors)
+pcoacor.genus <- pcoacorabund(abund.df, genera.names = abund.genus.bacteria)
 pcoacor.genus.df <- map_df(pcoacor.genus, ~as.data.frame(.x), .id="id")
@@
 pcoacor.genus.df %>%
-    arrange(-X3) %>%
-    head(10) %>% kable
+    arrange(-Axis.3) %>%
+    head(20) %>% kable
 ```

### 2019-06-03 comparepseqs.Rmd
Change R16S file

@@
             "ncbi" = "data/phfinrisk_custom70_species_all_2019-01-29.RDs",
-            "R16S" = "data/phfinrisk_16S_otu_raw_2019-02-19.RDs")
+            "R16S" = "data/phfinrisk_16S_otu_genus_2019-02-20.RDs")
 ```
@@
     if (length(missing) > 0) {
+
         pseq <- import_filter_data(rds[[key]],

### 2019-06-03 articletwo.Rmd
Transitioning to 3.6.0

@@
 library(reshape)
-library(Maaslin)
 library(parallel)
@@
 ```{r data}
-pseq.genus <- import_filter_data("data/phfinrisk_centrifuge_genus_all_drop50k_2018-11-30.RDs")
-pseq.species <- import_filter_data("data/phfinrisk_centrifuge_species_all_2018-11-30.RDs")
+pseq.genus <- import_filter_data("data/phfinrisk_centrifuge_genus_d50k_2019-06-02.rds")
+pseq.species <- import_filter_data("data/phfinrisk_centrifuge_species_all_drop50k_2018-11-30_2019-06-03.RDs")
 ```

### 2019-06-04 comparepseqs.Rmd
Ready for rmd

@@
 
+# Calculate alpha and beta diversity
+
 ```{R models}
@@
 
+```{r quit subruns}
+if (exists("subset.run")) {
+    quit()
+}
+```
+
+# Diversities for BP variables
+
 ```{r plot alpha-beta}
-names(rds)
-if (!exists("subset.run")) {
 plotlist <- list()
@@
 g <- arrangeGrob(grobs = plotlist, ncol=2)
-    ggsave(file="cache/plotlist.png", g, height=6*length(plotlist)/1.5, width=10*2)
-
-} else {
-    message("Skipping graphics")
-}
+ggsave(file="cache/plotlist.png", g, height=8*length(plotlist)/1.5, width=10*2)
 ```
 
-```{r debug, eval = FALSE}
-alphalist <- c()
-for (key in names(rds)) {
-    alphalist <- c(alphalist,
-              ret[[key]]$alphadiversity$min %>% pull(estimate),
-              ret[[key]]$alphadiversity$full %>% pull(estimate))
-}
-
-summary(alphalist)
-```
-
-```{r alpha size, eval = FALSE}
-pseq <- import_filter_data(rds[["gtdb"]])
-
-dset <- meta.merge.alphadiversity(pseq) %>%
-    mutate(diversities_shannon = scale(diversities_shannon))
-
-dset %>%
-    select(MAP, diversities_shannon) %>%
-    gather(key, value) %>%
-    group_by(key) %>%
-    summarise_all(funs(N = n(),
-                       min = min(as.numeric(value), na.rm = TRUE),
-                       max = max(as.numeric(value), na.rm = TRUE),
-                       Mean = mean(as.numeric(value), na.rm = TRUE),
-                       SD = sd(as.numeric(value), na.rm = TRUE),
-                       NAs = sum(is.na(value))))  %>%
-    kable()
-
-lm(sprintf(fo.alpha.full, "SYSTM"), data = dset) %>% summary
-```
+<img src="cache/plotlist.png"/>
 
@@
 
-```{r draw alpha distribution scatter plot, eval = FALSE}
+```{r draw alpha distribution scatter plot}
 scatterplot.df <- lapply(ret, function(x) x$alphadistribution) %>%
@@
        dpi = 300)
-
 ```
 
-Full model R2
+<img src="cache/ggpairs-alphas.png"/>
 
-```{r diff beta debug, eval = FALSE}
+# Beta diversity
+
+Beta diversity R2
+
+```{r beta diversity r2}
 betadf <- data.frame()
@@
 
-```{r debugging adonis, eval = FALSE}
-pseq <- import_filter_data(rds[["gtdb"]])
+Beta diversity p.values
 
-dset <- meta.merge.alphadiversity(pseq) %>%
-    mutate(diversities_shannon = scale(diversities_shannon))
-
-bray.dist.matrix <- readRDS(file = paste0("rds-comparepseqs/bray-gtdb.rds"))
-
-dset.sample <- dset %>% sample_n(50)
-
-dset.ids <- dset.sample %>% pull(Sample_ID)
-
-betadiversity <- list(min = calculateadonis(dset = dset.sample,
-                                            matrix = bray.dist.matrix[dset.ids, dset.ids],
-                                            covariates = covariates.min,
-                                            npermutations = 10000),
-                      full = calculateadonis(dset = dset.sample,
-                                      matrix = bray.dist.matrix[dset.ids, dset.ids],
-                                      covariates = covariates.full,
-                                      npermutations = 2))
-
-betadiversity$full
-
+```{r beta diversity pvalue}
 betadf <- data.frame()
 for (key in names(rds)) {
-    moddf.full <- Reduce(function(...) merge(..., all=T), betadiversity$full) %>%
+    moddf.full <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$full) %>%
         dplyr::filter(response == term) %>%
         mutate(model = key, covariates = "full") %>%
-        select(term, model, covariates, R2)
-    moddf.min <- Reduce(function(...) merge(..., all=T), betadiversity$min) %>%
+        select(term, model, covariates, `Pr(>F)`)
+    moddf.min <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$min) %>%
         dplyr::filter(response == term) %>%
         mutate(model = key, covariates = "min") %>%
-        select(term, model, covariates, R2)
+        select(term, model, covariates, `Pr(>F)`)
     betadf <- rbind(betadf, moddf.full, moddf.min)
 }
-betadf %>% arrange(term) %>% spread(model, R2) %>% kable
-
+betadf %>% arrange(term) %>% spread(model, `Pr(>F)`) %>% kable
 ```

### 2019-06-04 articletwo-ggplot.R
Beta plot show significance

@@ -60,8 +60,10 @@ plot.beta <- function(diversity,
                       fig.permar2.text = "R2 of variable \nin Bray-Curtis distance",
                       fig.textsize = 15,
                       ymax = 0.0016) {
-    plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2)) +
-        geom_bar(stat="identity", color="black") +
+    diversity <- diversity %>% mutate(color = as.factor(ifelse(R2.p < 0.05, 1, 0)))
+    plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2, color = color)) +
+        geom_bar(stat="identity") +
+        scale_color_manual(values = c("0" = "red", "1" = "black")) +
         coord_flip() +
         theme_classic(fig.textsize) +
         ylab(fig.permar2.text) +

### 2019-06-04 articletwo-rrbiome.R

@@ -208,7 +208,8 @@ mergediversities <- function(alphadiversity, betadiversity,
         beta.rbind <- betadiversity[[x]] %>%
             data.table::rbindlist(id = "model_name") %>%
             dplyr::filter(term %in% responses) %>%
-            dplyr::select(response, R2)
+            dplyr::rename(R2.p = `Pr(>F)`) %>%
+            dplyr::select(response, R2, R2.p)
         merge( alpha.select , beta.rbind, by = "response") %>%
             merge(names.dset %>% select(Covariate, Category, Name), by.x ="response", by.y = "Covariate") %>%
             mutate(Qstar = ifelse(p.value < 0.05, '*', ' '))

### 2019-06-04 articletwo.Rmd
Maaslin2

@@
 }
+
+#packrat::opts$ignored.packages
 ```
@@
 library(purrr)
+library(Maaslin2)
 ```
@@
 
-### EdgeR plot functions
-
-<details>
-  <summary>Open/Close</summary>
-
-```{r EdgeR functions, code=readLines("aaro/edgeR_plot_functions.R")}
-source("aaro/edgeR_plot_functions.R")
-```
-
-</details>
-
 ### Plot functions
@@
 
-## Models
+## Variables
+
+```{r variables}
+var.BP <- c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")
+var.CL.min <- c("BL_AGE", "SEX")
+var.CL <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
+            "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
+```
+
+## MODELS
 
@@
                                            matrix = bray.dist.m.species,
-                                           covariates = c("BL_AGE", "SEX"),
-                                           npermutations = 1000)
+                                           covariates = var.CL.min,
+                                           npermutations = 10000)
      adonis.species$max <- calculateadonis(dset = meta(pseq.species),
                                            matrix = bray.dist.m.species,
-                                           covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
-                                                          "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                                           npermutations = 1000)
+                                           covariates = var.CL,
+                                           npermutations = 10000)
      saveRDS(adonis.species, file = "rds/adonis.species.rds")
@@
 }
-```
-
-```{r check adonis, eval = FALSE}
-ad.min <- adonis(bray.dist.m.species ~ DIASM + BL_AGE + SEX,
-             data = meta(pseq.species),
-             permutations = 100)
-
-ad.max <- adonis(bray.dist.m.species ~ DIASM + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
-                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09,
-             data = meta(pseq.species),
-             permutations = 1000)
-
 ```
@@
 
-    maaslin <- list()
-    genus.core.bacteria <- pseq.genus.core %>%
-        abundances %>%
-        rownames %>%
-        replace.brackets %>%
-        mygrep(".Bacteria$", .)
-    maaslin$genus <- maaslinwrapper(pseq = pseq.genus.core,
-                                    looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE",
-                                               "HYPERTENSION"),
-                                    forced = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
-                                               "DIAB", "BL_USE_RX_C03",
-                                               "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                                    taxa = paste0(genus.core.bacteria, collapse = ","))
+    maaslin.pheno <- t(microbiome::meta(phyloseq::sample_data(pseq.genus))) %>%
+        t %>%
+        as.data.frame %>%
+        select(var.BP, var.CL)
+
+    maaslin.abund <- abundances(pseq.genus.core, transform = "compositional") %>%
+        t %>%
+        as.data.frame %>%
+        select(-contains("virus")) %>%
+        select(1:20)
+
+    maaslin <- Maaslin2(maaslin.abund,
+             maaslin.pheno,
+             paste0("cache/maaslin-", now),
+             transform = 'AST',
+             fixed_effects = paste0(var.CL, collapse = ","), 
+             cores = 10)
 

### 2019-06-04 articletwo.Rmd
Minor

@@
 
-```{r variables}
+```{r my variables}
 var.BP <- c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")

### 2019-06-04 aaro/edgeR_plot_functions.R
Removing files

@@ -2,141 +2,10 @@ library(reshape2)
 library(scales)
 library(ggpubr)
 # plot heatmap output of edgeR heatmap
-plot_edgeR_fc_heatmap <- function (tables.df, xlab.text = "Phyla", logFC.maxabs=NULL) {
-
-  if (nrow(tables.df) == 0) {
-    cat("empty heatmap\n")
-    p <- ggplot(data.frame()) + geom_blank()
-    return(p)
-  }
-
-  p <- ggplot(tables.df, aes(x = OTU, y = Name, fill=logFC)) + geom_tile()
-
-  if (is.null(logFC.maxabs)) {
-    p <- p + scale_fill_distiller(palette = "Spectral", name = 'log fold-change')
-  } else {
-    p <- p + scale_fill_distiller(palette = "Spectral", name = 'log fold-change',
-                                  limits=c(-logFC.maxabs, logFC.maxabs))
-  }
-  p <- p + xlab(xlab.text) + ylab("")
-  p <- p + theme_classic()
-  p <- p + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
-  p <- p + theme(axis.text.y = element_text(size=5))
-
-  p
-}
-
-plot_edgeR_fc_heatmap_reps <- function (tables.df, xlab.text = "Phyla", logFC.maxabs=NULL) {
-
-  if (nrow(tables.df) == 0) {
-    cat("empty heatmap\n")
-    p <- ggplot(data.frame()) + geom_blank()
-    return(p)
-  }
-
-  p <- ggplot(tables.df, aes(x = OTU, y = Name, fill=logFC)) + geom_tile()
-
-  p <- p + geom_text(aes(label = paste0("P ",
-                                        sprintf("%0.4f", round(PValue, digits=4)),
-                                        " rep ",
-                                        sprintf("%0.4f", round(PValue.reps, digits=4)),
-                                        "\nQ ",
-                                        sprintf("%0.4f", round(Qval, digits=4)),
-                                        " rep ",
-                                        sprintf("%0.4f", round(Qval.reps, digits=4)),
-                                        "\nlog FC ",
-                                        sprintf("%0.4f", round(logFC, digits=4)),
-                                        " rep ",
[…]

### 2019-06-04 articletwo-rrbiome.R

@@ -284,4 +284,99 @@ mykable <- function(x, ...) {
 }
 
 
+my_na_filter_aaro <- function(phf_f, variables) {
+  tmp_samples <- sample_data(phf_f)
+  tmp_samples$names <- rownames(tmp_samples)
+  sample_data(phf_f) <- tmp_samples
+  filtered_data <- filter_at(sample_data(phf_f),
+                                  variables,
+                                  all_vars(!is.na(.)))
+
+  rownames(filtered_data) <- filtered_data$names
+
+  sample_data(phf_f) <- filtered_data
+  phf_f
+}
+
+
+fit_asine_aaro <- function(phf_f,
+                           variables,
+                           intermediate.name="_asin1_",
+                           extra.printouts=FALSE,
+                           save.intermediate=FALSE,
+                           intermediate.list.file="data_private/processed_asine.txt",
+                           return.only.filename=FALSE) {
+
+  phf_f <- my_na_filter_aaro(phf_f, variables)
+  cat("\nafter na filter, fitting asin(sqrt(OTU))")
+
+  variables <- as.vector(remove_unnec_vars(variables, phf_f, last(variables)))
+
+  modelstr <- paste(variables, collapse=" + ")
+  modelstr <- paste0("~ ", modelstr)
+  cat(modelstr)
+  cat("\n")
+
+  ytaxa <- taxa(phf_f)
+  cat(paste("for ", length(ytaxa), " OTUs\n"))
+
+  pheno <- meta(sample_data(phf_f))[,variables]
+  pheno[,"sample.ids"] <- rownames(pheno)
+
+
+  fits <- lapply(1:length(ytaxa), function (gi) {
+                   g <- ytaxa[[gi]]
+                   if (extra.printouts) {
+                     cat(paste(as.character(g), gi, "/", length(ytaxa)))
+                     cat("\n")
+                   }
+
+                   x <- as.data.frame(t(otu_table(phf_f)[as.character(g),]))
+                   tmp <- rownames(x)
+                   genus.clear <- taxa2underscore(g)
+                   colnames(x)[[1]] <- genus.clear
+                   x[,"sample.ids"] <- tmp
+                   x <- merge(x, pheno, by="sample.ids", all.x=TRUE)
+                   modelstr <- paste0("asin(sqrt(", genus.clear, ")) ", modelstr)
+                   lmwrap <- function (modelstr, x) {
+                     lm(as.formula(modelstr), data=x)
+                   }
[…]

### 2019-06-04 fit_asine_aaro.R

@@ -1,233 +1,0 @@
-# some useful functions if importing finriskmetagcommon does not work outside FIMM Atlas
-
-# also I can't compile the package anyway as long Atlas is down
-
-
-fit_asine_aaro_my_na_filter <- function(phf_f, variables) {
-  tmp_samples <- sample_data(phf_f)
-  tmp_samples$names <- rownames(tmp_samples)
-  sample_data(phf_f) <- tmp_samples
-  filtered_data <- filter_at(sample_data(phf_f),
-                                  variables,
-                                  all_vars(!is.na(.)))
-
-  rownames(filtered_data) <- filtered_data$names
-
-  sample_data(phf_f) <- filtered_data
-  phf_f
-}
-
-
-#' Remove variables that have only one non-NA value when given fixed covariate is present.
-#'
-#' @importFrom phyloseq get_variable
-#' @importFrom dplyr n_distinct filter
-#' @importFrom stats na.omit
-#'
-#' @export
-fit_asine_aaro_remove_unnec_vars <- function(variables, phf_f, fixed.variable) {
-  if (class(phf_f) %in% c("phyloseq", "sample_data"))
-    sample.df <- meta(phf_f)
-  if (class(phf_f) == "data.frame")
-    sample.df <- phf_f
-  sample.df <- filter(sample.df, !is.na(sample.df[,fixed.variable]))
-
-  variables_tmp=c()
-  for (vi in 1:length(variables)) {
-    vr.content <- sample.df[,variables[vi]]
-    if (is.factor(vr.content) && (n_distinct(na.omit(vr.content)) < 2)) {
-      cat(paste("dropping factor", variables[vi], ", only one level\n"))
-    } else if (is.numeric(vr.content) && n_distinct(na.omit(vr.content)) < 2) {
-      cat(paste("dropping numeric", variables[vi], ", only one value\n"))
-    } else {
-      variables_tmp <- c(variables_tmp, variables[vi])
-    }
-  }
-  variables <- variables_tmp
-}
-
-
-
-fit_asine_aaro_fit_adonis <- function (phf.rel, variables, save.intermediate=FALSE, intermediate.name="phyla",
-                        dist.matrix=NULL, ...) {
-  # dist.matrix if supplied must be dist object and contain labels
-
-  # filter out rows with na's
-  phf.rel <- fit_asine_aaro_my_na_filter(phf.rel, variables)
-
-  variables <- as.vector(fit_asine_aaro_remove_unnec_vars(variables, phf.rel, last(variables)))
-
[…]

### 2019-06-04 aaro/comp_fig3_plots.R
Not limited to twenty taxa

@@ -1,106 +1,0 @@
-
-fig.height <- 4.0 + nrow(diversity) * 0.33
-fig.permar2.text <- "R2 of variable \nin Bray-Curtis distance" # TODO latexify expression?
-fig.textsize <- 15
-
-diversity <- diversity %>%
-    mutate(Qstar = ifelse(p.value < 0.5, '*', ' '))
-
-amaxb <- diversity %>% pull(estimate) %>% abs %>% max %>% signif(., digits = 2)
-
-
-plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2)) +
-    geom_bar(stat="identity", color='black') +
-    coord_flip() +
-    theme_classic(fig.textsize) +
-    ylab(fig.permar2.text) +
-    theme(axis.text.y = element_text(color="black"),
-          axis.text.x = element_text(color="black"),
-          legend.position="right",
-          legend.direction="vertical",
-          legend.box.margin=margin(0,0,0,0,"pt"),
-          legend.margin=margin(0,0,0,0,"pt"),
-          legend.justification="left",
-          legend.title.align = 0,
-          legend.text.align = 0)
-
-plot.gtable.beta.diversity <- ggplot_gtable(ggplot_build(plot.beta.diversity))
-plot.xlab.beta.diversity <- xlabb_extractor(gtable.beta.diversity)
-
-plot.beta.diversity <- plot.beta.diversity  +
-    theme(axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.title.x  =  element_blank(),
-          axis.title.y = element_blank(),
-          legend.position = "none")
-plot.gtable.beta.diversity.stripped <- ggplot_gtable(ggplot_build(plot.beta.diversity))
-
-
-
-plot.alpha.diversity <- ggplot(diversity, aes(x=reorder(Name, R2),  y = 1, fill = estimate)) +
-    geom_bar(stat="identity", color="black") +
-    coord_flip() +
-    scale_fill_gradientn(colours = c("darkblue", "blue", "white", "red", "darkred"),
-                         values = rescale(c(-amaxb,-0.1, 0, 0.1, amaxb)),
-                         name = 'Regression coefficient \nin linear model \nfor Shannon index',
-                         limits = c(-1.05*amaxb, 1.05*amaxb),
-                         na.value = "black") +
-    theme_classic(20) +
[…]

### 2019-06-04 aaro/comp_supp_pca_subnet_regres_results_heatmap.R

@@ -1,214 +1,0 @@
-library(biomformat)
-library(microbiome)
-library(phyloseq)
-library(data.table)
-library(RColorBrewer)
-library(dplyr)
-packageVersion('phyloseq')
-library(finriskmetagcommon)
-library(ggplot2)
-library(gridExtra)
-library(cowplot)
-library(ggpubr)
-library(grid)
-library(gtable)
-library(scales)
-
-source('code/phyloseq/generic_functions.R')
-source("code/phyloseq/taxa_names_underscored.R")
-source('code/phyloseq/netw_plot_funs.R')
-source('code/phyloseq/lm_comp_functions.R')
-
-source('code/phyloseq/netw_plot_funs.R')
-
-
-source('code/phyloseq/edgeR_functions.R')
-source('code/phyloseq/edgeR_plot_functions.R')
-
-maindir <- "/csc/fr_metagenome/microbiome_scratch/scratch/"
-datadir <- file.path(maindir, "data_private/")
-
-pltdir <- 'docs/assets/PC_subnet_pheno_association/'
-if (!dir.exists(pltdir)) {
-  dir.create(pltdir)
-}
-
-pseqdate <- "2018-12-21"
-phfg.pseq <- readRDS(paste0(datadir, "phfinrisk_genus_ok_drop50k_", pseqdate, ".RDs"))
-pseq.rel.core <- core(microbiome::transform(phfg.pseq, "compositional"), detection=0.1/100, prevalence=1/100)
-
-pseq.rel.core.rel.clr <- transform(transform(pseq.rel.core, "compositional"), "clr")
-
-
-# latest sample data
-phfg.pseq.all <- readRDS(paste0(datadir, "phfinrisk_genus_all_drop50k_", pseqdate, ".RDs"))
-
-data("phfinrisk_metadatadesc", package="finriskmetagcommon")
-names.df <- filter(phfinrisk_metadatadesc,
-                   Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115 == 0)
-phfg.pseq <- drop_unused_samplevars(phfg.pseq.all, names.df)
-
-# read subnet info + rel abundances of subnets:
-
-# from comp_fig2_plots_d50k.R
-n.components <- readRDS(paste0(datadir, "spieceasi_0_", pseqdate,
-                              "_components_fig2.RDs"))
-
-n.relabunds.core.genera <- abundances(pseq.rel.core)
[…]

### 2019-06-04 aaro/edgeR_plot_functions.R

@@ -1,270 +1,0 @@
-library(reshape2)
-library(scales)
-library(ggpubr)
-# plot heatmap output of edgeR heatmap
-
-
-
-
-
-plot_maaslin_heatmap_final <- function (tables.df, beta.maxabs=NULL, text.size=10,
-                                        cluster=TRUE, signif=NULL, signif.th=0.05,
-                                        fname='Feature', pname='Name',vname='beta',
-                                        qvalues=NULL, colours=NULL, breaks=NULL, names.df=NULL,
-                                        order.by.category=FALSE, ystr=NULL,
-                                        v.title.name="Effect size",
-                                        sign.name="Significance", no.sign.legend=FALSE,
-                                        discrete.colors=FALSE,
-                                        vname_discrete=NULL) {
-
-
-  # otherwise the same as edgeR heatmap plotter, but relevant df columns are different
-
-  if (is.null(beta.maxabs) && !discrete.colors) {
-    beta.maxabs <- max(abs(tables.df[,vname]))
-    cat("auto choose bet.maxbs: ")
-    cat(beta.maxabs)
-    cat("\n")
-  }
-
-  if (nrow(tables.df) == 0) {
-    cat("empty heatmap\n")
-    p <- ggplot(data.frame()) + geom_blank()
-    return(p)
-  }
-
-  if (!is.null(signif)) {
-    tables.df[,'Qlevel'] <- tables.df[,signif] < signif.th
-    tables.df[,'Qstar'] <- sapply(tables.df[,'Qlevel'], function (x) ifelse(x,'*',' '))
-  }
-
-  # ensure not significant pairs (no rows) are set to NA
-
-  all.otus <- unique(tables.df[,fname])
-  all.names <- unique(tables.df[,pname])
-
-  all.pairs <- expand.grid(Feature=all.otus, Name=all.names)
-
-  tables.df <- merge(tables.df, all.pairs, all.y=TRUE, by.y=c('Feature','Name'), by.x=c(fname,pname))
-  cat(colnames(tables.df))
-  if (!is.null(names.df)) {
-    for (pn in as.character(tables.df[[pname]])) {
-      tables.df[tables.df[,pname] == pn, 'Category'] <- as.character(names.df[names.df[,pname]==pn, 'Category'])
-    }
-  }
-
-
-
-  # plot
-  if (cluster) {
[…]

### 2019-06-04 aaro/plot_catcol_palette.R

@@ -1,51 +1,0 @@
-library(ggsci)
-library(dplyr)
-
-# the main function is scale_fill_my_catcol
-# which can be used as the usual ggplot scale_fill_*() functions
-
-get_catcol_palette <- function() {
-  if (!exists('phfinrisk_metadatadesc')) {
-    data("phfinrisk_metadatadesc", package="finriskmetagcommon")
-  }
-  my_catcol_palette <- c("#00aaaa",
-                         "#225522",
-                         "#cccccc",
-                         "#777777",
-                         "#bb9977",
-                         "#00aa00")
-  allcats <- unique(phfinrisk_metadatadesc$Category)
-  propercats <- allcats[!allcats %in% c(NA, "Unknown")] 
-  names(my_catcol_palette) <- propercats
-  my_catcol_palette
-}
-
-
-get_catcol_palette_grgr <- function() {
-  if (!exists('phfinrisk_metadatadesc')) {
-    data("phfinrisk_metadatadesc", package="finriskmetagcommon")
-  }
-
-
-  pal <- c("#00aa00", "#000000","#eeeeee")
-  my_catcol_palette <-  colorRampPalette(pal)(n_distinct(phfinrisk_metadatadesc$Category))
-  names(my_catcol_palette) <- unique(phfinrisk_metadatadesc$Category)
-  my_catcol_palette
-}
-
-
-get_catcol_palette_gray <- function() {
-  if (!exists('phfinrisk_metadatadesc')) {
-    data("phfinrisk_metadatadesc", package="finriskmetagcommon")
-  }
-  my_catcol_palette <- gray(seq(0.0, 0.9, length=n_distinct(phfinrisk_metadatadesc$Category)))
-  names(my_catcol_palette) <- unique(phfinrisk_metadatadesc$Category)
-  my_catcol_palette
-}
-
-
-scale_fill_my_catcol <- function() {
-  #scale_fill_manual(name="Metadata category", values=get_catcol_palette())
-  #scale_fill_uchicago(name="Metadata category")
-  scale_fill_startrek(name="Metadata category")
-}

### 2019-06-04 articletwo.Rmd
Not limited to twenty taxa
microbiome global renamed alpha

@@
         as.data.frame %>%
-        select(-contains("virus")) %>%
-        select(1:20)
+        select(-contains("virus"))
 
@@
              fixed_effects = paste0(var.CL, collapse = ","), 
-             cores = 10)
+             cores = detectCores())
 
@@
 dset.alpha <- meta.merge.alphadiversity(pseq.species) %>%
-    mutate(diversities_shannon = scale(diversities_shannon),
+    mutate(diversity_shannon = scale(diversity_shannon),
            MAP = oMAP,
@@
     min = calculateglm(dset.alpha,
-                       modelstring = "%s ~ BL_AGE + SEX  + diversities_shannon"),
+                       modelstring = sprintf("%%s ~ %s + diversity_shannon",
+                                             paste0(var.CL.min, collapse = " + ")),
+                       filterstr = "shannon"),
     max = calculateglm(dset.alpha,
-                       modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + diversities_shannon"))
-
-diversities <- mergediversities(alphadiversity, adonis.species)
+                       modelstring = sprintf("%%s ~ %s + diversity_shannon",
+                                             paste0(var.CL, collapse = " + ")),
+                       filterstr = "shannon"))
 ```
@@
 
+Difference of R2 between full and minimal models
+
 ```{r difference in beta diversity}
-merge(diversities$min %>% select(response, R2),
-      diversities$max %>% select(response, R2), by = "response") %>%
-    mutate(R2.diff = R2.y - R2.x) %>%
+merge(diversities$min %>%
+      select(response, R2, R2.p) %>%
+      dplyr::rename(R2.min = R2, p.min = R2.p),
+      diversities$max %>%
+      select(response, R2, R2.p) %>%
+      dplyr::rename(R2.full = R2, p.full = R2.p),
+      by = "response") %>%
+    mutate(R2.diff = R2.full - R2.min) %>%
     kable
@@
 
-Results for beta diversity
-
-```{r table beta diversity}
-adonis.species$max %>%
-    data.table::rbindlist(id = "model_name") %>%
-    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
-    dplyr::select(response, R2, "Pr(>F)") %>%
-    mutate(p.value = pub.p(`Pr(>F)`)) %>%
-    kable
-```
-
 ```{r save grob}
-alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 0.6)
-alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 0.6)
+alpha.diversity.min <- plot.alpha(diversities$min, alphabreaks = seq(-1,0,0.2))
+alpha.diversity.max <- plot.alpha(diversities$max, alphabreaks = seq(-1,0,0.2))
 
@@
 
-abund.genus.bacteria <- colnames(abund.genus)[-1] %>% mygrep(".bacteria$", .)
+abund.genus.bacteria <- abund.genus %>% select(-contains("virus")) %>% colnames
 
@@
 
+Significant associations between BP-variables and first three PCoA axis
+
 ```{r pcoa axis analysis axis 1}
+glm.pcoa <- lapply(c2l("Axis.1", "Axis.2", "Axis.3"), function(axis)
     calculateglm(pcoa.df,
-             modelstring = "%s ~ Axis.1 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
-                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
-             filterstr = "Axis") %>% 
+                 modelstring = sprintf("%%s ~ %s + %s", axis, paste(var.CL, collapse = "+")),
+                 filterstr = "Axis"))
+
+do.call(rbind, glm.pcoa) %>% 
     select(response, term, estimate, std.error, p.value) %>%
     mutate_if(is.numeric, round, 3) %>%
+    filter(p.value < 0.05) %>%
+    arrange(p.value) %>%
     kable
-```
-
-Testing associations to blood pressure variables on PCoA axis 2
-
-```{r pcoa axis analysis axis 2}
-calculateglm(pcoa.df,
-             modelstring = "%s ~ Axis.2 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
-                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
-             filterstr = "Axis") %>% 
-    select(response, term, estimate, std.error, p.value) %>%
-    mutate_if(is.numeric, round, 3) %>%
-    kable
-```
-
-Testing associations to blood pressure variables on PCoA axis 3
-
-```{r pcoa axis analysis axis 3}
-calculateglm(pcoa.df,
-             modelstring = "%s ~ Axis.3 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
-                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
-             filterstr = "Axis") %>%
-    select(response, term, estimate, std.error, p.value) %>%
-    mutate_if(is.numeric, round, 3) %>%
-    kable
-```
-
-```{r pcoa 3d}
-theta <- 130
-phi <- 10
-pcoa3d.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y = Axis.2, z=Axis.3, color = HYPERTENSION)) +
-  axes_3D(theta=theta, phi=phi) +
-  stat_3D(theta=theta, phi=phi, size = 0.01, alpha=0.4) +
-  labs_3D(theta=theta, phi=phi, 
-          hjust=c(1,0,0), vjust=c(1.5,1,-.2),
-          labs=c(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1]),
-                 sprintf("PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2]),
-                 sprintf("PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3]))) #+
-#  theme_void()
-
-ggsave(file = "cache/pcoa-species-3d.png", plot=pcoa3d.plot, height=4, width=7, units = "in", dpi = 300)
-
 ```

### 2019-06-04 articletwo-rrbiome.R

@@ -49,7 +49,7 @@ filter.phenotype.data <- function(pseq,
 }
 
 meta.merge.alphadiversity <- function(pseq, index = "shannon") {
-    alphadiversity  <- microbiome::global(pseq, index = index)
+    alphadiversity  <- microbiome::alpha(pseq, index = index)
     base::merge(meta(pseq), alphadiversity, by=0, all=TRUE) %>%
         dplyr::rename(Sample_ID = Row.names) %>%
             dplyr::mutate(BL_AGE_GROUP = group_age(BL_AGE))
@@ -68,7 +68,7 @@ calculateglm <- function(dset,
                                           model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
                          min_n_for_continuous = 10,
                          modelstring = "%s ~ BL_AGE + SEX",
-                         filterstr = "diversities_") {
+                         filterstr = ".") {
     glmlist <- lapply(responses, function(response) {
         fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
         fo <- sprintf(modelstring, response)

### 2019-06-05 articletwo.Rmd
Packrat, param

@@
 
- ```{r adonis calculation}
+ ```{r adonis calculation, eval = FALSE}
  if (!file.exists("rds/adonis.species.rds")) {
@@
              transform = 'AST',
+		normalization = "NONE",
+		random_effects = paste0(var.BP, collapse = ","),
              fixed_effects = paste0(var.CL, collapse = ","), 

### 2019-06-05 articletwo.Rmd
Debug

--- 2019-06-04 articletwo.Rmd
@@
 ```{r maaslin runs, results = 'hide'} 
+input_data <- system.file(
+    'extdata','example1_features.txt', package="Maaslin2")
+input_metadata <-system.file(
+    'extdata','example1_metadata.txt', package="Maaslin2")
+
+fit_data <- Maaslin2(
+    input_data, input_metadata,'maaslin2_example_output')
+
 if (!file.exists("rds/maaslin.rds")) {
@@
         as.data.frame %>%
-        select(var.BP, var.CL)
+        select(var.BP, var.CL) %>%
+        tibble::rownames_to_column("#")
 
@@
         as.data.frame %>%
-        select(-contains("virus"))
+        select(-contains("virus")) %>%
+        select(g_Yersinia, g_Zavarzinella, g_Zhongshania, g_Zobellia, g_Zunongwangia) %>% #DEBUG
+        tibble::rownames_to_column("#") %>%
+        sample_n(50)
 

### 2019-06-05 articletwo.Rmd
Aaro's asine

--- 2019-06-05 articletwo.Rmd
@@
 
+```{r fit asine}
+source("fit_asine.R")
+source("taxa_names_functions.R")
+```
+
+```{r maaslin}
+maaslin <- read.table("maaslin1/maaslin/maaslin.txt", header=TRUE, sep='\t')
+
+maaslin %>% filter(Q.value < 0.05)
+```
+
 ```{r maaslin runs, results = 'hide'} 
@@
         as.data.frame %>%
+        tibble::rownames_to_column("ids") %>%
+        tibble::column_to_rownames("ids") %>%
         select(var.BP, var.CL)
@@
         t %>%
-        as.data.frame %>%
-        select(-contains("virus"))
+        as.data.frame
+
+    maaslin <- Maaslin2(maaslin.abund %>% select(g_Sutterella, g_Adlercreutzia, g_Anaerotruncus),
+                        maaslin.pheno,
+                        "maaslin2_example_output",
+                        transform = "AST",
+                        normalization = "NONE",
+                        fixed_effects = paste0(var.CL, collapse = ","),
+                        random_effects = paste0(var.BP, collapse = ","))
+
+    random_formula = as.formula("expr ~ (1 | MAP) + (1 | SYSTM) + (1 | DIASM) + (1 | PULSEPRESSURE) + (1 | HYPERTENSION)")
+    formula = as.formula("expr ~  BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09")
+
+    formula2 <-
+                    paste(
+                        '. ~', 
+                        paste(all.vars(formula)[-1], collapse = ' + '), 
+                        '.', 
+                        sep = ' + ')
+
+    formula2
+
+    update(random_formula, formula2)
+
+    maaslin.pheno %>% head(n=1)
+
+    stats::glm(BL_AGE ~ SEX + (1|MAP), family = stats::gaussian, data = meta(pseq.genus.core))
+
+    maaslin$results %>% head(n=1)
+
+    maaslin.abund[1:10,1:10]
+    maaslin.pheno[1:10,1:10]
 
@@
              random_effects = paste0(var.BP, collapse = ","),
-             fixed_effects = paste0(var.CL, collapse = ","), 
+             fixed_effects = paste0(var.CL[1:3], collapse = ","), 
              cores = detectCores())

### 2019-06-05 fit_asine.R

@@ -1,0 +1,233 @@
+# some useful functions if importing finriskmetagcommon does not work outside FIMM Atlas
+
+# also I can't compile the package anyway as long Atlas is down
+
+
+my_na_filter <- function(phf_f, variables) {
+  tmp_samples <- sample_data(phf_f)
+  tmp_samples$names <- rownames(tmp_samples)
+  sample_data(phf_f) <- tmp_samples
+  filtered_data <- filter_at(sample_data(phf_f),
+                                  variables,
+                                  all_vars(!is.na(.)))
+
+  rownames(filtered_data) <- filtered_data$names
+
+  sample_data(phf_f) <- filtered_data
+  phf_f
+}
+
+
+#' Remove variables that have only one non-NA value when given fixed covariate is present.
+#'
+#' @importFrom phyloseq get_variable
+#' @importFrom dplyr n_distinct filter
+#' @importFrom stats na.omit
+#'
+#' @export
+remove_unnec_vars <- function(variables, phf_f, fixed.variable) {
+  if (class(phf_f) %in% c("phyloseq", "sample_data"))
+    sample.df <- meta(phf_f)
+  if (class(phf_f) == "data.frame")
+    sample.df <- phf_f
+  sample.df <- filter(sample.df, !is.na(sample.df[,fixed.variable]))
+
+  variables_tmp=c()
+  for (vi in 1:length(variables)) {
+    vr.content <- sample.df[,variables[vi]]
+    if (is.factor(vr.content) && (n_distinct(na.omit(vr.content)) < 2)) {
+      cat(paste("dropping factor", variables[vi], ", only one level\n"))
+    } else if (is.numeric(vr.content) && n_distinct(na.omit(vr.content)) < 2) {
+      cat(paste("dropping numeric", variables[vi], ", only one value\n"))
+    } else {
+      variables_tmp <- c(variables_tmp, variables[vi])
+    }
+  }
+  variables <- variables_tmp
+}
+
+
+
+fit_adonis <- function (phf.rel, variables, save.intermediate=FALSE, intermediate.name="phyla",
+                        dist.matrix=NULL, ...) {
+  # dist.matrix if supplied must be dist object and contain labels
+
+  # filter out rows with na's
+  phf.rel <- my_na_filter(phf.rel, variables)
+
+  variables <- as.vector(remove_unnec_vars(variables, phf.rel, last(variables)))
+
+  pheno <- meta(sample_data(phf.rel))
+
+  pheno <- pheno[,variables]
+
[…]

### 2019-06-05 taxa_names_functions.R

@@ -1,0 +1,87 @@
+# Helper functions for handling FR02 taxa names
+
+#' Remove some not-very-nice characters from taxa_names(pseq)
+#'
+#' spaces, hyphens, commas to underscores; parentheses to empty string
+#'
+#' @param pseq a phyloseq object
+#'
+#' @return processed taxa_names
+#'
+#' @export
+#'
+#' @importFrom phyloseq taxa_names
+#'
+#' @examples
+#' \dontrun{
+#' taxa_names(pseq) <- taxa_names_underscored(pseq)
+#' }
+taxa_names_underscored <- function (pseq) {
+  taxanames <- taxa_names(pseq)
+  taxanames <- characters_underscored(taxanames)
+  taxanames
+}
+
+
+#' Remove some not-very-nice characters from character vector
+#'
+#' Helper function, intended main interface for pseqs is
+#' taxa_names_underscored.
+#' spaces, hyphens, commas to underscores; parentheses to empty string
+#'
+#' @param taxanames eg output of taxa_names(pseq)
+#'
+#' @return processed taxa_names
+#'
+#' @export
+#'
+#' @examples
+#' \dontrun{
+#' taxa_names(pseq) <- taxa_names_underscored(pseq)
+#' }
+characters_underscored <- function (taxanames) {
+  taxanames <- unlist(lapply(taxanames, function(x) gsub(" ", "_", x)))
+  taxanames <- unlist(lapply(taxanames, function(x) gsub("-", "_", x)))
+  taxanames <- unlist(lapply(taxanames, function(x) gsub(",", "_", x)))
+  taxanames <- unlist(lapply(taxanames, function(x) gsub("\\(", "", x)))
+  taxanames <- unlist(lapply(taxanames, function(x) gsub("\\)", "", x)))
+}
+
+
+#' Return underscored version of a single taxa
+#'
+#' @param x character string
+#'
+#' @return underscored character string
+#'
+#' @export
+taxa2underscore <- function(x) {
+  gsub("\\)", "", gsub("\\(", "", gsub(",", "_", gsub("-", "_",  gsub(" ", "_", x)))))
+}
+
+
+#' Return 'pretty' taxa names for printing and plotting
+#'
+#' @param pseq a phyloseq object
+#'
+#' @return a vector of 'pretty' taxa_names
+#'
+#' @export
+#'
+#' @importFrom microbiome abundances
[…]

### 2019-06-05 articletwo.Rmd
Replacing maaslin?

@@
                       file.path = 'cache/', fig.path = 'cache/')
+
+options(max.print=1000)
 
@@
 source("taxa_names_functions.R")
+
+options(max.print=1000)
+
+pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
+pseq.genus.core.comp <- microbiome::transform(pseq.genus.core, "compositional")
+
+res <- lapply(c2l(var.BP), function(BP)
+    fit_asine(pseq.genus.core.comp, variables = c(BP, var.CL)))
+
+asine.df <- data.frame()
+
+res.trim <- lapply(c2l(var.BP), function(BP)
+    lapply(c2l(names(res[[BP]])), function(name) res[[BP]][[name]] %>%
+                                                 tidy %>%
+                                                 dplyr::filter(grepl(BP, term)) %>%
+                                                 mutate(taxa = name)))
+
+map_df(res.trim, ~do.call(rbind, .x)) %>%
+    dplyr::mutate(Qval = p.adjust(p.value, method="BH")) %>%
+    arrange(Qval)
+
+maaslin <- do.call(rbind, lapply(c2l(var.BP), function(BP) res[[BP]]))
+saveRDS(maaslin, file = "rds/maaslin.rds")
+```
+
+```{r quit}
+quit()
 ```

### 2019-06-05 fit_asine.R

@@ -91,69 +91,52 @@ fit_adonis <- function (phf.rel, variables, save.intermediate=FALSE, intermediat
 }
 
 
-fit_asine <- function(phf_f, variables, intermediate.name="_asin1_", extra.printouts=FALSE,
-                      save.intermediate=FALSE, intermediate.list.file="data_private/processed_asine.txt",
+fit_asine <- function(phf_f,
+                      variables,
+                      intermediate.name="_asin1_",
+                      extra.printouts=FALSE,
+                      save.intermediate=FALSE,
+                      intermediate.list.file="data_private/processed_asine.txt",
                       return.only.filename=FALSE) {
 
-  phf_f <- my_na_filter(phf_f, variables)
-  cat("\nafter na filter, fitting asin(sqrt(OTU))")
+    phf_f <- my_na_filter(phf_f, variables)
+    cat("\nafter na filter, fitting asin(sqrt(OTU))")
 
-  variables <- as.vector(remove_unnec_vars(variables, phf_f, last(variables)))
+    variables <- as.vector(remove_unnec_vars(variables, phf_f, last(variables)))
 
-  modelstr <- paste(variables, collapse=" + ")
-  modelstr <- paste0("~ ", modelstr)
-  cat(modelstr)
-  cat("\n")
+    modelstr <- paste(variables, collapse=" + ")
+    modelstr <- paste0("~ ", modelstr)
+    cat(modelstr)
+    cat("\n")
 
-  ytaxa <- taxa(phf_f)
-  cat(paste("for ", length(ytaxa), " OTUs\n"))
+    ytaxa <- taxa(phf_f)
+    cat(paste("for ", length(ytaxa), " OTUs\n"))
 
-  pheno <- meta(sample_data(phf_f))[,variables]
-  pheno[,"sample.ids"] <- rownames(pheno)
+    pheno <- meta(sample_data(phf_f))[,variables]
+    pheno[,"sample.ids"] <- rownames(pheno)
 
-
-  fits <- lapply(1:length(ytaxa), function (gi) {
-                   g <- ytaxa[[gi]]
-                   if (extra.printouts) {
-                     cat(paste(as.character(g), gi, "/", length(ytaxa)))
-                     cat("\n")
-                   }
-
-                   x <- as.data.frame(t(otu_table(phf_f)[as.character(g),]))
-                   tmp <- rownames(x)
-                   genus.clear <- taxa2underscore(g)
-                   colnames(x)[[1]] <- genus.clear
-                   x[,"sample.ids"] <- tmp
[…]

### 2019-06-05 articletwo.Rmd
Duplicate label
Saving wrong variable

@@
 
-```{r core taxa}
-pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
-```
-
 ```{r fit asine}
@@
 
-options(max.print=1000)
-
+if (!file.exists("rds/maaslin.rds")) {
     pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
@@
         fit_asine(pseq.genus.core.comp, variables = c(BP, var.CL)))
-
-asine.df <- data.frame()
 
@@
 
-map_df(res.trim, ~do.call(rbind, .x)) %>%
+    maaslin <- map_df(res.trim, ~do.call(rbind, .x)) %>%
         dplyr::mutate(Qval = p.adjust(p.value, method="BH")) %>%
@@
 
-maaslin <- do.call(rbind, lapply(c2l(var.BP), function(BP) res[[BP]]))
     saveRDS(maaslin, file = "rds/maaslin.rds")
+} else {
+    maaslin <- readRDS("rds/maaslin.rds")
+}
 ```
@@
 quit()
-```
-
-```{r maaslin}
-maaslin <- read.table("maaslin1/maaslin/maaslin.txt", header=TRUE, sep='\t')
-
-maaslin %>% filter(Q.value < 0.05)
 ```
@@
 
-Catabacter, Robinsoniella
-
-```{r compare maaslin to sun, eval = FALSE}
-maas.results.df.all <- maaslin$genus %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
-    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  Feature = gsub("_Bacteria", "", Feature),
-                  beta = Coefficient)
-
-maas.results.df.all %>% pull(Feature) %>% sort %>% unique %>% mygrep("Robin", .)
-
-genus.taxa <- pseq.genus.core %>% abundances %>% rownames
-
-genus.taxa %>% mygrep("Catabacter", .)
-
-maaslin.test <- maaslinwrapper(pseq = pseq.genus.core,
-                                    looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE",
-                                               "HYPERTENSION"),
-                                    forced = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
-                                               "DIAB", "BL_USE_RX_C03",
-                                               "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
-                               taxa = paste(c("Faecalicatena.Bacteria",
-                                        "Catabacter.Bacteria",
-                                        "Robinsoniella.Bacteria"), collapse = ", "))
-
-```
-
 ```{r scatter plots}

### 2019-06-05 articletwo.Rmd
fit_asine starts to be ok

@@
 
+xor(c(TRUE, TRUE, FALSE), TRUE)
+
 if (!file.exists("rds/maaslin.rds")) {
@@
 
+    pseq.genus.core.viruses <- taxa(pseq.genus.core.comp) %>% mygrep(word = "virus$", complement = TRUE)
+    pseq.genus.core.comp.pruned <- prune_taxa(pseq.genus.core.viruses, pseq.genus.core.comp)
+
     res <- lapply(c2l(var.BP), function(BP)
-        fit_asine(pseq.genus.core.comp, variables = c(BP, var.CL)))
+        fit_asine(pseq.genus.core.comp.pruned, variables = c(BP, var.CL)))
 
-    res.trim <- lapply(c2l(var.BP), function(BP)
-        lapply(c2l(names(res[[BP]])), function(name) res[[BP]][[name]] %>%
-                                                     tidy %>%
-                                                     dplyr::filter(grepl(BP, term)) %>%
-                                                     mutate(taxa = name)))
+    res.trim <- map_df(res, ~map_df(.x, ~tidy(.x), .id = "taxa"), .id = "model")
+
+    adonis <- res.trim %>%
+        filter(term %in% var.BP) %>%
+        group_by(model) %>%
+        dplyr::mutate(q.value = p.adjust(p.value, method="BH")) %>%
+        ungroup %>%
+        arrange(q.value)
+
+    res$MAP %>% names
+
+adonis %>% pull(taxa) %>% sort
+
+    adonis %>% dplyr::filter(taxa == "g_Sutterella")
+
+map_df(res.trim$HYPERTENSION, ~
 

### 2019-06-05 articletwo-rrbiome.R

@@ -252,8 +252,8 @@ maaslinwrapper <- function(pseq, looped, forced, taxa, tempstr = "%s/temp/maasli
     ret
 }
 
-mygrep <- function(word, list, ignorecase = TRUE) {
-    list[grepl(word, list, ignore.case = ignorecase)]
+mygrep <- function(..., word, ignorecase = TRUE, complement = FALSE) {
+    c(...)[xor(grepl(word, c(...), ignore.case = ignorecase), (complement == TRUE))]
 }
 
 cc <- function(...) {
@@ -298,75 +298,6 @@ my_na_filter_aaro <- function(phf_f, variables) {
   phf_f
 }
 
-
-fit_asine_aaro <- function(phf_f,
-                           variables,
-                           intermediate.name="_asin1_",
-                           extra.printouts=FALSE,
-                           save.intermediate=FALSE,
-                           intermediate.list.file="data_private/processed_asine.txt",
-                           return.only.filename=FALSE) {
-
-  phf_f <- my_na_filter_aaro(phf_f, variables)
-  cat("\nafter na filter, fitting asin(sqrt(OTU))")
-
-  variables <- as.vector(remove_unnec_vars(variables, phf_f, last(variables)))
-
-  modelstr <- paste(variables, collapse=" + ")
-  modelstr <- paste0("~ ", modelstr)
-  cat(modelstr)
-  cat("\n")
-
-  ytaxa <- taxa(phf_f)
-  cat(paste("for ", length(ytaxa), " OTUs\n"))
-
-  pheno <- meta(sample_data(phf_f))[,variables]
-  pheno[,"sample.ids"] <- rownames(pheno)
-
-
-  fits <- lapply(1:length(ytaxa), function (gi) {
-                   g <- ytaxa[[gi]]
-                   if (extra.printouts) {
-                     cat(paste(as.character(g), gi, "/", length(ytaxa)))
-                     cat("\n")
-                   }
-
-                   x <- as.data.frame(t(otu_table(phf_f)[as.character(g),]))
-                   tmp <- rownames(x)
-                   genus.clear <- taxa2underscore(g)
-                   colnames(x)[[1]] <- genus.clear
-                   x[,"sample.ids"] <- tmp
-                   x <- merge(x, pheno, by="sample.ids", all.x=TRUE)
-                   modelstr <- paste0("asin(sqrt(", genus.clear, ")) ", modelstr)
-                   lmwrap <- function (modelstr, x) {
[…]

### 2019-06-05 fit_asine.R
Limit cores 8

@@ -135,7 +135,7 @@ fit_asine <- function(phf_f,
             }
         )
         return(lmout)
-    }, mc.cores = parallel::detectCores(), mc.preschedule = FALSE)
+    }, mc.cores = min(8, parallel::detectCores()), mc.preschedule = FALSE)
     return(fits)
 }
 

### 2019-06-05 articletwo.Rmd
For night
Commenting

@@
 
-    res.trim <- map_df(res, ~map_df(.x, ~tidy(.x), .id = "taxa"), .id = "model")
+    saveRDS(res, file = "rds/res.rds")
 
-    adonis <- res.trim %>%
-        filter(term %in% var.BP) %>%
-        group_by(model) %>%
-        dplyr::mutate(q.value = p.adjust(p.value, method="BH")) %>%
-        ungroup %>%
-        arrange(q.value)
+##     res.trim <- map_df(res, ~map_df(.x, ~tidy(.x), .id = "taxa"), .id = "model")
 
-    res$MAP %>% names
+##     adonis <- res.trim %>%
+##         filter(term %in% var.BP) %>%
+##         group_by(model) %>%
+##         dplyr::mutate(q.value = p.adjust(p.value, method="BH")) %>%
+##         ungroup %>%
+##         arrange(q.value)
 
-adonis %>% pull(taxa) %>% sort
+##     res$MAP %>% names
 
-    adonis %>% dplyr::filter(taxa == "g_Sutterella")
+## adonis %>% pull(taxa) %>% sort
 
-map_df(res.trim$HYPERTENSION, ~
+##     adonis %>% dplyr::filter(taxa == "g_Sutterella")
 
-    maaslin <- map_df(res.trim, ~do.call(rbind, .x)) %>%
-        dplyr::mutate(Qval = p.adjust(p.value, method="BH")) %>%
-        arrange(Qval)
+## map_df(res.trim$HYPERTENSION, ~
 
-    saveRDS(maaslin, file = "rds/maaslin.rds")
+##     maaslin <- map_df(res.trim, ~do.call(rbind, .x)) %>%
+##         dplyr::mutate(Qval = p.adjust(p.value, method="BH")) %>%
+##         arrange(Qval)
+
+##     saveRDS(maaslin, file = "rds/maaslin.rds")
 } else {

### 2019-06-07 articletwo.Rmd
Run maaslin on R 3.4.3 by rscript

--- 2019-06-05 articletwo.Rmd
@@
 
-```{r fit asine}
-source("fit_asine.R")
-source("taxa_names_functions.R")
-
+```{r maaslin runs, results = 'hide'} 
 if (!file.exists("rds/maaslin.rds")) {
     pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
-    pseq.genus.core.comp <- microbiome::transform(pseq.genus.core, "compositional")
+    pseq.genus.core.viruses <- taxa(pseq.genus.core.comp) %>% mygrep(word = "virus$", complement = TRUE)
+    pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses, pseq.genus.core.comp)
 
-    res <- lapply(c2l(var.BP), function(BP)
-        fit_asine(pseq.genus.core.comp, variables = c(BP, var.CL)))
-
-    res.trim <- lapply(c2l(var.BP), function(BP)
-        lapply(c2l(names(res[[BP]])), function(name) res[[BP]][[name]] %>%
-                                                     tidy %>%
-                                                     dplyr::filter(grepl(BP, term)) %>%
-                                                     mutate(taxa = name)))
-
-    maaslin <- map_df(res.trim, ~do.call(rbind, .x)) %>%
-        dplyr::mutate(Qval = p.adjust(p.value, method="BH")) %>%
-        arrange(Qval)
+    maaslin <- maaslinwrapper(pseq = pseq.genus.core.pruned,
+                                    looped = var.BP,
+                                    forced = var.CL,
+                              taxa = taxa(pseq.genus.core.pruned))
 
@@
 } else {
-    maaslin <- readRDS("rds/maaslin.rds")
+    maaslin <- readRDS(file = "rds/maaslin.rds")
 }
-```
 
-```{r quit}
-quit()
-```
-
-```{r maaslin runs, results = 'hide'} 
-if (!file.exists("rds/maaslin.rds")) {
-
-    maaslin.pheno <- t(microbiome::meta(phyloseq::sample_data(pseq.genus))) %>%
-        t %>%
-        as.data.frame %>%
-        tibble::rownames_to_column("ids") %>%
-        tibble::column_to_rownames("ids") %>%
-        select(var.BP, var.CL)
-
-    maaslin.abund <- abundances(pseq.genus.core, transform = "compositional") %>%
-        t %>%
-        as.data.frame
-
-    maaslin <- Maaslin2(maaslin.abund %>% select(g_Sutterella, g_Adlercreutzia, g_Anaerotruncus),
-                        maaslin.pheno,
-                        "maaslin2_example_output",
-                        transform = "AST",
-                        normalization = "NONE",
-                        fixed_effects = paste0(var.CL, collapse = ","),
-                        random_effects = paste0(var.BP, collapse = ","))
-
-    random_formula = as.formula("expr ~ (1 | MAP) + (1 | SYSTM) + (1 | DIASM) + (1 | PULSEPRESSURE) + (1 | HYPERTENSION)")
-    formula = as.formula("expr ~  BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09")
-
-    formula2 <-
-                    paste(
-                        '. ~', 
-                        paste(all.vars(formula)[-1], collapse = ' + '), 
-                        '.', 
-                        sep = ' + ')
-
-    formula2
-
-    update(random_formula, formula2)
-
-    maaslin.pheno %>% head(n=1)
-
-    stats::glm(BL_AGE ~ SEX + (1|MAP), family = stats::gaussian, data = meta(pseq.genus.core))
-
-    maaslin$results %>% head(n=1)
-
-    maaslin.abund[1:10,1:10]
-    maaslin.pheno[1:10,1:10]
-
-    maaslin <- Maaslin2(maaslin.abund,
-             maaslin.pheno,
-             paste0("cache/maaslin-", now),
-             transform = 'AST',
-             normalization = "NONE",
-             random_effects = paste0(var.BP, collapse = ","),
-             fixed_effects = paste0(var.CL[1:3], collapse = ","), 
-             cores = detectCores())
-
-    saveRDS(maaslin, file = "rds/maaslin.rds")
-} else {
-    maaslin <- readRDS("rds/maaslin.rds")
-}
 ```

### 2019-06-07 articletwo-rrbiome.R

@@ -159,7 +159,7 @@ prepare.maaslin <- function (pseq,
                       paste(c("Read_PCL_Rows: ", paste(includedvars, collapse=",")), collapse = ""),
                       "",
                       "Matrix: Adundance",
-                      paste("Read_PCL_Rows:", readpclrows))
+                      paste(c("Read_PCL_Rows: ", paste(readpclrows, collapse=",")), collapse = ""))
     writeLines(pcl.contents, con=data)
     close(data)
 }
@@ -230,7 +230,7 @@ getdescriptions <- function() {
                                   "Sex is female, True for female, False for male")))
 }
 
-maaslinwrapper <- function(pseq, looped, forced, taxa, tempstr = "%s/temp/maaslin-%s")  {
+maaslinwrapper <- function(pseq, looped, forced, taxa, tempstr = "%s/maaslinruns/maaslin-%s")  {
     tempdir <- sprintf(tempstr, Sys.getenv("HOME"), format(Sys.time(), '%s'))
     dir.create(tempdir)
     cwd <- getwd()
@@ -240,16 +240,24 @@ maaslinwrapper <- function(pseq, looped, forced, taxa, tempstr = "%s/temp/maasli
                     conffile = "maaslin.config",
                     includedvars = c(looped, forced),
                     readpclrows = taxa)
-    Maaslin("maaslin.tsv",
-            "maaslin",
-            strInputConfig = "maaslin.config",
-            strModelSelection="none",
-            dSignificanceLevel = 1.0,
-            fAllvAll = TRUE,
-            strForcedPredictors = forced)
-    ret <- read.table("maaslin/maaslin.txt", header=TRUE, sep='\t')
+
+    fileConn<-file("run.R")
+    writeLines(c('.libPaths(c(.libPaths(), "~/maaslinruns/R-3.4"))',
+                 'library("Maaslin")',
+                 'Maaslin("maaslin.tsv",',
+                 '\t"maaslin",',
+                 '\tstrInputConfig = "maaslin.config",',
+                 '\tstrModelSelection="none",',
+                 '\tdSignificanceLevel = 1.0,',
+                 '\tfAllvAll = TRUE,',
+                 sprintf("\tstrForcedPredictors = \"%s\")", paste0(forced, collapse=","))), fileConn)
+    close(fileConn)
+
[…]

### 2019-06-07 articletwo.Rmd
Renaming

@@
     pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
-    pseq.genus.core.viruses <- taxa(pseq.genus.core.comp) %>% mygrep(word = "virus$", complement = TRUE)
-    pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses, pseq.genus.core.comp)
+    pseq.genus.core.viruses.names <- taxa(pseq.genus.core) %>% mygrep(word = "virus$", complement = TRUE)
+    pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses.names, pseq.genus.core.comp)
 

### 2019-06-07 articletwo.Rmd
Rename var

--- 2019-06-07 articletwo.Rmd
@@
     pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
-    pseq.genus.core.viruses <- taxa(pseq.genus.core.comp) %>% mygrep(word = "virus$", complement = TRUE)
-    pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses, pseq.genus.core.comp)
+    pseq.genus.core.viruses.names <- taxa(pseq.genus.core) %>% mygrep(word = "virus$", complement = TRUE)
+    pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses.names, pseq.genus.core)
 

### 2019-06-07 articletwo.Rmd
First full run on centrifuge

@@
 
-options(max.print=1000)
+options(max.print=100)
 
@@
                        filterstr = "shannon"))
+
+diversities <- mergediversities(alphadiversity, adonis.species)
 ```
@@
 diversities$max) %>% pull(estimate) %>% min`.
-
-Difference of R2 between full and minimal models
-
-```{r difference in beta diversity}
-merge(diversities$min %>%
-      select(response, R2, R2.p) %>%
-      dplyr::rename(R2.min = R2, p.min = R2.p),
-      diversities$max %>%
-      select(response, R2, R2.p) %>%
-      dplyr::rename(R2.full = R2, p.full = R2.p),
-      by = "response") %>%
-    mutate(R2.diff = R2.full - R2.min) %>%
-    kable
-```
 
@@
     arrange(p.value) %>%
+    kable
+```
+
+Difference of R2 between full and minimal models
+
+```{r difference in beta diversity}
+merge(diversities$min %>%
+      select(response, R2, R2.p) %>%
+      dplyr::rename(R2.min = R2, p.min = R2.p),
+      diversities$max %>%
+      select(response, R2, R2.p) %>%
+      dplyr::rename(R2.full = R2, p.full = R2.p),
+      by = "response") %>%
+    mutate(R2.diff = R2.full - R2.min) %>%
     kable
@@
                                    levels = c("11", "10", "01", "00")),
-           gMAP = cut(oMAP, seq(60, 160, 15)))
+           gMAP = cut(oMAP, seq(60, 160, 15))) %>%
+    gather(key, Axis, Axis.2, Axis.3)
 
-pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.3, color = HYPERTENSION)) +
+axis_labeller <- function(variable, value){
+    axis.labels <- list("Axis.2" = sprintf("A. PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2]),
+                        "Axis.3" = sprintf("B. PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3]))
+    return(axis.labels[value])
+}
+
+pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis, color = HYPERTENSION)) +
+    facet_wrap(~key, labeller = axis_labeller) +
     geom_point(size=0.1, alpha=0.4) +
     xlab(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1])) +
-    ylab(sprintf("PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3])) +
+    ylab(sprintf("Subsequent PCoA Axes", 100*eig.fracs[3])) +
     scale_colour_manual(name = "State and treatment",
@@
     theme(text = element_text(size=14),
-          legend.title = element_blank(),
-          legend.text=element_text(size=10),
-          legend.justification=c(1,0),
-          legend.position=c(1,0),
-          legend.margin=margin(t = 1, unit='mm'),
-          legend.background = element_rect(fill=alpha('gray', 0.1)))
+          legend.position = "none",
+          strip.background = element_blank(),
+          strip.placement = "outside")
 
-ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=4, width=5, units = "in", dpi = 300)
+ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=4, width=6, units = "in", dpi = 300)
 ```
@@
 
-Genera with top R2 with PCoA axes
-
-```{r top generas}
+```{r top generas, eval = FALSE, include = FLASE}
 abund.genus <- abundances(pseq.genus, transform = "compositional") %>%
@@
 ```{r maaslin}
-maas.results.df.signf <- maaslin$genus %>%
+maas.results.df.signf <- maaslin %>%
     merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
@@
     arrange(Feature)
-
-maas.results.df.signf %>% pull(Feature) %>% sort %>% unique %>% paste(., collapse=", ")
 
@@
 
-maaslinplot <- ggplot(maas.tables.df, aes(x = Feature,
-                                          y = reorder(Name, my.plot.order(Name)),
-                                          fill = p.groups)) +
-    geom_tile(color = "white", aes(fill = "NA")) +
-    geom_tile(color = "white") +
-    scale_fill_manual(values = mycolors, name = "P-value", na.translate = F) +
-    scale_x_discrete() +
+maas.tables.df %>% pull(beta)%>% summary
+
+maaslinplot <- ggplot(maas.tables.df, aes(x = Feature, y = reorder(Name, my.plot.order(Name)), fill = beta)) +
+    geom_tile(aes(fill = 1)) +
+    geom_tile(colour="white", size=0.25) +
+    scale_fill_distiller(palette = "RdBu", name = "Effect size",
+                         limits=c(-0.003, 0.003), breaks = c(-0.003, 0, 0.003), na.value="grey30") +
     coord_fixed() +
@@
           axis.text.y = element_text(size=10))
+
 ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 6, dpi = 300)
@@
 ```{r scatter plots}
-scatter.signf.bacteria <- paste0(maas.results.df.signf %>% pull(Feature) %>% unique %>% tolower, ".bacteria")
+scatter.signf.bacteria <- maas.results.df.signf %>% pull(Feature) %>% unique
 
@@
     tibble::rownames_to_column("id") %>%
-    setNames(tolower(replace.brackets(names(.)))) %>%
     select(id, scatter.signf.bacteria) %>%
@@
 
-    feature <- firstup(gsub(".bacteria", "", replace.brackets(abund)))
+    feature <- replace.brackets(abund)
+
     qval.maaslin <- maas.results %>%
@@
 
-```{r scatter plot at genus level for BP, fig.width=12, fig.height=20}
+```{r scatter plot at genus level for BP, fig.width=12, fig.height=10}
 scatterplot <- do.call("grid.arrange", c(scatterplots, ncol=4))
-ggsave(file = "cache/scatter.png", plot = scatterplot, height = 20, width = 10, dpi = 300)
+ggsave(file = "cache/scatter.png", plot = scatterplot, height = 10, width = 10, dpi = 300)
 ```
@@
 
-```{r check odds model assumptions}
-g.odds.autoplot <- autoplot(glm.hypertension.lactogroup,
-                            which = 1:4,
-                            nrow = 1,
-                            ncol = 2,
-                            label.size = 3,
-                            size = 0.1) +
-    theme_minimal()
-ggsave(file = "cache/odds-autoplot.png", plot=g.odds.autoplot, height=10, width=6, dpi = 300)
-```
-
 ## Twin-forest plot
@@
 ```{r forest plot lacto}
+odds.hypertension.lactogroup
 gsalt <- grid.arrange(
@@
                  subtitle = "A.",
-                 my_y_scale = scale_y_continuous(limits = c(2.5,3), breaks = seq(2.5, 3.0, 0.1))),
+                 my_y_scale = scale_y_continuous(limits = c(4.1,4.7), breaks = seq(4, 5, 0.1))),
     myforestplot(odds.hypertension.lactogroup,
@@
                  subtitle = "B.",
-                 my_y_scale = scale_y_log10(limits = c(0.5, 2), breaks = seq(0.5, 2.0, 0.5)),
+                 my_y_scale = scale_y_log10(limits = c(0.4, 1.5), breaks = seq(0.5, 2.0, 0.5)),
                  interceptone = TRUE),

### 2019-06-07 articletwo.Rmd
Removing eval false

@@
 
- ```{r adonis calculation, eval = FALSE}
+ ```{r adonis calculation}
  if (!file.exists("rds/adonis.species.rds")) {

### 2019-06-07 articletwo.Rmd
Typo

@@
 
-```{r top generas, eval = FALSE, include = FLASE}
+```{r top generas, eval = FALSE, include = FALSE}
 abund.genus <- abundances(pseq.genus, transform = "compositional") %>%

### 2019-06-07 articletwo.Rmd
data frame can not be edited there

@@
 eig.fracs <- pcoa.ordinate$values$Relative_eig
+
 pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2, Axis.3),
@@
                                    levels = c("11", "10", "01", "00")),
-           gMAP = cut(oMAP, seq(60, 160, 15))) %>%
-    gather(key, Axis, Axis.2, Axis.3)
+           gMAP = cut(oMAP, seq(60, 160, 15)))
 
@@
 
-pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis, color = HYPERTENSION)) +
+pcoa.plot <- ggplot(data=pcoa.df %>% gather(key, Axis, Axis.2, Axis.3),
+                    aes(x=Axis.1, y=Axis, color = HYPERTENSION)) +
     facet_wrap(~key, labeller = axis_labeller) +

### 2019-06-10 articletwo.Rmd
Also salt pseq from centrifuge and only one genus level import

@@
 ```{r lacto data frame}
-pseq.genus.salt <- import_salt_pseq("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+pseq.genus.salt <- pseq.genus %>%
+    prune_samples(rownames(subset(meta(.), !is.na(NA.))), .) %>%
+    prune_taxa(c("g_Lactobacillus", "g_Lactococcus"), .)
 
-abund.genus.salt <- subset_taxa(pseq.genus.salt, Genus == "Lactobacillus") %>%
+abund.genus.salt <- pseq.genus.salt %>%
     abundances(., transform = "clr") %>%
@@
     tibble::rownames_to_column("id") %>%
-    setNames(tolower(replace.brackets(names(.))))
+    select(-g_Lactococcus)
 
@@
     full_join(., abund.genus.salt, by = "id") %>%
-    dplyr::mutate(salt_groups = relevel(factor(dplyr::ntile(dUNA, 3)), ref='1'),
-                  lacto_groups = relevel(factor(dplyr::ntile(lactobacillus.bacteria, 3)), ref='1'))
+    dplyr::mutate(dUNA = myscale(NA.),
+                  salt_groups = relevel(factor(dplyr::ntile(dUNA, 3)), ref='1'),
+                  lacto_groups = relevel(factor(dplyr::ntile(g_Lactobacillus, 3)), ref='1'))
 ```
@@
 ```{r abund salt lm continuous}
-lm.lacto.duna <- lm(lactobacillus.bacteria ~ dUNA + BL_AGE + SEX, data = df.genus.salt)
+lm.lacto.duna <- lm(g_Lactobacillus ~ dUNA + BL_AGE + SEX, data = df.genus.salt)
 lm.lacto.duna %>% tidy %>% mutate_if(is.numeric, round, 3) %>% kable
@@
 ```{r abund salt lm group}
-lm.lacto.saltgroup <- lm(lactobacillus.bacteria ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
+lm.lacto.saltgroup <- lm(g_Lactobacillus ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
 lm.lacto.saltgroup %>% tidy %>% mutate_if(is.numeric, round, 3) %>% kable
@@
 Bartlett's test for equality of variance gives p = `r
-bartlett.test(lactobacillus.bacteria ~ salt_groups, data =
+bartlett.test(g_Lactobacillus ~ salt_groups, data =
 df.genus.salt)$p.value`.
@@
 One-way test assuming normal distributions gives for equal mean p = `r
-oneway.test(lactobacillus.bacteria ~ salt_groups, data =
+oneway.test(g_Lactobacillus ~ salt_groups, data =
 df.genus.salt, var.equal = TRUE)$p.value`.
@@
 ```{r salt anova}
-aov.lacto.saltgroup <- aov(lactobacillus.bacteria ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
+aov.lacto.saltgroup <- aov(g_Lactobacillus ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
 aov.lacto.saltgroup %>% xtable %>% kable
@@
 calculateglm(df.genus.salt,
-             modelstring = "%s ~ lactobacillus.bacteria + BL_AGE + SEX",
+             modelstring = "%s ~ g_Lactobacillus + BL_AGE + SEX",
              filterstr = "lactobacillus") %>% 
@@
 ```{r table map lactobacillus}
-gvlma.table.plot(lm(MAP ~ lactobacillus.bacteria + BL_AGE + SEX, data = df.genus.salt))$table %>% kable
+gvlma.table.plot(lm(MAP ~ g_Lactobacillus + BL_AGE + SEX, data = df.genus.salt))$table %>% kable
 ```
@@
 ```{r plot map lactobacillus, fig.width=12, fig.height=14}
-gvlma.table.plot(lm(MAP ~ lactobacillus.bacteria + BL_AGE + SEX, data = df.genus.salt))$plot %>% plot
+gvlma.table.plot(lm(MAP ~ g_Lactobacillus + BL_AGE + SEX, data = df.genus.salt))$plot %>% plot
 ```

### 2019-06-10 articletwo-rrbiome.R

@@ -16,11 +16,12 @@ filter.phenotype.data <- function(pseq,
                                   included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
                                                "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X",
                                                "BL_USE_RX_C03",
-                                               "BL_USE_RX_C07", "BL_USE_RX_C08")) {
+                                               "BL_USE_RX_C07", "BL_USE_RX_C08", "NA."),
+                                  allowna = c("NA.")) {
     meta(pseq) %>%
         tibble::rownames_to_column(var = "rowname") %>%
         dplyr::select(rowname, included) %>%
-        stats::na.omit() %>%
+        tidyr::drop_na(myin(included, allowna, complement = TRUE)) %>%
         dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
                                        BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
                       ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
@@ -87,7 +88,7 @@ import_filter_data <- function(file,
                                included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
                                             "CURR_SMOKE", "DIAB", "BL_USE_RX_C09",
                                             "Q57X", "BL_USE_RX_C03",
-                                            "BL_USE_RX_C07", "BL_USE_RX_C08")) {
+                                            "BL_USE_RX_C07", "BL_USE_RX_C08", "NA.")) {
     pseq.full <- readRDS(file)
     pseq.meta <- filter.phenotype.data(pseq.full, included = included)
     phyloseq::sample_data(pseq.full) <- phyloseq::sample_data(pseq.meta)
@@ -264,6 +265,11 @@ mygrep <- function(..., word, ignorecase = TRUE, complement = FALSE) {
     c(...)[xor(grepl(word, c(...), ignore.case = ignorecase), (complement == TRUE))]
 }
 
+
+myin <- function(x,y, complement = FALSE) {
+    x[xor(x %in% y, complement)]
+}
+
 cc <- function(...) {
     c2l(c(...))
 }
@@ -290,22 +296,6 @@ mykable <- function(x, ...) {
   capture.output(x <- print(x))
   knitr::kable(x, ...)
 }
-
-
[…]

### 2019-06-12 articletwo.Rmd
Subrun adonis

@@
         load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
+    if ("subset" %in% names(args))
+        subset.run <- as.numeric(args$subset)
 }
@@
 library(Maaslin2)
+library(pwr)
 ```
@@
                                            covariates = var.CL.min,
-                                           npermutations = 10000)
+                                           npermutations = 9999)
      adonis.species$max <- calculateadonis(dset = meta(pseq.species),
@@
                                            covariates = var.CL,
-                                           npermutations = 10000)
+                                           npermutations = 9999)
      saveRDS(adonis.species, file = "rds/adonis.species.rds")
@@
      adonis.species  <- readRDS("rds/adonis.species.rds")
+ }
+```
+
+```{r quit subruns, include = FALSE}
+if (exists("subset.run")) {
+    quit()
 }
@@
 
-maas.tables.df %>% pull(beta)%>% summary
-
 maaslinplot <- ggplot(maas.tables.df, aes(x = Feature, y = reorder(Name, my.plot.order(Name)), fill = beta)) +
@@
              modelstring = "%s ~ g_Lactobacillus + BL_AGE + SEX",
-             filterstr = "lactobacillus") %>% 
+             filterstr = "Lactobacillus") %>% 
     select(response, term, estimate, std.error, p.value) %>%
@@
 odds.hypertension.lactogroup %>% kable
+```
+
+Power analysis of odds of hypertension
+
+```{r power hypertention}
+power.htn <- pwr.f2.test(u = 3, v=NULL, f2 = 0.02, sig.level=0.05, power=0.8)
+round(power.htn$v + power.htn$u + 1)
+
+pwr.f2.test(u = 3, v=836, f2 = NULL, sig.level=0.05, power=0.95)
 ```

### 2019-06-12 articletwo.Rmd
Minor

@@
 library(purrr)
-library(Maaslin2)
 library(pwr)
+library(gtable)
 ```
@@
                                     looped = var.BP,
-                                    forced = var.CL,
+                                    forced = var.CL.min,
                               taxa = taxa(pseq.genus.core.pruned))
@@
 
-diversities <- mergediversities(alphadiversity, adonis.species)
+diversities <- mergediversities(alphadiversity, adonis.species, names.dset = names.dset)
 ```
@@
 
-pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2, Axis.3),
+pcoa.df <- merge(pcoa.vectors %>%
+                 select(Axis.1, Axis.2, Axis.3),
                  meta(pseq.species), by=0, all = TRUE) %>%
@@
                                    levels = c("11", "10", "01", "00")),
-           gMAP = cut(oMAP, seq(60, 160, 15)))
+           gMAP = cut(oMAP, seq(60, 160, 15))) %>%
+    mutate_at(vars(starts_with("Axis.")), .funs = funs(myscale(.)))
 
 axis_labeller <- function(variable, value){
-    axis.labels <- list("Axis.2" = sprintf("A. PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2]),
-                        "Axis.3" = sprintf("B. PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3]))
+    axis.labels <- list("Axis.2" = sprintf("PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2]),
+                        "Axis.3" = sprintf("PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3]))
     return(axis.labels[value])
@@
 
+plot_labeller <-  function(value) {
+        plot.labels <- list("Axis.2" = sprintf("A."),
+                            "Axis.3" = sprintf("B."))
+        return(plot.labels[value])
+}
+
+pcoa.df %>% pull(Axis.1) %>% summary
+
 pcoa.plot <- ggplot(data=pcoa.df %>% gather(key, Axis, Axis.2, Axis.3),
                     aes(x=Axis.1, y=Axis, color = HYPERTENSION)) +
-    facet_wrap(~key, labeller = axis_labeller) +
+    facet_wrap(~key, scales = "free_y", strip.position = "left", labeller = axis_labeller) +
     geom_point(size=0.1, alpha=0.4) +
     xlab(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1])) +
-    ylab(sprintf("Subsequent PCoA Axes", 100*eig.fracs[3])) +
+    ylab(NULL) +
     scale_colour_manual(name = "State and treatment",
@@
                         values = c("blue", "red")) +   
-    coord_fixed() +
-    scale_x_continuous(breaks=seq(-1, 1, .25)) +
-    scale_y_continuous(breaks=seq(-1, 1, .25)) +
+    scale_x_continuous(breaks=seq(-5, 5, 1)) +
+    scale_y_continuous(breaks=seq(-5, 5, 1)) +
     theme_classic() +
     guides(colour = guide_legend(override.aes = list(size=4))) +
-    theme(text = element_text(size=14),
+    coord_cartesian(clip = 'off') +
+    geom_text(aes(x = -Inf, y = Inf, label = plot_labeller(key)),
+              size = 5,
+              hjust = 2,
+              vjust = -1,
+              inherit.aes = FALSE) +
+    theme(text = element_text(size=12),
           legend.position = "none",
           strip.background = element_blank(),
-          strip.placement = "outside")
+          strip.placement = "outside",
+          aspect.ratio = 1,
+          strip.text = element_text(size = 14),
+          axis.title.x = element_text(size = 14))
 
-ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=4, width=6, units = "in", dpi = 300)
+ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=3.4, width=6, units = "in", dpi = 300)
 ```
@@
                  modelstring = sprintf("%%s ~ %s + %s", axis, paste(var.CL, collapse = "+")),
+                 responses = list(model_1 = "oMAP", model_2 = "oSYSTM", model_3 = "oDIASM",
+                                          model_4 = "oPULSEPRESSURE", model_5 = "HYPERTENSION"),
                  filterstr = "Axis"))
@@
 do.call(rbind, glm.pcoa) %>% 
-    select(response, term, estimate, std.error, p.value) %>%
+    select(response, term, estimate, std.error, p.value, conf.low, conf.high) %>%
     mutate_if(is.numeric, round, 3) %>%
@@
     kable
+
 ```
@@
 
-```{r maaslin}
+```{r maaslin clean results}
 maas.results.df.signf <- maaslin %>%
@@
     dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  Feature = gsub("_Bacteria", "", Feature),
+                  Feature = gsub("[gs]_", "", Feature),
                   beta = Coefficient) %>%
@@
     arrange(Feature)
+```
 
+```{r maaslin plot}
 my.plot.order <- function(x) {
@@
                          Name = maas.results.df.signf %>% pull(Name))
+
 maas.tables.df <- merge(maas.results.df.signf,
@@
 
-ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 6, dpi = 300)
+ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 30, width = 40, unit = "in", dpi = 300)
 ```
@@
     kable
+```
+
+```{r multi model}
+abund.genus.systm <- pseq.genus %>%
+    abundances(., transform = "compositional") %>%
+    t %>%
+    as.data.frame %>%
+    tibble::rownames_to_column("id")
+
+df.genus.systm <- meta(pseq.genus.systm) %>%
+    tibble::rownames_to_column("id") %>%
+    full_join(., abund.genus.systm, by = "id") %>%
+    mutate(Anaerotruncus = scale(asin(sqrt(g_Anaerotruncus))),
+           Sutterella = scale(asin(sqrt(g_Sutterella))))
+
+fo.genus.systm <- sprintf("oSYSTM ~ Anaerotruncus + Sutterella + Anaerotruncus:Sutterella + %s",
+                          paste0(var.CL, collapse = " + "))
+
+lm(as.formula(fo.genus.systm), data = df.genus.systm) %>% summary
+
+df.genus.systm %>%
+    mutate(diff = -0.75418 * Anaerotruncus + 0.89265 * Sutterella) %>%
+    pull(diff) %>%
+    summary
+
 ```
@@
 
+```{r forest lims, eval = FALSE, include = FALSE}
+adjmean.lacto.saltgroup
+```
+
 ```{r forest plot lacto}
@@
                  subtitle = "A.",
-                 my_y_scale = scale_y_continuous(limits = c(4.1,4.7), breaks = seq(4, 5, 0.1))),
+                 my_y_scale = scale_y_continuous(limits = c(1.15,1.42), breaks = seq(0, 5, 0.1))),
     myforestplot(odds.hypertension.lactogroup,

### 2019-06-12 articletwo-ggplot.R

@@ -63,7 +63,7 @@ plot.beta <- function(diversity,
     diversity <- diversity %>% mutate(color = as.factor(ifelse(R2.p < 0.05, 1, 0)))
     plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2, color = color)) +
         geom_bar(stat="identity") +
-        scale_color_manual(values = c("0" = "red", "1" = "black")) +
+        scale_color_manual(values = c("0" = "black", "1" = "red")) +
         coord_flip() +
         theme_classic(fig.textsize) +
         ylab(fig.permar2.text) +

### 2019-06-12 articletwo-rrbiome.R

@@ -71,17 +71,16 @@ calculateglm <- function(dset,
                          modelstring = "%s ~ BL_AGE + SEX",
                          filterstr = ".") {
     glmlist <- lapply(responses, function(response) {
-        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
+        is.logistic <- length(unique(pull(dset, response))) < min_n_for_continuous
+        fo.family <- ifelse(is.logistic, stats::binomial, stats::gaussian)
         fo <- sprintf(modelstring, response)
         stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
-            broom::tidy() %>%
+            broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = is.logistic) %>%
                 dplyr::filter(grepl(filterstr, term)) %>%
                 dplyr::mutate(response = response, fo = fo)
     })
     data.table::rbindlist(glmlist, id = "model_name") %>%
-        as.data.frame %>%
-        dplyr::mutate(conf.low = estimate - qnorm(1- 0.05/2) * std.error,
-                      conf.high = estimate + qnorm(1- 0.05/2) * std.error)
+        as.data.frame
 }
 
 import_filter_data <- function(file,
@@ -203,7 +202,8 @@ vectortolist <- function(c) {
 }
 
 mergediversities <- function(alphadiversity, betadiversity,
-                             responses = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) {
+                             responses = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
+                             names.dset) {
     lapply(vectortolist(names(alphadiversity)), function(x) {
         alpha.select <- alphadiversity[[x]] %>% select(estimate, p.value, response)
         beta.rbind <- betadiversity[[x]] %>%
@@ -309,3 +309,6 @@ myinstall.packages <- function(...) {
 }
 
 
+mydropna <- function(...) {
+    c(...)[!is.na(c(...))]
+}

### 2019-06-12 maaslinruns.Rmd
Comparing maaslin runs

---
title: "Maaslin runs"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(max.print=100)

dir.create("rds/", showWarnings = FALSE)
dir.create("session/", showWarnings = FALSE)
```

# Command line arguments

Calculations ran at

```{r Command line arguments}
now <- format(Sys.time(), '%Y%m%d-%H%M%S')

if (exists("args")) {
    if ("time" %in% names(args)) now <- args$time

    if("clean" %in% names(args)) {
        unlink("rds", recursive=TRUE)
        unlink("cache", recursive=TRUE)
    }
    if ("tags" %in% names(args)) 
        rtags('~/phd/research/articletwo/', recursive = FALSE, pattern = '\\.[RrSs](rw|md)?$',
              ofile = '~/phd/research/articletwo/TAGS', verbose = TRUE, append = FALSE)
    if ("loadlast" %in% names(args) )
        load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
    if ("subset" %in% names(args))
        subset.run <- as.numeric(args$subset)
}

#packrat::opts$ignored.packages
```

# Importing libraries:

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
```

### RR biome functions

<details>
  <summary>Open/Close</summary>

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```

</details>

```{r data}
pseq.genus <- import_filter_data("data/phfinrisk_centrifuge_genus_d50k_2019-06-02.rds")
pseq.species <- import_filter_data("data/phfinrisk_centrifuge_species_all_drop50k_2018-11-30_2019-06-03.RDs")
```

At species level meta has dimensions (`r dim(meta(pseq.species))`) and
there ntaxa is `r ntaxa(pseq.species)`, and at genus level meta has
dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is 
`r ntaxa(pseq.genus)`.

## Variables

```{r my variables}
var.BP <- c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")
var.CL.min <- c("BL_AGE", "SEX")
var.CL.add <- c("SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
            "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
```

```{r var combinations}
var.CL.list <- seq(length(var.CL.add)) %>% as.list
names(var.CL.list) <- var.CL.add
var.CL.comb <- lapply(var.CL.list, function(x) unique(c(var.CL.min, var.CL.add[1:x])))
var.CL.comb3rd <- lapply(var.CL.list, function(x) unique(c(var.CL.min, var.CL.add[x])))
```

```{r process genus}
pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
pseq.genus.core.viruses.names <- taxa(pseq.genus.core) %>% mygrep(word = "virus$", complement = TRUE)
pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses.names, pseq.genus.core)
```

```{r maaslin runs, results = 'hide', eval  = FALSE} 
ret <- mclapply(var.CL.comb, function(x)
    maaslinwrapper(pseq = pseq.genus.core.pruned,
                   looped = var.BP,
                   forced = x,
                   taxa = taxa(pseq.genus.core.pruned),
                   tempstr = sprintf("%%s/maaslinruns/maaslinruns/%%s_%s", paste0(x, collapse = "_"))),
    mc.cores = length(var.CL.comb))
```

```{r read combinations, eval = TRUE}
ret <- lapply(var.CL.comb, function(x) {
    file <- sprintf("/homes/jpalmu/maaslinruns/maaslinruns/1560334017_%s/maaslin/maaslin.txt",
                    paste0(x, collapse = "_"))
    read.table(file, header=TRUE, sep='\t') %>% as.data.frame
    })
```

```{r combine results}
ret.comb <- map_df(ret, ~as.data.frame(.x), .id="id")

ret.comb %>%
    dplyr::filter(Q.value < 0.05) %>%
    group_by(id) %>%
    summarize(n = n(), model = paste0(var.CL.comb[[id[1]]], collapse = "+")) %>%
    arrange(model) %>%
    kable
```

```{r adding third, eval = FALSE}
ret3rd <- mclapply(var.CL.comb3rd, function(x)
    maaslinwrapper(pseq = pseq.genus.core.pruned,
                   looped = var.BP,
                   forced = x,
                   taxa = taxa(pseq.genus.core.pruned),
                   tempstr = sprintf("%%s/maaslinruns/maaslinruns/%%s_%s", paste0(x, collapse = "_"))),
    mc.cores = length(var.CL.comb3rd))

```

```{r read combinations, eval = TRUE}
ret3rd <- lapply(var.CL.comb3rd, function(x) {
    file <- sprintf("/homes/jpalmu/maaslinruns/maaslinruns/1560338417_%s/maaslin/maaslin.txt",
                    paste0(x, collapse = "_"))
    read.table(file, header=TRUE, sep='\t') %>% as.data.frame
    })
```

```{r combine results}
ret3rd.comb <- map_df(ret3rd, ~as.data.frame(.x), .id="id")

ret3rd.comb %>%
    dplyr::filter(Q.value < 0.05) %>%
    group_by(id) %>%
    summarize(n = n(), model = paste0(var.CL.comb3rd[[id[1]]], collapse = "+")) %>%
    arrange(n) %>%
    kable
```

### 2019-06-12 articletwo.Rmd
Full
Renaming

@@
                                     looped = var.BP,
-                                    forced = var.CL.min,
+                                    forced = var.CL,
                               taxa = taxa(pseq.genus.core.pruned))
@@
 
-df.genus.systm <- meta(pseq.genus.systm) %>%
+df.genus.systm <- meta(pseq.genus) %>%
     tibble::rownames_to_column("id") %>%

### 2019-06-12 maaslinruns.Rmd
Duplicated labels

@@
 
-```{r read combinations, eval = TRUE}
+```{r read combinations 3rd, eval = TRUE}
 ret3rd <- lapply(var.CL.comb3rd, function(x) {
@@
 
-```{r combine results}
+```{r combine results 3rd}
 ret3rd.comb <- map_df(ret3rd, ~as.data.frame(.x), .id="id")

### 2019-06-12 articletwo.Rmd
eval false block

@@
 
-```{r multi model}
+```{r multi model, eval = FALSE, include = FALSE}
 abund.genus.systm <- pseq.genus %>%
@@
     summary
-
 ```

### 2019-06-12 articletwo.Rmd
Removing scatter plots
Scatter

@@
 
-```{r scatter plots}
+```{r scatter plots, eval = FALSE, include = FALSE}
 scatter.signf.bacteria <- maas.results.df.signf %>% pull(Feature) %>% unique
@@
                            scatter.plot(scatter.full, maas.results.df.signf, x$rr, x$abund))
-```
 
-```{r scatter plot at genus level for BP, fig.width=12, fig.height=10}
 scatterplot <- do.call("grid.arrange", c(scatterplots, ncol=4))

### 2019-06-12 articletwo.Rmd
Round 3

@@
                        NAs = sum(is.na(value))))  %>%
-    mutate_if(is.numeric, round, 2) %>%
+    mutate_if(is.numeric, round, 3) %>%
     kable()

### 2019-06-14 articletwo.Rmd
Cleaner salt data frame creation

@@
 library(gtable)
+library(car)
 ```
@@
 ```{r maaslin runs, results = 'hide'} 
-if (!file.exists("rds/maaslin.rds")) {
  pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
@@
 
+if (!file.exists("rds/maaslin.rds")) {
      maaslin <- maaslinwrapper(pseq = pseq.genus.core.pruned,
@@
 }
-
-pcoa.df %>% pull(Axis.1) %>% summary
 
@@
 
+```{r maaslin means}
+pseq.genus.core.pruned.table <- pseq.genus.core.pruned %>%
+    prune_taxa(c("g_Adlercreutzia", "g_Anaerotruncus", "g_Sutterella"), .) %>%
+    transform(transform = "compositional") %>%
+    psmelt %>%
+    mutate(map_groups = relevel(factor(dplyr::ntile(MAP, 3)), ref='1'),
+           systm_groups = relevel(factor(dplyr::ntile(SYSTM, 3)), ref='1'),
+           Abundance = myscale(Abundance))
+
+table_bp_abund <- function(pseq, group) {
+    pseq %>%
+        group_by(.dots = c(group, "OTU")) %>%
+        summarize(mean = mean(Abundance), sd = sd(Abundance)) %>%
+        mutate(beta_sd = sprintf("%.2f ± %.2f", mean, sd)) %>%
+        select(OTU, beta_sd) %>%
+        spread(OTU, beta_sd) %>%
+        ungroup %>%
+        dplyr::rename("grp" = group) %>%
+        mutate("grp" = paste0(group, "_", grp))
+}
+
+rbind(table_bp_abund(pseq.genus.core.pruned.table, "map_groups"),
+             table_bp_abund(pseq.genus.core.pruned.table, "systm_groups")) %>%
+    tibble::column_to_rownames("grp") %>%
+    t %>%
+    as.data.frame %>%
+    kable
+
+```
+
 ```{r Write figure 3, include=FALSE}
@@
 ```{r lacto data frame}
-pseq.genus.salt <- pseq.genus %>%
-    prune_samples(rownames(subset(meta(.), !is.na(NA.))), .) %>%
-    prune_taxa(c("g_Lactobacillus", "g_Lactococcus"), .)
-
-abund.genus.salt <- pseq.genus.salt %>%
-    abundances(., transform = "clr") %>%
-    t %>%
-    as.data.frame %>%
-    tibble::rownames_to_column("id") %>%
-    select(-g_Lactococcus)
-
-df.genus.salt <- meta(pseq.genus.salt) %>%
-    tibble::rownames_to_column("id") %>%
-    full_join(., abund.genus.salt, by = "id") %>%
+df.genus.salt <- pseq.genus %>%
+    microbiome::transform(transform = "clr") %>%
+    prune_taxa(c("g_Lactobacillus"), .) %>%
+    psmelt %>%
+    dplyr::filter(!is.na(NA.)) %>%
+    spread(OTU, Abundance) %>%
     dplyr::mutate(dUNA = myscale(NA.),
@@
                   lacto_groups = relevel(factor(dplyr::ntile(g_Lactobacillus, 3)), ref='1'))
+```
+
+Lactobacillus prevalence
+
+```{r lactobacillus prevalence}
+pseq.genus %>%
+    microbiome::transform(transform = "compositional") %>%
+    prune_taxa(c("g_Lactobacillus"), .) %>%
+    psmelt %>%
+    spread(OTU, Abundance) %>%
+    dplyr::mutate(lacto.prevalent = ifelse(g_Lactobacillus > 0.001, TRUE, FALSE),
+                  duna.present = ifelse(is.na(NA.), 0, 1)) %>%
+    group_by(duna.present) %>%
+    summarize(lacto_detected = sum(lacto.prevalent), n = n(), ratio = lacto_detected/n)
 ```
@@
 
-Bartlett's test for equality of variance gives p = `r
-bartlett.test(g_Lactobacillus ~ salt_groups, data =
-df.genus.salt)$p.value`.
-
-One-way test assuming normal distributions gives for equal mean p = `r
-oneway.test(g_Lactobacillus ~ salt_groups, data =
-df.genus.salt, var.equal = TRUE)$p.value`.
-
 Analysing also differences between salt groups using ANOVA.
@@
 aov.lacto.saltgroup %>% xtable %>% kable
+```
+
+Levene's test for equality of variance
+
+```{r salt levene}
+car::leveneTest(g_Lactobacillus ~ salt_groups, data = df.genus.salt) %>% kable
 ```

### 2019-06-14 articletwo.Rmd
Missing kable

@@
     group_by(duna.present) %>%
-    summarize(lacto_detected = sum(lacto.prevalent), n = n(), ratio = lacto_detected/n)
+    summarize(lacto_detected = sum(lacto.prevalent), n = n(), ratio = lacto_detected/n) %>%
+    kable
 ```

### 2019-06-14 articletwo.Rmd
Correct image size for maaslin plot
Variable missing

@@
 
-ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 30, width = 40, unit = "in", dpi = 300)
+ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 4, unit = "in", dpi = 300)
 ```
@@
 ```{r salt characteristics}
-pseq.genus.salt %>%
-    meta %>%
+df.genus.salt %>%
+    select(-Sample) %>% 
     gather(key,value) %>%

### 2019-06-14 articletwo.Rmd
Lacto y changed

@@
 
+Adjusted means
+
+```{r forest lims}
+adjmean.lacto.saltgroup %>% kable
+```
+
 ## Is there associations between lactobacillus abudance and hypertension?
@@
 
-```{r forest lims, eval = FALSE, include = FALSE}
-adjmean.lacto.saltgroup
-```
-
 ```{r forest plot lacto}
-odds.hypertension.lactogroup
 gsalt <- grid.arrange(
@@
                  subtitle = "A.",
-                 my_y_scale = scale_y_continuous(limits = c(1.15,1.42), breaks = seq(0, 5, 0.1))),
+                 my_y_scale = scale_y_continuous(limits = c(5.38, 5.67), breaks = seq(0, 10, 0.1))),
     myforestplot(odds.hypertension.lactogroup,

### 2019-08-05 articletwo.Rmd
PREVAL DIAB

@@
 library(car)
+library(M3C)
 ```
@@
 var.CL.min <- c("BL_AGE", "SEX")
-var.CL <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
+var.CL <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
             "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
@@
 
+```{r t-sne, eval = FALSE}
+sne.df <- pseq.genus %>%
+    microbiome::transform(transform = "clr") %>%
+    psmelt %>%
+    spread(OTU, Abundance)
+
+sne_names <- colnames(sne.df) %>% mygrep(word = "g_")
+
+sne.t <- sne.df %>% select(Sample, HYPERTENSION, sne_names[1:20]) %>%
+    mutate(HYPERTENSION = as.numeric(HYPERTENSION)) %>%
+    gather(variable, value, -Sample) %>% 
+    spread(Sample, value) %>%
+    column_to_rownames("variable")
+
+png("sample.png", 800, 800)
+tsne(sne.t[-1,], labels = as.factor(sne.t[1,]),
+     dotsize = 1)
+dev.off()
+
+### CONTINUOUS
+
+source("tsne.R")
+
+sne.t <- sne.df %>% select(Sample, oMAP, sne_names) %>%
+    gather(variable, value, -Sample) %>% 
+    spread(Sample, value) %>%
+    column_to_rownames("variable")
+
+tsne.g <- tsne(sne.t[rownames(sne.t) != "oMAP",],
+               labels = as.numeric(sne.t[rownames(sne.t) == "oMAP",]),
+               dotsize = 1,
+               controlscale = TRUE,
+               scale = 2,
+               low = "white")
+
+tsne.plot <- tsne.g +
+    scale_fill_gradientn(limits = c(-3,3),
+                         colours=c("navyblue", "darkmagenta", "darkorange1"),
+                         breaks=c(0,200), labels=format(c(0,200)))
+
+ggsave(file = "cache/tsne.png", plot=tsne.plot, height=6, width=9)
+
 ```{r maaslin runs, results = 'hide'} 
  pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
- pseq.genus.core.viruses.names <- taxa(pseq.genus.core) %>% mygrep(word = "virus$", complement = TRUE)
- pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses.names, pseq.genus.core)
 
 if (!file.exists("rds/maaslin.rds")) {
-     maaslin <- maaslinwrapper(pseq = pseq.genus.core.pruned,
+     maaslin <- maaslinwrapper(pseq = pseq.genus.core,
                                     looped = var.BP,
                                     forced = var.CL,
-                              taxa = taxa(pseq.genus.core.pruned))
+                              taxa = taxa(pseq.genus.core))
 
@@
 diversities <- mergediversities(alphadiversity, adonis.species, names.dset = names.dset)
+```
+
+Hypertension prevalence
+
+```{r test htn prevalence}
+dset.alpha %>%
+    dplyr::filter(BL_AGE < 60 + 12, BL_AGE > 60 - 12) %>%
+    summarize(htn = sum(ifelse(HYPERTENSION == 1, 1, 0)), n = n(), proportion = htn/n) %>%
+    kable
 ```
@@
 
-pcoa.df <- merge(pcoa.vectors %>%
-                 select(Axis.1, Axis.2, Axis.3),
+pcoa.df <- merge(pcoa.vectors,
                  meta(pseq.species), by=0, all = TRUE) %>%
@@
 
+```{r pcoa top axes}
+axis_top_labeller <- function(variable, value){
+    index <- as.integer(gsub("Axis.", "", value))
+    return(sprintf("Axis %i (%.1f%%)", index, 100*eig.fracs[index]))
+}
+
+pcoa.plot.top <- ggplot(data=pcoa.df %>% gather(key, Axis, paste0("Axis.", seq(2,43))) %>%
+                       mutate(key = factor(key, levels=paste0("Axis.", seq(2,43)))),
+                    aes(x=Axis.1, y=Axis, color = HYPERTENSION)) +
+    facet_wrap(~key, scales = "free_y", labeller = axis_top_labeller) +
+    geom_point(size=0.1, alpha=0.2) +
+    xlab("PCoA 1") +
+    ylab(NULL) +
+    scale_colour_manual(name = "State and treatment",
+                        labels = c("Normotensive", "Hypertensive"),
+                        breaks=c(0, 1),
+                        values = c("blue", "red")) + 
+    theme_classic() +
+    theme(legend.position = "none")
+
+ggsave(file = "cache/pcoa-species-top-axes.png", plot=pcoa.plot.top,
+       height=10, width=10, units = "in", dpi = 300)
+```
+
 First three axis explain portion (`r eig.fracs[1:3]`) of
@@
 
-Significant associations between BP-variables and first three PCoA axis
-
-```{r pcoa axis analysis axis 1}
-glm.pcoa <- lapply(c2l("Axis.1", "Axis.2", "Axis.3"), function(axis)
+Significant associations between BP-variables and top PCoA axes
+
+```{r pcoa analysis main axis}
+glm.pcoa <- lapply(c2l(paste0("Axis.", seq(1,5))), function(axis)
     calculateglm(pcoa.df,
@@
     mutate_if(is.numeric, round, 3) %>%
-    filter(p.value < 0.05) %>%
     arrange(p.value) %>%
@@
 Bray-Curtis distances are calculated for abundances after compositional transformation.
-First two axes explain %.1f%% of the variation.", round(100*sum(eig.fracs[c(1,3)]),1))
+First two axes explain %.1f%% of the variation.", round(100*sum(eig.fracs[1:3]),1))
 writeimage(doc, 2, "cache/pcoa-species.png", fig2head, fig2foot)
@@
     arrange(Feature)
+```
+
+```{r maaslin insignif, include = FALSE, eval = FALSE}
+ maaslin %>%
+    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
+                  Feature = gsub("[gs]_", "", Feature),
+                  beta = Coefficient) %>%
+     arrange(Feature) %>%
+     select(Variable, Feature, Qval) %>%
+     spread(Variable, Qval)
 ```
@@
                          abund,
-                         covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
+                         covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
                                         "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08",
@@
                Q57X = as.factor(1),
-               DIAB = as.factor(0),
+               PREVAL_DIAB = as.factor(0),
                BL_USE_RX_C03 = as.factor(0),
@@
                        Mean = mean(as.numeric(value), na.rm = TRUE),
+                       ones = sum(value == 1, na.rm = TRUE),
                        sd = sd(as.numeric(value), na.rm = TRUE),
@@
 df.genus.salt.unadjusted <- with(df.genus.salt, data.frame(salt_groups = as.factor(seq(3)),
-                                                           SEX = as.factor(0),
+                                                           SEX = as.factor(1),
                                                            BL_AGE = mean(BL_AGE)))
@@
                                df.genus.salt.unadjusted,
-                               interval = "confidence") %>%
+                               interval = "prediction") %>%
     as.data.frame %>%
@@
                  subtitle = "A.",
-                 my_y_scale = scale_y_continuous(limits = c(5.38, 5.67), breaks = seq(0, 10, 0.1))),
+                 my_y_scale = scale_y_continuous(limits = c(3.0, 8.0), breaks = seq(0, 10, 1))),
     myforestplot(odds.hypertension.lactogroup,
@@
 
-ggsave(file = "cache/salt-forest.png", plot=gsalt, height=4, width=6, dpi = 300)
+ggsave(file = "cache/salt-forest.png", plot=gsalt, height=3, width=5, dpi = 300)
 ```
@@
          response = "MAP",
-         loopcovariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
+         loopcovariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
                             "ANYDRUG"))

### 2019-08-05 articletwo-officer.R

@@ -34,7 +34,7 @@ tableone <- function(dset) {
                            "oMAP" = "Mean arterial pressure, mmHg (SD)",
                            "HYPERTENSION" = "Hypertension, N (%)",
                            "CURR_SMOKE" = "Current smoker, N (%)",
-                           "DIAB" = "Diabetes mellitus, N (%)",
+                           "PREVAL_DIAB" = "Diabetes mellitus, N (%)",
                            "Q57X" = "Exercise, N (%)",
                            "ANYDRUG" = "Antihypertensive medication, N (%)",
                            "BL_USE_RX_C03" = "  Diuretics, N (%)",
@@ -44,7 +44,7 @@ tableone <- function(dset) {
     extras <- list("1"  =  "  Light",
                            "2"  =  "  Moderate",
                            "3"  =  "  Heavy")
-    tableone.factors <- c("SEX", "HYPERTENSION",  "ANYDRUG", "CURR_SMOKE", "DIAB",
+    tableone.factors <- c("SEX", "HYPERTENSION",  "ANYDRUG", "CURR_SMOKE", "PREVAL_DIAB",
                           "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
     data <- characteristics(dset, tableone.names, tableone.factors, extras)
 

### 2019-08-05 articletwo-rrbiome.R

@@ -14,7 +14,7 @@ myscale <- function(x){
 
 filter.phenotype.data <- function(pseq,
                                   included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
-                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X",
+                                               "CURR_SMOKE", "PREVAL_DIAB", "BL_USE_RX_C09", "Q57X",
                                                "BL_USE_RX_C03",
                                                "BL_USE_RX_C07", "BL_USE_RX_C08", "NA."),
                                   allowna = c("NA.")) {
@@ -43,7 +43,6 @@ filter.phenotype.data <- function(pseq,
                       BL_USE_RX_C07 = as.factor(BL_USE_RX_C07),
                       BL_USE_RX_C08 = as.factor(BL_USE_RX_C08),
                       BL_USE_RX_C09 = as.factor(BL_USE_RX_C09)) %>%
-        mutate_at(vars(one_of("DIAB")), as.factor) %>%
         dplyr::select(-MEN) %>%
         tibble::remove_rownames() %>%
         tibble::column_to_rownames(var = "rowname")
@@ -85,7 +84,7 @@ calculateglm <- function(dset,
 
 import_filter_data <- function(file,
                                included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
-                                            "CURR_SMOKE", "DIAB", "BL_USE_RX_C09",
+                                            "CURR_SMOKE", "PREVAL_DIAB", "BL_USE_RX_C09",
                                             "Q57X", "BL_USE_RX_C03",
                                             "BL_USE_RX_C07", "BL_USE_RX_C08", "NA.")) {
     pseq.full <- readRDS(file)

### 2019-08-07 articletwo.Rmd
Renaming

@@
 ```{r maaslin means}
-pseq.genus.core.pruned.table <- pseq.genus.core.pruned %>%
+pseq.genus.core.table <- pseq.genus.core %>%
     prune_taxa(c("g_Adlercreutzia", "g_Anaerotruncus", "g_Sutterella"), .) %>%
@@
 
-rbind(table_bp_abund(pseq.genus.core.pruned.table, "map_groups"),
-             table_bp_abund(pseq.genus.core.pruned.table, "systm_groups")) %>%
+rbind(table_bp_abund(pseq.genus.core.table, "map_groups"),
+             table_bp_abund(pseq.genus.core.table, "systm_groups")) %>%
     tibble::column_to_rownames("grp") %>%

### 2019-08-07 maaslinruns.Rmd
Renaming

@@
 var.CL.min <- c("BL_AGE", "SEX")
-var.CL.add <- c("SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
+var.CL.add <- c("SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
             "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")

### 2019-08-07 partialrda.Rmd
Renaming

@@
     covariates <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
-                   "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07",
+                   "PREVAL_DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07",
                    "BL_USE_RX_C08", "BL_USE_RX_C09")
@@
 ex.rda <- rda(rda.abundances.species ~ SYSTM + Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X
-                                                        + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 +
+                                                        + PREVAL_DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 +
                                                         + BL_USE_RX_C08 + BL_USE_RX_C09),

### 2019-08-07 articletwo.Rmd
Fixing core level calculations

@@
 At species level meta has dimensions (`r dim(meta(pseq.species))`) and
-there ntaxa is `r ntaxa(pseq.species)`, and at genus level meta has
+there ntaxa is `r ntaxa(pseq.species)`. At genus level meta has
 dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is 
 `r ntaxa(pseq.genus)`.
+
+```{r defining genus core}
+pseq.genus.core <- microbiome::transform(pseq.genus, "compositional") %>%
+    core(detection = 0.0001, prevalence = 0.01)
+```
+
+At genus-core level meta has dimensions (`r dim(meta(pseq.genus.core))`) 
+and there ntaxa is `r ntaxa(pseq.genus.core)`.
 
@@
 ```{r maaslin runs, results = 'hide'} 
- pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
-
 if (!file.exists("rds/maaslin.rds")) {

### 2019-08-07 articletwo-rrbiome.R

@@ -140,17 +140,10 @@ prepare.maaslin <- function (pseq,
                              tsvfile = "merged_phfinrisk_genus.tsv",
                              conffile = "maaslin_config.config",
                              includedvars = c("BL_AGE", "BMI", "SEX"),
-                             readpclrows = "Archaea.Archaea-",
-                             core = FALSE,
-                             core.detection = 0.001,
-                             core.prevalence = 0.01) {
-    pseq.comp <- microbiome::transform(pseq, "compositional")
-    if (core) {
-         pseq.comp  <- microbiome::core(pseq.comp, detection = core.detection, prevalence = core.prevalence)
-    }
-    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq.comp)))
+                             readpclrows = "") {
+    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
     sample.ids <- colnames(pheno)
-    abud.rel <- abundances(pseq.comp)
+    abud.rel <- abundances(pseq)
     rownames(abud.rel) <- sapply(rownames(abud.rel), replace.brackets)
     write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
     data <- file(conffile , open='wt')

### 2019-08-07 maaslinruns.Rmd
Fixing maaslin runs

@@
 pseq.genus <- import_filter_data("data/phfinrisk_centrifuge_genus_d50k_2019-06-02.rds")
-pseq.species <- import_filter_data("data/phfinrisk_centrifuge_species_all_drop50k_2018-11-30_2019-06-03.RDs")
 ```
@@
 `r ntaxa(pseq.genus)`.
+
+```{r defining genus core}
+pseq.genus.core <- microbiome::transform(pseq.genus, "compositional") %>%
+    core(detection = 0.0001, prevalence = 0.01)
+```
 
@@
 
-```{r process genus}
-pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
-pseq.genus.core.viruses.names <- taxa(pseq.genus.core) %>% mygrep(word = "virus$", complement = TRUE)
-pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses.names, pseq.genus.core)
-```
-
 ```{r maaslin runs, results = 'hide', eval  = FALSE} 
 ret <- mclapply(var.CL.comb, function(x)
-    maaslinwrapper(pseq = pseq.genus.core.pruned,
+    maaslinwrapper(pseq = pseq.genus.core,
                    looped = var.BP,
                    forced = x,
-                   taxa = taxa(pseq.genus.core.pruned),
+                   taxa = taxa(pseq.genus.core),
                    tempstr = sprintf("%%s/maaslinruns/maaslinruns/%%s_%s", paste0(x, collapse = "_"))),
@@
 ret3rd <- mclapply(var.CL.comb3rd, function(x)
-    maaslinwrapper(pseq = pseq.genus.core.pruned,
+    maaslinwrapper(pseq = pseq.genus.core,
                    looped = var.BP,
                    forced = x,
-                   taxa = taxa(pseq.genus.core.pruned),
+                   taxa = taxa(pseq.genus.core),
                    tempstr = sprintf("%%s/maaslinruns/maaslinruns/%%s_%s", paste0(x, collapse = "_"))),

### 2019-08-07 maaslinruns.Rmd
Minor

@@
 
-```{r maaslin runs, results = 'hide', eval  = FALSE} 
+```{r maaslin runs, results = 'hide', eval  = TRUE} 
 ret <- mclapply(var.CL.comb, function(x)
@@
                    taxa = taxa(pseq.genus.core),
-                   tempstr = sprintf("%%s/maaslinruns/maaslinruns/%%s_%s", paste0(x, collapse = "_"))),
+                   tempstr = sprintf("%%s/combinations/maaslinruns/%%s_%s", paste0(x, collapse = "_"))),
     mc.cores = length(var.CL.comb))
+```
+
+```{r quit}
+quit()
 ```

### 2019-08-07 maaslinruns.Rmd
Removing unneeded variable

@@
 pseq.genus <- import_filter_data("data/phfinrisk_centrifuge_genus_d50k_2019-06-02.rds")
-```
-
-At species level meta has dimensions (`r dim(meta(pseq.species))`) and
-there ntaxa is `r ntaxa(pseq.species)`, and at genus level meta has
-dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is 
-`r ntaxa(pseq.genus)`.
-
-```{r defining genus core}
 pseq.genus.core <- microbiome::transform(pseq.genus, "compositional") %>%

### 2019-08-07 maaslinruns.Rmd
Typo

@@
                    taxa = taxa(pseq.genus.core),
-                   tempstr = sprintf("%%s/combinations/maaslinruns/%%s_%s", paste0(x, collapse = "_"))),
+                   tempstr = sprintf("%%s/maaslinruns/combinations/%%s_%s", paste0(x, collapse = "_"))),
     mc.cores = length(var.CL.comb))

### 2019-08-07 maaslinruns.Rmd
Run full

@@
                    taxa = taxa(pseq.genus.core),
-                   tempstr = sprintf("%%s/maaslinruns/combinations/%%s_%s", paste0(x, collapse = "_"))),
+                   tempstr = sprintf("%%s/maaslinruns/combinations/single_%s", paste0(x, collapse = "_"))),
     mc.cores = length(var.CL.comb))
@@
 
-```{r quit}
-quit()
+```{r adding third, eval = TRUE}
+ret3rd <- mclapply(var.CL.comb3rd, function(x)
+    maaslinwrapper(pseq = pseq.genus.core,
+                   looped = var.BP,
+                   forced = x,
+                   taxa = taxa(pseq.genus.core),
+                   tempstr = sprintf("%%s/maaslinruns/combinations/all3rd_%s", paste0(x, collapse = "_"))),
+    mc.cores = length(var.CL.comb3rd))
+
 ```
 
-```{r read combinations, eval = TRUE}
+```{r read combinations}
 ret <- lapply(var.CL.comb, function(x) {
-    file <- sprintf("/homes/jpalmu/maaslinruns/maaslinruns/1560334017_%s/maaslin/maaslin.txt",
+    file <- sprintf("/homes/jpalmu/maaslinruns/maaslinruns/combinations/single_%s/maaslin/maaslin.txt",
                     paste0(x, collapse = "_"))
@@
 
-```{r adding third, eval = FALSE}
-ret3rd <- mclapply(var.CL.comb3rd, function(x)
-    maaslinwrapper(pseq = pseq.genus.core,
-                   looped = var.BP,
-                   forced = x,
-                   taxa = taxa(pseq.genus.core),
-                   tempstr = sprintf("%%s/maaslinruns/maaslinruns/%%s_%s", paste0(x, collapse = "_"))),
-    mc.cores = length(var.CL.comb3rd))
-
-```
-
-```{r read combinations 3rd, eval = TRUE}
+```{r read combinations 3rd}
 ret3rd <- lapply(var.CL.comb3rd, function(x) {
-    file <- sprintf("/homes/jpalmu/maaslinruns/maaslinruns/1560338417_%s/maaslin/maaslin.txt",
+    file <- sprintf("/homes/jpalmu/maaslinruns/combinations/all3rd_%s/maaslin/maaslin.txt",
                     paste0(x, collapse = "_"))

### 2019-08-26 articletwo-officer.R
Sensitivity analysis

@@ -66,7 +66,7 @@ tableone <- function(dset) {
         flextable::align(align = "left", part = "body", j = 1)
 }
 
-typologyformatter <- function(data, font = 12, typology) {
+typologyformatter <- function(data, font = 12, typology, left = c(1), hleft = c(1)) {
   flex <- flextable(data = data) %>%
     flextable::theme_booktabs() %>%
     flextable::border(border = fp_border(width=0), part="body") %>%
@@ -74,21 +74,24 @@ typologyformatter <- function(data, font = 12, typology) {
     flextable::border(part="header", border.bottom = fp_border(width=1))
 
   if (!missing(typology)) {
-    flex <- flex %>%
-      set_header_df(mapping = typology, key = "col_keys") %>%
-      merge_h(part = "header") %>%
-      flextable::border(part="header", border.bottom = fp_border(width=1))
+      flex <- flex %>%
+          set_header_df(mapping = typology, key = "col_keys") %>%
+          merge_h(part = "header") %>%
+          flextable::border(part="header", border.bottom = fp_border(width=1))
+      if (missing(hleft)) {
+          hleft <- c(2)
+      }
   }
 
   flex %>%
-    flextable::border(i = nrow(data), part="body", border.bottom = fp_border(width=1)) %>%
-    flextable::bold(bold = FALSE, part = "header") %>%
-    flextable::bold(bold = FALSE, part = "body") %>%
-    flextable::fontsize(size = font, part = "header") %>%
-    flextable::fontsize(size = font, part = "body") %>%
-    flextable::align(align = "center", part = "all") %>%
-    flextable::align(align = "left", part = "header", j = 1) %>%
-    flextable::align(align = "left", part = "body", j = 1)
+      flextable::border(i = nrow(data), part="body", border.bottom = fp_border(width=1)) %>%
+      flextable::bold(bold = FALSE, part = "header") %>%
+      flextable::bold(bold = FALSE, part = "body") %>%
+      flextable::fontsize(size = font, part = "header") %>%
+      flextable::fontsize(size = font, part = "body") %>%
+      flextable::align(align = "center", part = "all") %>%
+      flextable::align(align = "left", part = "header", j = left, i = hleft) %>%
+      flextable::align(align = "left", part = "body", j = left)
 }
 
 xtableformatter <- function(table, font = 12) {

### 2019-08-26 articletwo-sensitivityanalysis.Rmd

---
title: "Hypertension and microbiome"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(max.print=100)
```

# Importing libraries:

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
```

# Import previous session

```{r Import previous session}
load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
```

## Variables

```{r my variables}
var.CL.opt <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB", "BP_TREAT")
```

# Sensitivity analysis

```{r sensitivity analysis bp treat}
pseq.species.askbp <- subset_samples(pseq.species,
                                       PREVAL_DIAB == 0 &
                                       PREVAL_CHD == 0 &
                                       PREVAL_CR_ANYCANC == 0)

bray.dist.m.species.askbp <- bray.dist.m.species[pseq.species.askbp %>% meta %>% rownames,
                                                   pseq.species.askbp %>% meta %>% rownames]

alphadiversity.askbp <- calculate.alphadiversity(pseq.species.askbp,
                                                   list("max" = var.CL.opt,
                                                        "min" = var.CL.min))

betadiversity.askbp <- calculate.betadiversity(pseq = pseq.species.askbp,
                                                matrix = bray.dist.m.species.askbp,
                                                vars = list("max" = var.CL.opt,
                                                            "min" = var.CL.min))

diversities.helthy <- mergediversities(alphadiversity.askbp,
                                       betadiversity.askbp,
                                       names.dset = names.dset)

plot.diversities(diversities.askbp)
```

Subjects with diabetes, cancer, kidney disease, and coronary disease are dropped.

```{r sensitivity analysis only healthy}
pseq.species.healthy <- subset_samples(pseq.species,
                                       PREVAL_DIAB == 0 &
                                       PREVAL_CHD == 0 &
                                       PREVAL_CR_ANYCANC == 0)

bray.dist.m.species.healthy <- bray.dist.m.species[pseq.species.healthy %>% meta %>% rownames,
                                                   pseq.species.healthy %>% meta %>% rownames]

alphadiversity.healthy <- calculate.alphadiversity(pseq.species.healthy,
                                                   list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                        "min" = var.CL.min))

betadiversity.healthy <- calculate.betadiversity(pseq = pseq.species.healthy,
                                                matrix = bray.dist.m.species.healthy,
                                                vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                            "min" = var.CL.min))

diversities.helthy <- mergediversities(alphadiversity.healthy,
                                       betadiversity.healthy,
                                       names.dset = names.dset)

plot.diversities(diversities.healthy)
```

### 2019-08-26 articletwo.Rmd

@@
 library(M3C)
+library(emmeans)
 ```
@@
 
-Loading phyloseq object
+eLoading phyloseq object
 
@@
  if (!file.exists("rds/adonis.species.rds")) {
-     adonis.species = list()
-     adonis.species$min <- calculateadonis(dset = meta(pseq.species),
+     adonis.species <- calculate.betadiversity(pseq = pseq.species,
                                                matrix = bray.dist.m.species,
-                                           covariates = var.CL.min,
-                                           npermutations = 9999)
-     adonis.species$max <- calculateadonis(dset = meta(pseq.species),
-                                           matrix = bray.dist.m.species,
-                                           covariates = var.CL,
-                                           npermutations = 9999)
+                                               vars = list("max" = var.CL,
+                                                            "min" = var.CL.min))
+
      saveRDS(adonis.species, file = "rds/adonis.species.rds")
@@
 ```{r alphabeta definitions}
-dset.alpha <- meta.merge.alphadiversity(pseq.species) %>%
-    mutate(diversity_shannon = scale(diversity_shannon),
-           MAP = oMAP,
-           PULSEPRESSURE = oPULSEPRESSURE,
-           SYSTM = oSYSTM,
-           DIASM = oDIASM)
-
-alphadiversity <- list(
-    min = calculateglm(dset.alpha,
-                       modelstring = sprintf("%%s ~ %s + diversity_shannon",
-                                             paste0(var.CL.min, collapse = " + ")),
-                       filterstr = "shannon"),
-    max = calculateglm(dset.alpha,
-                       modelstring = sprintf("%%s ~ %s + diversity_shannon",
-                                             paste0(var.CL, collapse = " + ")),
-                       filterstr = "shannon"))
+alphadiversity <- calculate.alphadiversity(pseq.species,
+                                           list("max" = var.CL,
+                                                "min" = var.CL.min))
 
@@
 ```{r save grob}
-alpha.diversity.min <- plot.alpha(diversities$min, alphabreaks = seq(-1,0,0.2))
-alpha.diversity.max <- plot.alpha(diversities$max, alphabreaks = seq(-1,0,0.2))
-
-beta.diversity.min <- plot.beta(diversities$min)
-beta.diversity.max <- plot.beta(diversities$max)
-
-gs4 <- list(alpha.diversity.min$yaxis,
-            alpha.diversity.min$plot,
-            beta.diversity.min$plot,
-            alpha.diversity.min$legend,
-            alpha.diversity.min$xaxis,
-            beta.diversity.min$xaxis,
-            beta.diversity.min$xlab,
-            text_grob("A.", size = 18),
-            alpha.diversity.max$yaxis,
-            alpha.diversity.max$plot,
-            beta.diversity.max$plot,  
-            alpha.diversity.max$xaxis,
-            beta.diversity.max$xaxis,
-            beta.diversity.max$xlab,
-            text_grob("B.", size = 18))
-
-g4 <- arrangeGrob(grobs = gs4,
-                   layout_matrix = rbind(
-                      c(8,    NA,    NA, NA,   NA, NA, NA),
-                      c(NA, 1,    2,    NA, 3, NA, NA),
-                      c(NA, 1,    2,    NA, 3, NA, 4),
-                      c(NA, NA, 5,    NA,6, NA, NA),
-                      c(NA, NA, NA, NA,7, NA, NA),
-                      c(15,    NA,    NA, NA,   NA, NA, NA),
-                      c(NA, 9,    10,  NA,  11, NA, NA),
-                      c(NA, 9,    10,  NA,  11, NA, NA),
-                      c(NA, NA, 12,   NA, 13, NA, NA),
-                      c(NA, NA, NA, NA,14, NA, NA)),
-                  widths=c(0.7, 4, 0.7, 0.1, 5, 0.7, 3),
-                  heights=rep(c(1, 1, 6, 1, 2), 2))
-
+g4 <- plot.diversities(diversities)
 ggsave(file = "cache/alpha-beta.png", plot=g4, height=6, width=9)
@@
     return(axis.labels[value])
-}
-
-plot_labeller <-  function(value) {
-        plot.labels <- list("Axis.2" = sprintf("A."),
-                            "Axis.3" = sprintf("B."))
-        return(plot.labels[value])
 }
@@
     coord_cartesian(clip = 'off') +
-    geom_text(aes(x = -Inf, y = Inf, label = plot_labeller(key)),
-              size = 5,
-              hjust = 2,
-              vjust = -1,
-              inherit.aes = FALSE) +
     theme(text = element_text(size=12),
-          legend.position = "none",
+          legend.position = c(0.35, 0.95),
+          legend.title = element_blank(),
+          legend.text = element_text(size = 10),
+          legend.key.size = unit(5, "mm"),
+          legend.background = element_blank(),
           strip.background = element_blank(),
@@
 
-```{r top generas, eval = FALSE, include = FALSE}
-abund.genus <- abundances(pseq.genus, transform = "compositional") %>%
-    t %>%
+```{r top pcoa genera}
+df.genus.abund <- abundances(pseq.genus, transform = "compositional") %>%
     as.data.frame %>%
-    tibble::rownames_to_column("id") %>%
-    setNames(tolower(replace.brackets(names(.))))
-
-abund.genus.bacteria <- abund.genus %>% select(-contains("virus")) %>% colnames
-
-abund.df <- merge(pcoa.vectors %>%
-                  select("Axis.1", "Axis.2", "Axis.3") %>%
-                  tibble::rownames_to_column("id"),
-                  abund.genus %>% select(id, abund.genus.bacteria),
-                  by = "id") %>%
-    merge(.,  meta(pseq.genus) %>% tibble::rownames_to_column("id"), by = "id")
+    tibble::rownames_to_column() %>%
+    dplyr::mutate(rowname = make.names(rowname)) %>%
+    gather(id, value, -rowname) %>% 
+    spread(rowname, value)
+
+df.genus.pcoa <- merge(pcoa.vectors %>% tibble::rownames_to_column("id"),
+                         df.genus.abund,
+                         by = "id")
 
 pcoacorabund <- function(df,
-                         genera.names,
+                         taxas,
                          axis.names = c("Axis.1", "Axis.2", "Axis.3")) {
-    lapply(c2l(genera.names), function (taxa) 
+    mclapply(c2l(taxas), function (taxa) 
         lapply(c2l(axis.names), function(axis) {
@@
             summary(model)$r.squared
-        }))
-}
-
-pcoacor.genus <- pcoacorabund(abund.df, genera.names = abund.genus.bacteria)
-pcoacor.genus.df <- map_df(pcoacor.genus, ~as.data.frame(.x), .id="id")
-
-pcoacor.genus.df %>%
-    arrange(-Axis.3) %>%
-    head(20) %>% kable
+        }), mc.cores = 20)
+}
+
+pcoa.correlations <- pcoacorabund(df.genus.pcoa, taxas = colnames(df.genus.abund)[-1])
+
+pcoa.correlations.df <- map_df(pcoa.correlations, ~as.data.frame(.x), .id="id")
+
+pcoa.correlations.df %>% head
+
+pcoa.top.names <- lapply(c2l("Axis.1", "Axis.2", "Axis.3"), function(x) {
+    top.names <- pcoa.correlations.df %>% arrange(desc(!!sym(x))) %>% head(10) %>% pull(id)
+    pcoa.correlations.df %>%
+        dplyr::filter(id %in% top.names) %>%
+        select(id, x) %>%
+        tibble::rownames_to_column(var = "number") %>%
+        dplyr::mutate(!! sprintf("taxa_%s", x) := gsub("g_", "", id)) %>%
+        select(one_of(sprintf("taxa_%s", x), x, "number"))
+    })
+
+pcoa.top.table <- pcoa.top.names %>%
+    reduce(left_join, by = "number") %>%
+    select(-number) %>%
+    mutate_if(is.numeric, round, 3)
 ```
@@
 ```{r pcoa analysis main axis}
-glm.pcoa <- lapply(c2l(paste0("Axis.", seq(1,5))), function(axis)
+glm.pcoa <- lapply(c2l(paste0("Axis.", seq(1,3))), function(axis)
     calculateglm(pcoa.df,
@@
     select(response, term, estimate, std.error, p.value, conf.low, conf.high) %>%
+    dplyr::mutate(Qval = p.adjust(p.value, method="BH")) %>%
     mutate_if(is.numeric, round, 3) %>%
-    arrange(p.value) %>%
+    arrange(Qval) %>%
     kable
@@
 ```{r lacto adjusted means}
-df.genus.salt.unadjusted <- with(df.genus.salt, data.frame(salt_groups = as.factor(seq(3)),
-                                                           SEX = as.factor(1),
-                                                           BL_AGE = mean(BL_AGE)))
-
-adjmean.lacto.saltgroup <- predict(lm.lacto.saltgroup,
-                               df.genus.salt.unadjusted,
-                               interval = "prediction") %>%
+adjmean.lacto.saltgroup <- emmeans::emmeans(lm.lacto.saltgroup, specs = c("salt_groups")) %>%
     as.data.frame %>%
-    dplyr::rename(estimate = fit,
-                  conf.low = lwr,
-                  conf.high = upr) %>%
-    tibble::rownames_to_column("term")
+    dplyr::rename(term = salt_groups, estimate = emmean, conf.low = lower.CL, conf.high = upper.CL)
 ```
@@
     myforestplot(adjmean.lacto.saltgroup,
-                 xlab = "Dietary salt",
+                 xlab = "Tertile of 24-hour urinary sodium",
                  ylab = "Lactobacillus abundance",
                  subtitle = "A.",
-                 my_y_scale = scale_y_continuous(limits = c(3.0, 8.0), breaks = seq(0, 10, 1))),
+                 my_y_scale = scale_y_continuous(limits = c(5.4, 5.65), breaks = seq(0, 10, 0.1))),
     myforestplot(odds.hypertension.lactogroup,
-                 xlab = "Lactobacillus abundance",
+                 xlab = "Tertile of Lactobacillus abundance",
                  ylab = "Odds of hypertension",
@@
 
-ggsave(file = "cache/salt-forest.png", plot=gsalt, height=3, width=5, dpi = 300)
+ggsave(file = "cache/salt-forest.png", plot=gsalt, height=3, width=5.5, dpi = 300)
 ```
@@
 writeimage(doc, 4, "cache/salt-forest.png", fig4head, fig4foot)
+```
+
+Testing salt model using maaslin
+
+```{r salt model maaslin}
+taxa(pseq.genus.core) %>% mygrep(word = "lacto")
+
+pseq.genus.salt <- prune_taxa(c("g_Lactobacillus"), pseq.genus.core)
+
+pseq.genus.salt.meta <- meta(pseq.genus.salt) %>%
+    tibble::rownames_to_column(var = "rowname") %>%
+    dplyr::filter(!is.na(NA.)) %>%
+    dplyr::mutate(dUNA = myscale(NA.),
+                  salt_groups = relevel(factor(dplyr::ntile(dUNA, 3)), ref='1')) %>%
+    tibble::remove_rownames() %>%
+    tibble::column_to_rownames(var = "rowname") %>%
+    select(var.BP, var.CL.min, starts_with("salt_groups"))
+
+phyloseq::sample_data(pseq.genus.salt) <- phyloseq::sample_data(pseq.genus.salt.meta)
+
+maaslin.salt <- maaslinwrapper(pseq = pseq.genus.salt,
+                               looped = var.BP,
+                               forced = c(var.CL.min, "salt_groups"),
+                               taxa = "g_Lactobacillus")
+
+maaslin.salt %>%
+    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
+    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
+                  Feature = gsub("[gs]_", "", Feature),
+                  beta = Coefficient) %>%
+    arrange(Feature)
 ```
@@
 
+```{r write supplement}
+core.names <- pseq.genus.core %>%
+    taxa %>%
+    gsub("g_", "", .) %>%
+    as.data.frame %>%
+    setNames("name") %>%
+    arrange(name) %>%
+    mutate(name = as.character(name)) %>%
+    mutate(class = case_when(row_number()-1 < n()/4 ~ 1,
+                             row_number()-1 < 2*n()/4 ~ 2,
+                             row_number()-1 < 3*n()/4 ~ 3,
+                             TRUE ~ 4))
+
+core.names %>% pull(class) %>% as.factor %>% summary
+
+core.names.fourcols <- cbind(
+    core.names %>% filter(class == 1) %>% pull(name),
+    core.names %>% filter(class == 2) %>% pull(name),
+    core.names %>% filter(class == 3) %>% pull(name),
+    core.names %>% filter(class == 4) %>% pull(name)) %>%
+    as.data.frame
+
+core.names.fourcols %>% unlist %>% length
+
+core.names.flextable <- flextable(core.names.fourcols) %>%
+    flextable::fontsize(size = 10, part = "header") %>%
+    flextable::fontsize(size = 10, part = "body") %>%
+    flextable::width(j=1:4, width = 1.8) %>%
+    flextable::align(align = "left", part = "header") %>%
+    flextable::align(align = "left")
+
+doc.s <- read_docx(path = "style/articlestyle.docx") %>%
+  body_remove()
+writetable(doc.s, core.names.flextable, number = "S1", "Core genera", "") %>%
+print(doc.s, target = paste0("report/articletwo-supplement.docx"))
+```
+
 ```{r Write docx to file, include=FALSE}

### 2019-08-26 articletwo-sensitivityanalysis.Rmd
Fix model 1

@@
 ```{r sensitivity analysis bp treat}
-pseq.species.askbp <- subset_samples(pseq.species,
-                                       PREVAL_DIAB == 0 &
-                                       PREVAL_CHD == 0 &
-                                       PREVAL_CR_ANYCANC == 0)
-
-bray.dist.m.species.askbp <- bray.dist.m.species[pseq.species.askbp %>% meta %>% rownames,
-                                                   pseq.species.askbp %>% meta %>% rownames]
-
-alphadiversity.askbp <- calculate.alphadiversity(pseq.species.askbp,
+alphadiversity.askbp <- calculate.alphadiversity(pseq.species,
                                                    list("max" = var.CL.opt,
@@
 
-betadiversity.askbp <- calculate.betadiversity(pseq = pseq.species.askbp,
-                                                matrix = bray.dist.m.species.askbp,
+betadiversity.askbp <- calculate.betadiversity(pseq = pseq.species,
+                                                matrix = bray.dist.m.species,
                                                 vars = list("max" = var.CL.opt,

### 2019-08-28 articletwo-sensitivityanalysis.Rmd
Smaller number of permutations

@@
                                                 vars = list("max" = var.CL.opt,
-                                                            "min" = var.CL.min))
+                                                            "min" = var.CL.min),
+                                               npermutations = 999)
 
@@
                                                 vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
-                                                            "min" = var.CL.min))
+                                                            "min" = var.CL.min),
+                                                npermutations = 999)
 
@@
 ```
+
+## BMI
+
+Normal weigth
+
+```{r sensitivity analysis normal weight}
+pseq.species.normalweight <- subset_samples(pseq.species, BMI > 18.5 & BMI <= 25)
+
+bray.dist.m.species.normalweight <-
+    bray.dist.m.species[pseq.species.normalweight %>% meta %>% rownames,
+                        pseq.species.normalweight %>% meta %>% rownames]
+
+alphadiversity.normalweight <- calculate.alphadiversity(pseq.species.normalweight,
+                                                   list("max" = var.CL[var.CL != "BMI"],
+                                                        "min" = var.CL.min))
+
+betadiversity.normalweight <- calculate.betadiversity(pseq = pseq.species.normalweight,
+                                                matrix = bray.dist.m.species.normalweight,
+                                                vars = list("max" = var.CL[var.CL != "BMI"],
+                                                            "min" = var.CL.min),
+                                                npermutations = 999)
+
+diversities.helthy <- mergediversities(alphadiversity.normalweight,
+                                       betadiversity.normalweight,
+                                       names.dset = names.dset)
+
+plot.diversities(diversities.normalweight)
+```
+
+Obese
+
+```{r sensitivity analysis obese}
+pseq.species.obese <- subset_samples(pseq.species, BMI > 30 & BMI <= 40)
+
+bray.dist.m.species.obese <-
+    bray.dist.m.species[pseq.species.obese %>% meta %>% rownames,
+                        pseq.species.obese %>% meta %>% rownames]
+
+alphadiversity.obese <- calculate.alphadiversity(pseq.species.obese,
+                                                   list("max" = var.CL[var.CL != "BMI"],
+                                                        "min" = var.CL.min))
+
+betadiversity.obese <- calculate.betadiversity(pseq = pseq.species.obese,
+                                                matrix = bray.dist.m.species.obese,
+                                                vars = list("max" = var.CL[var.CL != "BMI"],
+                                                            "min" = var.CL.min),
+                                               npermutations = 999)
+
+diversities.helthy <- mergediversities(alphadiversity.obese,
+                                       betadiversity.obese,
+                                       names.dset = names.dset)
+
+plot.diversities(diversities.obese)
+```

### 2019-08-28 articletwo.Rmd
Deseq code and logistic model sample size calculation

@@
 }
-
-#packrat::opts$ignored.packages
 ```
@@
 library(emmeans)
+library(DESeq2)
 ```
@@
 
-```{r t-sne, eval = FALSE}
-sne.df <- pseq.genus %>%
-    microbiome::transform(transform = "clr") %>%
-    psmelt %>%
-    spread(OTU, Abundance)
-
-sne_names <- colnames(sne.df) %>% mygrep(word = "g_")
-
-sne.t <- sne.df %>% select(Sample, HYPERTENSION, sne_names[1:20]) %>%
-    mutate(HYPERTENSION = as.numeric(HYPERTENSION)) %>%
-    gather(variable, value, -Sample) %>% 
-    spread(Sample, value) %>%
-    column_to_rownames("variable")
-
-png("sample.png", 800, 800)
-tsne(sne.t[-1,], labels = as.factor(sne.t[1,]),
-     dotsize = 1)
-dev.off()
-
-### CONTINUOUS
-
-source("tsne.R")
-
-sne.t <- sne.df %>% select(Sample, oMAP, sne_names) %>%
-    gather(variable, value, -Sample) %>% 
-    spread(Sample, value) %>%
-    column_to_rownames("variable")
-
-tsne.g <- tsne(sne.t[rownames(sne.t) != "oMAP",],
-               labels = as.numeric(sne.t[rownames(sne.t) == "oMAP",]),
-               dotsize = 1,
-               controlscale = TRUE,
-               scale = 2,
-               low = "white")
-
-tsne.plot <- tsne.g +
-    scale_fill_gradientn(limits = c(-3,3),
-                         colours=c("navyblue", "darkmagenta", "darkorange1"),
-                         breaks=c(0,200), labels=format(c(0,200)))
-
-ggsave(file = "cache/tsne.png", plot=tsne.plot, height=6, width=9)
-
 ```{r maaslin runs, results = 'hide'} 
@@
  }
+```
+
+```{r deseq2, eval = FALSE}
+pseq.genus.core.deseq <- pseq.genus %>% core(detection = 0.0001, prevalence = 0.01)
+
+dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq,
+                               as.formula(sprintf("~ %s + %s", "MAP", paste0(var.CL, collapse = "+"))))
+
+dds = DESeq(dds.data, test="Wald", fitType="parametric")
+resultsNames(dds)
+res = results(dds)
 ```
@@
 
+Required sub sample size
+
+```{r required sodium subsample size}
+lapply(c2l(seq(1, 1.7, 0.05)), function(or) 
+    powerMediation::powerLogisticCon(n = nrow(df.genus.salt),
+                                     p1 = 0.4,
+                                     OR = or)) %>%
+    map_df(., ~as.data.frame(.x), .id = "or") %>%
+    kable
+```
+
 ### How sodium affects lactobacillus abundances?

### 2019-08-28 articletwo-sensitivityanalysis.Rmd
Revert "Deseq2"
This reverts commit b47c65788968223cf674043ebbb895a53b7e1867.

@@
 
+### RR biome functions
+
+<details>
+  <summary>Open/Close</summary>
+
+```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
+source("articletwo-rrbiome.R")
+```
+
+</details>
+
+### Plot functions
+
+<details>
+  <summary>Open/Close</summary>
+
+```{r Plot functions, code=readLines("articletwo-ggplot.R")}
+source("articletwo-ggplot.R")
+```
+
+</details>
+
 # Import previous session
@@
 ```{r Import previous session}
-load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
+pseq.species <- import_filter_data("data/phfinrisk_centrifuge_species_all_drop50k_2018-11-30_2019-06-03.RDs")
+bray.dist.m.species  <- readRDS("rds-20190828/bray.dist.m.species.rds")
+names.dset <- getdescriptions()
 ```
@@
 ```{r my variables}
+var.BP <- c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")
+var.CL.min <- c("BL_AGE", "SEX")
+var.CL <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
+            "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
 var.CL.opt <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB", "BP_TREAT")
+permutations <- 3
 ```
@@
                                                             "min" = var.CL.min),
-                                               npermutations = 999)
+                                               npermutations = permutations)
 
-diversities.helthy <- mergediversities(alphadiversity.askbp,
+diversities.askbp <- mergediversities(alphadiversity.askbp,
                                        betadiversity.askbp,
@@
 
-Subjects with diabetes, cancer, kidney disease, and coronary disease are dropped.
+Subjects with diabetes, cancer, and coronary disease are dropped. No
+easy variable was found for kidney disesase.
 
@@
                                                             "min" = var.CL.min),
-                                                npermutations = 999)
+                                                npermutations = permutations)
 
-diversities.helthy <- mergediversities(alphadiversity.healthy,
+diversities.healthy <- mergediversities(alphadiversity.healthy,
                                        betadiversity.healthy,
@@
                                                             "min" = var.CL.min),
-                                                npermutations = 999)
+                                                npermutations = permutations)
 
-diversities.helthy <- mergediversities(alphadiversity.normalweight,
+diversities.normalweight <- mergediversities(alphadiversity.normalweight,
                                        betadiversity.normalweight,
@@
                                                             "min" = var.CL.min),
-                                               npermutations = 999)
+                                               npermutations = permutations)
 
-diversities.helthy <- mergediversities(alphadiversity.obese,
+diversities.obese <- mergediversities(alphadiversity.obese,
                                        betadiversity.obese,
@@
 ```
+
+```{r save session}
+save.image(file = "session/session-sensitivity.Rdata")
+```

### 2019-08-28 articletwo-rrbiome.R

@@ -334,7 +334,7 @@ calculate.alphadiversity <- function(pseq, vars, modelstr = "%%s ~ %s + diversit
 calculate.betadiversity <- function(pseq, matrix, vars, npermutations = 9999) {
     lapply(vars, function(var)
         calculateadonis(dset = meta(pseq),
-                                      matrix = matrix,
-                                      covariates = var,
-                                      npermutations = npermutations))
+                        matrix = matrix,
+                        covariates = var,
+                        npermutations = npermutations))
 }

### 2019-08-28 articletwo-sensitivityanalysis.Rmd
99 permutaions on sensitivity

@@
 var.CL.opt <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB", "BP_TREAT")
-permutations <- 3
+permutations <- 99
 ```

### 2019-08-29 articletwo.Rmd
Deseq2
Supplemet S2, S3. Decrease permutations to 999

@@
               ofile = '~/phd/research/articletwo/TAGS', verbose = TRUE, append = FALSE)
-    if ("loadlast" %in% names(args) )
-        load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
+    if ("loadlast" %in% names(args) ) {
+        file <- paste0("session/", sort(list.files("session"), decreasing = TRUE)[1])
+        message("Loading variables from file", file)
+        load(file)
+    }
     if ("subset" %in% names(args))
@@
 
-dds = DESeq(dds.data, test="Wald", fitType="parametric")
-resultsNames(dds)
-res = results(dds)
+#dds = DESeq(dds.data, test="Wald", fitType="parametric")
+#resultsNames(dds)
+#res = results(dds)
+
+ds2 <- estimateSizeFactors(dds.data, type="poscounts")
+dds <- DESeq(ds2)
+
+res <- results(dds)
+df <- as.data.frame(res)
+df$taxon <- rownames(df)
+  #df <- df %>% arrange(pvalue) # TODO
+
+  df$model.str <- modelstr
+  df$covariate <- variables[nv]
+
 ```
@@
 diversities <- mergediversities(alphadiversity, adonis.species, names.dset = names.dset)
+```
+
+```{r bptreat numbers}
+pseq.species %>% meta %>% pull(BP_TERAT) %>% summary
 ```
@@
 
-core.names %>% pull(class) %>% as.factor %>% summary
-
 core.names.fourcols <- cbind(
@@
 
-core.names.fourcols %>% unlist %>% length
-
 core.names.flextable <- flextable(core.names.fourcols) %>%
@@
     flextable::align(align = "left")
-
+```
+
+```{r supplement table for alpha diversity}
+alphadiversity.table <- cbind(alphadiversity$min %>% dplyr::rename_all(paste0, ".min"),
+                                  alphadiversity$max %>% dplyr::rename_all(paste0, ".max")) %>%
+    dplyr::filter(model_name.min == model_name.max) %>%
+    merge(names.dset, by.x ="response.min", by.y = "Covariate") %>%
+    mutate(mean_ci.min = sprintf("%.2f (%.2f - %.2f)", estimate.min, conf.low.min, conf.high.min),
+           pubp.min = pub.p(p.value.min),
+           mean_ci.max = sprintf("%.2f (%.2f - %.2f)", estimate.max, conf.low.max, conf.high.max),
+           pubp.max = pub.p(p.value.max)) %>%
+    arrange(match(response.min, c("SYSTM", "DIASM", "MAP", "PULSEPRESSURE", "HT"))) %>%
+    select(Name, mean_ci.min, pubp.min, mean_ci.max, pubp.max)
+
+typology.tbls2 <- data.frame(
+    col_keys = colnames(alphadiversity.table),
+    what = c("", rep("Age- and sex adjusted model", 2), rep("Multivaraible adjusted model", 2)),
+    measure = c("", rep(c("beta (95%-CI)", "p"), 2)),
+    stringsAsFactors = FALSE)
+
+alphadiversity.flextable <-
+    typologyformatter(data = alphadiversity.table, font = 12, typology = typology.tbls2) %>%
+    flextable::width(j=1, width = 1.9) %>%
+    flextable::width(j=c(2:2,4:4), width = 1.8) %>%
+    flextable::width(j=c(3:3,5:5), width = 0.7)
+```
+
+```{r supplement table for beta diversity}
+betadiversity.table <- merge(diversities$min %>% dplyr::rename_all(paste0, ".min"),
+      diversities$max %>% dplyr::rename_all(paste0, ".max")) %>%
+    dplyr::filter(response.min == response.max) %>%
+    merge(names.dset, by.x ="response.min", by.y = "Covariate") %>%
+    mutate(R2.min = sprintf("%.3f%%", R2.min*100),
+           pubp.min = pub.p(R2.p.min),
+           R2.max = sprintf("%.3f%%", R2.max*100),
+           pubp.max = pub.p(R2.p.max)) %>%
+    arrange(match(response.min, c("SYSTM", "DIASM", "MAP", "PULSEPRESSURE", "HT"))) %>%
+    select(Name, R2.min, pubp.min, R2.max, pubp.max)
+
+typology.tbls3 <- data.frame(
+    col_keys = colnames(betadiversity.table),
+    what = c("", rep("Age- and sex adjusted model", 2), rep("Multivaraible adjusted model", 2)),
+    measure = c("", rep(c("R2", "p"), 2)),
+    stringsAsFactors = FALSE)
+
+betadiversity.flextable <-
+    typologyformatter(data = betadiversity.table, font = 12, typology = typology.tbls3) %>%
+    flextable::width(j=1, width = 1.9) %>%
+    flextable::width(j=c(2:5), width = 0.8)
+
+```
+
+```{r Write supplement}
 doc.s <- read_docx(path = "style/articlestyle.docx") %>%
   body_remove()
-writetable(doc.s, core.names.flextable, number = "S1", "Core genera", "") %>%
+writetable(doc.s, core.names.flextable, number = "S1", "Core genera", "")
+writetable(doc.s, alphadiversity.flextable, number = "S2", "Alpha diverstity", "")
+writetable(doc.s, betadiversity.flextable, number = "S3", "Beta diverstity", "")
 print(doc.s, target = paste0("report/articletwo-supplement.docx"))

### 2019-08-29 articletwo-rrbiome.R

@@ -286,7 +286,7 @@ c2l <- function(...) {
 pub.p <- function(p, Nfdr = FALSE) {
     p <- as.numeric(p)
     if (Nfdr) p <- p.adjust(p, method="BH", n = Nfdr)
-    ifelse(p < 0.01, ifelse(p<0.001, sprintf("%.2e", p), sprintf("%.3f", p)), sprintf("%.2f", p))
+    ifelse(p < 0.01, ifelse(p<0.001, "<0.001", sprintf("%.3f", p)), sprintf("%.2f", p))
 }
 
 firstup <- function(x) {
@@ -331,7 +331,7 @@ calculate.alphadiversity <- function(pseq, vars, modelstr = "%%s ~ %s + diversit
 
 }
 
-calculate.betadiversity <- function(pseq, matrix, vars, npermutations = 9999) {
+calculate.betadiversity <- function(pseq, matrix, vars, npermutations = 999) {
     lapply(vars, function(var)
         calculateadonis(dset = meta(pseq),
                         matrix = matrix,

### 2019-09-02 articletwo.Rmd
Minor

@@
 ```{r bptreat numbers}
-pseq.species %>% meta %>% pull(BP_TERAT) %>% summary
-```
-
-Hypertension prevalence
-
-```{r test htn prevalence}
-dset.alpha %>%
-    dplyr::filter(BL_AGE < 60 + 12, BL_AGE > 60 - 12) %>%
-    summarize(htn = sum(ifelse(HYPERTENSION == 1, 1, 0)), n = n(), proportion = htn/n) %>%
-    kable
-```
-
-Minimum coefficient for alpha diversity is `r rbind(diversities$min,
-diversities$max) %>% pull(estimate) %>% min`.
+pseq.species %>% meta %>% pull(BP_TREAT) %>% summary
+```
 
@@
 
-```{r top pcoa genera}
+```{r top pcoa genera, eval = FALSE}
 df.genus.abund <- abundances(pseq.genus, transform = "compositional") %>%

### 2019-09-02 articletwo.Rmd
Run error

@@
 
-```{r maaslin means}
+```{r maaslin means, eval = FALSE}
 pseq.genus.core.table <- pseq.genus.core %>%

### 2019-09-02 articletwo.Rmd
Deseq2

@@
 
-```{r deseq2, eval = FALSE}
+```{r deseq2}
 pseq.genus.core.deseq <- pseq.genus %>% core(detection = 0.0001, prevalence = 0.01)
@@
                                as.formula(sprintf("~ %s + %s", "MAP", paste0(var.CL, collapse = "+"))))
-
-#dds = DESeq(dds.data, test="Wald", fitType="parametric")
-#resultsNames(dds)
-#res = results(dds)
 
@@
 df$taxon <- rownames(df)
-  #df <- df %>% arrange(pvalue) # TODO
 
@@
 
+df %>% kable
 ```

### 2019-09-04 articletwo.Rmd
Parallel
Minor

@@
 ```{r deseq2}
-pseq.genus.core.deseq <- pseq.genus %>% core(detection = 0.0001, prevalence = 0.01)
+pseq.genus.core.deseq <-  prune_taxa(taxa(pseq.genus.core), pseq.genus)
 
@@
 ds2 <- estimateSizeFactors(dds.data, type="poscounts")
-dds <- DESeq(ds2)
-
-res <- results(dds)
-df <- as.data.frame(res)
-df$taxon <- rownames(df)
-
-df$model.str <- modelstr
-df$covariate <- variables[nv]
-
-df %>% kable
+
+dds <- DESeq(ds2, fitType='mean', parallel = TRUE, BPPARAM=MulticoreParam(16))
+
+results(dds) %>%
+    as.data.frame %>%
+    kable
 ```

### 2019-09-12 articletwo.Rmd
save rds
Minor

@@
 dds <- DESeq(ds2, fitType='mean', parallel = TRUE, BPPARAM=MulticoreParam(16))
+saveRDS(dds, file = "rds/dds.rds")
+
+dds <- readRDS(file = "rds/dds.rds")
 
@@
     as.data.frame %>%
+    tibble::rownames_to_column("Feature") %>%
+    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
+    dplyr::filter(qval < 0.05) %>%
     kable

### 2019-09-12 articletwo.Rmd
HT instead of MAP

--- 2019-09-04 articletwo.Rmd
@@
 dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq,
-                               as.formula(sprintf("~ %s + %s", "MAP", paste0(var.CL, collapse = "+"))))
+                               as.formula(sprintf("~ %s + %s", "HT", paste0(var.CL, collapse = "+"))))
 
@@
 dds <- DESeq(ds2, fitType='mean', parallel = TRUE, BPPARAM=MulticoreParam(16))
+
+saveRDS(dds, file = "rds/dds.rds")
+#dds <- readRDS(file = "rds/dds.rds")
 

### 2019-09-30 articletwo.Rmd
Typo
Save dds

@@
 ```{r deseq2}
+if (!file.exists("rds/dds.rds")) {
     pseq.genus.core.deseq <-  prune_taxa(taxa(pseq.genus.core), pseq.genus)
-
     dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq,
-                               as.formula(sprintf("~ %s + %s", "HT", paste0(var.CL, collapse = "+"))))
-
+                                   as.formula(sprintf("~ %s + %s", "HYPERTENSION", paste0(var.CL, collapse = "+"))))
     ds2 <- estimateSizeFactors(dds.data, type="poscounts")
-
     dds <- DESeq(ds2, fitType='mean', parallel = TRUE, BPPARAM=MulticoreParam(16))
-
     saveRDS(dds, file = "rds/dds.rds")
-#dds <- readRDS(file = "rds/dds.rds")
+} else {
+    dds <- readRDS(file = "rds/dds.rds")
+}
 

### 2019-09-30 articletwo.Rmd
Deseq

--- 2019-09-12 articletwo.Rmd
@@
 ```{r deseq2}
+if (!file.exists("rds/dds.rds")) {
     pseq.genus.core.deseq <-  prune_taxa(taxa(pseq.genus.core), pseq.genus)
-
     dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq,
-                               as.formula(sprintf("~ %s + %s", "HT", paste0(var.CL, collapse = "+"))))
-
-ds2 <- estimateSizeFactors(dds.data, type="poscounts")
-
-dds <- DESeq(ds2, fitType='mean', parallel = TRUE, BPPARAM=MulticoreParam(16))
-
+                                   as.formula(sprintf("~ %s + %s",
+                                                      paste0(var.CL, collapse = "+"),
+                                                      "HYPERTENSION")))
+
+    dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq, ~HYPERTENSION)
+
+    dds <- DESeq(dds.data, test="Wald", fitType="parametric", parallel = TRUE, BPPARAM=MulticoreParam(16))
     saveRDS(dds, file = "rds/dds.rds")
-#dds <- readRDS(file = "rds/dds.rds")
-
-results(dds) %>%
+} else {
+    dds <- readRDS(file = "rds/dds.rds")
+}
+```
+
+## DeSeq
+
+```{r deseq results table}
+results(dds, contrast = c("HYPERTENSION", 0, 1)) %>%
     as.data.frame %>%
@@
 ```
+
+```{r deseq plot}
+g.deseq <- results(dds, contrast = c("HYPERTENSION", 0, 1)) %>%
+    as.data.frame %>%
+    tibble::rownames_to_column("Feature") %>%
+    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
+    dplyr::filter(qval < 0.05) %>%
+    ggplot(aes(x=Feature, y=log2FoldChange, color=qval)) +
+    geom_point(size=1) +
+    geom_hline(yintercept = 0, color = "gray") +
+    theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) +
+    coord_flip() +
+    theme_classic()
+
+ggsave(file = "cache/deseq.png", plot=g.deseq, height=10, width=5)
+```
+
+<img src = "cache/deseq.png" />
 

### 2019-10-04 comparepseqs.Rmd
Beta plot show significance
Minor

@@
     if (length(missing) > 0) {
-
         pseq <- import_filter_data(rds[[key]],

### 2019-10-04 articletwo-rrbiome.R

@@ -97,9 +97,13 @@ calculateglm <- function(dset,
         as.data.frame
 }
 
-import_filter_data <- function(file) {
+import_filter_data <- function(file, included) {
     pseq.full <- readRDS(file)
-    pseq.meta <- filter.phenotype.data(pseq.full)
+    if (missing(included)) {
+        pseq.meta <- filter.phenotype.data(pseq.full)
+    } else {
+        pseq.meta <- filter.phenotype.data(pseq.full, included = included)
+    }
     phyloseq::sample_data(pseq.full) <- phyloseq::sample_data(pseq.meta)
     return(pseq.full)
 }

### 2019-10-04 comparepseqs.Rmd
Rewrote

--- 2019-06-04 comparepseqs.Rmd
@@
 for (key in names(rds)) {
-    missing = list()
-
-    if (!file.exists(paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))) {
-        missing$alphadistribution = TRUE
-    } else {
-        alphadistribution <- readRDS(file =
-                                         paste0("rds-comparepseqs/alphadistribution-",
-                                                key , ".rds"))
-    }
-
-    if (!file.exists(paste0("rds-comparepseqs/alpha-", key , ".rds"))) {
-        missing$alpha = TRUE
-    } else {
-        alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
-    }
-
-    if (!file.exists(paste0("rds-comparepseqs/bray-", key , ".rds"))) {
-        missing$bray = TRUE
-    } else {
-        bray.dist.matrix <- readRDS(file = paste0("rds-comparepseqs/bray-", key , ".rds"))
-    }
-
-    if (!file.exists(paste0("rds-comparepseqs/beta-", key , ".rds"))) {
-        missing$beta = TRUE
-    } else {
-        betadiversity <- readRDS(file = paste0("rds-comparepseqs/beta-", key , ".rds"))
-    }
-
-    if (length(missing) > 0) {
-
     pseq <- import_filter_data(rds[[key]],
@@
                                             "BL_USE_RX_C07", "BL_USE_RX_C08"))
-        dset <- meta.merge.alphadiversity(pseq) %>%
-            mutate(diversities_shannon = as.numeric(scale(diversities_shannon)))
 
-        if ("alphadistribution" %in% names(missing)) {
-            message("Missing alpha distribution")
-            alphadistribution <- dset %>%
-                mutate(!!key := diversities_shannon) %>%
-                select(Sample_ID, key)
-            saveRDS(alphadistribution,
-                    file = paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))
-        }
+    bray.dist.m.species <- calculate.beta.matrix(pseq)
 
-        if ("alpha" %in% names(missing)) {
-            message("Missing alpha")
-            alphadiversity <- list(min =  calculateglm(dset, modelstring = fo.alpha.min),
-                                   full = calculateglm(dset, modelstring = fo.alpha.full))
-            saveRDS(alphadiversity, file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
-        }
+    adonis.species <- calculate.betadiversity(pseq = pseq.species,
+                                               matrix = bray.dist.m.species,
+                                               vars = list("max" = var.CL,
+                                                            "min" = var.CL.min))
 
-        if ("bray" %in% names(missing)) {
-            message("Missing Bray")
-            bray.dist.matrix <- calculate.beta.matrix(pseq)
-            saveRDS(bray.dist.matrix, file = paste0("rds-comparepseqs/bray-", key , ".rds"))
-        }
+    alphadiversity <- calculate.alphadiversity(pseq.species,
+                                               list("max" = var.CL,
+                                                    "min" = var.CL.min))
 
-        if ("beta" %in% names(missing)) {
-            message("Missing beta")
-            betadiversity <- list(min = calculateadonis(dset = dset,
-                                                        matrix = bray.dist.matrix,
-                                                        covariates = covariates.min,
-                                                        npermutations = 1000),
-                                  full = calculateadonis(dset = dset,
-                                                         matrix = bray.dist.matrix,
-                                                         covariates = covariates.full,
-                                                         npermutations = 1000))
-            saveRDS(betadiversity, file = paste0("rds-comparepseqs/beta-", key , ".rds"))
-        }
-    }
+    alphadiversity <- calculate.alphadiversity(pseq.species,
+                                               list("max" = var.CL,
+                                                    "min" = var.CL.min))
 
-    ret[[key]] = list("alphadiversity" = alphadiversity,
-                      "betadiversity" = betadiversity,
-                      "alphadistribution" = alphadistribution)
+    diversities <- mergediversities(alphadiversity, adonis.species, names.dset = names.dset)
+
+    g4 <- plot.diversities(diversities)
+    ggsave(file = sprintf("cache/comparepseqs-%s.png", key), plot=g4, height=6, width=9)
 }
 ```
-
-```{r quit subruns}
-if (exists("subset.run")) {
-    quit()
-}
-```
-
-# Diversities for BP variables
-
-```{r plot alpha-beta}
-plotlist <- list()
-
-for (key in names(rds)) {
-    diversities <- mergediversities(ret[[key]][["alphadiversity"]],
-                                    ret[[key]][["betadiversity"]])
-
-    alpha.diversity.min <- plot.alpha(diversities$min, alphabreaks = c(-0.15, -0.10, -0.05, -0.01, 0))
-    alpha.diversity.full <- plot.alpha(diversities$full, alphabreaks = c(-0.15, -0.10, -0.05, -0.01, -0))
-
-    beta.diversity.min <- plot.beta(diversities$min)
-    beta.diversity.full <- plot.beta(diversities$full)
-
-    gs4 <- list(alpha.diversity.min$yaxis,
-                alpha.diversity.min$plot,
-                beta.diversity.min$plot,
-                alpha.diversity.min$legend,
-                alpha.diversity.min$xaxis,
-                beta.diversity.min$xaxis,
-                beta.diversity.min$xlab,
-                text_grob(sprintf("A. %s (%s)", key, paste(covariates.min, collapse = ", ")),
-                          size = 16,
-                          just = "left"),
-                alpha.diversity.full$yaxis,
-                alpha.diversity.full$plot,
-                beta.diversity.full$plot,  
-                alpha.diversity.full$xaxis,
-                beta.diversity.full$xaxis,
-                beta.diversity.full$xlab,
-                text_grob(sprintf("B. %s (%s)", key, paste(covariates.full %>%
-                                                           gsub("BL_USE_", "", .), collapse = ", ")),
-                          size = 16, just = "left"))
-
-    g4 <- arrangeGrob(grobs = gs4,
-                      layout_matrix = rbind(
-                          c(8,    NA,    NA, NA,   NA, NA, NA),
-                          c(NA, 1,    2,    NA, 3, NA, NA),
-                          c(NA, 1,    2,    NA, 3, NA, 4),
-                          c(NA, NA, 5,    NA,6, NA, NA),
-                          c(NA, NA, NA, NA,7, NA, NA),
-                          c(15,    NA,    NA, NA,   NA, NA, NA),
-                          c(NA, 9,    10,  NA,  11, NA, NA),
-                          c(NA, 9,    10,  NA,  11, NA, NA),
-                          c(NA, NA, 12,   NA, 13, NA, NA),
-                          c(NA, NA, NA, NA,14, NA, NA)),
-                      widths=c(0.7, 4, 0.7, 0.1, 5, 0.7, 3),
-                      heights=rep(c(1, 1, 6, 1, 2), 2))
-
-    ggsave(file = paste0("cache/alpha-beta-", key , ".png"), plot=g4, height=6, width=9)
-
-    plotlist[[key]] = g4
-}
-
-g <- arrangeGrob(grobs = plotlist, ncol=2)
-ggsave(file="cache/plotlist.png", g, height=8*length(plotlist)/1.5, width=10*2)
-```
-
-<img src="cache/plotlist.png"/>
-
-# Alpha scatter
-
-```{r draw alpha distribution scatter plot}
-scatterplot.df <- lapply(ret, function(x) x$alphadistribution) %>%
-    Reduce(function(x, y) merge(x, y, all=TRUE, by="Sample_ID"), .) %>%
-    select(-Sample_ID) %>%
-    na.omit()
-
-galphascatter <- ggpairs(scatterplot.df,
-                         lower = list(continuous = wrap("points",
-                                                        alpha = 0.3,
-                                                        size=0.1), 
-                                      combo = wrap("dot",
-                                                   alpha = 0.4,
-                                                   size=0.2))) +
-    theme(legend.position = "none", 
-          panel.grid.major = element_blank(), 
-          axis.ticks = element_blank(), 
-          panel.border = element_rect(linetype = "dashed", colour = "black", fill = NA))
-
-ggsave(file = paste0("cache/ggpairs-alphas.png"),
-       plot= galphascatter,
-       height=6,
-       width=9,
-       unit = "in",
-       dpi = 300)
-```
-
-<img src="cache/ggpairs-alphas.png"/>
-
-# Beta diversity
-
-Beta diversity R2
-
-```{r beta diversity r2}
-betadf <- data.frame()
-for (key in names(rds)) {
-    moddf.full <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$full) %>%
-        dplyr::filter(response == term) %>%
-        mutate(model = key, covariates = "full") %>%
-        select(term, model, covariates, R2)
-    moddf.min <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$min) %>%
-        dplyr::filter(response == term) %>%
-        mutate(model = key, covariates = "min") %>%
-        select(term, model, covariates, R2)
-    betadf <- rbind(betadf, moddf.full, moddf.min)
-}
-
-betadf %>% arrange(term) %>% spread(model, R2) %>% kable
-```
-
-Beta diversity p.values
-
-```{r beta diversity pvalue}
-betadf <- data.frame()
-for (key in names(rds)) {
-    moddf.full <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$full) %>%
-        dplyr::filter(response == term) %>%
-        mutate(model = key, covariates = "full") %>%
-        select(term, model, covariates, `Pr(>F)`)
-    moddf.min <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$min) %>%
-        dplyr::filter(response == term) %>%
-        mutate(model = key, covariates = "min") %>%
-        select(term, model, covariates, `Pr(>F)`)
-    betadf <- rbind(betadf, moddf.full, moddf.min)
-}
-betadf %>% arrange(term) %>% spread(model, `Pr(>F)`) %>% kable
-```

### 2019-10-04 comparepseqs.Rmd
Typo

@@
 for (key in names(rds)) {
-    pseq <- import_filter_data(rds[[key]],
+    pseq.species <- import_filter_data(rds[[key]],
                                included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
@@
 
-    bray.dist.m.species <- calculate.beta.matrix(pseq)
+    bray.dist.m.species <- calculate.beta.matrix(pseq.species)
 

### 2019-10-04 comparepseqs.Rmd
MC

@@
 ```{R models}
-ret <- list()
-
-for (key in names(rds)) {
+mclapply(c2l(names(rds)), function(key) {
     pseq.species <- import_filter_data(rds[[key]],
@@
     ggsave(file = sprintf("cache/comparepseqs-%s.png", key), plot=g4, height=6, width=9)
-}
+}, mc.cores = length(rds))
 ```

### 2019-10-04 comparepseqs.Rmd
Unnceccessary code removed

@@
 
-```{r subset runs}
-if (exists("subset.run")) {
-    message(sprintf("Subset %s of %s", subset.run, rds[[subset.run]]))
-    rds <- rds[subset.run]
-}
-```
-
-```{r alpha scatter, include = FALSE, eval = FALSE}
-rds = rds[c(1,2,4)]
-ret <- list()
-
-for (key in names(rds)) {
-    missing = list()
-
-    if (!file.exists(paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))) {
-        missing$alphadistribution = TRUE
-    } else {
-        alphadistribution <- readRDS(file =
-                                         paste0("rds-comparepseqs/alphadistribution-",
-                                                key , ".rds"))
-    }
-
-    ret[[key]] = list("alphadistribution" = alphadistribution)
-}
-```
-
 # Calculate alpha and beta diversity

### 2019-10-07 articletwo.Rmd
Adjusted

@@
 
-    dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq, ~HYPERTENSION)
-
     dds <- DESeq(dds.data, test="Wald", fitType="parametric", parallel = TRUE, BPPARAM=MulticoreParam(16))
@@
     dplyr::filter(qval < 0.05) %>%
-    ggplot(aes(x=Feature, y=log2FoldChange, color=qval)) +
+    dplyr::mutate(group = dplyr::ntile(row_number(), 3)) %>%
+    ggplot(aes(x=Feature, y=log2FoldChange, color=qval, group = group)) +
+    facet_wrap(~group, scales = "free_y") +
     geom_point(size=1) +
@@
 
-ggsave(file = "cache/deseq.png", plot=g.deseq, height=10, width=5)
+ggsave(file = "cache/deseq.png", plot=g.deseq, height=10, width=15)
 ```

### 2019-10-09 articletwo.Rmd
Decrease core

--- 2019-09-30 articletwo.Rmd
@@
 pseq.genus.core <- microbiome::transform(pseq.genus, "compositional") %>%
-    core(detection = 0.0001, prevalence = 0.01)
+    core(detection = 0.001, prevalence = 0.01)
 ```
@@
 
-```{r deseq results table}
-results(dds, contrast = c("HYPERTENSION", 0, 1)) %>%
+```{r deseq results}
+dset.deseq.signf <- results(dds, contrast = c("HYPERTENSION", 0, 1)) %>%
     as.data.frame %>%
@@
     dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
-    dplyr::filter(qval < 0.05) %>%
-    kable
+    dplyr::filter(qval < 0.05)
+
+dset.deseq.boxplot <- pseq.genus %>%
+    microbiome::transform(., transform = "compositional") %>%
+    prune_taxa(pull(dset.deseq.signf, Feature), .) %>%
+    psmelt %>%
+    mutate(Abundance = sqrt(Abundance))
+```
+
+```{r deseq results numerical}
+dset.deseq.signf %>% kable
 ```
@@
 ```{r deseq plot}
-g.deseq <- results(dds, contrast = c("HYPERTENSION", 0, 1)) %>%
-    as.data.frame %>%
-    tibble::rownames_to_column("Feature") %>%
-    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
-    dplyr::filter(qval < 0.05) %>%
-    ggplot(aes(x=Feature, y=log2FoldChange, color=qval)) +
+g.deseq <- dset.deseq.signf %>%
+    ggplot(aes(x=gsub("g_", "", Feature), y=log2FoldChange, color=qval)) +
     geom_point(size=1) +
@@
     coord_flip() +
+    xlab("Feature") +
     theme_classic()
-
-ggsave(file = "cache/deseq.png", plot=g.deseq, height=10, width=5)
+ggsave(file = "cache/deseq.png", plot=g.deseq, height=8, width=5)
 ```
@@
 <img src = "cache/deseq.png" />
+
+```{r deseq boxplot}
+g.boxplot <- dset.deseq.boxplot %>% 
+    ggplot(aes(x=HYPERTENSION, y=Abundance, color = OTU)) +
+    geom_boxplot(lwd=0.1, outlier.size = 0.1) +
+    scale_y_continuous(limits = c(0, 0.5)) +
+    scale_x_discrete(labels=c("0" = "Normotensive", "1" = "Hypertensive")) +
+    ylab("Sqrt For Relative Abundances") +
+    theme_classic() +
+    guides(shape = guide_legend(override.aes = list(size = 0.1)),
+           color = guide_legend(override.aes = list(size = 0.1))) +
+    theme(legend.title = element_text(size = 4), 
+          legend.text  = element_text(size = 4),
+          legend.key.size = unit(1, "lines"),
+          legend.position="bottom")
+ggsave(file = "cache/dds-test.png", plot=g.boxplot, height=7, width=6, units = "in", dpi = 300)
+```
+
+<img src = "cache/dds-test.png" />
 

### 2019-10-09 articletwo.Rmd
Jitter to boxplot

@@
 g.boxplot <- dset.deseq.boxplot %>%
+    filter(OTU %in% pull(dset.deseq.signf  %>% dplyr::filter(abs(log2FoldChange) > 1), Feature)) %>%
+    group_by(OTU) %>%
+    mutate(outlier = abs(Abundance) > median(Abundance) + IQR(Abundance) * 1.5) %>%
     ggplot(aes(x=HYPERTENSION, y=Abundance, color = OTU)) +
-    geom_boxplot(lwd=0.1, outlier.size = 0.1) +
-    scale_y_continuous(limits = c(0, 0.5)) +
+    geom_boxplot(lwd=0.1, outlier.shape = NA) +
+    geom_point(data = function(x) dplyr::filter_(x, ~ outlier),
+               position = position_jitterdodge(jitter.width = 0.2), size = 0.01, alpha = 0.4) +
+    scale_y_continuous(limits = c(0, 0.1)) +
     scale_x_discrete(labels=c("0" = "Normotensive", "1" = "Hypertensive")) +
@@
           legend.position="bottom")
-ggsave(file = "cache/dds-test.png", plot=g.boxplot, height=7, width=6, units = "in", dpi = 300)
+ggsave(file = "cache/dds-test.png", plot=g.boxplot, height=4, width=6, units = "in", dpi = 300)
 ```

### 2019-10-09 articletwo.Rmd
Minor
Multiple deseqs

--- 2019-10-07 articletwo.Rmd
@@
 pseq.genus.core <- microbiome::transform(pseq.genus, "compositional") %>%
-    core(detection = 0.0001, prevalence = 0.01)
+    core(detection = 0.1/100, prevalence = 1/100)
 ```
@@
     pseq.genus.core.deseq <-  prune_taxa(taxa(pseq.genus.core), pseq.genus)
+
+    dds <- lapply(c2l(var.BP), function(x) {
         dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq,
                                        as.formula(sprintf("~ %s + %s",
-                                                      paste0(var.CL, collapse = "+"),
-                                                      "HYPERTENSION")))
-
-    dds <- DESeq(dds.data, test="Wald", fitType="parametric", parallel = TRUE, BPPARAM=MulticoreParam(16))
+                                                          paste0(var.CL, collapse = "+"), x)))
+        DESeq(dds.data,
+              test="Wald",
+              fitType="parametric",
+                     parallel = TRUE,
+                     BPPARAM=MulticoreParam(16))})
+
     saveRDS(dds, file = "rds/dds.rds")
@@
 
-```{r deseq results table}
-results(dds, contrast = c("HYPERTENSION", 0, 1)) %>%
+```{r deseq results}
+dset.deseq.signf <- results(dds$HYPERTENSION, contrast = c("HYPERTENSION", 0, 1)) %>%
     as.data.frame %>%
@@
 ```{r deseq plot}
-g.deseq <- results(dds, contrast = c("HYPERTENSION", 0, 1)) %>%
-    as.data.frame %>%
-    tibble::rownames_to_column("Feature") %>%
-    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
-    dplyr::filter(qval < 0.05) %>%
-    dplyr::mutate(group = dplyr::ntile(row_number(), 3)) %>%
-    ggplot(aes(x=Feature, y=log2FoldChange, color=qval, group = group)) +
-    facet_wrap(~group, scales = "free_y") +
+g.deseq <- dset.deseq.signf %>%
+    ggplot(aes(x=gsub("g_", "", Feature), y=log2FoldChange, color=qval)) +
     geom_point(size=1) +
@@
     coord_flip() +
+    xlab("Feature") +
     theme_classic()
-
-ggsave(file = "cache/deseq.png", plot=g.deseq, height=10, width=15)
+ggsave(file = "cache/deseq.png", plot=g.deseq, height=4, width=5)
 ```
@@
 <img src = "cache/deseq.png" />
+
+```{r deseq boxplot}
+g.boxplot <- dset.deseq.boxplot %>%
+    group_by(OTU) %>%
+    mutate(outlier = abs(Abundance) > median(Abundance) + IQR(Abundance) * 1.5) %>%
+    ggplot(aes(x = HYPERTENSION, y = Abundance, color = OTU)) +
+    geom_boxplot(lwd=0.1, outlier.shape = NA) +
+    geom_point(data = function(x) dplyr::filter_(x, ~ outlier),
+               position = position_jitterdodge(jitter.width = 0.02),
+               size = 0.1,
+               alpha = 0.4, shape = ".") +
+    scale_y_continuous(limits = c(0, 0.2), expand = c(0, 0)) +
+    scale_x_discrete(labels=c("0" = "Normotensive", "1" = "Hypertensive")) +
+    ylab("Abundance (Hellinger)") +
+    xlab("") +
+    theme_classic() +
+    guides(shape = guide_legend(override.aes = list(size = 0.1)),
+           color = guide_legend(override.aes = list(size = 0.1))) +
+    theme(legend.title = element_text(size = 4), 
+          legend.text  = element_text(size = 4),
+          legend.key.size = unit(1, "lines"),
+          legend.position="bottom")
+ggsave(file = "cache/dds-test.png", plot=g.boxplot, height=4, width=6, units = "in", dpi = 300)
+```
+
+<img src = "cache/dds-test.png" />
 

### 2019-10-11 articletwo.Rmd
Deseq analyses

@@
 
-## DeSeq
-
-```{r deseq results}
-dset.deseq.signf <- results(dds$HYPERTENSION, contrast = c("HYPERTENSION", 0, 1)) %>%
-    as.data.frame %>%
-    tibble::rownames_to_column("Feature") %>%
-    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
-    dplyr::filter(qval < 0.05)
-
-dset.deseq.boxplot <- pseq.genus %>%
-    microbiome::transform(., transform = "compositional") %>%
-    prune_taxa(pull(dset.deseq.signf, Feature), .) %>%
-    psmelt %>%
-    mutate(Abundance = sqrt(Abundance))
-```
-
-```{r deseq results numerical}
-dset.deseq.signf %>% kable
-```
-
-```{r deseq plot}
-g.deseq <- dset.deseq.signf %>%
-    ggplot(aes(x=gsub("g_", "", Feature), y=log2FoldChange, color=qval)) +
-    geom_point(size=1) +
-    geom_hline(yintercept = 0, color = "gray") +
-    theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) +
-    coord_flip() +
-    xlab("Feature") +
-    theme_classic()
-ggsave(file = "cache/deseq.png", plot=g.deseq, height=4, width=5)
-```
-
-<img src = "cache/deseq.png" />
-
-```{r deseq boxplot}
-g.boxplot <- dset.deseq.boxplot %>%
-    group_by(OTU) %>%
-    mutate(outlier = abs(Abundance) > median(Abundance) + IQR(Abundance) * 1.5) %>%
-    ggplot(aes(x = HYPERTENSION, y = Abundance, color = OTU)) +
-    geom_boxplot(lwd=0.1, outlier.shape = NA) +
-    geom_point(data = function(x) dplyr::filter_(x, ~ outlier),
-               position = position_jitterdodge(jitter.width = 0.02),
-               size = 0.1,
-               alpha = 0.4, shape = ".") +
-    scale_y_continuous(limits = c(0, 0.2), expand = c(0, 0)) +
-    scale_x_discrete(labels=c("0" = "Normotensive", "1" = "Hypertensive")) +
-    ylab("Abundance (Hellinger)") +
-    xlab("") +
-    theme_classic() +
-    guides(shape = guide_legend(override.aes = list(size = 0.1)),
-           color = guide_legend(override.aes = list(size = 0.1))) +
-    theme(legend.title = element_text(size = 4), 
-          legend.text  = element_text(size = 4),
-          legend.key.size = unit(1, "lines"),
-          legend.position="bottom")
-ggsave(file = "cache/dds-test.png", plot=g.boxplot, height=4, width=6, units = "in", dpi = 300)
-```
-
-<img src = "cache/dds-test.png" />
-
 ## Characteristics of the study sample. (Table 1.)
@@
 
-For magnitude, age, sex, and BMI explain `r round(summary(lm(MAP ~ BL_AGE +
-SEX + BMI, data = meta(pseq.genus)))$adj.r.squared * 100, 1)`% of MAP.
-
-```{r maaslin clean results}
-maas.results.df.signf <- maaslin %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
-    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  Feature = gsub("[gs]_", "", Feature),
-                  beta = Coefficient) %>%
-    dplyr::filter(Qval < 0.05) %>%
-    arrange(Feature)
-```
-
-```{r maaslin insignif, include = FALSE, eval = FALSE}
- maaslin %>%
-    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  Feature = gsub("[gs]_", "", Feature),
-                  beta = Coefficient) %>%
-     arrange(Feature) %>%
-     select(Variable, Feature, Qval) %>%
-     spread(Variable, Qval)
-```
-
-```{r maaslin plot}
-my.plot.order <- function(x) {
-    y <- seq(length(x))
-    y[x == "Systolic BP"] = 4
-    y[x == "Diastolic BP"] = 3
-    y[x == "Pulse pressure"] = 2
-    y[x == "Mean arterial pressure"] = 1
-    y[x == "Hypertension"] = 0
-    y
-}
-
-maas.all.pairs <- expand.grid(Feature = maas.results.df.signf %>% pull(Feature),
-                         Name = maas.results.df.signf %>% pull(Name))
-
-maas.tables.df <- merge(maas.results.df.signf,
-                        maas.all.pairs,
-                        all.y=TRUE,
-                        by.y=c('Feature', 'Name'),
-                        by.x=c('Feature', 'Name')) %>%
-    mutate(p.groups = cut(P.value*ifelse(Coefficient < 0, -1, 1),
-                          breaks = c(-0.05, -0.01, -0.001, 0, 0.001, 0.01, 0.05),
-                          labels = c("beta < 0, p < 0.05",
-                                     "beta < 0, p < 0.01",
-                                     "beta < 0, p < 0.001",
-                                     "beta > 0, p < 0.001",
-                                     "beta > 0, p < 0.01",
-                                     "beta > 0, p < 0.05")))
-
-mycolors <- c("beta < 0, p < 0.05" = "#B2182B",
-              "beta < 0, p < 0.01" = "#EF8A62",
-              "beta < 0, p < 0.001" = "#FDDBC7",
-              "beta > 0, p < 0.001" = "#D1E5F0",
-              "beta > 0, p < 0.01" = "#67A9CF",
-              "beta > 0, p < 0.05" = "#2166AC",
-              "NA" = "gray50")
-
-maaslinplot <- ggplot(maas.tables.df, aes(x = Feature, y = reorder(Name, my.plot.order(Name)), fill = beta)) +
-    geom_tile(aes(fill = 1)) +
-    geom_tile(colour="white", size=0.25) +
-    scale_fill_distiller(palette = "RdBu", name = "Effect size",
-                         limits=c(-0.003, 0.003), breaks = c(-0.003, 0, 0.003), na.value="grey30") +
-    coord_fixed() +
+```{r deseq continuous}
+dset.deseq.signf <- deseqresults(dds, names.dset)
+```
+
+```{r deseq heatmap}
+g.deseq <- deseqhaetmap(dset.deseq.signf)
+ggsave(file = "cache/deseq.png", plot = g.deseq, height = 3, width = 12, unit = "in", dpi = 300)
+```
+
+<img src = "cache/deseq.png" />
+
+```{r deseq abundances for boxplot}
+dset.deseq.boxplot <- pseq.genus %>%
+    microbiome::transform(., transform = "compositional") %>%
+    prune_taxa(unique(pull(dset.deseq.signf, Feature)), .) %>%
+    psmelt %>%
+    mutate(Abundance = sqrt(Abundance))
+```
+
+```{r deseq boxplot}
+g.boxplot <- dset.deseq.boxplot %>%
+    group_by(OTU) %>%
+    mutate(outlier = abs(Abundance) > median(Abundance) + IQR(Abundance) * 1.5) %>%
+    ggplot(aes(x = HYPERTENSION, y = Abundance, color = OTU)) +
+    geom_boxplot(lwd=0.1, outlier.shape = NA) +
+    geom_point(data = function(x) dplyr::filter_(x, ~ outlier),
+               position = position_jitterdodge(jitter.width = 0.02),
+               size = 0.1,
+               alpha = 0.4, shape = ".") +
+    scale_y_continuous(limits = c(0, 0.2), expand = c(0, 0)) +
+    scale_x_discrete(labels=c("0" = "Normotensive", "1" = "Hypertensive")) +
+    ylab("Abundance (Hellinger)") +
     xlab("") +
-    ylab("") +
-    theme_classic(10) +  
-    theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust=0.5, size=10),
-          axis.text.y = element_text(size=10))
-
-ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 4, unit = "in", dpi = 300)
-```
-
-Max value of abs beta is `r maas.results.df.signf %>% pull(Coefficient) %>% abs %>% max`.
-
-```{r maaslin table}
-maas.results.df.signf %>%
-    select(Feature, Value, Coefficient, Qval) %>%
-    arrange(Qval) %>%
-    kable
-```
-
-```{r multi model, eval = FALSE, include = FALSE}
-abund.genus.systm <- pseq.genus %>%
-    abundances(., transform = "compositional") %>%
-    t %>%
-    as.data.frame %>%
-    tibble::rownames_to_column("id")
-
-df.genus.systm <- meta(pseq.genus) %>%
-    tibble::rownames_to_column("id") %>%
-    full_join(., abund.genus.systm, by = "id") %>%
-    mutate(Anaerotruncus = scale(asin(sqrt(g_Anaerotruncus))),
-           Sutterella = scale(asin(sqrt(g_Sutterella))))
-
-fo.genus.systm <- sprintf("oSYSTM ~ Anaerotruncus + Sutterella + Anaerotruncus:Sutterella + %s",
-                          paste0(var.CL, collapse = " + "))
-
-lm(as.formula(fo.genus.systm), data = df.genus.systm) %>% summary
-
-df.genus.systm %>%
-    mutate(diff = -0.75418 * Anaerotruncus + 0.89265 * Sutterella) %>%
-    pull(diff) %>%
-    summary
-```
-
-```{r scatter plots, eval = FALSE, include = FALSE}
-scatter.signf.bacteria <- maas.results.df.signf %>% pull(Feature) %>% unique
-
-scatter.meta <- meta(pseq.genus) %>%
-    tibble::rownames_to_column("id") %>%
-      mutate(MAP = oMAP,
-           PULSEPRESSURE = oPULSEPRESSURE,
-           SYSTM = oSYSTM,
-           DIASM = oDIASM)
-
-scatter.abund <- abundances(pseq.genus, transform = "clr") %>%
-    t %>%
-    as.data.frame %>%
-    tibble::rownames_to_column("id") %>%
-    select(id, scatter.signf.bacteria) %>%
-    mutate_at(.vars = vars(scatter.signf.bacteria), .funs = funs(myscale(.)))
-
-scatter.full <- full_join(scatter.meta, scatter.abund, by = "id")
-
-scatter.combinations <- expand.grid(cc("SYSTM", "DIASM", "PULSEPRESSURE", "MAP"),
-                                    c2l(scatter.signf.bacteria)) %>%
-    setNames(c("rr", "abund"))
-
-scatter.plot <- function(df,
-                         maas.results,
-                         rr,
-                         abund,
-                         covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
-                                        "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08",
-                                        "BL_USE_RX_C09"),
-                         prediction = TRUE,
-                         plots = TRUE) {
-    fo <- sprintf("%s ~ %s + %s", rr, abund, paste0(covariates, collapse="+"))
-    model <- lm(as.formula(fo), data = df)
-
-    if (prediction == FALSE) return(model %>% summary)
-
-    df.predict <- data.frame(BPV = seq(-5,5,0.1)) %>%
-        mutate(SEX = as.factor(0),
-               BL_AGE = df %>% pull(BL_AGE) %>% mean,
-               BMI = df %>% pull(BMI) %>% mean,
-               CURR_SMOKE = as.factor(0),
-               Q57X = as.factor(1),
-               PREVAL_DIAB = as.factor(0),
-               BL_USE_RX_C03 = as.factor(0),
-               BL_USE_RX_C07 = as.factor(0),
-               BL_USE_RX_C08 = as.factor(0),
-               BL_USE_RX_C09 = as.factor(0),
-               !!abund := seq(-5,5,0.1)
-               )
-    prediction <- cbind(df.predict, predict(model, df.predict, interval = 'confidence'))
-
-    if (plots == FALSE) return(prediction)
-
-    feature <- replace.brackets(abund)
-
-    qval.maaslin <- maas.results %>%
-        dplyr::filter(Feature == feature, Value == rr) %>%
-        nrow
-
-    if (qval.maaslin == 0) { return (ggplot() + ggplot2::theme_minimal()) }
-    pval <- pub.p(summary(model)$coef[2,4])
-    rsqr <- signif(summary(model)$adj.r.squared, 2)
-    slope <- signif(model$coef[[2]], 2)
-
-    ggplot(prediction, aes_string(x = abund, y = "fit")) +
-        geom_line() +
-        geom_point(size=0.1, shape=1, aes_string(x = abund, y = rr, color = "SEX"),
-                   data = df) +
-        labs(title = sprintf("Adj R2 = %s, Slope = %s, p = %s", rsqr, slope, pval)) +
-        xlab(feature) +
-        ylab(rr) +
-        ylim(0,200) +
-        ggplot2::theme_minimal() +
-        theme(plot.title = element_text(size=8)) +
-        theme(legend.position="none")
-}
-
-scatterplots <- lapply(data.frame(t(scatter.combinations)),
-                       function(x)
-                           scatter.plot(scatter.full, maas.results.df.signf, x$rr, x$abund))
-
-scatterplot <- do.call("grid.arrange", c(scatterplots, ncol=4))
-ggsave(file = "cache/scatter.png", plot = scatterplot, height = 10, width = 10, dpi = 300)
-```
-
-```{r maaslin means, eval = FALSE}
-pseq.genus.core.table <- pseq.genus.core %>%
-    prune_taxa(c("g_Adlercreutzia", "g_Anaerotruncus", "g_Sutterella"), .) %>%
-    transform(transform = "compositional") %>%
-    psmelt %>%
-    mutate(map_groups = relevel(factor(dplyr::ntile(MAP, 3)), ref='1'),
-           systm_groups = relevel(factor(dplyr::ntile(SYSTM, 3)), ref='1'),
-           Abundance = myscale(Abundance))
-
-table_bp_abund <- function(pseq, group) {
-    pseq %>%
-        group_by(.dots = c(group, "OTU")) %>%
-        summarize(mean = mean(Abundance), sd = sd(Abundance)) %>%
-        mutate(beta_sd = sprintf("%.2f ± %.2f", mean, sd)) %>%
-        select(OTU, beta_sd) %>%
-        spread(OTU, beta_sd) %>%
-        ungroup %>%
-        dplyr::rename("grp" = group) %>%
-        mutate("grp" = paste0(group, "_", grp))
-}
-
-rbind(table_bp_abund(pseq.genus.core.table, "map_groups"),
-             table_bp_abund(pseq.genus.core.table, "systm_groups")) %>%
-    tibble::column_to_rownames("grp") %>%
-    t %>%
-    as.data.frame %>%
-    kable
-
-```
+    theme_classic() +
+    guides(shape = guide_legend(override.aes = list(size = 0.1)),
+           color = guide_legend(override.aes = list(size = 0.1))) +
+    theme(legend.title = element_text(size = 4), 
+          legend.text  = element_text(size = 4),
+          legend.key.size = unit(1, "lines"),
+          legend.position="bottom")
+
+ggsave(file = "cache/dds-test.png", plot=g.boxplot, height=4, width=6, units = "in", dpi = 300)
+```
+
+<img src = "cache/dds-test.png" />
 
@@
 Dark gray color represents insignificant associations."
-writeimage(doc, 3, "cache/maaslin.png", fig3head, fig3foot)
+writeimage(doc, 3, "cache/deseq.png", fig3head, fig3foot)
 ```

### 2019-10-11 articletwo-ggplot.R

@@ -173,3 +173,36 @@ plot.diversities <- function(diversities) {
                       heights=rep(c(1.5, 1, 6, 1, 2), 2))
     return(g4)
 }
+
+deseqhaetmap <- function(df, legend.name = "Log Base Two\nFold Change") {
+    ggplot(df,
+           aes(x = Feature,
+               y = reorder(Name, deseqplotorder(Name)),
+               fill = log2FoldChange)) +
+        geom_tile(aes(fill = 1)) +
+        geom_tile(colour="white", size=0.25) +
+        scale_fill_distiller(palette = "RdBu",
+                             name = legend.name,
+                             limits=c(-1, 1),
+                             breaks = c(-1, -0.5, 0.5, 0, 1),
+                             na.value="grey30") +
+        coord_fixed() +
+        xlab("") +
+        ylab("") +
+        theme_classic() +  
+        theme(axis.text.x = element_text(angle=90, size=10, hjust=1.0,vjust=0.5),
+              axis.text.y = element_text(size=10),
+              legend.title=element_text(size=10),
+              legend.position = c(-0.09, -1))
+}
+
+
+deseqplotorder <- function(x) {
+    y <- seq(length(x))
+    y[x == "Systolic BP"] = 4
+    y[x == "Diastolic BP"] = 3
+    y[x == "Pulse pressure"] = 2
+    y[x == "Mean arterial pressure"] = 1
+    y[x == "Hypertension"] = 0
+    y
+}

### 2019-10-11 articletwo-rrbiome.R

@@ -342,3 +342,17 @@ calculate.betadiversity <- function(pseq, matrix, vars, npermutations = 999) {
                         covariates = var,
                         npermutations = npermutations))
 }
+
+deseqresults <- function(modellist, names.dset) {
+    vars <- c2l(names(modellist))
+    lapply(c2l(vars), function(x, models = modellist) {
+        name <- ifelse(x %in% c("HYPERTENSION"), sprintf("%s_1_vs_0", x), x)
+        results(models[[x]], name = name) %>%
+            as.data.frame %>%
+            tibble::rownames_to_column("Feature") }) %>%    
+        map_df(., ~as.data.frame(.x), .id="name") %>%
+        dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
+                      Feature = gsub("g_", "", Feature)) %>%
+        dplyr::filter(qval < 0.05) %>%
+        merge(names.dset %>% select(Covariate, Name), by.x ="name", by.y = "Covariate")
+}

### 2019-10-14 articletwo.Rmd
Salt plot

@@
 library(DESeq2)
+library(ggeffects)
 ```
@@
 
-    dds <- lapply(c2l(var.BP), function(x) {
-        dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq,
-                                       as.formula(sprintf("~ %s + %s",
-                                                          paste0(var.CL, collapse = "+"), x)))
+    lapply(c2l(var.BP), function(x) {
+        dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq, deseq.formula(x, var.CL))
         DESeq(dds.data,
@@
 dset.deseq.signf <- deseqresults(dds, names.dset)
+dset.deseq.signf %>% kable
 ```
@@
 
-```{r lacto data frame}
-df.genus.salt <- pseq.genus %>%
-    microbiome::transform(transform = "clr") %>%
-    prune_taxa(c("g_Lactobacillus"), .) %>%
-    psmelt %>%
-    dplyr::filter(!is.na(NA.)) %>%
-    spread(OTU, Abundance) %>%
-    dplyr::mutate(dUNA = myscale(NA.),
-                  salt_groups = relevel(factor(dplyr::ntile(dUNA, 3)), ref='1'),
-                  lacto_groups = relevel(factor(dplyr::ntile(g_Lactobacillus, 3)), ref='1'))
+```{r salt phyloseq object}
+rownames.salt <- pseq.genus %>% meta %>% subset(!is.na(NA.)) %>% rownames
+pseq.genus.salt <- pseq.genus %>%
+    prune_samples(rownames.salt, .) %>%
+    prune_taxa(taxa(pseq.genus.core), .)
+dset.genus.salt <- prune_lactobacillus(pseq.genus.salt, transform = "clr")
 ```
 
-Lactobacillus prevalence
-
-```{r lactobacillus prevalence}
-pseq.genus %>%
-    microbiome::transform(transform = "compositional") %>%
-    prune_taxa(c("g_Lactobacillus"), .) %>%
-    psmelt %>%
-    spread(OTU, Abundance) %>%
-    dplyr::mutate(lacto.prevalent = ifelse(g_Lactobacillus > 0.001, TRUE, FALSE),
-                  duna.present = ifelse(is.na(NA.), 0, 1)) %>%
-    group_by(duna.present) %>%
-    summarize(lacto_detected = sum(lacto.prevalent), n = n(), ratio = lacto_detected/n) %>%
+```{r salt characteristics}
+dset.genus.salt %>%
+    mytableone() %>%
     kable
@@
 
-Characteristics in salt analysis. Analyses are limited by the number
-of participants with collected dU-NA.
-
-```{r salt characteristics}
-df.genus.salt %>%
-    select(-Sample) %>% 
-    gather(key,value) %>%
-    group_by(key) %>%
-    summarise_all(funs(N = n(),
-                       Mean = mean(as.numeric(value), na.rm = TRUE),
-                       ones = sum(value == 1, na.rm = TRUE),
-                       sd = sd(as.numeric(value), na.rm = TRUE),
-                       NAs = sum(is.na(value))))  %>%
-    mutate_if(is.numeric, round, 3) %>%
-    kable()
-```
-
-Required sub sample size
-
-```{r required sodium subsample size}
-lapply(c2l(seq(1, 1.7, 0.05)), function(or) 
-    powerMediation::powerLogisticCon(n = nrow(df.genus.salt),
-                                     p1 = 0.4,
-                                     OR = or)) %>%
-    map_df(., ~as.data.frame(.x), .id = "or") %>%
+```{r lactobacillus prevalence}
+dset.genus.salt %>%
+    summarize(lacto_detected = sum(lacto.prevalent),
+              n = n(),
+              ratio = lacto_detected/n) %>%
     kable
@@
 
-### How sodium affects lactobacillus abundances?
-
-Associations between continuous dU-NA and lactobacillus abundances.
-
-```{r abund salt lm continuous}
-lm.lacto.duna <- lm(g_Lactobacillus ~ dUNA + BL_AGE + SEX, data = df.genus.salt)
-lm.lacto.duna %>% tidy %>% mutate_if(is.numeric, round, 3) %>% kable
-```
-
-Testing assumptions for linear regression analysis
-
-```{r abund salt lm continuous checks table}
-gvlma.table.plot(lm.lacto.duna)$table %>% kable
-```
-
-Diagnostic plots for linear model
-
-```{r abund salt lm continuous checks plot, fig.width=12, fig.height=14}
-gvlma.table.plot(lm.lacto.duna)$plot %>% plot
-```
-
-### Is there associations between salt groups and abudances?
-
-Associations between categorical dU-NA split in three equal size groups and lactobacillus abundances
-
-```{r abund salt lm group}
-lm.lacto.saltgroup <- lm(g_Lactobacillus ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
-lm.lacto.saltgroup %>% tidy %>% mutate_if(is.numeric, round, 3) %>% kable
-```
-
-Testing assumptions for linear regression analysis
-
-```{r abund salt lm group checks table}
-gvlma.table.plot(lm.lacto.saltgroup)$table %>% kable
-```
-
-Diagnostic plots for linear model
-
-```{r abund salt lm group checks plot, fig.width=12, fig.height=14}
-gvlma.table.plot(lm.lacto.saltgroup)$plot %>% plot
-```
-
-Analysing also differences between salt groups using ANOVA.
-
-```{r salt anova}
-aov.lacto.saltgroup <- aov(g_Lactobacillus ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
-aov.lacto.saltgroup %>% xtable %>% kable
-```
-
-Levene's test for equality of variance
-
-```{r salt levene}
-car::leveneTest(g_Lactobacillus ~ salt_groups, data = df.genus.salt) %>% kable
-```
-
-Tukey's range for pairwise test between salt groups
-
-```{r salt HSD}
-hsd.lacto.saltgroup <- TukeyHSD(aov.lacto.saltgroup, "salt_groups")
-hsd.lacto.saltgroup$salt_groups %>% kable
-```
-
-### Adjusting means for forest plot
-
-Calculate adjusted means for lactobacillus assuming male sex and mean age.
-
-```{r lacto adjusted means}
-adjmean.lacto.saltgroup <- emmeans::emmeans(lm.lacto.saltgroup, specs = c("salt_groups")) %>%
-    as.data.frame %>%
-    dplyr::rename(term = salt_groups, estimate = emmean, conf.low = lower.CL, conf.high = upper.CL)
-```
-
-Adjusted means
-
-```{r forest lims}
-adjmean.lacto.saltgroup %>% kable
-```
-
-## Is there associations between lactobacillus abudance and hypertension?
-
-Studying associations between blood pressure variables and lactobacillus abundances
-
-```{r lactobacillus RR}
-calculateglm(df.genus.salt,
-             modelstring = "%s ~ g_Lactobacillus + BL_AGE + SEX",
-             filterstr = "Lactobacillus") %>% 
-    select(response, term, estimate, std.error, p.value) %>%
-    mutate_if(is.numeric, round, 3) %>%
-    kable
-```
-
-Checking assumptions for linear regression for MAP
-
-```{r table map lactobacillus}
-gvlma.table.plot(lm(MAP ~ g_Lactobacillus + BL_AGE + SEX, data = df.genus.salt))$table %>% kable
-```
-
-Diagnostic plots for linear model
-
-```{r plot map lactobacillus, fig.width=12, fig.height=14}
-gvlma.table.plot(lm(MAP ~ g_Lactobacillus + BL_AGE + SEX, data = df.genus.salt))$plot %>% plot
-```
-
-Odds for hypertension by categorical lactobacillus abundance.
-
-```{r glm htn lacto groups}
-glm.hypertension.lactogroup <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
-                  family = binomial(link=logit),
-                  data = df.genus.salt)
-
-odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
-    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
-    filter(!grepl("Intercept", term)) %>%
-    select(term, estimate, conf.low, conf.high) %>%
-    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
-    dplyr::filter(grepl("lacto", term))
-
-glm.hypertension.lactogroup %>%
-    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
-    mutate_if(is.numeric, round, 3) %>%
-    kable
-```
-
-Odds for hypertension
-
-```{r odds htn lacto}
-odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
-    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
-    filter(!grepl("Intercept", term)) %>%
-    select(term, estimate, p.value, conf.low, conf.high) %>%
-    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
-    dplyr::filter(grepl("lacto", term))
-odds.hypertension.lactogroup %>% kable
-```
-
-Power analysis of odds of hypertension
-
-```{r power hypertention}
-power.htn <- pwr.f2.test(u = 3, v=NULL, f2 = 0.02, sig.level=0.05, power=0.8)
-round(power.htn$v + power.htn$u + 1)
-
-pwr.f2.test(u = 3, v=836, f2 = NULL, sig.level=0.05, power=0.95)
-```
-
-## Twin-forest plot
-
-```{r forest plot lacto}
-gsalt <- grid.arrange(
-    myforestplot(adjmean.lacto.saltgroup,
-                 xlab = "Tertile of 24-hour urinary sodium",
-                 ylab = "Lactobacillus abundance",
-                 subtitle = "A.",
-                 my_y_scale = scale_y_continuous(limits = c(5.4, 5.65), breaks = seq(0, 10, 0.1))),
-    myforestplot(odds.hypertension.lactogroup,
-                 xlab = "Tertile of Lactobacillus abundance",
-                 ylab = "Odds of hypertension",
-                 subtitle = "B.",
-                 my_y_scale = scale_y_log10(limits = c(0.4, 1.5), breaks = seq(0.5, 2.0, 0.5)),
-                 interceptone = TRUE),
-    ncol=2)
-
-ggsave(file = "cache/salt-forest.png", plot=gsalt, height=3, width=5.5, dpi = 300)
+```{r lacto scatter plot}
+g.salt <- saltplot(dset.genus.salt)
+ggsave(file = "cache/gsalt.png", plot=g.salt, height=4, width=4, dpi = 300)
 ```
@@
 A 95 % CI is drawn in both panels."
-writeimage(doc, 4, "cache/salt-forest.png", fig4head, fig4foot)
+writeimage(doc, 4, "cache/gsalt.png", fig4head, fig4foot)
 ```
 
-Testing salt model using maaslin
+Deseq for lactobacillus
 
-```{r salt model maaslin}
-taxa(pseq.genus.core) %>% mygrep(word = "lacto")
+```{r lacto deseq}
+dds.data.salt <- phyloseq_to_deseq2(pseq.genus.salt, ~ BL_AGE + SEX + NA.)
 
-pseq.genus.salt <- prune_taxa(c("g_Lactobacillus"), pseq.genus.core)
+dds.salt <- DESeq(dds.data.salt, test="Wald", fitType="parametric")
 
-pseq.genus.salt.meta <- meta(pseq.genus.salt) %>%
-    tibble::rownames_to_column(var = "rowname") %>%
-    dplyr::filter(!is.na(NA.)) %>%
-    dplyr::mutate(dUNA = myscale(NA.),
-                  salt_groups = relevel(factor(dplyr::ntile(dUNA, 3)), ref='1')) %>%
-    tibble::remove_rownames() %>%
-    tibble::column_to_rownames(var = "rowname") %>%
-    select(var.BP, var.CL.min, starts_with("salt_groups"))
-
-phyloseq::sample_data(pseq.genus.salt) <- phyloseq::sample_data(pseq.genus.salt.meta)
-
-maaslin.salt <- maaslinwrapper(pseq = pseq.genus.salt,
-                               looped = var.BP,
-                               forced = c(var.CL.min, "salt_groups"),
-                               taxa = "g_Lactobacillus")
-
-maaslin.salt %>%
-    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
-    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
-                  Feature = gsub("[gs]_", "", Feature),
-                  beta = Coefficient) %>%
-    arrange(Feature)
+results(dds.salt, name = "NA.") %>%
+    as.data.frame %>%
+    tibble::rownames_to_column("Feature") %>%
+    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
+    filter(Feature %in% "g_Lactobacillus") %>%
+    kable
 ```
 
-```{r squared clinical, eval = FALSE, include = FALSE}
-mylmrsqr <- function(df, response, loopcovariates) {
-    model.list <- list()
-    for (i in seq(length(loopcovariates))) {
-        fo <- sprintf("%s ~ %s", response, paste(loopcovariates, collapse = " + "))
-        message(sprintf("formula : %s", fo))
-        model <- lm(as.formula(fo), data = df)
-        model$call <- as.formula(fo)
-        model.list[[i]] <- model
-        loopcovariates <- loopcovariates[-length(loopcovariates)]
-    }
-    return(model.list)
-}
-
-af.models <- mylmrsqr(df = meta(pseq.species),
-         response = "MAP",
-         loopcovariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
-                            "ANYDRUG"))
-
-af <- do.call(anova, af.models)
-
-af %>%
-    as.data.frame %>%
-        mutate(Rsqr =  `Sum of Sq`/sum(`Sum of Sq`, na.rm = T)*100)
-
-c(summary(lm(SYSTM ~ BL_AGE + SEX, meta(pseq.species)))$adj.r.squared,
-summary(lm(DIASM ~ BL_AGE + SEX, meta(pseq.species)))$adj.r.squared,
-summary(lm(PULSEPRESSURE ~ BL_AGE + SEX, meta(pseq.species)))$adj.r.squared,
-summary(lm(MAP ~ BL_AGE + SEX, meta(pseq.species)))$adj.r.squared) %>% mean
-
-c(summary(lm(SYSTM ~ BL_AGE + SEX + BMI, meta(pseq.species)))$adj.r.squared,
-summary(lm(DIASM ~ BL_AGE + SEX + BMI, meta(pseq.species)))$adj.r.squared,
-summary(lm(PULSEPRESSURE ~ BL_AGE + SEX + BMI, meta(pseq.species)))$adj.r.squared,
-summary(lm(MAP ~ BL_AGE + SEX + BMI, meta(pseq.species)))$adj.r.squared) %>% sort
-
-```
+# Supplements
 

### 2019-10-14 articletwo-ggplot.R

@@ -206,3 +206,23 @@ deseqplotorder <- function(x) {
     y[x == "Hypertension"] = 0
     y
 }
+
+saltplot <- function(dset) {
+    ggplot(data = dset, aes(x = NA., y = g_Lactobacillus)) +
+        geom_point(aes(color = SEX), size = 0.05, position = "jitter") +
+        geom_line(data = lm_ribbon(dset)) +
+        geom_ribbon(data = lm_ribbon(dset), aes(ymin=conf.low, ymax=conf.high), alpha=0.15) +
+        scale_x_continuous(limits=c(20, 205), expand = c(0, 0)) +
+        scale_y_continuous(limits=c(4.5, 8), expand = c(0, 0)) +
+        scale_colour_manual(name = "SEX",
+                            labels = c("Male", "Female"),
+                            breaks=c("0", "1"),
+                            values = c("blue", "red")) +
+        xlab("24-hour urinary sodium") +
+        ylab("CLR-transformed Lactobacillus abundance") +
+        guides(colour = guide_legend(override.aes = list(size=4, alpha = 0.4))) +
+        theme_classic() +
+        theme(plot.title = element_blank(),
+              legend.title = element_blank(),
+              legend.position = c(0.9,0.95))
+}

### 2019-10-14 articletwo-rrbiome.R

@@ -356,3 +356,39 @@ deseqresults <- function(modellist, names.dset) {
         dplyr::filter(qval < 0.05) %>%
         merge(names.dset %>% select(Covariate, Name), by.x ="name", by.y = "Covariate")
 }
+
+prune_lactobacillus <- function(pseq, transform = "compositional", drop = TRUE) {
+    microbiome::transform(pseq, transform = transform) %>%
+        prune_taxa(c("g_Lactobacillus"), .) %>%
+        psmelt %>%
+        spread(OTU, Abundance) %>%
+        dplyr::mutate(lacto.prevalent = ifelse(g_Lactobacillus > 0.001, TRUE, FALSE),
+                      duna.present = ifelse(is.na(NA.), 0, 1)) %>%
+        { if (drop == TRUE) dplyr::filter(., duna.present == 1) else . }
+}
+
+lm_ribbon <- function(dset, model = "g_Lactobacillus ~ NA. + BL_AGE + SEX", term = "NA.", taxon = "g_Lactobacillus") {
+    ret <- lm(as.formula(model), data=dset)
+    ggpredict(ret, terms = term) %>%
+        dplyr::rename(!!term := x, !!taxon := predicted)
+}
+
+mytableone <- function(dset, variables) {
+    { if (!missing(variables)) dplyr::select(dset, variables) else dset } %>%
+        select(-Sample) %>% 
+        gather(key,value) %>%
+        group_by(key) %>%
+        summarise_all(funs(N = n(),
+                           Mean = mean(as.numeric(value), na.rm = TRUE),
+                           ones = sum(value == 1, na.rm = TRUE),
+                           sd = sd(as.numeric(value), na.rm = TRUE),
+                           NAs = sum(is.na(value))))  %>%
+        mutate_if(is.numeric, round, 3)
+}
+
+deseq.formula <- function(..., fo = "~ %s") {
+    vars <- c(...)
+    if ("HYPERTENSION" %in% vars)
+        vars <- vars[!grepl("BL_USE", vars)]
+    as.formula(sprintf(fo, paste0(vars, collapse = "+")))
+}

### 2019-10-16 articletwo.Rmd
remove medication from hypertension models

@@
 pander(sessionInfo(), compact = TRUE)
-```
-
-```{r Initialize docx, include = FALSE, cache = FALSE}
-doc <- read_docx(path = "style/articlestyle.docx") %>%
-  body_remove()
-dir.create("rds", showWarnings = FALSE)
-dir.create("cache", showWarnings = FALSE)
 ```
@@
 
-```{r quit subruns, include = FALSE}
-if (exists("subset.run")) {
-    quit()
-}
-```
-
 ```{r pcoa calculate}
@@
     pcoa.ordinate <- readRDS("rds/pcoa.ordinate.rds")
-}
-```
-
-```{r maaslin runs, results = 'hide'} 
-if (!file.exists("rds/maaslin.rds")) {
-     maaslin <- maaslinwrapper(pseq = pseq.genus.core,
-                                    looped = var.BP,
-                                    forced = var.CL,
-                              taxa = taxa(pseq.genus.core))
-
-    saveRDS(maaslin, file = "rds/maaslin.rds")
-} else {
-    maaslin <- readRDS(file = "rds/maaslin.rds")
 }
@@
 }
+```
+
+## Officer
+
+```{r Initialize docx, include = FALSE, cache = FALSE}
+doc <- read_docx(path = "style/articlestyle.docx") %>%
+  body_remove()
+dir.create("rds", showWarnings = FALSE)
+dir.create("cache", showWarnings = FALSE)
 ```
@@
 
-Significant associations between BP-variables and top PCoA axes
-
-```{r pcoa analysis main axis}
-glm.pcoa <- lapply(c2l(paste0("Axis.", seq(1,3))), function(axis)
-    calculateglm(pcoa.df,
-                 modelstring = sprintf("%%s ~ %s + %s", axis, paste(var.CL, collapse = "+")),
-                 responses = list(model_1 = "oMAP", model_2 = "oSYSTM", model_3 = "oDIASM",
-                                          model_4 = "oPULSEPRESSURE", model_5 = "HYPERTENSION"),
-                 filterstr = "Axis"))
-
-do.call(rbind, glm.pcoa) %>% 
-    select(response, term, estimate, std.error, p.value, conf.low, conf.high) %>%
-    dplyr::mutate(Qval = p.adjust(p.value, method="BH")) %>%
-    mutate_if(is.numeric, round, 3) %>%
-    arrange(Qval) %>%
-    kable
-
-```
-
 ```{r Write figure 2, include=FALSE}
@@
 dset.deseq.signf <- deseqresults(dds, names.dset)
-dset.deseq.signf %>% kable
+dset.deseq.signf %>% pull(qval) %>% length
 ```
@@
 
+```{r salt hypertension}
+ret.salt <- calculateglm(dset.genus.salt, covariates = c("NA.", var.CL), filterstr = "NA.")
+ret.salt %>% select(response, term, estimate, p.value) %>% kable
+```
+
 ```{r salt characteristics}
@@
 ```{r lacto deseq}
-dds.data.salt <- phyloseq_to_deseq2(pseq.genus.salt, ~ BL_AGE + SEX + NA.)
+#dds.data.salt <- phyloseq_to_deseq2(pseq.genus.salt, ~ BL_AGE + SEX + NA.)
+dds.data.salt <- phyloseq_to_deseq2(pseq.genus.salt, deseq.formula("NA.", var.CL))
 
@@
 
+```{r supplement table deseq}
+deseq.table <- deseqresults(dds, names.dset) %>%
+    mutate(lfc_se = sprintf("%.2f±%.2f", log2FoldChange, lfcSE),
+           p.value = pub.p(qval)) %>%
+    myspread %>%
+    select(Feature, starts_with("Systolic"), starts_with("Diastolic"), starts_with("PULSE PRESSURE"),
+           starts_with("Mean arterial pressure"), starts_with("Hypertension"))
+
+typology.tbls4 <- data.frame(
+    col_keys = colnames(deseq.table),
+    what = c("", rep("Systolic BP", 2), rep("Diastolic BP", 2),
+             rep("Pulse pressure", 2), rep("Mean arterial pressure", 2), rep("Hypertension", 2)),
+    measure = c("", rep(c("Log2FC±SE", "p"), 5)),
+    stringsAsFactors = FALSE)
+
+deseq.flextable <-
+    typologyformatter(data = deseq.table, font = 9, typology = typology.tbls4) %>%
+    flextable::width(j = 1, width = 1.2) %>%
+    flextable::width(j = seq(2, 11, 2), width = 0.75) %>%
+    flextable::width(j = seq(3, 11, 2), width = 0.55)
+
+deseqresults(dds, names.dset) %>% pull(qval) %>% max
+
+```
+
 ```{r Write supplement}
@@
   body_remove()
-writetable(doc.s, core.names.flextable, number = "S1", "Core genera", "")
-writetable(doc.s, alphadiversity.flextable, number = "S2", "Alpha diverstity", "")
-writetable(doc.s, betadiversity.flextable, number = "S3", "Beta diverstity", "")
+#writetable(doc.s, core.names.flextable, number = "S1", "Core genera", "")
+#writetable(doc.s, alphadiversity.flextable, number = "S2", "Alpha diverstity", "")
+#writetable(doc.s, betadiversity.flextable, number = "S3", "Beta diverstity", "")
+writetable(doc.s, deseq.flextable, number = "S4", "Deseq", "")
 print(doc.s, target = paste0("report/articletwo-supplement.docx"))

### 2019-10-16 articletwo-officer.R

@@ -173,3 +173,12 @@ text.normal <- fp_text(font.size = 12, font.family = "Arial")
     body_add_par(footer, style = "footnote text") %>%
     body_add_break(pos = "after")
 }
+
+myspread <- function(ret, list = c2l("lfc_se", "p.value"), term  = "Feature", key = "Name") {
+    lapply(list, function(x) ret %>%
+                             select(term, key, x) %>%
+                             spread(key, x) %>%
+                             rename_at(vars(-term), funs(paste0(., "_", x)))) %>%
+        Reduce(function(...) full_join(..., by = term), .) %>%
+        dplyr::select(term, noquote(order(colnames(.))))
+}

### 2019-10-16 articletwo-rrbiome.R

@@ -82,12 +82,15 @@ calculateglm <- function(dset,
                          responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
                                           model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
                          min_n_for_continuous = 10,
-                         modelstring = "%s ~ BL_AGE + SEX",
+                         covariates = c(),
                          filterstr = ".") {
     glmlist <- lapply(responses, function(response) {
+        vars <- covariates
+        if (response == "HYPERTENSION")
+            vars <- vars[!grepl("BL_USE", vars)]
         is.logistic <- length(unique(pull(dset, response))) < min_n_for_continuous
         fo.family <- ifelse(is.logistic, stats::binomial, stats::gaussian)
-        fo <- sprintf(modelstring, response)
+        fo <- sprintf("%s ~ %s", response, paste0(vars, collapse = "+"))
         stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
             broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = is.logistic) %>%
                 dplyr::filter(grepl(filterstr, term)) %>%
@@ -142,7 +145,10 @@ calculateadonis <- function(dset,
                             npermutations = 99,
                             maxcores = 100) {
     mclapply(responses, function(response) {
-        fo <- sprintf("matrix ~ %s + %s", paste(covariates, collapse = " + "), response)
+        vars <- covariates
+        if ("HYPERTENSION" %in% vars)
+            vars <- vars[!grepl("BL_USE", vars)]
+        fo <- sprintf("matrix ~ %s + %s", paste(vars, collapse = " + "), response)
         ad <- adonis(formula = as.formula(fo), data = dset, permutations = npermutations)
         ad$aov.tab %>%
             as.data.frame %>%
@@ -330,7 +336,7 @@ calculate.alphadiversity <- function(pseq, vars, modelstr = "%%s ~ %s + diversit
 
     lapply(vars, function(var)
         min = calculateglm(dset,
-                           modelstring = sprintf(modelstr, paste0(var, collapse = " + ")),
+                           covariates = var,
                            filterstr = "shannon"))
 
 }

### 2019-10-16 articletwo.Rmd
Medication with hypertension

@@
 
+dir.create("cache/", showWarnings = FALSE)
 dir.create("rds/", showWarnings = FALSE)
@@
 ```{r alphabeta definitions}
-alphadiversity <- calculate.alphadiversity(pseq.species,
-                                           list("max" = var.CL,
-                                                "min" = var.CL.min))
-
-diversities <- mergediversities(alphadiversity, adonis.species, names.dset = names.dset)
-```
-
-```{r bptreat numbers}
-pseq.species %>% meta %>% pull(BP_TREAT) %>% summary
+diversities <- diversities(pseq = pseq.species,
+                           vars = list("max" = var.CL, "min" = var.CL.min),
+                           betadiversity = adonis.species,
+                           names.dset = names.dset)
 ```
@@
 ```{r talbe alpha div min}
-alphadiversity$min %>%
-    select(response, estimate, p.value, conf.low, conf.high) %>%
-    mutate(mean_ci = sprintf("p=%s; 95%%-CI %.2f-%.2f",
-                              pub.p(p.value), conf.low, conf.high)) %>%
+diversities %>%
+    map_df(., ~as.data.frame(.x), .id = "covariates") %>%
     mutate_if(is.numeric, round, 3) %>%
-    arrange(p.value) %>%
-    kable
-```
-
-Results for full model of alpha diversity
-
-```{r talbe alpha div max}
-alphadiversity$max %>%
-    select(response, estimate, p.value, conf.low, conf.high) %>%
-    mutate(mean_ci = sprintf("p=%s; 95%%-CI %.2f-%.2f",
-                              pub.p(p.value), conf.low, conf.high)) %>%
-    mutate_if(is.numeric, round, 3) %>%
-    arrange(p.value) %>%
-    kable
-```
-
-Difference of R2 between full and minimal models
-
-```{r difference in beta diversity}
-merge(diversities$min %>%
-      select(response, R2, R2.p) %>%
-      dplyr::rename(R2.min = R2, p.min = R2.p),
-      diversities$max %>%
-      select(response, R2, R2.p) %>%
-      dplyr::rename(R2.full = R2, p.full = R2.p),
-      by = "response") %>%
-    mutate(R2.diff = R2.full - R2.min) %>%
     kable
@@
 
+General linear models for associations between dU-NA and hypertension indices
+
 ```{r salt hypertension}
@@
 dds.salt <- DESeq(dds.data.salt, test="Wald", fitType="parametric")
+
+dset.genus.salt %>% pull(NA.) %>% mean
 

### 2019-10-16 articletwo-ggplot.R

@@ -1,8 +1,11 @@
 plot.alpha <- function(diversity,
                        alphabreaks = c(-Inf, -1, -0.1, -0.01, 0)) {
     diversity <- diversity %>%
-        mutate(estimate_fac =  cut(estimate, breaks = alphabreaks))
-    plot.alpha.diversity <- ggplot(diversity, aes(x=reorder(Name, R2),
+        mutate(estimate = ifelse(response == "HYPERTENSION", log(alpha.effect), alpha.effect),
+               estimate_fac =  cut(estimate, breaks = alphabreaks),
+               Qstar = ifelse(alpha.p < 0.05, '*', ' '))
+
+    plot.alpha.diversity <- ggplot(diversity, aes(x = reorder(Name, deseqplotorder(Name)),
                                                   y = 1,
                                                   fill = estimate_fac)) +
         geom_bar(stat="identity", color="black") +
@@ -11,12 +14,6 @@ plot.alpha <- function(diversity,
                           name = 'Regression coefficient \nin linear model \nfor Shannon index',
                           drop=FALSE,
                           direction = -1) +
-        #scale_color_discrete_sequential(palette = "Blues", nmax = 6, order = 2:6) +
-#        scale_fill_gradientn(colours = c("blue", "white"),
-#                             breaks = c(-colorlimit, 0),
-#                             name = 'Regression coefficient \nin linear model \nfor Shannon index',
-#                             limits = c(-colorlimit, 0),
-#                             na.value = "black") +
         theme_classic(20) +
         geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=FALSE, color='black', size=20) +
         scale_shape_manual(name="",
@@ -57,13 +54,15 @@ plot.alpha <- function(diversity,
 }
 
 plot.beta <- function(diversity,
-                      fig.permar2.text = "R2 of variable \nin Bray-Curtis distance",
+                      fig.permar2.text = "Coefficient of determination\nfor beta diversity",
                       fig.textsize = 15,
                       ymax = 0.0016) {
-    diversity <- diversity %>% mutate(color = as.factor(ifelse(R2.p < 0.05, 1, 0)))
-    plot.beta.diversity <- ggplot(diversity, aes(x = reorder(Name, R2), y= R2, color = color)) +
-        geom_bar(stat="identity") +
[…]

### 2019-10-16 articletwo-rrbiome.R

@@ -67,7 +67,11 @@ meta.merge.alphadiversity <- function(pseq, index = "shannon") {
     alphadiversity  <- microbiome::alpha(pseq, index = index)
     base::merge(meta(pseq), alphadiversity, by=0, all=TRUE) %>%
         dplyr::rename(Sample_ID = Row.names) %>%
-            dplyr::mutate(BL_AGE_GROUP = group_age(BL_AGE))
+        dplyr::mutate(diversity_shannon = scale(diversity_shannon),
+                      MAP = oMAP,
+                      PULSEPRESSURE = oPULSEPRESSURE,
+                      SYSTM = oSYSTM,
+                      DIASM = oDIASM)
 }
 
 calculate.beta.matrix <- function(pseq) {
@@ -85,12 +89,9 @@ calculateglm <- function(dset,
                          covariates = c(),
                          filterstr = ".") {
     glmlist <- lapply(responses, function(response) {
-        vars <- covariates
-        if (response == "HYPERTENSION")
-            vars <- vars[!grepl("BL_USE", vars)]
         is.logistic <- length(unique(pull(dset, response))) < min_n_for_continuous
         fo.family <- ifelse(is.logistic, stats::binomial, stats::gaussian)
-        fo <- sprintf("%s ~ %s", response, paste0(vars, collapse = "+"))
+        fo <- sprintf("%s ~ %s", response, paste0(covariates, collapse = "+"))
         stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
             broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = is.logistic) %>%
                 dplyr::filter(grepl(filterstr, term)) %>%
@@ -145,10 +146,7 @@ calculateadonis <- function(dset,
                             npermutations = 99,
                             maxcores = 100) {
     mclapply(responses, function(response) {
-        vars <- covariates
-        if ("HYPERTENSION" %in% vars)
-            vars <- vars[!grepl("BL_USE", vars)]
-        fo <- sprintf("matrix ~ %s + %s", paste(vars, collapse = " + "), response)
+        fo <- sprintf("matrix ~ %s + %s", paste(covariates, collapse = " + "), response)
         ad <- adonis(formula = as.formula(fo), data = dset, permutations = npermutations)
         ad$aov.tab %>%
             as.data.frame %>%
@@ -212,22 +210,6 @@ vectortolist <- function(c) {
   l <- as.list(c)
[…]

### 2019-10-17 comparepseqs.Rmd
Duplicated line
Atlas changes

@@
 
-covariates.min <- c("BL_AGE", "SEX")
-covariates.full <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE",
-                         "Q57X", "BL_USE_RX_C03", "BL_USE_RX_C07",
-                         "BL_USE_RX_C08", "BL_USE_RX_C09")
+var.CL.min <- c("BL_AGE", "SEX")
+var.CL <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", 
+            "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
 ```
@@
 ```{R models}
-mclapply(c2l(names(rds)), function(key) {
+plots <- mclapply(c2l(names(rds)), function(key) {
     pseq.species <- import_filter_data(rds[[key]],
@@
 
-    alphadiversity <- calculate.alphadiversity(pseq.species,
-                                               list("max" = var.CL,
-                                                    "min" = var.CL.min))
-
     diversities <- mergediversities(alphadiversity, adonis.species, names.dset = names.dset)
@@
     ggsave(file = sprintf("cache/comparepseqs-%s.png", key), plot=g4, height=6, width=9)
+	g4
 }, mc.cores = length(rds))
 ```
+
+```{r save session}
+save.image(file = "session/comparepseq.Rdata")
+```

### 2019-10-17 articletwo-rrbiome.R

@@ -319,7 +319,7 @@ diversities <- function(pseq, vars, betadiversity, names.dset) {
                                      dplyr::filter(term %in% var.BP) %>%
                                      select(response, beta.R2 = R2, beta.p=`Pr(>F)`)))
         full_join(alphadiversity, betadiversity, by = "response") %>%
-            merge(names.dset %>% select(Covariate, Name), by.x = "response", by.y = "Covariate") %>%
+            merge(names.dset %>% select(Covariate, Name), by.x = "response", by.y = "Covariate")
     })
 }
 

### 2019-10-17 bin/reports.sh

@@ -1,13 +1,0 @@
-#!/bin/bash
-
-cd "${HOME}/phd/research/articletwo/reports" || exit 1
-HOSTPATH="utu:articletwo"
-FILE="rrbiome.html"
-BACKUP="rrbiome_$(date +%Y%m%d).html"
-
-RET=$(rsync -aEim $HOSTPATH/$FILE $FILE)
-
-if [ $? -eq 0 -a -n "${RET}" ]; then
-	cp $FILE $BACKUP
-	open $FILE
-fi

### 2019-10-18 articletwo.Rmd
Comparing how number of associations are changed in deseq

@@
 
+```{r comparing deseq2}
+if (!file.exists("rds/dds3.rds")) {
+    pseq.genus.core.deseq <-  prune_taxa(taxa(pseq.genus.core), pseq.genus)
+    dds3 <- lapply(deseq.list(var.CL, var.CL.min), function(x, dds) {
+        dds.data <- phyloseq_to_deseq2(dds, deseq.formula(x, "HYPERTENSION"))
+        DESeq(dds.data,
+              test="Wald",
+              fitType="parametric",
+              parallel = TRUE,
+              BPPARAM=MulticoreParam(16))}, dds = pseq.genus.core.deseq)
+    saveRDS(dds3, file = "rds/dds3.rds")
+} else {
+    dds3 <- readRDS(file = "rds/dds3.rds")
+}
+```
+
 ## Officer
@@
 diversities %>%
-    map_df(., ~as.data.frame(.x), .id = "covariates") %>%
+    map_df(~as.data.frame(.x), .id = "covariates") %>%
     mutate_if(is.numeric, round, 3) %>%
@@
 ## Direct associations between genus level and blood pressure variables (figure 3)
+
+Comparing how adding additional covariate to basic model for age, sex,
+and hypertension changes number of significant associations found for
+hypertension.
+
+```{r results compare deseq2}
+map(dds3, ~sum(results(.x, name = "HYPERTENSION_1_vs_0")$padj < 0.05)) %>%
+    map_df(~as.data.frame(.x) %>% dplyr::rename(qval = .x), .id = "included") %>%
+    kable
+```
 

### 2019-10-18 articletwo-ggplot.R

@@ -15,7 +15,7 @@ plot.alpha <- function(diversity,
                           drop=FALSE,
                           direction = -1) +
         theme_classic(20) +
-        geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=FALSE, color='black', size=20) +
+        geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=FALSE, color='black', size=14) +
         scale_shape_manual(name="",
                            values=c('*'='*', ' '=' '),
                            labels=c("*"="significant at\nFDR 0.05", ' '=' '),
@@ -54,22 +54,19 @@ plot.alpha <- function(diversity,
 }
 
 plot.beta <- function(diversity,
-                      fig.permar2.text = "Coefficient of determination\nfor beta diversity",
                       fig.textsize = 15,
                       ymax = 0.0016) {
     diversity <- diversity %>% mutate(Qstar = ifelse(beta.p < 0.05, "*", "-"))
     plot.beta.diversity <- ggplot(diversity,
                                   aes(x = reorder(Name, deseqplotorder(Name)), y = beta.R2)) +
-        geom_bar(stat="identity", color = "black", fill = "white") +
-        geom_point(aes(y = 0.0001, shape=Qstar), show.legend=FALSE, color='black', size=20) +
+        geom_bar(stat="identity", color = "black", fill = "gray96") +
+        geom_point(aes(y = beta.R2 - 0.00007, shape=Qstar), show.legend=FALSE, color='black', size=14) +
         scale_shape_manual(values=c('*'='*', '-'='')) +
         coord_flip() +
         theme_classic(fig.textsize) +
-        ylab(fig.permar2.text) +
+        ylab("Coefficient of determination for\nbeta diversity") +
         scale_y_continuous(limits = c(0, ymax)) +
-        theme(axis.text.y = element_text(color="black"),
-              axis.text.x = element_text(color="black"),
-              legend.position="right",
+        theme(legend.position="right",
               legend.direction="vertical",
               legend.box.margin=margin(0,0,0,0,"pt"),
               legend.margin=margin(0,0,0,0,"pt"),

### 2019-10-18 articletwo-rrbiome.R

@@ -377,3 +377,10 @@ mytableone <- function(dset, variables) {
 deseq.formula <- function(..., fo = "~ %s") {
     as.formula(sprintf(fo, paste0(c(...), collapse = "+")))
 }
+
+deseq.list <- function(var.CL, var.CL.min) {
+    l <- lapply(c2l(var.CL), function(x) {
+        unique(c(var.CL.min, x))
+    })
+    l[!duplicated(l)]
+}

### 2019-10-18 articletwo.Rmd
Diagnostic plot

@@
 ```{r lacto deseq}
-#dds.data.salt <- phyloseq_to_deseq2(pseq.genus.salt, ~ BL_AGE + SEX + NA.)
 dds.data.salt <- phyloseq_to_deseq2(pseq.genus.salt, deseq.formula("NA.", var.CL))
+dds.salt <- DESeq(dds.data.salt, test="Wald", fitType="parametric") 
+```
 
-dds.salt <- DESeq(dds.data.salt, test="Wald", fitType="parametric")
-
-dset.genus.salt %>% pull(NA.) %>% mean
-
-results(dds.salt, name = "NA.") %>%
+```{r lacto deseq results}
+dds.salt %>%
+    results(., name = "NA.") %>%
     as.data.frame %>%
@@
 ```
+
+```{r lacto deseq diagnostic}
+png("cache/salt-ma.png", width = 1200, height = 800, pointsize = 20)
+par(mfrow=c(1,2))
+plotMA(dds.salt)
+plotDispEsts(dds.salt)
+dev.off()
+```
+
+<img src = "cache/salt-ma.png" />
 

### 2019-10-21 articletwo.Rmd
Minor

@@
     facet_wrap(~key, scales = "free_y", strip.position = "left", labeller = axis_labeller) +
-    geom_point(size=0.1, alpha=0.4) +
+    geom_jitter(size=4, alpha=0.4, shape = ".") +
     xlab(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1])) +
@@
     theme_classic() +
-    guides(colour = guide_legend(override.aes = list(size=4))) +
+    guides(colour = guide_legend(override.aes = list(size=4, shape = 19))) +
     coord_cartesian(clip = 'off') +
-    theme(text = element_text(size=12),
+    theme(text = element_text(size=10),
           legend.position = c(0.35, 0.95),
@@
           aspect.ratio = 1,
-          strip.text = element_text(size = 14),
-          axis.title.x = element_text(size = 14))
+          strip.text = element_text(size = 12),
+          axis.title.x = element_text(size = 12))
 
-ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=3.4, width=6, units = "in", dpi = 300)
+ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=3.0, width=6, units = "in", dpi = 300)
 ```
@@
 
-First three axis explain portion (`r eig.fracs[1:3]`) of
+First three axis explain portion (`r sum(eig.fracs[1:3])`) of
 variation. Testing associations to blood pressure variables on PCoA
@@
 ```{r results compare deseq2}
-map(dds3, ~sum(results(.x, name = "HYPERTENSION_1_vs_0")$padj < 0.05)) %>%
+map(dds3, ~sum(results(.x, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)) %>%
     map_df(~as.data.frame(.x) %>% dplyr::rename(qval = .x), .id = "included") %>%
@@
 dset.deseq.signf <- deseqresults(dds, names.dset)
-dset.deseq.signf %>% pull(qval) %>% length
 ```
@@
 <img src = "cache/deseq.png" />
-
-```{r deseq abundances for boxplot}
-dset.deseq.boxplot <- pseq.genus %>%
-    microbiome::transform(., transform = "compositional") %>%
-    prune_taxa(unique(pull(dset.deseq.signf, Feature)), .) %>%
-    psmelt %>%
-    mutate(Abundance = sqrt(Abundance))
-```
-
-```{r deseq boxplot}
-g.boxplot <- dset.deseq.boxplot %>%
-    group_by(OTU) %>%
-    mutate(outlier = abs(Abundance) > median(Abundance) + IQR(Abundance) * 1.5) %>%
-    ggplot(aes(x = HYPERTENSION, y = Abundance, color = OTU)) +
-    geom_boxplot(lwd=0.1, outlier.shape = NA) +
-    geom_point(data = function(x) dplyr::filter_(x, ~ outlier),
-               position = position_jitterdodge(jitter.width = 0.02),
-               size = 0.1,
-               alpha = 0.4, shape = ".") +
-    scale_y_continuous(limits = c(0, 0.2), expand = c(0, 0)) +
-    scale_x_discrete(labels=c("0" = "Normotensive", "1" = "Hypertensive")) +
-    ylab("Abundance (Hellinger)") +
-    xlab("") +
-    theme_classic() +
-    guides(shape = guide_legend(override.aes = list(size = 0.1)),
-           color = guide_legend(override.aes = list(size = 0.1))) +
-    theme(legend.title = element_text(size = 4), 
-          legend.text  = element_text(size = 4),
-          legend.key.size = unit(1, "lines"),
-          legend.position="bottom")
-
-ggsave(file = "cache/dds-test.png", plot=g.boxplot, height=4, width=6, units = "in", dpi = 300)
-```
-
-<img src = "cache/dds-test.png" />
 

### 2019-10-21 articletwo.Rmd
Missing assingment

--- 2019-10-18 articletwo.Rmd
@@
 
-    lapply(c2l(var.BP), function(x) {
+    dds <- lapply(c2l(var.BP), function(x) {
         dds.data <- phyloseq_to_deseq2(pseq.genus.core.deseq, deseq.formula(x, var.CL))

### 2019-10-21 articletwo.Rmd
Remove table one

@@
     body_remove()
+
 dir.create("rds", showWarnings = FALSE)
@@
 
-```{r Write characteristics, include=FALSE}
+```{r Write characteristics, include=FALSE, eval = FALSE}
 tbl1head <- "Characteristics of the study sample."

### 2019-10-21 articletwo-officer.R

@@ -10,6 +10,7 @@ characteristics <- function(dset, tableone.names, tableone.factors, extras = lis
                       digits = 1,
                       pDigits = 3,
                       contDigits=1)
+
     tableone.fullnames <- c(tableone.names, extras)
     tablecsv %>%
         as.data.frame %>%
@@ -152,14 +153,14 @@ riskmodel.joincolumns <- function(rset, riskmodels, idforn = "fwdbonf_riskclass"
 }
 
 writetable <- function(doc, tbl, number = 0, head, foot) {
-	text.bold <- fp_text(font.size = 12, bold = TRUE, font.family = "Arial")
-	text.normal <- fp_text(font.size = 12, font.family = "Arial")
+    text.bold <- fp_text(font.size = 12, bold = TRUE, font.family = "Arial")
+    text.normal <- fp_text(font.size = 12, font.family = "Arial")
 
-  doc <- doc %>%
-    body_add_fpar(fpar(ftext(paste0("Table ", number, ". "), prop = text.bold), ftext(head, prop = text.normal)), style = "Table Caption") %>%
-    body_add_flextable(tbl, align = "left") %>%
-    body_add_par(foot, style = "footnote text") %>%
-    body_add_break(pos = "after")
+    doc <- doc %>%
+        body_add_fpar(fpar(ftext(paste0("Table ", number, ". "), prop = text.bold), ftext(head, prop = text.normal)), style = "Table Caption") %>%
+        body_add_flextable(tbl, align = "left") %>%
+        body_add_par(foot, style = "footnote text") %>%
+        body_add_break(pos = "after")
 }
 
 writeimage <- function(doc, number = 0, filename, header, footer, width = 7) {

### 2019-10-21 articletwo.Rmd
Minor

@@
 ```{r alphabeta definitions}
-diversities <- diversities(pseq = pseq.species,
+diversity <- diversities(pseq = pseq.species,
                            vars = list("max" = var.CL, "min" = var.CL.min),
@@
 ```{r talbe alpha div min}
-diversities %>%
+diversity %>%
     map_df(~as.data.frame(.x), .id = "covariates") %>%
@@
 ```{r save grob}
-g4 <- plot.diversities(diversities)
+g4 <- plot.diversities(diversity)
 ggsave(file = "cache/alpha-beta.png", plot=g4, height=6, width=9)
@@
 and hypertension changes number of significant associations found for
-hypertension.
+hypertension. Full model has
+
+`r sum(results(dds$HYPERTENSION, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)`
+
+associations for hypertension.
 
@@
 
+Listing our "core" genera
+
 ```{r write supplement}
@@
 
+Listing results for alpha diversity
+
 ```{r supplement table for alpha diversity}
-alphadiversity.table <- cbind(alphadiversity$min %>% dplyr::rename_all(paste0, ".min"),
-                                  alphadiversity$max %>% dplyr::rename_all(paste0, ".max")) %>%
-    dplyr::filter(model_name.min == model_name.max) %>%
-    merge(names.dset, by.x ="response.min", by.y = "Covariate") %>%
-    mutate(mean_ci.min = sprintf("%.2f (%.2f - %.2f)", estimate.min, conf.low.min, conf.high.min),
-           pubp.min = pub.p(p.value.min),
-           mean_ci.max = sprintf("%.2f (%.2f - %.2f)", estimate.max, conf.low.max, conf.high.max),
-           pubp.max = pub.p(p.value.max)) %>%
-    arrange(match(response.min, c("SYSTM", "DIASM", "MAP", "PULSEPRESSURE", "HT"))) %>%
-    select(Name, mean_ci.min, pubp.min, mean_ci.max, pubp.max)
+alphadiversity.table <- diversity %>% map_df(~as.data.frame(.x), .id = "model") %>%
+    mutate(mean_ci = sprintf("%.2f (%.2f - %.2f)", alpha.effect, alpha.low, alpha.high),
+           pubp = pub.p(alpha.p)) %>%
+    myspread(list = c2l("mean_ci", "pubp"), term = "Name", key = "model") %>%
+    arrange(match(Name, c("Systolic BP", "Diastolic BP", "Mean arterial pressure", "Pulse pressure", "Hypertension"))) %>%
+        select(Name, min_mean_ci, min_pubp, max_mean_ci, max_pubp)
 
@@
 ```{r supplement table for beta diversity}
-betadiversity.table <- merge(diversities$min %>% dplyr::rename_all(paste0, ".min"),
-      diversities$max %>% dplyr::rename_all(paste0, ".max")) %>%
-    dplyr::filter(response.min == response.max) %>%
-    merge(names.dset, by.x ="response.min", by.y = "Covariate") %>%
-    mutate(R2.min = sprintf("%.3f%%", R2.min*100),
-           pubp.min = pub.p(R2.p.min),
-           R2.max = sprintf("%.3f%%", R2.max*100),
-           pubp.max = pub.p(R2.p.max)) %>%
-    arrange(match(response.min, c("SYSTM", "DIASM", "MAP", "PULSEPRESSURE", "HT"))) %>%
-    select(Name, R2.min, pubp.min, R2.max, pubp.max)
+betadiversity.table <- diversity %>% map_df(~as.data.frame(.x), .id = "model") %>%
+    mutate(rsquared = sprintf("%.3f%%", beta.R2*100),
+           pubp = pub.p(beta.p)) %>%
+    myspread(list = c2l("rsquared", "pubp"), term = "Name", key = "model") %>%
+    arrange(match(Name, c("Systolic BP", "Diastolic BP", "Mean arterial pressure", "Pulse pressure", "Hypertension"))) %>%
+    select(Name, min_rsquared, min_pubp, max_rsquared, max_pubp)
 
@@
   body_remove()
-#writetable(doc.s, core.names.flextable, number = "S1", "Core genera", "")
-#writetable(doc.s, alphadiversity.flextable, number = "S2", "Alpha diverstity", "")
-#writetable(doc.s, betadiversity.flextable, number = "S3", "Beta diverstity", "")
+writetable(doc.s, core.names.flextable, number = "S1", "Core genera", "")
+writetable(doc.s, alphadiversity.flextable, number = "S2", "Alpha diverstity", "")
+writetable(doc.s, betadiversity.flextable, number = "S3", "Beta diverstity", "")
 writetable(doc.s, deseq.flextable, number = "S4", "Deseq", "")

### 2019-10-21 articletwo-rrbiome.R

@@ -313,7 +313,8 @@ diversities <- function(pseq, vars, betadiversity, names.dset) {
         alphadiversity <- calculateglm(meta.merge.alphadiversity(pseq),
                                        covariates = c("diversity_shannon", vars[[var]]),
                                        filterstr = "shannon") %>%
-            select(response, alpha.p = p.value, alpha.effect = estimate)
+            select(response, alpha.p = p.value, alpha.effect = estimate,
+                   alpha.low = conf.low, alpha.high = conf.high)
         betadiversity <- betadiversity[[var]] %>%
             map_df(., ~as.data.frame(.x %>%
                                      dplyr::filter(term %in% var.BP) %>%

### 2019-10-21 articletwo.Rmd
Disable supplement

@@
 
-```{r Write supplement}
+```{r Write supplement, eval = FALSE}
 doc.s <- read_docx(path = "style/articlestyle.docx") %>%

### 2019-10-21 articletwo.Rmd
Minor

--- 2019-10-21 articletwo.Rmd
@@
 tbl1 <- tableone(meta(pseq.species))
+tbl1
 ```
@@
     flextable::align(align = "left")
+core.names.flextable
 ```
@@
     flextable::width(j=c(3:3,5:5), width = 0.7)
+alphadiversity.flextable 
 ```
@@
 
+betadiversity.flextable
 ```
@@
 
-deseqresults(dds, names.dset) %>% pull(qval) %>% max
-
+deseq.flextable
 ```
 
-```{r Write supplement}
+```{r Write supplement, eval = FALSE}
 doc.s <- read_docx(path = "style/articlestyle.docx") %>%
@@
 writetable(doc.s, deseq.flextable, number = "S4", "Deseq", "")
+
 print(doc.s, target = paste0("report/articletwo-supplement.docx"))

### 2019-10-21 articletwo.Rmd
Minor

@@
 
+Number of significant associations `r sum(pull(dset.deseq.signf, qval) < 0.05)` for 
+`r pull(dset.deseq.signf, Feature)  %>% unique %>% length` features. Largest p.value is 
+`r max(pull(dset.deseq.signf, qval))`
+
 ```{r deseq heatmap}
@@
 ```{r lactobacillus prevalence}
-dset.genus.salt %>%
+prune_lactobacillus(pseq.genus.salt, drop = FALSE) %>%
     summarize(lacto_detected = sum(lacto.prevalent),

### 2019-10-21 articletwo-rrbiome.R

@@ -351,7 +351,7 @@ prune_lactobacillus <- function(pseq, transform = "compositional", drop = TRUE) 
         prune_taxa(c("g_Lactobacillus"), .) %>%
         psmelt %>%
         spread(OTU, Abundance) %>%
-        dplyr::mutate(lacto.prevalent = ifelse(g_Lactobacillus > 0.001, 1, 0),
+        dplyr::mutate(lacto.prevalent = ifelse(g_Lactobacillus > 0.1/100, 1, 0),
                       duna.present = ifelse(is.na(NA.), 0, 1)) %>%
         { if (drop == TRUE) dplyr::filter(., duna.present == 1) else . }
 }

### 2019-12-20 articletwo-ggplot.R
Fig 4 from deseq2

@@ -172,8 +172,8 @@ plot.diversities <- function(diversities) {
 
 deseqhaetmap <- function(df, legend.name = "Log Base Two\nFold Change") {
     ggplot(df,
-           aes(x = Feature,
-               y = reorder(Name, deseqplotorder(Name)),
+           aes(x = reorder(Name, deseqplotorder(Name)),
+               y = ordered(Feature, levels=rev(unique(Feature))),
                fill = log2FoldChange)) +
         geom_tile(aes(fill = 1)) +
         geom_tile(colour="white", size=0.25) +
@@ -182,40 +182,38 @@ deseqhaetmap <- function(df, legend.name = "Log Base Two\nFold Change") {
                              limits=c(-1, 1),
                              breaks = c(-1, -0.5, 0.5, 0, 1),
                              na.value="grey30") +
-        coord_fixed() +
         xlab("") +
         ylab("") +
         theme_classic() +  
         theme(axis.text.x = element_text(angle=90, size=10, hjust=1.0,vjust=0.5),
               axis.text.y = element_text(size=10),
-              legend.title=element_text(size=10),
-              legend.position = c(-0.09, -1))
+              legend.title = element_text(size=10),
+              legend.position = "right")
 }
 
 
 deseqplotorder <- function(x) {
     y <- seq(length(x))
-    y[x == "Systolic BP"] = 4
-    y[x == "Diastolic BP"] = 3
+    y[x == "Systolic BP"] = 0
+    y[x == "Diastolic BP"] = 1
     y[x == "Pulse pressure"] = 2
-    y[x == "Mean arterial pressure"] = 1
-    y[x == "Hypertension"] = 0
+    y[x == "MAP"] = 3
+    y[x == "Hypertension"] = 4
     y
 }
 
 saltplot <- function(dset, ymax = 2) {
     ggplot(data = dset, aes(x = NA., y = g_Lactobacillus)) +
         geom_point(aes(color = SEX), size = 0.05, position = "jitter") +
-        geom_line(data = lm_ribbon(dset)) +
-        geom_ribbon(data = lm_ribbon(dset), aes(ymin=conf.low, ymax=conf.high), alpha=0.15) +
+        geom_smooth(method='lm', formula= y~x, colour = "black", size = 0.4) +
         scale_x_continuous(limits=c(20, 205), expand = c(0, 0)) +
-        scale_y_continuous(limits=c(-1, ymax), expand = c(0, 0)) +
+        scale_y_log10(limits = c(100, 10000), expand = c(0, 0)) +
[…]

### 2019-12-20 articletwo-rrbiome.R

@@ -218,7 +218,7 @@ getdescriptions <- function() {
     bind_rows(names.dset,
               data.frame(Covariate = c("MAP", "PULSEPRESSURE", "HYPERTENSION", "SEX"),
                          Category = rep("Physical", 4),
-                         Name =c( "Mean arterial pressure", "Pulse pressure", "Hypertension", "Female"),
+                         Name = c("MAP", "Pulse pressure", "Hypertension", "Female"),
                          Desc = c("Mean arterial presure, 2./3.*DIASM + 1./3.*SYSTM)",
                                   "Pulse pressure, SYSTM - DIASM",
                                   "Hypertension, SYSTM >=140 or DIAS >= 90 or BP mediaction",
@@ -354,12 +354,6 @@ prune_lactobacillus <- function(pseq, transform = "compositional", drop = TRUE) 
         dplyr::mutate(lacto.prevalent = ifelse(g_Lactobacillus > 0.1/100, 1, 0),
                       duna.present = ifelse(is.na(NA.), 0, 1)) %>%
         { if (drop == TRUE) dplyr::filter(., duna.present == 1) else . }
-}
-
-lm_ribbon <- function(dset, model = "g_Lactobacillus ~ NA. + BL_AGE + SEX", term = "NA.", taxon = "g_Lactobacillus") {
-    ret <- lm(as.formula(model), data=dset)
-    ggpredict(ret, terms = term) %>%
-        dplyr::rename(!!term := x, !!taxon := predicted)
 }
 
 mytableone <- function(dset, variables) {

### 2019-12-20 articletwo.Rmd
Init

@@
 library(rvg)
-library(tableone)
+!library(tableone)
 library(scales)
@@
 library(png)
-library(pander)
 library(ggpubr)
@@
 library(purrr)
-library(pwr)
 library(gtable)
@@
 
-### Officer functions
+<details><summary>Functions</summary>
 
-<details>
-  <summary>Open/Close</summary>
-
-```{r Officer importer, code=readLines("articletwo-officer.R")}
-source("articletwo-officer.R")
+```{r Functions}
+sourcefiles <- c("articletwo-officer.R",
+                 "articletwo-rrbiome.R",
+                 "articletwo-ggplot.R")
 ```
 
-</details>
-
-### RR biome functions
-
-<details>
-  <summary>Open/Close</summary>
-
-```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
-source("articletwo-rrbiome.R")
+```{r import files, echo = FALSE}
+for (f in sourcefiles) {
+    source(f)
+}
 ```
 
-</details>
-
-### Plot functions
-
-<details>
-  <summary>Open/Close</summary>
-
-```{r Plot functions, code=readLines("articletwo-ggplot.R")}
-source("articletwo-ggplot.R")
+```{r embbed files, echo = FALSE}
+xfun::embed_files(c("rrnmr.Rmd", sourcefiles))
 ```
@@
 ```
+
+Species has average number of reads `r pseq.species %>% sample_sums %>% mean` and
+genus `r pseq.genus %>% sample_sums %>% mean`.
 
@@
 tbl1 <- tableone(meta(pseq.species))
-tbl1
 ```
@@
 and hypertension changes number of significant associations found for
-hypertension. Full model has
-
-`r sum(results(dds$HYPERTENSION, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)`
-
-associations for hypertension.
+hypertension.
 
@@
 g.deseq <- deseqhaetmap(dset.deseq.signf)
-ggsave(file = "cache/deseq.png", plot = g.deseq, height = 3, width = 12, unit = "in", dpi = 300)
+ggsave(file = "cache/deseq.png", plot = g.deseq, height = 8, width = 4, unit = "in", dpi = 300)
 ```
@@
 
-```{r salt phyloseq object}
-rownames.salt <- pseq.genus %>% meta %>% subset(!is.na(NA.)) %>% rownames
-pseq.genus.salt <- pseq.genus %>%
-    prune_samples(rownames.salt, .) %>%
-    prune_taxa(taxa(pseq.genus.core), .)
-dset.genus.salt <- prune_lactobacillus(pseq.genus.salt, transform = "clr")
-```
-
-General linear models for associations between dU-NA and hypertension indices
-
-```{r salt hypertension}
-ret.salt <- calculateglm(dset.genus.salt, covariates = c("NA.", var.CL), filterstr = "NA.")
-ret.salt %>% select(response, term, estimate, p.value) %>% kable
-```
-
-```{r salt characteristics}
-dset.genus.salt %>%
-    mytableone() %>%
-    kable
-```
-
-```{r lactobacillus prevalence}
-prune_lactobacillus(pseq.genus.salt, drop = FALSE) %>%
-    summarize(lacto_detected = sum(lacto.prevalent),
-              n = n(),
-              ratio = lacto_detected/n) %>%
-    kable
-```
-
-```{r lacto scatter plot}
-g.salt <- saltplot(dset.genus.salt)
-ggsave(file = "cache/gsalt.png", plot=g.salt, height=4, width=4, dpi = 300)
-```
-
-```{r Write figure 4, include=FALSE}
-fig4head <- "Associations between dietary salt, lactobacillus abundance, and odds for hypertension."
-fig4foot <- "In panel A we see lactobacillus abundances in different dietary salt groups.
-The mean value is adjusted using age and sex.
-In panel B we split participants in four groups by lactobacillus abundances and calculate odds ratio for hypertension.
-Age and sex are included in logistic model.
-A 95 % CI is drawn in both panels."
-writeimage(doc, 4, "cache/gsalt.png", fig4head, fig4foot)
-```
-
-Deseq for lactobacillus
-
 ```{r lacto deseq}
@@
 
-```{r lacto deseq diagnostic}
-png("cache/salt-ma.png", width = 1200, height = 800, pointsize = 20)
-par(mfrow=c(1,2))
-plotMA(dds.salt)
-plotDispEsts(dds.salt)
-dev.off()
+```{r scatter plot from deseq}
+df.mu <- assays(dds.salt)[["mu"]] %>%
+    as.data.frame %>%
+    tibble::rownames_to_column("rowname") %>%
+    gather(sampleid, value, -rowname) %>%
+    spread(rowname, value)
+
+df.countmu <- full_join(meta(pseq.genus.salt) %>% tibble::rownames_to_column("sampleid"),
+                        df.mu %>% select(sampleid, g_Lactobacillus),
+                        by = "sampleid")
+
+g.salt <- saltplot(df.countmu)
+ggsave(file = "cache/gsalt.png", plot=g.salt, height=4, width=4, dpi = 300)
 ```
 
-<img src = "cache/salt-ma.png" />
+```{r Write figure 4, include=FALSE}
+fig4head <- "Associations between dietary salt, lactobacillus abundance, and odds for hypertension."
+fig4foot <- "In panel A we see lactobacillus abundances in different dietary salt groups.
+The mean value is adjusted using age and sex.
+In panel B we split participants in four groups by lactobacillus abundances and calculate odds ratio for hypertension.
+Age and sex are included in logistic model.
+A 95 % CI is drawn in both panels."
+writeimage(doc, 4, "cache/gsalt.png", fig4head, fig4foot)
+```
 

### 2019-12-20 articletwo-ggplot.R

@@ -1,0 +1,222 @@
+plot.alpha <- function(diversity,
+                       alphabreaks = c(-Inf, -1, -0.1, -0.01, 0)) {
+    diversity <- diversity %>%
+        mutate(estimate = ifelse(response == "HYPERTENSION", log(alpha.effect), alpha.effect),
+               estimate_fac =  cut(estimate, breaks = alphabreaks),
+               Qstar = ifelse(alpha.p < 0.05, '*', ' '))
+
+    plot.alpha.diversity <- ggplot(diversity, aes(x = reorder(Name, deseqplotorder(Name)),
+                                                  y = 1,
+                                                  fill = estimate_fac)) +
+        geom_bar(stat="identity", color="black") +
+        coord_flip() +
+        scale_fill_brewer(palette="Blues",
+                          name = 'Regression coefficient \nin linear model \nfor Shannon index',
+                          drop=FALSE,
+                          direction = -1) +
+        theme_classic(20) +
+        geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=FALSE, color='black', size=14) +
+        scale_shape_manual(name="",
+                           values=c('*'='*', ' '=' '),
+                           labels=c("*"="significant at\nFDR 0.05", ' '=' '),
+                           breaks=c("*", ' ')) +
+        xlab("") +
+        ylab("") +
+        theme(legend.position="right",
+              legend.title=element_text(size = 14),
+              legend.direction="vertical",
+              legend.text=element_text(size = floor(0.75*20)),
+              legend.box.margin=margin(t = 0, unit = "cm"),
+              legend.margin=margin(0,0,0,0,"pt"),
+              legend.justification="left",
+              axis.line.y = element_line(linetype="blank"),
+              axis.line.x = element_line(linetype="blank"),
+              legend.title.align = 0,
+              legend.text.align = 0,
+              axis.text.y = element_text(color="black"),
+              axis.text.x = element_text(color="black"))+
+        scale_y_continuous(breaks=c(0.5), labels=c(expression(alpha)))
+
+    plot.alpha.diversity.gtable <- ggplot_gtable(ggplot_build(plot.alpha.diversity))
+    plot.alpha.diversity.legend <- legend_extractor(plot.alpha.diversity.gtable)
[…]

### 2019-12-20 articletwo-officer.R

@@ -1,0 +1,185 @@
+characteristics <- function(dset, tableone.names, tableone.factors, extras = list()) {
+    title <- "Characteristics"
+    overall <- paste0("Cases, n=", dim(dset)[1])
+    tableobject <- tableone::CreateTableOne(data = dset, vars = names(tableone.names), factorVars = tableone.factors)
+    tablecsv <- print(tableobject,
+                      exact = "stage",
+                      quote = FALSE,
+                      noSpaces = TRUE,
+                      printToggle = FALSE,
+                      digits = 1,
+                      pDigits = 3,
+                      contDigits=1)
+
+    tableone.fullnames <- c(tableone.names, extras)
+    tablecsv %>%
+        as.data.frame %>%
+        tibble::rownames_to_column(var = "rowname") %>%
+        dplyr::filter(row_number() > 1) %>%
+        format(justify = "left", trim = TRUE) %>%
+        rowwise() %>%
+        mutate(id = gsub("^ *([A-Za-z_0-9]+).*", "\\1", rowname)) %>%
+        mutate(present = id %in%  names(tableone.fullnames)) %>%
+        mutate(rowname = ifelse(present == TRUE, tableone.fullnames[[id]], rowname)) %>%
+        select(rowname, Overall)
+}
+
+tableone <- function(dset) {
+
+    tableone.names <- list("BL_AGE" = "Age, y (SD)",
+                           "SEX" = "Female, N (%)",
+                           "BMI" = "BMI, kg/m² (SD)",
+                           "oSYSTM" = "Systolic BP, mmHg (SD)",
+                           "oDIASM" = "Diastolic BP, mmHg (SD)",
+                           "oPULSEPRESSURE" = "Pulse pressure, mmHg (SD)",
+                           "oMAP" = "Mean arterial pressure, mmHg (SD)",
+                           "HYPERTENSION" = "Hypertension, N (%)",
+                           "CURR_SMOKE" = "Current smoker, N (%)",
+                           "PREVAL_DIAB" = "Diabetes mellitus, N (%)",
+                           "Q57X" = "Exercise, N (%)",
+                           "ANYDRUG" = "Antihypertensive medication, N (%)",
+                           "BL_USE_RX_C03" = "  Diuretics, N (%)",
+                           "BL_USE_RX_C07" = "  Beta blockers, N (%)",
[…]

### 2019-12-20 articletwo-rrbiome.R

@@ -1,0 +1,381 @@
+# Helper functions for articletwo
+
+replace.brackets <- function (genus) gsub('(.*) \\(+(.+)\\)', '\\1.\\2', genus)
+
+rename.genus  <- function (genus, markword = "Plasmid", mark = "*") {
+    name <- gsub('(.*) \\(.*\\)', '\\1', genus)
+    star <- ifelse(grepl(markword, genus), mark, "")
+    paste0(gsub("_", " ", name), star)
+}
+
+myscale <- function(x){
+  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
+}
+
+filter.phenotype.data <- function(pseq,
+                                  included = c("MEN",
+                                               "BL_AGE",
+                                               "SYSTM",
+                                               "DIASM",
+                                               "BMI",
+                                               "CURR_SMOKE",
+                                               "PREVAL_DIAB",
+                                               "Q57X",
+                                               "BL_USE_RX_C09",
+                                               "BL_USE_RX_C03",
+                                               "BL_USE_RX_C07",
+                                               "BL_USE_RX_C08",
+                                               "NA.",
+                                               "BP_TREAT",
+                                               "PREVAL_CHD",
+                                               "PREVAL_CR_ANYCANC"),
+                                  allowna = c("NA.",
+                                              "BP_TREAT",
+                                              "PREVAL_CHD",
+                                              "PREVAL_CR_ANYCANC")) {
+    meta(pseq) %>%
+        tibble::rownames_to_column(var = "rowname") %>%
+        dplyr::select(rowname, included) %>%
+        tidyr::drop_na(myin(included, allowna, complement = TRUE)) %>%
+        dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
+                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
+                      ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
[…]

### 2020-01-08 articletwo.Rmd
Back to SHOGUN

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis',
+knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, results='asis',
                       cache=FALSE, warning=FALSE)
@@
 
-```{r data}
-pseq.genus <- import_filter_data("data/phfinrisk_centrifuge_genus_d50k_2019-06-02.rds")
-pseq.species <- import_filter_data("data/phfinrisk_centrifuge_species_all_drop50k_2018-11-30_2019-06-03.RDs")
+```{r data
+pseq.genus <- import_filter_data("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
+pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
 ```

### 2020-01-08 articletwo.Rmd
subtitles

@@
 ```{r data
+pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
 pseq.genus <- import_filter_data("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
-pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
+pseq.genus.coretaxa <- coretaxa(pseq.genus, detection = 0.1/100, prevalence = 1/100)
 ```
@@
 Species has average number of reads `r pseq.species %>% sample_sums %>% mean` and
-genus `r pseq.genus %>% sample_sums %>% mean`.
+genus `r pseq.genus %>% sample_sums %>% mean`. Core has length 
+`r length(pseq.genus.coretaxa)`.
 
@@
 `r ntaxa(pseq.genus)`.
-
-```{r defining genus core}
-pseq.genus.core <- microbiome::transform(pseq.genus, "compositional") %>%
-    core(detection = 0.1/100, prevalence = 1/100)
-```
-
-At genus-core level meta has dimensions (`r dim(meta(pseq.genus.core))`) 
-and there ntaxa is `r ntaxa(pseq.genus.core)`.
 
@@
 
+### Bray curtis distance matrix
+
 ```{r matrix calculation}
@@
 ```
+
+### Beta diversity
 
@@
 
+### PCoA
+
 ```{r pcoa calculate}
@@
 
+### DeSeq2
+
 ```{r deseq2}
 if (!file.exists("rds/dds.rds")) {
-    pseq.genus.core.deseq <-  prune_taxa(taxa(pseq.genus.core), pseq.genus)
+    pseq.genus.core.deseq <-  prune_taxa(pseq.genus.coretaxa, pseq.genus)
 
@@
               parallel = TRUE,
-              BPPARAM=MulticoreParam(16))})
+              BPPARAM=MulticoreParam(16))
+    })
 
@@
 if (!file.exists("rds/dds3.rds")) {
-    pseq.genus.core.deseq <-  prune_taxa(taxa(pseq.genus.core), pseq.genus)
+    pseq.genus.core.deseq <-  prune_taxa(pseq.genus.coretaxa, pseq.genus)
     dds3 <- lapply(deseq.list(var.CL, var.CL.min), function(x, dds) {

### 2020-01-08 articletwo-rrbiome.R

@@ -53,7 +53,7 @@ filter.phenotype.data <- function(pseq,
                       PULSEPRESSURE = myscale(PULSEPRESSURE),
                       MAP = myscale(MAP),
                       SEX = as.factor(SEX),
-                      Q57X = as.factor(Q57X),
+                      Q57X = factor(Q57X, ordered = FALSE),
                       BL_USE_RX_C03 = as.factor(BL_USE_RX_C03),
                       BL_USE_RX_C07 = as.factor(BL_USE_RX_C07),
                       BL_USE_RX_C08 = as.factor(BL_USE_RX_C08),
@@ -378,4 +378,10 @@ deseq.list <- function(var.CL, var.CL.min) {
         unique(c(var.CL.min, x))
     })
     l[!duplicated(l)]
+}
+
+coretaxa <- function(pseq, detection = 0.1/100, prevalence = 1/100) {
+    microbiome::transform(pseq, "compositional") %>%
+        core(detection = detection, prevalence = prevalence) %>%
+        taxa
 }
